### 2018-11-24 manuscript/manuscript.Rmd
initial commit

---
title             : "Plant life history strategies predicted by satellite-detected drought"
shorttitle        : ""

author: 
  - name          : "J Grey Author"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "307 University Ave, Fort Collins, CO 80523"
    email         : "monroejg@colostate.edu"
  - name          : "Brian Gill"
    affiliation   : "3"
  - name          : "Brian Gill"
    affiliation   : "4"

affiliation:
  - id            : "1"
    institution   : "Graduate Degree Program in Ecology, Colorado State University"
  - id            : "2"
    institution   : "College of Agriculture, Colorado State University"
  - id            : "3"
    institution   : "Department of Ecology and Evolutionary Biology"
  - id            : "4"
    institution   : "Biology Department, Penn State University"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

abstract: |

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "drought, life history, remote sensing, phylogeography"
wordcount         : ""

bibliography      : ["r-references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
figsintext        : yes

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library(raster)
library(tidyverse)
library(ggplot2)
```

#Introduction

Life history 
Annual vs perennial
Hypotheses
Summary

# Methods
## Data
### Life history
### Phyologeny
### Herbarium Specimens
```{r raw_GBIF}
GBIF<-read.csv("../data/0006178-160311141623029/occurrence.csv")
```
We downloaded `r nrow(GBIF)` records from GBIF.

### Drought

## Analyses
### Phylogeny
### Drought frequency
### Contrast
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

(ref:foo) A scatterplot of the data `cars` using **base** R graphics.

```{r foo, fig.cap='(ref:foo)',out.width = "\\textwidth", fig.pos = "!h"}
plot(cars)  # a scatterplot
```

First we did this (Fig. \@ref(fig:foo).)

# Discussion
The best paper ever [@monroe2016adaptation]

There is a package installer at pandoc’s download page. If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page. If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page. If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page. If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page. If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page. If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

### 2018-11-24 manuscript/manuscript.tex
initial commit

\documentclass[man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Plant life history strategies predicted by satellite-detected drought},
            pdfauthor={J Grey Author, Brian Gill, \& Brian Gill},
            pdfkeywords={drought, life history, remote sensing, phylogeography},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Plant life history strategies predicted by satellite-detected drought}
    \author{J Grey Author\textsuperscript{1,2}, Brian Gill\textsuperscript{3}, \&
Brian Gill\textsuperscript{4}}
    \date{}

\shorttitle{}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University\\\textsuperscript{2} College of Agriculture, Colorado State University\\\textsuperscript{3} Department of Ecology and Evolutionary Biology\\\textsuperscript{4} Biology Department, Penn State University}
\keywords{drought, life history, remote sensing, phylogeography\newline\indent Word count: }
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\usepackage{lineno}

\linenumbers

\authornote{Add complete departmental affiliations for each
author here. Each new line herein must be indented, like this line.

Correspondence concerning this article should be addressed to J Grey
Author, 307 University Ave, Fort Collins, CO 80523. E-mail:
\href{mailto:monroejg@colostate.edu}{\nolinkurl{monroejg@colostate.edu}}}

\abstract{

}

\begin{document}
\maketitle

\#Introduction

Life history Annual vs perennial Hypotheses Summary

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

\hypertarget{life-history}{%
\subsubsection{Life history}\label{life-history}}

\hypertarget{phyologeny}{%
\subsubsection{Phyologeny}\label{phyologeny}}

\hypertarget{herbarium-specimens}{%
\subsubsection{Herbarium Specimens}\label{herbarium-specimens}}

We downloaded 8670 records from GBIF. (Fig. @ref(tab:raw\_GBIF).)

\hypertarget{drought}{%
\subsubsection{Drought}\label{drought}}

\hypertarget{analyses}{%
\subsection{Analyses}\label{analyses}}

\hypertarget{phylogeny}{%
\subsubsection{Phylogeny}\label{phylogeny}}

\hypertarget{drought-frequency}{%
\subsubsection{Drought frequency}\label{drought-frequency}}

\hypertarget{contrast}{%
\subsubsection{Contrast}\label{contrast}}

We used R (Version 3.5.1; R Core Team, 2018) and the R-packages
\emph{dplyr} (Version 0.7.8; Wickham et al., 2018), \emph{forcats}
(Version 0.3.0; Wickham, 2018a), \emph{ggplot2} (Version 3.1.0; Wickham,
2016), \emph{papaja} (Version 0.1.0.9842; Aust \& Barth, 2018),
\emph{purrr} (Version 0.2.5; Henry \& Wickham, 2018), \emph{raster}
(Version 2.8.4; Hijmans, 2018), \emph{readr} (Version 1.1.1; Wickham et
al., 2017), \emph{sp} (Version 1.3.1; Pebesma \& Bivand, 2005),
\emph{stringr} (Version 1.3.1; Wickham, 2018b), \emph{tibble} (Version
1.4.2; Müller \& Wickham, 2018), \emph{tidyr} (Version 0.8.2; Wickham \&
Henry, 2018), and \emph{tidyverse} (Version 1.2.1; Wickham, 2017) for
all our analyses.

\hypertarget{results}{%
\section{Results}\label{results}}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{manuscript_files/figure-latex/foo-1} \caption{A scatterplot of the data \texttt{cars} using \textbf{base} R
graphics.}\label{fig:foo}
\end{figure}

First we did this (Fig. \ref{fig:foo}.)

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The best paper ever (Monroe et al., 2016)

There is a package installer at pandoc's download page. If you later
want to uninstall the package, you can do so by downloading this script
and running it with perl uninstall-pandoc.pl.There is a package
installer at pandoc's download page. If you later want to uninstall the
package, you can do so by downloading this script and running it with
perl uninstall-pandoc.pl.There is a package installer at pandoc's
download page. If you later want to uninstall the package, you can do so
by downloading this script and running it with perl
uninstall-pandoc.pl.There is a package installer at pandoc's download
page. If you later want to uninstall the package, you can do so by
downloading this script and running it with perl
uninstall-pandoc.pl.There is a package installer at pandoc's download
page. If you later want to uninstall the package, you can do so by
downloading this script and running it with perl
uninstall-pandoc.pl.There is a package installer at pandoc's download
page. If you later want to uninstall the package, you can do so by
downloading this script and running it with perl uninstall-pandoc.pl.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-R-papaja}{}%
Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with
R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}

\leavevmode\hypertarget{ref-R-purrr}{}%
Henry, L., \& Wickham, H. (2018). \emph{Purrr: Functional programming
tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}

\leavevmode\hypertarget{ref-R-raster}{}%
Hijmans, R. J. (2018). \emph{Raster: Geographic data analysis and
modeling}. Retrieved from
\url{https://CRAN.R-project.org/package=raster}

\leavevmode\hypertarget{ref-monroe2016adaptation}{}%
Monroe, J. G., McGovern, C., Lasky, J. R., Grogan, K., Beck, J., \&
McKay, J. K. (2016). Adaptation to warmer climates by parallel
functional evolution of cbf genes in arabidopsis thaliana.
\emph{Molecular Ecology}, \emph{25}(15), 3632--3644.

\leavevmode\hypertarget{ref-R-tibble}{}%
Müller, K., \& Wickham, H. (2018). \emph{Tibble: Simple data frames}.
Retrieved from \url{https://CRAN.R-project.org/package=tibble}

\leavevmode\hypertarget{ref-R-sp}{}%
Pebesma, E. J., \& Bivand, R. S. (2005). Classes and methods for spatial
data in R. \emph{R News}, \emph{5}(2), 9--13. Retrieved from
\url{https://CRAN.R-project.org/doc/Rnews/}

\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2018). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\leavevmode\hypertarget{ref-R-ggplot2}{}%
Wickham, H. (2016). \emph{Ggplot2: Elegant graphics for data analysis}.
Springer-Verlag New York. Retrieved from \url{http://ggplot2.org}

\leavevmode\hypertarget{ref-R-tidyverse}{}%
Wickham, H. (2017). \emph{Tidyverse: Easily install and load the
'tidyverse'}. Retrieved from
\url{https://CRAN.R-project.org/package=tidyverse}

\leavevmode\hypertarget{ref-R-forcats}{}%
Wickham, H. (2018a). \emph{Forcats: Tools for working with categorical
variables (factors)}. Retrieved from
\url{https://CRAN.R-project.org/package=forcats}

\leavevmode\hypertarget{ref-R-stringr}{}%
Wickham, H. (2018b). \emph{Stringr: Simple, consistent wrappers for
common string operations}. Retrieved from
\url{https://CRAN.R-project.org/package=stringr}

\leavevmode\hypertarget{ref-R-dplyr}{}%
Wickham, H., François, R., Henry, L., \& Müller, K. (2018). \emph{Dplyr:
A grammar of data manipulation}. Retrieved from
\url{https://CRAN.R-project.org/package=dplyr}

\leavevmode\hypertarget{ref-R-tidyr}{}%
Wickham, H., \& Henry, L. (2018). \emph{Tidyr: Easily tidy data with
'spread()' and 'gather()' functions}. Retrieved from
\url{https://CRAN.R-project.org/package=tidyr}

\leavevmode\hypertarget{ref-R-readr}{}%
Wickham, H., Hester, J., \& Francois, R. (2017). \emph{Readr: Read
rectangular text data}. Retrieved from
\url{https://CRAN.R-project.org/package=readr}

\endgroup

\end{document}

### 2018-11-24 src/GBIF_records_drought_frequency.R

@@ -1,0 +1,396 @@
+#Heliophila data cleaning
+#9/6/2018
+
+#accompanies/partially replicates JGM script "GBIF_records_drought_frequency.R" rows 1-19
+#things to clean/check:
+#basis of specimen
+#names through TNRS
+#duplicates?
+
+library(raster)
+library(tidyverse)
+library(ggplot2)
+#library(plotly)
+#library(rgbif) #0.9.8
+#library(taxize)
+# library(dplyr)
+
+# Read occurance data 
+GBIF<-read.csv("data/0006178-160311141623029/occurrence.csv");cat("\nrows in raw GBIF data file:", nrow(GBIF))
+# GBIF<-read.csv("Heliophila_occurrence.csv");cat("\nrows in raw GBIF data file:", nrow(GBIF), file = "analysis log")
+# summary(GBIF$species)
+#261 rows have blanks for species, NA for speciesKey
+# colnames(GBIF)
+
+# select species with life history data 
+species_life_history<-read.csv("data/Heliophila_life_history.csv")
+species_life_history$species<-gsub("_", " ", species_life_history$species)
+# unique(species_life_history$species)#42
+species_life_history$species <- as.factor(species_life_history$species)
+# levels(species_life_history$species) #42
+
+# class(GBIF$species)
+# [1] "factor"
+# class(species_life_history$species)
+# [1] "factor"
+GBIF_JGMcleaned<-subset(GBIF, species %in% species_life_history$species);cat("\nrows in raw GBIF data after choosing species with life history JGM:", nrow(GBIF_JGMcleaned), file = "analysis log", append=T)
+#6634
+GBIF_JGMcleaned<-merge(GBIF_JGMcleaned, species_life_history)
+
+####check names through tnrs, see how JGM cleaned and tnrs cleaned lists vary####
+# # RESULT: original JGM cleaned list fine for this stage
+# tnrsnames <- sapply(species_life_history$species, function(x) name_suggest(x)$key[1], USE.NAMES=FALSE)
+# unique(tnrsnames)#42
+# class(tnrsnames)#integer
+# class(GBIF$speciesKey)#integer
+# 
[…]

### 2018-11-24 manuscript/manuscript.Rmd
initial commit

@@
 title             : "Plant life history strategies predicted by satellite-detected drought"
-shorttitle        : ""
+shorttitle        : "Drought and life history"
 
 author: 
-  - name          : "J Grey Author"
+  - name          : "J Grey Monroe"
     affiliation   : "1,2"
@@
     affiliation   : "3"
-  - name          : "Brian Gill"
+  - name          : "Kathryn Turner"
     affiliation   : "4"
@@
   - id            : "1"
-    institution   : "Graduate Degree Program in Ecology, Colorado State University"
+    institution   : "Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO"
   - id            : "2"
-    institution   : "College of Agriculture, Colorado State University"
+    institution   : "College of Agriculture, Colorado State University, Fort Collins, CO"
   - id            : "3"
-    institution   : "Department of Ecology and Evolutionary Biology"
+    institution   : "Department of Ecology and Evolutionary Biology, Brown University, Provience, RI"
   - id            : "4"
-    institution   : "Biology Department, Penn State University"
+    institution   : "Biology Department, Pennsylvania State University, State College, PA"
 
-authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
+abstract: Identifying the environmental factors that predict the evolution of annual and perennial life history strategies in plants is important for understanding ecosystem functioning, perennial cropping systems, and is a long-standing goal of evolutionary ecology.
+A classic hypothesis is that annual and perennial strategies reflect adaptation to environments that differ in relative drought frequency.
+We test this hypothesis in Heliophila (Brassicaceae), a genus of flowering plants native to Southern Africa using herbarium occurrence records and satellite-detected drought histories.
+We find that perennial Heliophila species are observed in environments where droughts are significantly less frequent compared to annuals, especially with regards to summer drought.
+These correlations remain predictive while controlling for phylogeny, lending support to the hypothesis that drought associated natural selection has shaped differences in the distributions of these strategies.
+Additionally, the collection dates of herbarium records are consistent with a scenario in which annual species escape drought prone seasons.
+Together these results point to a model of evolution that supports classical hypotheses favoring perenniality in environments that experience less drought, compared to annuals which escape drought prone seasons as seed.
+Thus, changing drought regimens during the Anthropocene may threaten locally adapted species.
 
-abstract: |
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "drought, life history, remote sensing, phylogeography"
+keywords          : "drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records"
 wordcount         : ""
@@
 library(ggplot2)
+library(logistf)
+library(gee) 
+library(phytools) 
+library(geiger) 
+library(ape) 
+library(caper) 
+library(MCMCglmm) 
+library(phylolm)
+library(logistf)
 ```
 
-#Introduction
+# Introduction
 
-Life history 
-Annual vs perennial
-Hypotheses
-Summary
+## Life history
+
+Plants exhibit extraordinary diversity in life history strategies, from herbaceous species that complete the entire seed to seed life cycle in a number of weeks {Guo, 2012 #73} to trees that live for thousands of years {Brown 1996 http://www.rmtrr.org/oldlist.htm }.
+Along this continuum an important division exists, distinguishing annuals which complete their seed to seed life cycle within a single calendar year from perennials which can persist over multiple years and therefore experience the entire range of seasonal conditions.
+The ecological factors that explain the evolution of these alternative strategies remains poorly resolved {Friedman, 2015 #523}.
+Understanding the drivers of selection for these alternative strategies is important because annuals and perennials have differing impacts on ecosystem functioning, such as higher nitrogen concentration in perennials and larger specific root length in annuals that affect nutrient cycling {Garnier, 1997 #280}{Roumet, 2006 #738}.
+Furthermore, predicting the conditions that favor these strategies in nature is useful for developing more sustainable perennial cropping systems {Lelievre and Volaire, 2009}.
+
+## Hypotheses
+
+Seasonal drought conditions can represent a severe challenge to the persistence of individual organisms.
+Well-established hypotheses suggest that an annual life history is adaptive in environments where frequent droughts makes an escape from stressful drought conditions in the form of seed advantageous {Stearns 1992}{Silvertown and Charlesworth, 2001}.
+These hypotheses are based on models of life history evolution in which short life cycles are favored in unpredictable environments or those with frequent stressful conditions.
+Plants are expected to be more likely to survive a drought that cooccurs with the seed phase of a plant’s life cycle in which they are most resistant to desiccation.
+Previous efforts to address this hypothesis have yielded mixed or qualitative results.
+Transitions to annuality in Oenothera were associated with warmer summers and drier winters, but not with increased drought directly {Evans, 2005 #260}.
+Annual and perennial species of Nemesia were qualitatively associated with winter rather and summer rainfall environments respectively {Datson, 2008}.
+Similarly, annual species of Scorzoneroides were associated with environments classified as “unpredictable” {Cruz-Mazo, 2009 #598}.
+Annual and perennial varieties of the wild rice species Oryza rufipogon were observed more frequently in dry and humid climates, respectively {Morishima et al.
+1984}.
+While these reports suggest the prominence of annuals in environments considered more arid, a quantitative assessment of how historic drought frequency may favor annual or perennial life history strategies remains lacking.
+
+## Summary  
+Here we combine a long-term dataset of global imaging with metadata from natural history collections to test classic hypotheses about the evolution of life history strategies within the African endemic mustard genus, Heliophila L.
+(Brassicaceae).
+If annual species temporally escape drought as seed, then drought frequency should be an important determinant of the distribution of life history strategies across the landscape, and annual species should be more commonly associated with drought prone regions than perennial species.
+Furthermore, if annual species have adapted to escape drought prone seasons as seeds, observations of annual species should be rare during drought prone seasons.
+Phylogenetic relatedness can have significant non-random effects on species distributions and life history traits (cite), and therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetically controlled background.
 
@@
 ## Data
+### Drought
+```{r Droughtdata}
+Winter<-raster("../data/drought/Summer_drought_freq")
+Spring<-raster("../data/drought/Fall_drought_freq")
+Summer<-raster("../data/drought/Winter_drought_freq")
+Fall<-raster("../data/drought/Spring_drought_freq")
+```
+
+Remotely sensed data is a valuable tool for characterizing seasonal patterns in drought {AghaKouchak 2015}.
+With these data, we can exploit the fact that satellite imagery can detect reductions in plant cover and moisture at landscape scales that are characteristic of drought conditions.
+Unlike weather station data, which is limited in geographic coverage, and manual observation, which is limited in scope, satellite detected data provides a long term global historical record of drought events at fine temporal and spatial resolutions.
+One such measure, the Vegetative Health Index (VHI), has been collected since 1981 by NOAA AVHRR satellites and is a combined measure of thermal and vegetative parameters.
+VHI quantifies drought induced vegetative stress weekly by detecting reduced plant cover based on the normalized difference vegetation index (Vegetative Condition Index) combined with reductions in moisture associated with anomalies in thermal spectra (Temperature Condition Index).
+Put most simply, the VHI detects drought by identifying environments that appear unusually brown and dry for that location and time of the year.
+
+This data provides a global, long term, and quantitative perspective on drought variability.
+The VHI database (see methods) presents a valuable resource to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes.
+As such, it has been validated as a tool for detecting drought and predicting crop yields {Kogan 1997}.
+It has also proven useful for predicting infraspecific variation in drought tolerance traits and genes {Mojica et al 2016}{Dittberner et al 2018}{Monroe et al 2018}.
+
+VHI quantifies drought induced vegetative stress weekly by detecting reduced plant cover based on the normalized difference vegetation index (Vegetative Condition Index) combined with reductions in moisture associated with anomalies in thermal spectra (Temperature Condition Index).
+Put most simply, the VHI detects drought by identifying environments that appear unusually brown and dry for that location and time of the year.
+
+The Vegetative Health Index (VHI) was used to examine seasonal drought frequency at the collection locations of Heliophila herbarium specimens.
+The VHI is a satellite-detected drought measurement method based on observations of vegetative stress caused by drought {Kogan, 1995 #534}, combining deviations from historic climatic (Temperature Condition Index) and vegetative conditions (Vegetative Condition Index).
+The VHI has been measured weekly at 16km2 spatial resolution since XXXX.
+The frequencies of observing drought conditions (VHI<40, standard recommended by NOAA developers of VHI) during the southern hemisphere spring (quarter surrounding spring equinox), summer (quarter surrounding summer solstice), fall (quarter surrounding fall equinox), and winter (quarter surrounding winter solstice) were calculated globally from 1981 to 2015.
+These values were calculated for the collection location of each Heliophila herbarium specimen in our filtered dataset.
+
+NDVI = (Ch2 - Chl)/(Ch2 + Chl)
+
+$$NDVI = \frac{NIR - Red}{NIR + Red}$$
+
+$$VCI = 100\frac{NDVI - NDVI_{min}}{NDVI_{max} - NDVI_{min}}$$
+
+$$TCI = 100\frac{T - T{min}}{T{max} - T{min}}$$
+
+$$VHI = 0.5(VCI) + 0.5(TCI)$$
+This data provides a global, long term, and quantitative perspective on drought variability.
+The VHI database presents a valuable resource to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes.
+As such, it has been validated as a tool for detecting drought and predicting crop yields [@Kogan1997].
+It has also proven useful for predicting infraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought].
+
 ### Life history
+Heliophila is a genus of flowering plants endemic to the southern portion of Africa including the Cape Floristic and Succulent Karoo Regions which are among the most botanically diverse environments on Earth.
+Indeed, the estimated ~50 Heliophila species are considered to represent the most diverse genus of the family Brassicaceae {Mummenhoff, 2005 #494}.
+This genus includes both perennial and annual species and this change in life history strategy has likely arisen multiple independent times {Appel & Al-Shehbaz, 1997}{Mummenhoff, 2005 #494}.
+These multiple transitions between life history strategy and it’s well documented record in global herbaria provided the ideal system to address classic hypotheses about the evolutionary drivers of annual and perennial life history strategies.
+
 ### Phyologeny
+Herbaria records and satellite detected drought provide data with which the distributions of annual and perennial species can be compared with respect to historic drought frequency.
+However, it is necessary to control for the demographic history caused by common ancestry of species if evolutionary processes such as natural selection are to be invoked as explanations for any differences observed.
+If, for example, annual and perennial species show significantly different ranges with respect to historic drought this could be confounded by common ancestry if annuals originated from a common ancestor and vice versa.
+In this case, it would be challenging to distinguish between natural selection and demographic history.
+On the other hand, if annual and perennial life history strategies arose independently in multiple species, controlling for phylogenetic relationships allows us to better account for demography and make stronger assertions about the importance of processes such as natural selection to explain patterns.
+
+Heliophila is a charismatic genus of flowering plants from the Brassicaceae provides a valuable model to study the evolution of annual and perennial life history strategies because each has arisen independently multiple times within the genus.
+These independent origins allow for phylogenetically constrained analyses comparing the climate distributions of annual and perennial species, to gain greater insight into the role of selection.
+
+Aligned Heliophila ITS sequences were obtained from previous work by Mandáková et al.
+{Mandáková, 2012 #339}.
+Aethionema, Alliaria, Cardamine, Chamira, and Rorippa ITS records from were downloaded from Genbank (link) and aligned together with Heliophila ITS sequences using MAFFT as outgroups?
+(cite).
+Model selection for construction of phylogeny was performed in jModeltest2 with CIPRES (cite).
+Based on this analysis, $GTR + L$ were selected.
+Ultrametric phylogeny was estimated with branch lengths as relative time (details).
+Life history character states were used based on habit reported by Mummemhoff et al.
+{Mummenhoff, 2005 #494}.
+Species reported to have any form of perennial life history (such as biennial…) were classified for analyses here as perennial, where species reported to have strictly annual life histories were classified for analyses here as annual.
+
 ### Herbarium Specimens
@@
 ```
+Botanists have collected and maintained over 350 million botanical specimens worldwide over the past 300 years (Lang et al 2018).
+These collections, housed in herbaria, comprise an enormous, yet largely untapped, ecological dataset.
+These specimens are increasingly being recognized as an invaluable and underutilized source of data pertaining to biological responses to past environmental conditions (Pyke et al.
+2010).
+Herbarium specimens and their associated metadata have been used since the 1960s to study species’ geographical distributions (reviewed in Lang et al.
+2018).
+Herbarium specimens have been used to track plant responses to climate change using herbarium time-series, including the relationships between traits, geography, and climate.
+{Wolf et al.
+2016, Davis et al.
+2015, Stropp et al.
+2017}.
+A potentially fruitful approach combines herbarium data with remotely sensed data sets that characterize the environmental and climatic conditions across landscapes.
+For example, remotely sensed global ultraviolet-B radiation exposure may explain differences in trichome phenotypes, observed from herbarium specimens, between the native and invasive ranges of two herbaceous plants (Václavík et al.
+2017).
+
+Herbarium specimen records for all Heliophila species were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018.
+Several filtering steps were employed to clean this data (see Sup Mat.
+R code).
+For these analyses, records were restricted to those which are associated with a physical herbarium specimen, rather than other types of observations.
+Consequentially, this sample represents individuals in a narrow range of vegetative and/or flowering states – for example, herbarium specimens cannot be collected for individuals existing as seed.
+Analyses were restricted to Heliophila species with previously-reported life history information {Mummenhoff, 2005 #494}; 42 of ~50 named species in this genus were used.
+Only records including geo-referenced information (latitude and longitude of collection site) were used for downstream analyses of historic drought frequency.
+To discard records likely to be erroneous and to restrict analyses to plants in their native, non-cultivated, habitats, records occurring outside of the native range (roughly, the southern half of Africa) were removed.
+This excluded specimens such as those growing in likely cultivation in Australia or terrestrial plants recorded from non-terrestrial oceanic locations..
+Finally, duplicate records, based on species, collection location coordinates, and date, were removed.
+
 We downloaded `r nrow(GBIF)` records from GBIF.
 
-### Drought
+## Analyses
+### Drought frequency
 
-## Analyses
+### Herbarium clean up
+```{r cleanup}
+species_life_history<-read.csv("../data/Heliophila_life_history.csv")
+species_life_history$species<-gsub("_", " ", species_life_history$species)
+species_life_history$species <- as.factor(species_life_history$species)
+
+GBIF <- GBIF %>%
+  mutate(Winter=raster::extract(Winter, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(Spring=raster::extract(Spring, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(Summer=raster::extract(Summer, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(Fall=raster::extract(Fall, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(ymd_lat_lon_sp = paste0(year,"_",month,"_",day,"_", decimalLatitude,"_", decimalLongitude,"_", species))
+
+GBIF_cleaned <- GBIF %>%
+  filter(species %in% species_life_history$species) %>% 
+  #filter(basisOfRecord %in% "PRESERVED_SPECIMEN") %>%
+  filter(!is.na(decimalLatitude)) %>%
+  filter(hasGeospatialIssues==FALSE) %>%
+  filter(decimalLatitude<(-20)) %>%
+  filter(decimalLongitude<40) %>%
+  filter(!duplicated(ymd_lat_lon_sp)) %>%
+  merge(species_life_history) %>%
+  filter(!is.na(Summer))
+```
+
+We filtered raw GBIF by restricting our analyses to
+
+* Species with reported life history  
+* Records with geospatial data  
+* Records from collection sites classified as land pixels
+* Records for perserved specimens 
+* Records from Africa 
+* Non-duplicate records (ie. identical species, location, collection date)
+
 ### Phylogeny
-### Drought frequency
+
 ### Contrast
+```{r speciesmeans}
+species_means <- GBIF_cleaned %>%
+  dplyr::select(species, life_history, Winter, Summer, Spring, Fall) %>%
+  group_by(species, life_history) %>%
+  summarise(n=n(), Winter_mean=mean(Winter), Spring_mean=mean(Spring), Summer_mean=mean(Summer), Fall_mean=mean(Fall)) %>% 
+  as.data.frame()
+
+```
+```{r firthmodels, echo=F,  results="hide"}
+
+firthmodelstable<-data.frame()
+
+# non-phylogenetic logistic regression firth penalized
+species_means$life_history_num<-as.numeric(species_means$life_history)-1
+row.names(species_means)<-gsub(" ", "_", species_means$species)
+seasons<-c("Winter_mean", "Spring_mean", "Summer_mean","Fall_mean")
+for(s in 1:length(seasons)){
+  fit1 <- logistf(species_means$life_history_num~(species_means[,seasons[s]]), firth = TRUE)
+  sum<-summary(fit1)
+  row.names(sum$coefficients)<-NULL
+  sum_df<-data.frame(Predictor=c("Intercept", gsub("_mean", " drought freq.", seasons[s])), 
+                     Estimate=sum$coefficients, 
+                     p.value=sum$prob)
+  row.names(sum_df)<-NULL
+  firthmodelstable<-rbind(firthmodelstable, sum_df)
+}
+
+```
+```{r phylomodels, results="hide"}
+# phylogentic constraint
+tree<-read.nexus("../data/TrimmedMCCwoPolytomies.nex")
+tree<- rescale(tree, "depth", 1)
+
+phylomodelstable<-data.frame()
+
+for(s in 1:length(seasons)){
+  fit1 <-  phyloglm(species_means$life_history_num~species_means[,seasons[s]], data=species_means, phy=tree)
+  sum<-summary(fit1)
+  row.names(sum$coefficients)<-NULL
+  sum_df<-data.frame(Predictor=c("Intercept", gsub("_mean", " drought freq.", seasons[s])), sum$coefficients)
+  phylomodelstable<-rbind(phylomodelstable, sum_df)
+}
+
+phylomodelstable <- phylomodelstable[,c( "Predictor","Estimate","p.value")]
+
+```
+To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimes, we tested the corresponding prediction that the observed distributions of annual and perennial Heliophila species would be significantly associated with historic drought frequency.
+To do so, the average drought frequency for each season was calculated across the herbarium specimen collection locations for each species.
+The relationship between drought frequencies across each taxon’s range and life habitat (annual or perennial) was evaluated using Firth's penalized-likelihood logistic regression and phylogenetic logistic regression.
+
+### Collection dates
+```{r makecollectiondates, results="hide", message=F, warning=F}
+GBIF_cleaned$date_calc<-as.Date(paste(
+  GBIF_cleaned$month,
+  GBIF_cleaned$day, sep="-"),"%m-%d")
+GBIF_cleaned$days<-lubridate::yday(GBIF_cleaned$date_calc)
+GBIF_cleaned$days_adjusted<-sapply(GBIF_cleaned$days, function(x)
+  if(is.na(x)){ return(NA)
+  }else if(x>135) {
+    return (x-135)
+  }else return (x+230))
+
+pdf("../figures/collection_dates.pdf", width=5, height=3)
+p<-ggplot(GBIF_cleaned, aes(x=days_adjusted, fill=life_history)) +
+  geom_density(alpha=.5)+
+  scale_fill_manual(values=c( "#56B4E9","#E69F00"), name="Life history", labels=c("annual", "perennial"))+
+  theme_classic()+
+  scale_x_continuous(name="Season", breaks=c(45,135,225, 315), labels=c("Winter","Spring","Summer","Fall"))+
+  labs(title="Collection date\nof herbaria specimens", y="Density")+
+  theme(plot.title = element_text(hjust = 0.5))
+p
+dev.off()
+
+# TEST ANNUAL AND PERENNIAL DATES -----------------------------------------
+annuals<-subset(GBIF_cleaned, life_history=="a")
+perennials<-subset(GBIF_cleaned, life_history=="p")
+
+ks.test(annuals$days,perennials$days)
+t.test(days~life_history, GBIF_cleaned)
+bartlett.test(days~life_history, GBIF_cleaned)
+```
+To test the hypothesis that annual species have adapted to escape drought prone seasons as seeds, collection dates for herbarium specimens were compared between annual and perennial species.
+Comparisons of distributions were made by Two-sample Kolmogorov-Smirnov test, t-test, and Barlett variance test using R (supplemental script 2) {cite R}.
+### Software
 We used `r cite_r("r-references.bib")` for all our analyses.
@@
 
-(ref:foo) A scatterplot of the data `cars` using **base** R graphics.
+To test the hypothesis that annual and perennial plants reflect adaptation to alternative drought environments we examined the landscape distribution of life history strategies in the large and diverse mustard genus, _Heliophila_ Figure \@ref(fig:phylogeny).
+Using both herbarium specimen metadata and a 30 year dataset of satellite generated climate information, we tested the prediction that annual species  are more often observed in drought-prone locations than perennial species, when controlling for phylogenetic relatedness.
+We found that drought frequency is significantly different between the distributions of annual and perennial species, with annuals being found in environments with significantly more frequent drought, and that this signal is strongest during the summer.
+These results remain significant while controlling for the phylogenetic relationships of Heliophila species, yielding support for the role that natural selection has played in driving contemporary distributions of these alternatives strategies in relation to drought regimes.
 
-```{r foo, fig.cap='(ref:foo)',out.width = "\\textwidth", fig.pos = "!h"}
-plot(cars)  # a scatterplot
+(ref:phylogeny) Phylogeny of Heliophila. 
+```{r makephylogeny, eval=T, message=F, results="hide"}
+
+pdf("../figures/phylogeny.pdf", height=6, width=5)
+par(mar=c(0,0,0,0))
+tree<-read.nexus("../data/TrimmedMCCwoPolytomies.nex")
+tree<- rescale(tree, "depth", 1)
+tree$tip.label<-gsub("_"," ", tree$tip.label)
+plot(tree, label.offset = .05)
+tiplabels(pch = 21, bg = ifelse(species_life_history$life_history[match(tree$tip.label,species_life_history$species)]=="a", "dodgerblue", "orange2")
+, cex = 1.6, adj = 0.5)
+
+dev.off()
+
+```
+```{r phylogeny, fig.cap='(ref:phylogeny)', out.width = "\\textwidth", fig.pos = "!h"}
+knitr::include_graphics("../figures/phylogeny.pdf", dpi = 108)
 ```
 
-First we did this (Fig. \@ref(fig:foo).)
+### GBIF records
+
+Out of `r nrow(GBIF)` _Heliophila_ GBIF recrods,  `r sum(GBIF$species %in% species_life_history$species)` were for species with reported life history [@mummenhoff2005phylogeny], `r sum(!is.na(GBIF$decimalLatitude))` had geospatial data, `r sum(!(is.na(GBIF$Summer)) & !(is.na(GBIF$decimalLatitude)))` were located on pixels classified as land having drought measurements, `r sum(GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude))` were located in Africa, `r sum(!duplicated(GBIF$ymd_lat_lon_sp) & !is.na(GBIF$decimalLatitude)) ` were not duplicated.
+After all filtering steps, `r nrow(GBIF_cleaned)` records for `r length(unique(GBIF_cleaned$species))` species (Figure \@ref(tab:gbifrecordsmap), Table \@ref(tab:speciesmeanstable)) passed.
+
+(ref:gbifrecordsmap) Map of `r nrow(GBIF_cleaned)` GBIF records that passed quality filtering.  
+```{r makegbifrecordsmap, message=F, results=F}
+
+area <- rgdal::readOGR("../data/ne_10m_ocean/ne_10m_ocean.shp")
+
+pdf("../figures/gbifrecords.pdf", width=5, height=3.5)
+par(mar=c(4,4,1,1))
+plot(area, xlab="Longitude (deg)", ylab="Latitude, (deg)", xlim=c(10,40), ylim=c(-35, -20), bg="gray90", col="white")
+axis(1);axis(2)
+points(GBIF_cleaned$decimalLongitude, GBIF_cleaned$decimalLatitude, pch=3, col="black")
+dev.off()
+```
+
+```{r gbifrecordsmap, fig.cap='(ref:gbifrecordsmap)'}
+knitr::include_graphics("../figures/gbifrecords.pdf", dpi = 108)
+```
+
+```{r speciesmeanstable, results = 'asis', echo = F}
+species_means_table<-species_means
+row.names(species_means_table) <- NULL
+species_means_table<-species_means_table %>% dplyr::select(-life_history_num)
+colnames(species_means_table)<-c("Species","LH","n","Winter","Spring","Summer","Fall")
+apa_table(
+  species_means_table
+  , caption = "Heliophila species records and the mean drought frequencies during different seasons at the location of records "
+  , note = "LH = Life history (a = annual, p = perennial). n=sample size of GBIF records. Seasons are mean drought frequencies observed at locations of records.",
+  digits=2,
+  font_size="small",
+  longtable = TRUE
+)
+```
+
+### Drought frequency.
+
+(ref:boxplots) Comparison of drought frequency across seasons measured at the GBIF records of annual and perennial species of Heliophila.
+
+```{r makeboxplots, eval=T, message=F, results="hide"}
+
+GBIF_cleaned_long<- GBIF_cleaned %>%
+  dplyr::select(species, life_history, Winter, Spring, Summer,Fall) %>%
+  gather(season, drought_freq, Winter, Spring, Summer,Fall)
+
+pdf("../figures/life_history_drought_boxplots.pdf", height=3, width=5)
+
+p<-ggplot(GBIF_cleaned_long, aes(x=life_history, col=life_history, y=drought_freq))+
+    facet_wrap(~season, ncol=4 )+
+    geom_jitter(width = .3, alpha=0.1, pch=3)+
+    geom_boxplot(outlier.colour = "white", outlier.fill = "white", fill=NA, width=.3)+
+    scale_color_manual(values=c("dodgerblue", "orange2"))+
+    theme_classic()+
+    theme(legend.position = "none")+
+    scale_x_discrete(name="life history", labels=c("annual","perennial"))+
+    labs(y="Drought frequency")
+
+  plot(p)
+
+dev.off()
+
+```
+
+```{r boxplots, fig.cap='(ref:boxplots)'}
+knitr::include_graphics("../figures/life_history_drought_boxplots.pdf", dpi = 108)
+```
+
+```{r firthmodelstable, results = 'asis', echo = F}
+apa_table(firthmodelstable
+  , caption = "Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species. "
+  , note = "Firth's penalized logistic regression. Annual species were scored as 0 and perennial species as 1.",
+  midrules=c(2,4,6,8),
+  digits=4
+)
+
+```
+
+```{r phylomodelstable, results = 'asis', echo = F}
+apa_table(
+  phylomodelstable
+  , caption = "Phlyogenetically constrained logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species. "
+  , note = "Annual species were scored as 0 and perennial species as 1.",
+  midrules=c(2,4,6,8),
+  digits=4
+)
+
+```
+
+(ref:lineplots) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of Heliophila.
+
+```{r makelineplots, eval=T, message=F, results="hide"}
+life_history_drought_means<- species_means %>% 
+  dplyr::select(-species, -n, -life_history_num)  %>%
+  gather(variable, value, -life_history) %>%
+  group_by(life_history, variable) %>%
+  summarise(mean = mean(value), se = sd(value)/sqrt(length(value))) %>%
+  mutate(variable = reorder(variable, c(4,2,3,1)))
+
+pdf("../figures/life_history_drought_means_line.pdf", height=3, width=5)
+
+p<-ggplot(life_history_drought_means, aes(x=variable, col=life_history,group=life_history, y=mean))+
+  theme_classic()+
+  geom_point(size=2)+
+  geom_line(lwd=0.5, lty=1)+
+  geom_errorbar(width=0.25,lwd=1, aes(ymin=mean-se, ymax=mean+se))+
+  scale_x_discrete(labels=c("Winter","Spring","Summer","Fall"))+
+  scale_color_manual(values=c( "#56B4E9","#E69F00"), name="Life history", labels=c("annual", "perennial"))+
+  labs(x="Season", y="Drought Frequency")+
+  annotate("text", x=2, y=0.45, label="*", size=7) +
+  annotate("text", x=3:4, y=0.45, label="**", size=7)
+
+plot(p)
+dev.off()
+
+```
+
+```{r lineplots, fig.cap='(ref:lineplots)'}
+knitr::include_graphics("../figures/life_history_drought_means_line.pdf", dpi = 108)
+```
+
+### Collection dates
+
+(ref:collectiondates) Collection dates of GBIF records of annual and perennial species of Heliophila.
+
+```{r collectiondates, fig.cap='(ref:collectiondates)'}
+knitr::include_graphics("../figures/collection_dates.pdf", dpi = 108)
+```
 
 # Discussion
-The best paper ever [@monroe2016adaptation]
+## Summary
+We found that the distribution of annual and perennial species of Heliophila is significantly predicted by satellite detected historic drought frequencies.
+Annual species are found in environments that experience more frequent drought during the summer and fall quarters compared to perennials.
+This relationship was consistent while controlling for phylogenetic relatedness among the taxa studied, indicating that these distributions cannot be explained entirely by common ancestry., These results support the hypothesis that natural selection has played a role in shaping the contemporary distributions of these alternative life-history strategies.
 
-There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.
+## Relationship to previous hypotheses/work
+These findings support classical theoretical predictions about the adaptive value of annual and perennial life history strategies.
+Taken together, they suggest that in Heliophila, annual species are adapted to environments with increased summer droughts by avoiding these seasons in a dormant seed phase of their life cycle.
+Indeed, we found that very few annuals are collected during this season, supporting the prediction that they are not in a vegetative and/or reproductive phase at this time.
+Traditionally, the focus has been on the evolutionary origins of annual life histories {citations}.
+However, we also find evidence that the transition to perenniality could be explained by historical drought regimes.
+The phylogeny reveals several transitions from annual to perennial life history.
+Perennials may be able to out complete annual relatives in environments where the infrequency of drought favors strategies that allow plants to benefit from growth over many seasons.
 
-\newpage
+## Caveats
+Correlation does not prove causation. But it does indicate predictive power and is consistent with adaptive hypotheses.
+Herbarium collections and their associated data do not represent systematic or random sampling of a species distribution.
+Significant biases in collecting exist, which we have not necessarily controlled for here, and may have some effect on our findings, such as a bias toward collecting near roads or near the locations of natural history collections (Daru et al.
+2018, Heberling, in press).
+Despite these biases, the Cape Floristic region is a biodiversity hotspot and one of the most botanically well sampled regions on Earth (Daru et al.
+2018, Heberling, in press?), suggesting that this may currently be the optimal region for our analyses of life history distribution.
+Future research will benefit from systematic sampling efforts to avoid these noted biases.
+The climate data used here are assessed only from 198X-XXXX and do not reflect conditions at the estimated divergence dates of these species (XXX million of years ago).
+Rather, the results suggest that the current distributions of annual and perennial species reflect a history of environmental filtration and ongoing natural selection.
+That is, their distributions are non-random with respect to historic drought and this is not explained by phylogeny.
+
+## Broader implications 
+These findings suggest that rapidly changing drought regimes threaten species adapted to current environments.
+Studies predict changing drought regimes.
+This could have impacts on ecosystem functioning.
+This should also be considered when thinking about using perennial crops.
+Studies predict changing drought regimes.
+
+## Conclusions 
+Perenniality appears to be adaptive in environments with less frequent drought. This work demonstrates the power of emerging data to address outstanding classic hypotheses in ecology and evolution.
+
+# Acknowledgments 
+This work was supported by NSF no. USDA grant no. XXX to J.G.M.
 
 # References
+\newpage
 ```{r create_r-references}
@@
 ```
-
 \begingroup

### 2018-11-24 manuscript/manuscript.tex
initial commit

@@
             pdftitle={Plant life history strategies predicted by satellite-detected drought},
-            pdfauthor={J Grey Author, Brian Gill, \& Brian Gill},
-            pdfkeywords={drought, life history, remote sensing, phylogeography},
+            pdfauthor={J Grey Monroe, Brian Gill, \& Kathryn Turner},
+            pdfkeywords={drought adaptation, life history evolution, remote sensing,
+phylogeography, herbaria records},
             pdfborder={0 0 0},
@@
   \title{Plant life history strategies predicted by satellite-detected drought}
-    \author{J Grey Author\textsuperscript{1,2}, Brian Gill\textsuperscript{3}, \&
-Brian Gill\textsuperscript{4}}
+    \author{J Grey Monroe\textsuperscript{1,2}, Brian Gill\textsuperscript{3}, \&
+Kathryn Turner\textsuperscript{4}}
     \date{}
 
-\shorttitle{}
+\shorttitle{Drought and life history}
 \affiliation{
 \vspace{0.5cm}
-\textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University\\\textsuperscript{2} College of Agriculture, Colorado State University\\\textsuperscript{3} Department of Ecology and Evolutionary Biology\\\textsuperscript{4} Biology Department, Penn State University}
-\keywords{drought, life history, remote sensing, phylogeography\newline\indent Word count: }
+\textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO\\\textsuperscript{3} Department of Ecology and Evolutionary Biology, Brown University, Provience, RI\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA}
+\keywords{drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records\newline\indent Word count: }
 \usepackage{csquotes}
@@
 
-\authornote{Add complete departmental affiliations for each
-author here. Each new line herein must be indented, like this line.
+\authornote{
 
 Correspondence concerning this article should be addressed to J Grey
-Author, 307 University Ave, Fort Collins, CO 80523. E-mail:
+Monroe, 307 University Ave, Fort Collins, CO 80523. E-mail:
 \href{mailto:monroejg@colostate.edu}{\nolinkurl{monroejg@colostate.edu}}}
@@
 \abstract{
-
+Identifying the environmental factors that predict the evolution of
+annual and perennial life history strategies in plants is important for
+understanding ecosystem functioning, perennial cropping systems, and is
+a long-standing goal of evolutionary ecology. A classic hypothesis is
+that annual and perennial strategies reflect adaptation to environments
+that differ in relative drought frequency. We test this hypothesis in
+Heliophila (Brassicaceae), a genus of flowering plants native to
+Southern Africa using herbarium occurrence records and
+satellite-detected drought histories. We find that perennial Heliophila
+species are observed in environments where droughts are significantly
+less frequent compared to annuals, especially with regards to summer
+drought. These correlations remain predictive while controlling for
+phylogeny, lending support to the hypothesis that drought associated
+natural selection has shaped differences in the distributions of these
+strategies. Additionally, the collection dates of herbarium records are
+consistent with a scenario in which annual species escape drought prone
+seasons. Together these results point to a model of evolution that
+supports classical hypotheses favoring perenniality in environments that
+experience less drought, compared to annuals which escape drought prone
+seasons as seed. Thus, changing drought regimens during the Anthropocene
+may threaten locally adapted species. 
 }
 
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}{Lemma}[section]
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{definition}
+\newtheorem{example}{Example}[section]
+\theoremstyle{definition}
+\newtheorem{exercise}{Exercise}[section]
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
+\newtheorem*{solution}{Solution}
 \begin{document}
@@
 
-\#Introduction
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-Life history Annual vs perennial Hypotheses Summary
+\hypertarget{life-history}{%
+\subsection{Life history}\label{life-history}}
+
+Plants exhibit extraordinary diversity in life history strategies, from
+herbaceous species that complete the entire seed to seed life cycle in a
+number of weeks \{Guo, 2012 \#73\} to trees that live for thousands of
+years \{Brown 1996 \url{http://www.rmtrr.org/oldlist.htm} \}. Along this
+continuum an important division exists, distinguishing annuals which
+complete their seed to seed life cycle within a single calendar year
+from perennials which can persist over multiple years and therefore
+experience the entire range of seasonal conditions. The ecological
+factors that explain the evolution of these alternative strategies
+remains poorly resolved \{Friedman, 2015 \#523\}. Understanding the
+drivers of selection for these alternative strategies is important
+because annuals and perennials have differing impacts on ecosystem
+functioning, such as higher nitrogen concentration in perennials and
+larger specific root length in annuals that affect nutrient cycling
+\{Garnier, 1997 \#280\}\{Roumet, 2006 \#738\}. Furthermore, predicting
+the conditions that favor these strategies in nature is useful for
+developing more sustainable perennial cropping systems \{Lelievre and
+Volaire, 2009\}.
+
+\hypertarget{hypotheses}{%
+\subsection{Hypotheses}\label{hypotheses}}
+
+Seasonal drought conditions can represent a severe challenge to the
+persistence of individual organisms. Well-established hypotheses suggest
+that an annual life history is adaptive in environments where frequent
+droughts makes an escape from stressful drought conditions in the form
+of seed advantageous \{Stearns 1992\}\{Silvertown and Charlesworth,
+2001\}. These hypotheses are based on models of life history evolution
+in which short life cycles are favored in unpredictable environments or
+those with frequent stressful conditions. Plants are expected to be more
+likely to survive a drought that cooccurs with the seed phase of a
+plant's life cycle in which they are most resistant to desiccation.
+Previous efforts to address this hypothesis have yielded mixed or
+qualitative results. Transitions to annuality in Oenothera were
+associated with warmer summers and drier winters, but not with increased
+drought directly \{Evans, 2005 \#260\}. Annual and perennial species of
+Nemesia were qualitatively associated with winter rather and summer
+rainfall environments respectively \{Datson, 2008\}. Similarly, annual
+species of Scorzoneroides were associated with environments classified
+as \enquote{unpredictable} \{Cruz-Mazo, 2009 \#598\}. Annual and
+perennial varieties of the wild rice species Oryza rufipogon were
+observed more frequently in dry and humid climates, respectively
+\{Morishima et al.~1984\}. While these reports suggest the prominence of
+annuals in environments considered more arid, a quantitative assessment
+of how historic drought frequency may favor annual or perennial life
+history strategies remains lacking.
+
+\hypertarget{summary}{%
+\subsection{Summary}\label{summary}}
+
+Here we combine a long-term dataset of global imaging with metadata from
+natural history collections to test classic hypotheses about the
+evolution of life history strategies within the African endemic mustard
+genus, Heliophila L. (Brassicaceae). If annual species temporally escape
+drought as seed, then drought frequency should be an important
+determinant of the distribution of life history strategies across the
+landscape, and annual species should be more commonly associated with
+drought prone regions than perennial species. Furthermore, if annual
+species have adapted to escape drought prone seasons as seeds,
+observations of annual species should be rare during drought prone
+seasons. Phylogenetic relatedness can have significant non-random
+effects on species distributions and life history traits (cite), and
+therefore we assessed the relationship between life history distribution
+and drought frequency in a phylogenetically controlled background.
 
@@
 
-\hypertarget{life-history}{%
-\subsubsection{Life history}\label{life-history}}
+\hypertarget{drought}{%
+\subsubsection{Drought}\label{drought}}
+
+Remotely sensed data is a valuable tool for characterizing seasonal
+patterns in drought \{AghaKouchak 2015\}. With these data, we can
+exploit the fact that satellite imagery can detect reductions in plant
+cover and moisture at landscape scales that are characteristic of
+drought conditions. Unlike weather station data, which is limited in
+geographic coverage, and manual observation, which is limited in scope,
+satellite detected data provides a long term global historical record of
+drought events at fine temporal and spatial resolutions. One such
+measure, the Vegetative Health Index (VHI), has been collected since
+1981 by NOAA AVHRR satellites and is a combined measure of thermal and
+vegetative parameters. VHI quantifies drought induced vegetative stress
+weekly by detecting reduced plant cover based on the normalized
+difference vegetation index (Vegetative Condition Index) combined with
+reductions in moisture associated with anomalies in thermal spectra
+(Temperature Condition Index). Put most simply, the VHI detects drought
+by identifying environments that appear unusually brown and dry for that
+location and time of the year.
+
+This data provides a global, long term, and quantitative perspective on
+drought variability. The VHI database (see methods) presents a valuable
+resource to study seasonal patterns in the frequency of drought across
+environments and to test hypotheses about the effect of drought on
+ecological and evolutionary processes. As such, it has been validated as
+a tool for detecting drought and predicting crop yields \{Kogan 1997\}.
+It has also proven useful for predicting infraspecific variation in
+drought tolerance traits and genes \{Mojica et al 2016\}\{Dittberner et
+al 2018\}\{Monroe et al 2018\}.
+
+VHI quantifies drought induced vegetative stress weekly by detecting
+reduced plant cover based on the normalized difference vegetation index
+(Vegetative Condition Index) combined with reductions in moisture
+associated with anomalies in thermal spectra (Temperature Condition
+Index). Put most simply, the VHI detects drought by identifying
+environments that appear unusually brown and dry for that location and
+time of the year.
+
+The Vegetative Health Index (VHI) was used to examine seasonal drought
+frequency at the collection locations of Heliophila herbarium specimens.
+The VHI is a satellite-detected drought measurement method based on
+observations of vegetative stress caused by drought \{Kogan, 1995
+\#534\}, combining deviations from historic climatic (Temperature
+Condition Index) and vegetative conditions (Vegetative Condition Index).
+The VHI has been measured weekly at 16km2 spatial resolution since XXXX.
+The frequencies of observing drought conditions (VHI\textless{}40,
+standard recommended by NOAA developers of VHI) during the southern
+hemisphere spring (quarter surrounding spring equinox), summer (quarter
+surrounding summer solstice), fall (quarter surrounding fall equinox),
+and winter (quarter surrounding winter solstice) were calculated
+globally from 1981 to 2015. These values were calculated for the
+collection location of each Heliophila herbarium specimen in our
+filtered dataset.
+
+NDVI = (Ch2 - Chl)/(Ch2 + Chl)
+
+\[NDVI = \frac{NIR - Red}{NIR + Red}\]
+
+\[VCI = 100\frac{NDVI - NDVI_{min}}{NDVI_{max} - NDVI_{min}}\]
+
+\[TCI = 100\frac{T - T{min}}{T{max} - T{min}}\]
+
+\[VHI = 0.5(VCI) + 0.5(TCI)\] This data provides a global, long term,
+and quantitative perspective on drought variability. The VHI database
+presents a valuable resource to study seasonal patterns in the frequency
+of drought across environments and to test hypotheses about the effect
+of drought on ecological and evolutionary processes. As such, it has
+been validated as a tool for detecting drought and predicting crop
+yields ({\textbf{???}}). It has also proven useful for predicting
+infraspecific variation in drought tolerance traits and genes
+(Dittberner et al., 2018; Mojica et al., 2016; Monroe et al., 2018).
+
+\hypertarget{life-history-1}{%
+\subsubsection{Life history}\label{life-history-1}}
+
+Heliophila is a genus of flowering plants endemic to the southern
+portion of Africa including the Cape Floristic and Succulent Karoo
+Regions which are among the most botanically diverse environments on
+Earth. Indeed, the estimated \textasciitilde{}50 Heliophila species are
+considered to represent the most diverse genus of the family
+Brassicaceae \{Mummenhoff, 2005 \#494\}. This genus includes both
+perennial and annual species and this change in life history strategy
+has likely arisen multiple independent times \{Appel \& Al-Shehbaz,
+1997\}\{Mummenhoff, 2005 \#494\}. These multiple transitions between
+life history strategy and it's well documented record in global herbaria
+provided the ideal system to address classic hypotheses about the
+evolutionary drivers of annual and perennial life history strategies.
 
@@
 
+Herbaria records and satellite detected drought provide data with which
+the distributions of annual and perennial species can be compared with
+respect to historic drought frequency. However, it is necessary to
+control for the demographic history caused by common ancestry of species
+if evolutionary processes such as natural selection are to be invoked as
+explanations for any differences observed. If, for example, annual and
+perennial species show significantly different ranges with respect to
+historic drought this could be confounded by common ancestry if annuals
+originated from a common ancestor and vice versa. In this case, it would
+be challenging to distinguish between natural selection and demographic
+history. On the other hand, if annual and perennial life history
+strategies arose independently in multiple species, controlling for
+phylogenetic relationships allows us to better account for demography
+and make stronger assertions about the importance of processes such as
+natural selection to explain patterns.
+
+Heliophila is a charismatic genus of flowering plants from the
+Brassicaceae provides a valuable model to study the evolution of annual
+and perennial life history strategies because each has arisen
+independently multiple times within the genus. These independent origins
+allow for phylogenetically constrained analyses comparing the climate
+distributions of annual and perennial species, to gain greater insight
+into the role of selection.
+
+Aligned Heliophila ITS sequences were obtained from previous work by
+Mandáková et al.~\{Mandáková, 2012 \#339\}. Aethionema, Alliaria,
+Cardamine, Chamira, and Rorippa ITS records from were downloaded from
+Genbank (link) and aligned together with Heliophila ITS sequences using
+MAFFT as outgroups? (cite). Model selection for construction of
+phylogeny was performed in jModeltest2 with CIPRES (cite). Based on this
+analysis, \(GTR + L\) were selected. Ultrametric phylogeny was estimated
+with branch lengths as relative time (details). Life history character
+states were used based on habit reported by Mummemhoff et
+al.~\{Mummenhoff, 2005 \#494\}. Species reported to have any form of
+perennial life history (such as biennial\ldots{}) were classified for
+analyses here as perennial, where species reported to have strictly
+annual life histories were classified for analyses here as annual.
+
 \hypertarget{herbarium-specimens}{%
@@
 
-We downloaded 8670 records from GBIF. (Fig. @ref(tab:raw\_GBIF).)
+Botanists have collected and maintained over 350 million botanical
+specimens worldwide over the past 300 years (Lang et al 2018). These
+collections, housed in herbaria, comprise an enormous, yet largely
+untapped, ecological dataset. These specimens are increasingly being
+recognized as an invaluable and underutilized source of data pertaining
+to biological responses to past environmental conditions (Pyke et
+al.~2010). Herbarium specimens and their associated metadata have been
+used since the 1960s to study species' geographical distributions
+(reviewed in Lang et al.~2018). Herbarium specimens have been used to
+track plant responses to climate change using herbarium time-series,
+including the relationships between traits, geography, and climate.
+\{Wolf et al.~2016, Davis et al.~2015, Stropp et al.~2017\}. A
+potentially fruitful approach combines herbarium data with remotely
+sensed data sets that characterize the environmental and climatic
+conditions across landscapes. For example, remotely sensed global
+ultraviolet-B radiation exposure may explain differences in trichome
+phenotypes, observed from herbarium specimens, between the native and
+invasive ranges of two herbaceous plants (Václavík et al.~2017).
 
-\hypertarget{drought}{%
-\subsubsection{Drought}\label{drought}}
+Herbarium specimen records for all Heliophila species were downloaded
+from the Global Biodiversity Information Facility (gbif.org) on July 21,
+2018. Several filtering steps were employed to clean this data (see Sup
+Mat. R code). For these analyses, records were restricted to those which
+are associated with a physical herbarium specimen, rather than other
+types of observations. Consequentially, this sample represents
+individuals in a narrow range of vegetative and/or flowering states --
+for example, herbarium specimens cannot be collected for individuals
+existing as seed. Analyses were restricted to Heliophila species with
+previously-reported life history information \{Mummenhoff, 2005 \#494\};
+42 of \textasciitilde{}50 named species in this genus were used. Only
+records including geo-referenced information (latitude and longitude of
+collection site) were used for downstream analyses of historic drought
+frequency. To discard records likely to be erroneous and to restrict
+analyses to plants in their native, non-cultivated, habitats, records
+occurring outside of the native range (roughly, the southern half of
+Africa) were removed. This excluded specimens such as those growing in
+likely cultivation in Australia or terrestrial plants recorded from
+non-terrestrial oceanic locations.. Finally, duplicate records, based on
+species, collection location coordinates, and date, were removed.
+
+We downloaded 8670 records from GBIF.
 
@@
 
+\hypertarget{drought-frequency}{%
+\subsubsection{Drought frequency}\label{drought-frequency}}
+
+\hypertarget{herbarium-clean-up}{%
+\subsubsection{Herbarium clean up}\label{herbarium-clean-up}}
+
+We filtered raw GBIF by restricting our analyses to
+
+\begin{itemize}
+\tightlist
+\item
+  Species with reported life history\\
+\item
+  Records with geospatial data\\
+\item
+  Records from collection sites classified as land pixels
+\item
+  Records for perserved specimens
+\item
+  Records from Africa
+\item
+  Non-duplicate records (ie. identical species, location, collection
+  date)
+\end{itemize}
+
 \hypertarget{phylogeny}{%
 \subsubsection{Phylogeny}\label{phylogeny}}
-
-\hypertarget{drought-frequency}{%
-\subsubsection{Drought frequency}\label{drought-frequency}}
 
@@
 
-We used R (Version 3.5.1; R Core Team, 2018) and the R-packages
-\emph{dplyr} (Version 0.7.8; Wickham et al., 2018), \emph{forcats}
-(Version 0.3.0; Wickham, 2018a), \emph{ggplot2} (Version 3.1.0; Wickham,
-2016), \emph{papaja} (Version 0.1.0.9842; Aust \& Barth, 2018),
-\emph{purrr} (Version 0.2.5; Henry \& Wickham, 2018), \emph{raster}
-(Version 2.8.4; Hijmans, 2018), \emph{readr} (Version 1.1.1; Wickham et
-al., 2017), \emph{sp} (Version 1.3.1; Pebesma \& Bivand, 2005),
-\emph{stringr} (Version 1.3.1; Wickham, 2018b), \emph{tibble} (Version
-1.4.2; Müller \& Wickham, 2018), \emph{tidyr} (Version 0.8.2; Wickham \&
-Henry, 2018), and \emph{tidyverse} (Version 1.2.1; Wickham, 2017) for
-all our analyses.
+To evaluate the hypothesis that annual and perennial life history
+strategies reflect adaptations to alternative drought regimes, we tested
+the corresponding prediction that the observed distributions of annual
+and perennial Heliophila species would be significantly associated with
+historic drought frequency. To do so, the average drought frequency for
+each season was calculated across the herbarium specimen collection
+locations for each species. The relationship between drought frequencies
+across each taxon's range and life habitat (annual or perennial) was
+evaluated using Firth's penalized-likelihood logistic regression and
+phylogenetic logistic regression.
+
+\hypertarget{collection-dates}{%
+\subsubsection{Collection dates}\label{collection-dates}}
+
+To test the hypothesis that annual species have adapted to escape
+drought prone seasons as seeds, collection dates for herbarium specimens
+were compared between annual and perennial species. Comparisons of
+distributions were made by Two-sample Kolmogorov-Smirnov test, t-test,
+and Barlett variance test using R (supplemental script 2) \{cite R\}.
+\#\#\# Software We used R (Version 3.5.1; R Core Team, 2018) and the
+R-packages \emph{ape} (Version 5.2; Paradis \& Schliep, 2018; Orme et
+al., 2018; Soetaert, 2018), \emph{bindrcpp} (Version 0.2.2; Müller,
+2018), \emph{caper} (Version 1.0.1; Orme et al., 2018), \emph{coda}
+(Version 0.19.2; Plummer, Best, Cowles, \& Vines, 2006), \emph{diagram}
+(Version 1.6.4; Soetaert, 2017), \emph{dplyr} (Version 0.7.8; Wickham et
+al., 2018), \emph{forcats} (Version 0.3.0; Wickham, 2018a), \emph{gee}
+(Version 4.13.19; R by Thomas Lumley \& author., 2015), \emph{geiger}
+(Version 2.0.6; Alfaro et al., 2009; Eastman, Alfaro, Joyce, Hipp, \&
+Harmon, 2011; Harmon, Weir, Brock, Glor, \& Challenger, 2008; Slater et
+al., 2012), \emph{ggplot2} (Version 3.1.0; Wickham, 2016),
+\emph{logistf} (Version 1.23; Heinze \& Ploner, 2018), \emph{maps}
+(Version 3.3.0; Richard A. Becker, Ray Brownrigg. Enhancements by Thomas
+P Minka, \& Deckmyn., 2018), \emph{MASS} (Version 7.3.51.1; Venables \&
+Ripley, 2002), \emph{Matrix} (Version 1.2.15; Bates \& Maechler, 2018),
+\emph{MCMCglmm} (Version 2.26; Hadfield, 2010), \emph{mvtnorm} (Version
+1.0.8; Genz \& Bretz, 2009), \emph{papaja} (Version 0.1.0.9842; Aust \&
+Barth, 2018), \emph{phylolm} (Version 2.6; Ho \& Ane, 2014),
+\emph{phytools} (Version 0.6.60; Revell, 2012), \emph{purrr} (Version
+0.2.5; Henry \& Wickham, 2018), \emph{raster} (Version 2.8.4; Hijmans,
+2018), \emph{readr} (Version 1.2.1; Wickham et al., 2017), \emph{shape}
+(Version 1.4.4; Soetaert, 2018), \emph{sp} (Version 1.3.1; Pebesma \&
+Bivand, 2005), \emph{stringr} (Version 1.3.1; Wickham, 2018b),
+\emph{tibble} (Version 1.4.2; Müller \& Wickham, 2018), \emph{tidyr}
+(Version 0.8.2; Wickham \& Henry, 2018), and \emph{tidyverse} (Version
+1.2.1; Wickham, 2017) for all our analyses.
 
@@
 
+To test the hypothesis that annual and perennial plants reflect
+adaptation to alternative drought environments we examined the landscape
+distribution of life history strategies in the large and diverse mustard
+genus, \emph{Heliophila} Figure \ref{fig:phylogeny}. Using both
+herbarium specimen metadata and a 30 year dataset of satellite generated
+climate information, we tested the prediction that annual species are
+more often observed in drought-prone locations than perennial species,
+when controlling for phylogenetic relatedness. We found that drought
+frequency is significantly different between the distributions of annual
+and perennial species, with annuals being found in environments with
+significantly more frequent drought, and that this signal is strongest
+during the summer. These results remain significant while controlling
+for the phylogenetic relationships of Heliophila species, yielding
+support for the role that natural selection has played in driving
+contemporary distributions of these alternatives strategies in relation
+to drought regimes.
+
 \begin{figure}[!h]
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/foo-1} \caption{A scatterplot of the data \texttt{cars} using \textbf{base} R
-graphics.}\label{fig:foo}
+\includegraphics[width=\textwidth]{../figures/phylogeny} \caption{Phylogeny of Heliophila.}\label{fig:phylogeny}
 \end{figure}
 
-First we did this (Fig. \ref{fig:foo}.)
+\hypertarget{gbif-records}{%
+\subsubsection{GBIF records}\label{gbif-records}}
+
+Out of 8670 \emph{Heliophila} GBIF recrods, 6634 were for species with
+reported life history (Mummenhoff, Al-Shehbaz, Bakker, Linder, \&
+Mühlhausen, 2005), 3653 had geospatial data, 3460 were located on pixels
+classified as land having drought measurements, 3457 were located in
+Africa, 3162 were not duplicated. After all filtering steps, 2192
+records for 42 species (Figure \ref{tab:gbifrecordsmap}, Table
+\ref{tab:speciesmeanstable}) passed.
+
+\begin{figure}
+\centering
+\includegraphics{../figures/gbifrecords.pdf}
+\caption{\label{fig:gbifrecordsmap}Map of 2192 GBIF records that passed quality
+filtering.}
+\end{figure}
+
+\begin{center}
+\begin{ThreePartTable}
+\begin{TableNotes}[para]
+\normalsize{\textit{Note.} LH = Life history (a = annual, p = perennial). n=sample size of GBIF records. Seasons are mean drought frequencies observed at locations of records.}
+\end{TableNotes}
+\small{
+\begin{longtable}{lllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:speciesmeanstable}Heliophila species records and the mean drought frequencies during different seasons at the location of records }\\
+\toprule
+Species & \multicolumn{1}{c}{LH} & \multicolumn{1}{c}{n} & \multicolumn{1}{c}{Winter} & \multicolumn{1}{c}{Spring} & \multicolumn{1}{c}{Summer} & \multicolumn{1}{c}{Fall}\\
+\midrule
+Heliophila acuminata & a & 28 & 0.32 & 0.38 & 0.41 & 0.36\\
+Heliophila africana & a & 91 & 0.33 & 0.35 & 0.34 & 0.34\\
+Heliophila amplexicaulis & a & 60 & 0.32 & 0.36 & 0.39 & 0.33\\
+Heliophila arenaria & a & 65 & 0.34 & 0.37 & 0.38 & 0.34\\
+Heliophila carnosa & p & 129 & 0.33 & 0.37 & 0.39 & 0.31\\
+Heliophila cedarbergensis & p & 3 & 0.40 & 0.43 & 0.32 & 0.27\\
+Heliophila collina & a & 16 & 0.35 & 0.47 & 0.48 & 0.45\\
+Heliophila cornellsbergia & a & 2 & 0.33 & 0.42 & 0.35 & 0.21\\
+Heliophila cornuta & p & 101 & 0.35 & 0.40 & 0.40 & 0.34\\
+Heliophila coronopifolia & a & 40 & 0.37 & 0.42 & 0.40 & 0.37\\
+Heliophila crithmifolia & a & 97 & 0.35 & 0.42 & 0.45 & 0.38\\
+Heliophila descurva & a & 12 & 0.36 & 0.38 & 0.38 & 0.29\\
+Heliophila deserticola & a & 133 & 0.48 & 0.48 & 0.46 & 0.45\\
+Heliophila digitata & a & 30 & 0.33 & 0.38 & 0.44 & 0.38\\
+Heliophila dregeana & p & 17 & 0.33 & 0.37 & 0.33 & 0.32\\
+Heliophila elongata & p & 82 & 0.26 & 0.32 & 0.30 & 0.25\\
+Heliophila ephemera & a & 3 & 0.14 & 0.27 & 0.31 & 0.26\\
+Heliophila esterhuyseniae & p & 3 & 0.21 & 0.30 & 0.37 & 0.27\\
+Heliophila eximia & p & 12 & 0.42 & 0.41 & 0.32 & 0.34\\
+Heliophila gariepina & a & 12 & 0.50 & 0.53 & 0.48 & 0.41\\
+Heliophila glauca & p & 35 & 0.29 & 0.35 & 0.34 & 0.33\\
+Heliophila juncea & p & 150 & 0.32 & 0.37 & 0.39 & 0.35\\
+Heliophila linearis & p & 94 & 0.32 & 0.33 & 0.28 & 0.30\\
+Heliophila macowaniana & a & 31 & 0.33 & 0.38 & 0.44 & 0.39\\
+Heliophila macra & p & 22 & 0.30 & 0.30 & 0.32 & 0.29\\
+Heliophila macrosperma & p & 5 & 0.28 & 0.36 & 0.35 & 0.25\\
+Heliophila minima & p & 35 & 0.36 & 0.45 & 0.51 & 0.39\\
+Heliophila namaquana & a & 16 & 0.39 & 0.46 & 0.48 & 0.39\\
+Heliophila nubigena & p & 19 & 0.31 & 0.36 & 0.43 & 0.38\\
+Heliophila pectinata & a & 16 & 0.27 & 0.34 & 0.50 & 0.34\\
+Heliophila polygaloides & p & 12 & 0.40 & 0.48 & 0.42 & 0.34\\
+Heliophila pubescens & a & 9 & 0.31 & 0.40 & 0.48 & 0.39\\
+Heliophila pusilla & a & 45 & 0.32 & 0.38 & 0.38 & 0.34\\
+Heliophila rigidiuscula & p & 201 & 0.30 & 0.33 & 0.28 & 0.24\\
+Heliophila scoparia & p & 106 & 0.31 & 0.37 & 0.36 & 0.31\\
+Heliophila seselifolia & a & 80 & 0.36 & 0.42 & 0.45 & 0.40\\
+Heliophila suavissima & p & 92 & 0.30 & 0.39 & 0.42 & 0.31\\
+Heliophila subulata & p & 103 & 0.29 & 0.33 & 0.31 & 0.29\\
+Heliophila tricuspidata & p & 8 & 0.28 & 0.33 & 0.38 & 0.30\\
+Heliophila trifurca & a & 77 & 0.45 & 0.48 & 0.48 & 0.43\\
+Heliophila tulbaghensis & p & 3 & 0.36 & 0.41 & 0.36 & 0.35\\
+Heliophila variabilis & a & 97 & 0.35 & 0.41 & 0.40 & 0.37\\
+\bottomrule
+\addlinespace
+\insertTableNotes
+\end{longtable}
+}
+\end{ThreePartTable}
+\end{center}
+
+\hypertarget{drought-frequency.}{%
+\subsubsection{Drought frequency.}\label{drought-frequency.}}
+
+\begin{figure}
+\centering
+\includegraphics{../figures/life_history_drought_boxplots.pdf}
+\caption{\label{fig:boxplots}Comparison of drought frequency across seasons measured
+at the GBIF records of annual and perennial species of Heliophila.}
+\end{figure}
+
+\begin{table}[tbp]
+\begin{center}
+\begin{threeparttable}
+\caption{\label{tab:firthmodelstable}Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species. }
+\begin{tabular}{lll}
+\toprule
+Predictor & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{p.value}\\
+\midrule
+Intercept & 2.2575 & 0.1739\\
+Winter drought freq. & -6.7484 & 0.1661\\ \midrule
+Intercept & 4.5594 & 0.0443\\
+Spring drought freq. & -11.7895 & 0.0423\\ \midrule
+Intercept & 7.1742 & 0.0011\\
+Summer drought freq. & -18.2999 & 0.0010\\ \midrule
+Intercept & 6.4226 & 0.0029\\
+Fall drought freq. & -19.0512 & 0.0026\\ \midrule
+\bottomrule
+\addlinespace
+\end{tabular}
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Firth's penalized logistic regression. Annual species were scored as 0 and perennial species as 1.}
+\end{tablenotes}
+\end{threeparttable}
+\end{center}
+\end{table}
+
+\begin{table}[tbp]
+\begin{center}
+\begin{threeparttable}
+\caption{\label{tab:phylomodelstable}Phlyogenetically constrained logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species.
+}
+\begin{tabular}{lll}
+\toprule
+Predictor & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{p.value}\\
+\midrule
+Intercept & 0.7231 & 0.6636\\
+Winter drought freq. & -1.5452 & 0.7274\\ \midrule
+Intercept & 5.0107 & 0.0534\\
+Spring drought freq. & -12.9014 & 0.0464\\ \midrule
+Intercept & 7.7093 & 0.0054\\
+Summer drought freq. & -19.9056 & 0.0042\\ \midrule
+Intercept & 7.0162 & 0.0082\\
+Fall drought freq. & -20.8174 & 0.0067\\ \midrule
+\bottomrule
+\addlinespace
+\end{tabular}
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Annual species were scored as 0 and perennial species as 1.}
+\end{tablenotes}
+\end{threeparttable}
+\end{center}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{../figures/life_history_drought_means_line.pdf}
+\caption{\label{fig:lineplots}Comparison (mean +- SE) of drought frequency across
+seasons measured at the GBIF records of annual and perennial species of
+Heliophila.}
+\end{figure}
+
+\hypertarget{collection-dates-1}{%
+\subsubsection{Collection dates}\label{collection-dates-1}}
+
+\begin{figure}
+\centering
+\includegraphics{../figures/collection_dates.pdf}
+\caption{\label{fig:collectiondates}Collection dates of GBIF records of annual and
+perennial species of Heliophila.}
+\end{figure}
 
@@
 
-The best paper ever (Monroe et al., 2016)
+\hypertarget{summary-1}{%
+\subsection{Summary}\label{summary-1}}
 
-There is a package installer at pandoc's download page. If you later
-want to uninstall the package, you can do so by downloading this script
-and running it with perl uninstall-pandoc.pl.There is a package
-installer at pandoc's download page. If you later want to uninstall the
-package, you can do so by downloading this script and running it with
-perl uninstall-pandoc.pl.There is a package installer at pandoc's
-download page. If you later want to uninstall the package, you can do so
-by downloading this script and running it with perl
-uninstall-pandoc.pl.There is a package installer at pandoc's download
-page. If you later want to uninstall the package, you can do so by
-downloading this script and running it with perl
-uninstall-pandoc.pl.There is a package installer at pandoc's download
-page. If you later want to uninstall the package, you can do so by
-downloading this script and running it with perl
-uninstall-pandoc.pl.There is a package installer at pandoc's download
-page. If you later want to uninstall the package, you can do so by
-downloading this script and running it with perl uninstall-pandoc.pl.
+We found that the distribution of annual and perennial species of
+Heliophila is significantly predicted by satellite detected historic
+drought frequencies. Annual species are found in environments that
+experience more frequent drought during the summer and fall quarters
+compared to perennials. This relationship was consistent while
+controlling for phylogenetic relatedness among the taxa studied,
+indicating that these distributions cannot be explained entirely by
+common ancestry., These results support the hypothesis that natural
+selection has played a role in shaping the contemporary distributions of
+these alternative life-history strategies.
 
-\newpage
+\hypertarget{relationship-to-previous-hypotheseswork}{%
+\subsection{Relationship to previous
+hypotheses/work}\label{relationship-to-previous-hypotheseswork}}
+
+These findings support classical theoretical predictions about the
+adaptive value of annual and perennial life history strategies. Taken
+together, they suggest that in Heliophila, annual species are adapted to
+environments with increased summer droughts by avoiding these seasons in
+a dormant seed phase of their life cycle. Indeed, we found that very few
+annuals are collected during this season, supporting the prediction that
+they are not in a vegetative and/or reproductive phase at this time.
+Traditionally, the focus has been on the evolutionary origins of annual
+life histories \{citations\}. However, we also find evidence that the
+transition to perenniality could be explained by historical drought
+regimes. The phylogeny reveals several transitions from annual to
+perennial life history. Perennials may be able to out complete annual
+relatives in environments where the infrequency of drought favors
+strategies that allow plants to benefit from growth over many seasons.
+
+\hypertarget{caveats}{%
+\subsection{Caveats}\label{caveats}}
+
+Correlation does not prove causation. But it does indicate predictive
+power and is consistent with adaptive hypotheses. Herbarium collections
+and their associated data do not represent systematic or random sampling
+of a species distribution. Significant biases in collecting exist, which
+we have not necessarily controlled for here, and may have some effect on
+our findings, such as a bias toward collecting near roads or near the
+locations of natural history collections (Daru et al.~2018, Heberling,
+in press). Despite these biases, the Cape Floristic region is a
+biodiversity hotspot and one of the most botanically well sampled
+regions on Earth (Daru et al.~2018, Heberling, in press?), suggesting
+that this may currently be the optimal region for our analyses of life
+history distribution. Future research will benefit from systematic
+sampling efforts to avoid these noted biases. The climate data used here
+are assessed only from 198X-XXXX and do not reflect conditions at the
+estimated divergence dates of these species (XXX million of years ago).
+Rather, the results suggest that the current distributions of annual and
+perennial species reflect a history of environmental filtration and
+ongoing natural selection. That is, their distributions are non-random
+with respect to historic drought and this is not explained by phylogeny.
+
+\hypertarget{broader-implications}{%
+\subsection{Broader implications}\label{broader-implications}}
+
+These findings suggest that rapidly changing drought regimes threaten
+species adapted to current environments. Studies predict changing
+drought regimes. This could have impacts on ecosystem functioning. This
+should also be considered when thinking about using perennial crops.
+Studies predict changing drought regimes.
+
+\hypertarget{conclusions}{%
+\subsection{Conclusions}\label{conclusions}}
+
+Perenniality appears to be adaptive in environments with less frequent
+drought. This work demonstrates the power of emerging data to address
+outstanding classic hypotheses in ecology and evolution.
+
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
+
+This work was supported by NSF no. USDA grant no. XXX to J.G.M.
 
@@
 \section{References}\label{references}}
+
+\newpage
 
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-R-geiger_a}{}%
+Alfaro, M., Santini, F., Brock, C., Alamillo, H., Dornburg, A., Rabosky,
+D., \ldots{} Harmon, L. (2009). Nine exceptional radiations plus high
+turnover explain species diversity in jawed vertebrates.
+\emph{Proceedings of the National Academy of Sciences of the United
+States of America}, \emph{106}, 13410--13414.
+
 \leavevmode\hypertarget{ref-R-papaja}{}%
@@
 R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}
+
+\leavevmode\hypertarget{ref-R-Matrix}{}%
+Bates, D., \& Maechler, M. (2018). \emph{Matrix: Sparse and dense matrix
+classes and methods}. Retrieved from
+\url{https://CRAN.R-project.org/package=Matrix}
+
+\leavevmode\hypertarget{ref-dittberner2018natural}{}%
+Dittberner, H., Korte, A., Mettler-Altmann, T., Weber, A., Monroe, G.,
+\& Meaux, J. de. (2018). Natural variation in stomata size contributes
+to the local adaptation of water-use efficiency in arabidopsis thaliana.
+\emph{bioRxiv}, 253021.
+
+\leavevmode\hypertarget{ref-R-geiger_b}{}%
+Eastman, J., Alfaro, M., Joyce, P., Hipp, A., \& Harmon, L. (2011). A
+novel comparative method for identifying shifts in the rate of character
+evolution on trees. \emph{Evolution}, \emph{65}, 3578--3589.
+
+\leavevmode\hypertarget{ref-R-mvtnorm}{}%
+Genz, A., \& Bretz, F. (2009). \emph{Computation of multivariate normal
+and t probabilities}. Heidelberg: Springer-Verlag.
+
+\leavevmode\hypertarget{ref-R-MCMCglmm}{}%
+Hadfield, J. D. (2010). MCMC methods for multi-response generalized
+linear mixed models: The MCMCglmm R package. \emph{Journal of
+Statistical Software}, \emph{33}(2), 1--22. Retrieved from
+\url{http://www.jstatsoft.org/v33/i02/}
+
+\leavevmode\hypertarget{ref-R-geiger_d}{}%
+Harmon, L., Weir, J., Brock, C., Glor, R., \& Challenger, W. (2008).
+GEIGER: Investigating evolutionary radiations. \emph{Bioinformatics},
+\emph{24}, 129--131.
+
+\leavevmode\hypertarget{ref-R-logistf}{}%
+Heinze, G., \& Ploner, M. (2018). \emph{Logistf: Firth's bias-reduced
+logistic regression}. Retrieved from
+\url{https://CRAN.R-project.org/package=logistf}
 
@@
 
-\leavevmode\hypertarget{ref-monroe2016adaptation}{}%
-Monroe, J. G., McGovern, C., Lasky, J. R., Grogan, K., Beck, J., \&
-McKay, J. K. (2016). Adaptation to warmer climates by parallel
-functional evolution of cbf genes in arabidopsis thaliana.
-\emph{Molecular Ecology}, \emph{25}(15), 3632--3644.
+\leavevmode\hypertarget{ref-R-phylolm}{}%
+Ho, L. S. T., \& Ane, C. (2014). A linear-time algorithm for gaussian
+and non-gaussian trait evolution models. \emph{Systematic Biology},
+\emph{63}, 397--408.
+
+\leavevmode\hypertarget{ref-Mojica2016}{}%
+Mojica, J. P., Mullen, J., Lovell, J. T., Monroe, J. G., Paul, J. R.,
+Oakley, C. G., \& McKay, J. K. (2016). Genetics of water use physiology
+in locally adapted Arabidopsis thaliana. \emph{Plant Science}.
+doi:\href{https://doi.org/10.1016/j.plantsci.2016.03.015}{10.1016/j.plantsci.2016.03.015}
+
+\leavevmode\hypertarget{ref-monroe2018drought}{}%
+Monroe, J. G., Powell, T., Price, N., Mullen, J., Howard, A., Evans, K.,
+\ldots{} McKay, J. (2018). Drought adaptation in nature by extensive
+genetic loss-of-function. \emph{bioRxiv}, 372854.
+
+\leavevmode\hypertarget{ref-mummenhoff2005phylogeny}{}%
+Mummenhoff, K., Al-Shehbaz, I. A., Bakker, F. T., Linder, H. P., \&
+Mühlhausen, A. (2005). Phylogeny, morphological evolution, and
+speciation of endemic brassicaceae genera in the cape flora of southern
+africa. \emph{Annals of the Missouri Botanical Garden}, 400--424.
+
+\leavevmode\hypertarget{ref-R-bindrcpp}{}%
+Müller, K. (2018). \emph{Bindrcpp: An 'rcpp' interface to active
+bindings}. Retrieved from
+\url{https://CRAN.R-project.org/package=bindrcpp}
 
@@
 Retrieved from \url{https://CRAN.R-project.org/package=tibble}
+
+\leavevmode\hypertarget{ref-R-caper}{}%
+Orme, D., Freckleton, R., Thomas, G., Petzoldt, T., Fritz, S., Isaac,
+N., \& Pearse, W. (2018). \emph{Caper: Comparative analyses of
+phylogenetics and evolution in r}. Retrieved from
+\url{https://CRAN.R-project.org/package=caper}
+
+\leavevmode\hypertarget{ref-R-ape}{}%
+Paradis, E., \& Schliep, K. (2018). Ape 5.0: An environment for modern
+phylogenetics and evolutionary analyses in R. \emph{Bioinformatics},
+\emph{xx}, xxx--xxx.
 
@@
 
+\leavevmode\hypertarget{ref-R-coda}{}%
+Plummer, M., Best, N., Cowles, K., \& Vines, K. (2006). CODA:
+Convergence diagnosis and output analysis for mcmc. \emph{R News},
+\emph{6}(1), 7--11. Retrieved from
+\url{https://journal.r-project.org/archive/}
+
+\leavevmode\hypertarget{ref-R-gee}{}%
+R by Thomas Lumley, V. J. C. P. to, \& author., B. R. N. that
+maintainers are not available to give advice on using a package they did
+not. (2015). \emph{Gee: Generalized estimation equation solver}.
+Retrieved from \url{https://CRAN.R-project.org/package=gee}
+
 \leavevmode\hypertarget{ref-R-base}{}%
@@
 Retrieved from \url{https://www.R-project.org/}
+
+\leavevmode\hypertarget{ref-R-phytools}{}%
+Revell, L. J. (2012). Phytools: An r package for phylogenetic
+comparative biology (and other things). \emph{Methods in Ecology and
+Evolution}, \emph{3}, 217--223.
+
+\leavevmode\hypertarget{ref-R-maps}{}%
+Richard A. Becker, O. S. code by, Ray Brownrigg. Enhancements by Thomas
+P Minka, A. R. W. R. version by, \& Deckmyn., A. (2018). \emph{Maps:
+Draw geographical maps}. Retrieved from
+\url{https://CRAN.R-project.org/package=maps}
+
+\leavevmode\hypertarget{ref-R-geiger_c}{}%
+Slater, G., Harmon, L., Wegmann, D., Joyce, P., Revell, L., \& Alfaro,
+M. (2012). Fitting models of continuous trait evolution to incompletely
+sampled comparative data using approximate bayesian computation.
+\emph{Evolution}, \emph{66}, 752--762.
+
+\leavevmode\hypertarget{ref-R-diagram}{}%
+Soetaert, K. (2017). \emph{Diagram: Functions for visualising simple
+graphs (networks), plotting flow diagrams}. Retrieved from
+\url{https://CRAN.R-project.org/package=diagram}
+
+\leavevmode\hypertarget{ref-R-shape}{}%
+Soetaert, K. (2018). \emph{Shape: Functions for plotting graphical
+shapes, colors}. Retrieved from
+\url{https://CRAN.R-project.org/package=shape}
+
+\leavevmode\hypertarget{ref-R-MASS}{}%
+Venables, W. N., \& Ripley, B. D. (2002). \emph{Modern applied
+statistics with s} (Fourth.). New York: Springer. Retrieved from
+\url{http://www.stats.ox.ac.uk/pub/MASS4}
 

### 2018-11-24 src/GBIF_records_drought_frequency.R

@@ -332,7 +332,7 @@ model_summaries<-data.frame()
 # non-phylogenetic logistic regression firth penalized
 species_means$life_history_num<-as.numeric(species_means$life_history)-1
 row.names(species_means)<-gsub(" ", "_", species_means$species)
-
+seasons<-c("Winter","Spring","Summer","Fall")
 for(s in 1:length(seasons)){
   fit1 <- logistf(species_means$life_history_num~species_means[,seasons[s]], firth = TRUE)
   sum<-summary(fit1)

### 2018-11-26 manuscript/manuscript.Rmd
round 2

--- 2018-11-24 manuscript/manuscript.Rmd
@@
 ---
-title             : "Plant life history strategies predicted by satellite-detected drought"
-shorttitle        : ""
+title             : "Satellite-detected drought frequencies predict plant life history strategies"
+shorttitle        : "Drought and life history"
 
 author: 
-  - name          : "J Grey Author"
+  - name          : "J. Grey Monroe"
     affiliation   : "1,2"
@@
     affiliation   : "3"
-  - name          : "Brian Gill"
+  - name          : "Kathryn Turner"
     affiliation   : "4"
+  - name          : "John K. McKay"
+    affiliation   : "2"
 
@@
   - id            : "1"
-    institution   : "Graduate Degree Program in Ecology, Colorado State University"
+    institution   : "Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80523, USA"
   - id            : "2"
-    institution   : "College of Agriculture, Colorado State University"
+    institution   : "College of Agriculture, Colorado State University, Fort Collins, CO 80523, USA"
   - id            : "3"
-    institution   : "Department of Ecology and Evolutionary Biology"
+    institution   : "Institute for Environment and Society, Brown University, Providence, RI 02912, USA"
   - id            : "4"
-    institution   : "Biology Department, Penn State University"
+    institution   : "Biology Department, Pennsylvania State University, State College, PA 16802, USA"
 
-authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
+abstract: Explaining variation in life history strategies is a long-standing goal of evolutionary biology.
+For plants, annual and perennial life histories are thought to reflect adaptation to environments that differ in the frequency of environmental stress such as drought.
+Here we test this hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Southern Africa using herbarium occurrence records and satellite-detected drought histories.
+We find that perennial \textit{Heliophila} species are found in environments where droughts are significantly less frequent compared to annuals.
+These associations are predictive while controlling for phylogeny, lending support to the hypothesis that drought related natural selection has influenced the  distributions of these strategies.
+Additionally, the difference in drought frequency between annual and perennial species distributions is greatest during the summer and fall, which also appears to be when annuals are in the seed phase of their life cycle based on collection dates of annual species.
+Together, these findings support traditional hypotheses about the drivers of life history strategy in plants - that perenniality is favored in environments with less frequent drought and annuality in enviroments with more frequent drought by allowing them to escape drought prone seasons as seeds.
 
-abstract: |
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "drought, life history, remote sensing, phylogeography"
-wordcount         : ""
+keywords          : "drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records"
 
@@
 output            : papaja::apa6_pdf
+
+header-includes   : 
+  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+
 ---
@@
 library(ggplot2)
+library(logistf)
+#library(gee) 
+#library(phytools) 
+library(geiger) 
+library(ape) 
+#library(caper) 
+#library(MCMCglmm) 
+library(phylolm)
+library(gridExtra)
+library(cowplot)
+library(rasterVis)
+library(ggtree)
+library(grid)
+library(png)
 ```
 
-#Introduction
+# Introduction
 
-Life history 
-Annual vs perennial
-Hypotheses
-Summary
+## Life history variation in plants
 
-# Methods
+Understanding the causes and consequences of life history variation is a longstanding goal of ecology and evolutionary biology  [@cole1954population].
+In plants, life histories are especially diverse, with herbaceous species that complete their life cycle in a number of weeks to trees that live for thousands of years [@brown1996oldlist].
+Along this continuum an important division exists, distinguishing annuals (i.e.
+monocarpic or semelparous) which complete their seed to seed life cycle within a single calendar year from perennials (i.e.
+polycarpic or iteroparous] which can persist over multiple years.
+Annual plants flower once, set seed, senesce, and then die, spending at least some portion of the year as a seed.
+In contrast, perennial plants can continue vegetative growth after reproduction and experience all seasons.
+These represent fundamentally different life history strategies, but the ecological factors that explain their evolution remain uresolved [@friedman2015all].
+
+Identifying the drivers of selection for these alternative strategies is important because annuals and perennials have differing impacts on ecosystem functioning.
+For example, perennials have higher nitrogen concentration and larger specific root length in annuals, which both affect nutrient cycling [@garnier1997specific; @roumet2006suites].
+Furthermore, predicting the conditions that favor these strategies in nature is useful for developing more sustainable perennial cropping systems [@lelievre2009current].
+
+Classical theory predict shorter life spans and annuality in environments where adult mortality is high [@charnov1973life; @stearns1992evolution;@franco1996life].
+In plants, this has been extended to the hypothessis that annuality is adaptive when it allows plants to escape drought [@schaffer1975selection].
+Lack of water is perhaps the greatest threat to survival during vegetatitive or reproductive growth and annuals can remain dormant (and protected as a seed) during drought.
+Thus, environments with greater seasonal drought frequency may select for annual life histories that complete reproduction prior to drought prone seasons.
+Conversely, environments with less frequent drought may select for perennial species, which may benefit from multiple bouts of reproduction and competitive advantage [@corbin2004competition].
+These prediction has been supported by the association of annualilty with arid environments in wild rice [@morishima1984differentiation] and _Oenothera_ [@evans2005climate].
+Additionally, annual and perennial species of Nemesia were qualitatively associated with winter rather and summer rainfall environments respectively [@datson2008climate] and annual species of Scorzoneroides were associated with environments classified as unpredictable [@cruz2009molecular].
+These findings provide qualitative evidence supporting the prediction that annual species are found in environments that experience more frequent drought, and vice versa, but whether the history frequency of drought events indeed predicts the distributions annual or perennial life history strategies has yet to be tested.
+
+Here we combine a long-term global dataset of satellite detected drought events with metadata from natural collections to test these classic hypotheses about the evolution of life history strategies within the African endemic mustard genus, _Heliophila_ L.
+(Brassicaceae).
+If annuallity is an adaptive strategy allowing plants to escape drought prone seasons, then drought frequency should predict the distribution of life history strategies across landscapes, and annual species should be more commonly associated with drought prone regions than perennial species.
+Furthermore, if annual species have adapted to escape drought prone seasons as seeds, observations of annual species should be rare during drought prone seasons.
+Phylogenetic relatedness can have significant non-random effects on species distributions and life history traits [], and therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetically controlled background.
+
+# Materials and Methods
+
+## Availability
+All data and the source code to produce this manuscript are available at https://github.com/greymonroe/heliophila.
+
 ## Data
-### Life history
-### Phyologeny
-### Herbarium Specimens
+### Satellite-detected drought data
+```{r Droughtdata}
+Winter<-raster("../data/drought/Summer_drought_freq")
+Spring<-raster("../data/drought/Fall_drought_freq")
+Summer<-raster("../data/drought/Winter_drought_freq")
+Fall<-raster("../data/drought/Spring_drought_freq")
+```
+
+Remotely sensed data is a powerful tool for characterizing seasonal patterns in drought because it is less limited in spatial and temporal scope and resolution than weather stations or field observations [@aghakouchak2015remote].
+To quantify the frequency of drought during different seasons across landscapes, we used the remotely sensed Vegetative Health Index (VHI), which measures landscape scale reductions in plant cover and temperature conditions characteristic of drought [@kogan2001operational].
+Generated from data collected by NOAA AVHRR satellites since 1981, the VHI combines normalized difference vegetation index (NDVI) derived measures of vegetative stress (Vegetative Condition Index - VCI) with temperature stress indicated by anomalies in thermal spectra (Temperature Condition Index - TCI).
+
+The VHI of year $y$ during week $w$ of $[1,52]$ at pixel $i$ is derived from the following equations, where $n$ is the number of years observed.
+
+$$VCI_{y,w,i} = 100\frac{NDVI_{y,w,i} - NDVI_{min}}{NDVI_{max} - NDVI_{min}}$$
+
+where $NDVI_{min} = min(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})$ and $NDVI_{max} = max(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})$
+
+$$TCI_{y,w,i} = 100\frac{T_{y,w,i} - T_{min}}{T_{max} - T_{min}}$$
+
+where $T_{min} = min(T_{1981,w,i}...T_{1981+n,w,i})$ and $T_{max} = max(T_{1981,w,i}...T_{1981+n,w,i})$
+
+$$VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})$$
+
+Thus, VHI measurements are standardized according to conditions historically observed at each locations.
+These measurements have been validated and generally used for evaluating drought risk and predicting crop yields in agriculture [e.g., @rojas2011assessing; @kogan2016modelling].
+But they also present a tool to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes.
+Indeed, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting infraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought].
+Here, we accessed VHI data at $16km^2$ resolution from 1981 to 2015 (https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php) to characterize the seasonal drought frequencies experienced by annual and perennial _Heliophila_ sepecies.
+
+### Life history data for Heliophila
+_Heliophila_ is a genus of flowering plants endemic to the southern portion of Africa including the Cape Floristic and Succulent Karoo Regions.
+These are among the most botanically diverse environments on Earth and the estimated ~50 _Heliophila_ species are considered to be the most diverse genus of the family Brassicaceae [@mummenhoff2005phylogeny].
+This genus includes both perennial and annual species and this change in life history strategy has likely arisen multiple independent times [@appel1997generic;@mummenhoff2005phylogeny].
+We used life histories reported by @mummenhoff2005phylogeny, grouping species with annual or perennial life histories.
+Perenniality was defined based on the traditional defition and included any form of perennial life history (e.g., herbs, shrubs, mixed, etc).
+
+### Heliophila herbarium specimens
 ```{r raw_GBIF}
@@
 ```
-We downloaded `r nrow(GBIF)` records from GBIF.
+Botanists have collected and maintained over 350 million botanical specimens worldwide over the past 300 years.
+Herbarium specimens and their associated metadata have been used since the 1960s to study species’ geographical distributions (reviewed by  @willis2017old and @lang2018using).
+And as they become digitized [@soltis2017digitization], these collections have been used to study relationships between trait distributions, geography, and climate [@davis2015herbarium;@wolf2016altitudinal;@stropp2016mapping;@vaclavik2017effects].
+To characterize the diributions of annual and perennial _Heliophila_ species, all records for the genus _Heliophila_ were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018.
 
-### Drought
+### Sequence data for phylogeny
+Aligned Heliophila ITS sequences were obtained from previous work by @mandakova2012whole.
+Aethionema, Alliaria, Cardamine, Chamira, and Rorippa ITS records from were downloaded from Genbank [] and aligned together with Heliophila ITS sequences using MAFFT as outgroups [].
 
 ## Analyses
-### Phylogeny
-### Drought frequency
-### Contrast
-We used `r cite_r("r-references.bib")` for all our analyses.
+### Drought frequency calculations
+
+To characterize drought regimens across the distrubtions of annual and perennial species of _Heliohpila_, we calculated drought during different seasons at the location of observations for _Heliophila_ records using the VHI.
+Specifically, we created global maps of the frequencies of observing drought conditions (VHI<40, NOAA) during the winter (quarter surrounding winter solstice), spring (quarter surrounding spring equinox), summer (quarter surrounding summer solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015.
+From these maps, the drought frequency during the winter, spring, summer, and fall were extracted for the locations of all GBIF records.
+
+### Herbarium record quality control
+```{r cleanup}
+species_life_history<-read.csv("../data/Heliophila_life_history.csv")
+species_life_history$species<-gsub("_", " ", species_life_history$species)
+species_life_history$species <- as.factor(species_life_history$species)
+
+GBIF <- GBIF %>%
+  mutate(Winter=raster::extract(Winter, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(Spring=raster::extract(Spring, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(Summer=raster::extract(Summer, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(Fall=raster::extract(Fall, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(ymd_lat_lon_sp = paste0(year,"_",month,"_",day,"_", decimalLatitude,"_", decimalLongitude,"_", species))
+
+GBIF_cleaned <- GBIF %>%
+  filter(species %in% species_life_history$species) %>% 
+  filter(!is.na(decimalLatitude)) %>%
+  filter(hasGeospatialIssues==FALSE) %>%
+  filter(decimalLatitude<(-20)) %>%
+  filter(decimalLongitude<40) %>%
+  filter(!duplicated(ymd_lat_lon_sp)) %>%
+  merge(species_life_history) %>%
+  filter(!is.na(Summer))
+```
+
+To remove instance with spurious location data, we filtered raw GBIF by restricting our analyses to include only:  
+* Species with reported life history  
+* Records with geospatial data  
+* Records without known geospatial issues  
+* Records from collection sites classified as land pixels   
+* Records from Africa  
+* Non-duplicate records (ie. identical species, location, collection date)
+
+### Phylogeny construction
+Model selection for construction of phylogeny was performed in jModeltest2 with CIPRES (cite).
+Based on this analysis, $GTR + L$ were selected.
+Ultrametric phylogeny was estimated with branch lengths as relative time (details).
+
+### Comparison of drought frequency between annual and perennial
+```{r speciesmeans}
+species_means <- GBIF_cleaned %>%
+  dplyr::select(species, life_history, Winter, Summer, Spring, Fall) %>%
+  group_by(species, life_history) %>%
+  summarise(n=n(), Winter_mean=mean(Winter), Spring_mean=mean(Spring), Summer_mean=mean(Summer), Fall_mean=mean(Fall)) %>% 
+  as.data.frame()
+
+```
+```{r firthmodels, echo=F,  results="hide"}
+
+firthmodelstable<-data.frame()
+
+# non-phylogenetic logistic regression firth penalized
+species_means$life_history_num<-as.numeric(species_means$life_history)-1
+row.names(species_means)<-gsub(" ", "_", species_means$species)
+seasons<-c("Winter_mean", "Spring_mean", "Summer_mean","Fall_mean")
+for(s in 1:length(seasons)){
+  fit1 <- logistf(species_means$life_history_num~(species_means[,seasons[s]]), firth = TRUE)
+  sum<-summary(fit1)
+  row.names(sum$coefficients)<-NULL
+  sum_df<-data.frame(Predictor=c("Intercept", gsub("_mean", " drought freq.", seasons[s])), 
+                     Estimate=sum$coefficients, 
+                     p.value=sum$prob)
+  row.names(sum_df)<-NULL
+  firthmodelstable<-rbind(firthmodelstable, sum_df)
+}
+
+```
+```{r phylomodels, results="hide"}
+# phylogentic constraint
+tree<-read.nexus("../data/TrimmedMCCwoPolytomies.nex")
+tree<- rescale(tree, "depth", 1)
+
+phylomodelstable<-data.frame()
+
+for(s in 1:length(seasons)){
+  fit1 <-  phyloglm(species_means$life_history_num~species_means[,seasons[s]], data=species_means, phy=tree)
+  sum<-summary(fit1)
+  row.names(sum$coefficients)<-NULL
+  sum_df<-data.frame(Predictor=c("Intercept", gsub("_mean", " drought freq.", seasons[s])), sum$coefficients)
+  phylomodelstable<-rbind(phylomodelstable, sum_df)
+}
+
+phylomodelstable <- phylomodelstable[,c( "Predictor","Estimate","p.value")]
+
+```
+To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimes, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency.
+The relationship between drought frequencies across each taxon’s range and life habitat (annual or perennial) was evaluated using Firth's penalized-likelihood logistic regression and phylogenetic logistic regression.
+
+### Collection dates
+
+To test the hypothesis that annual species have adapted to escape drought prone seasons as seeds, collection dates for herbarium specimens were compared between annual and perennial species.
+Comparisons of distributions were made by Two-sample Kolmogorov-Smirnov test, t-test, and Barlett variance test.
 
@@
 
-(ref:foo) A scatterplot of the data `cars` using **base** R graphics.
+To test the hypothesis that annual and perennial plants reflect adaptation to alternative drought environments we examined the landscape distribution of life history strategies in the large and diverse mustard genus, _Heliophila_ Figure \@ref(fig:phylogeny).
+Using both herbarium specimen metadata and a 30 year dataset of satellite generated climate information, we tested the prediction that annual species  are more often observed in drought-prone locations than perennial species, when controlling for phylogenetic relatedness.
+We found that drought frequency is significantly different between the distributions of annual and perennial species, with annuals being found in environments with significantly more frequent drought, and that this signal is strongest during the summer.
+These results remain significant while controlling for the phylogenetic relationships of Heliophila species, yielding support for the role that natural selection has played in driving contemporary distributions of these alternatives strategies in relation to drought regimes.
 
-```{r foo, fig.cap='(ref:foo)',out.width = "\\textwidth", fig.pos = "!h"}
-plot(cars)  # a scatterplot
+(ref:phylogeny) Phylogeny of Heliophila.
+
+```{r makephylogeny, eval=T, message=F, warning="hide", results="hide"}
+
+pdf("../figures/phylogeny.pdf", height=6, width=5)
+
+tree<-read.nexus("../data/TrimmedMCCwoPolytomies.nex")
+tree<- rescale(tree, "depth", 1)
+tree$tip.label<-gsub("_"," ", tree$tip.label)
+
+treeplot<-ggtree(tree, layout = "rectangular")+
+  xlim(c(0,2))+
+  geom_tippoint(size=3,col=ifelse(species_life_history$life_history[match(tree$tip.label,species_life_history$species)]=="a", "orange3", "dodgerblue3"))+
+  geom_tiplab(offset = .05 , size=3)
+
+minima <- readPNG("../pictures/minima.png")
+minima <- rasterGrob(minima, interpolate=TRUE)
+#https://www.gbif.org/occurrence/1099023487
+
+deserticola <- readPNG("../pictures/deserticola.png")
+deserticola <- rasterGrob(deserticola, interpolate=TRUE)
+#https://www.gbif.org/occurrence/1057389408
+
+coronopifolia<-readPNG("../pictures/coronopifolia.png")
+coronopifolia <- rasterGrob(coronopifolia, interpolate=TRUE)
+#https://www.gbif.org/occurrence/1099023562
+
+ephemera<-readPNG("../pictures/ephemera.png")
+ephemera <- rasterGrob(ephemera, interpolate=TRUE)
+#https://www.gbif.org/occurrence/1099023490
+
+pics<-plot_grid(minima, deserticola, coronopifolia, ephemera,
+                labels = c("b","c","d","e"), ncol = 1)
+
+plot_grid(treeplot, pics, ncol=2,labels=c("a",""), rel_widths = c(1,0.3))
+
+dev.off()
+
+```
+```{r phylogeny, fig.cap='(ref:phylogeny)', out.width = "\\textwidth", fig.pos = "!h"}
+knitr::include_graphics("../figures/phylogeny.pdf", dpi = 108)
 ```
 
-First we did this (Fig. \@ref(fig:foo).)
+### GBIF records
+
+Out of `r nrow(GBIF)` _Heliophila_ GBIF recrods,  `r sum(GBIF$species %in% species_life_history$species)` were for species with reported life history [@mummenhoff2005phylogeny], `r sum(!is.na(GBIF$decimalLatitude))` had geospatial data, `r sum(!(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)))` did not have geospatial issues, `r sum(!(is.na(GBIF$Summer)) & !(is.na(GBIF$decimalLatitude)))` were located on pixels classified as land having drought measurements, `r sum(GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude))` were located in Africa, `r sum(!duplicated(GBIF$ymd_lat_lon_sp) & !is.na(GBIF$decimalLatitude)) ` were not duplicated.
+After all filtering steps, `r nrow(GBIF_cleaned)` records for `r length(unique(GBIF_cleaned$species))` species (Figure \@ref(fig:mapsboxplots), Table \@ref(tab:speciesmeanstable)) passed.
+
+### Drought frequency
+
+Drought frequency was signicantly different between annual and perennial species.
+
+(ref:mapsboxplots) Locations of annual and perennial Heliophila.
+Drought frequency during the winter, spring, summer and fall.
+Drought frequencies during each season observed at the collection sites of Heliophila records.
+
+```{r make_maps_boxplots, eval=T, message=F, results="hide"}
+
+pdf("../figures/maps_boxplots.pdf", width=5, height=7.5)
+world <- map_data("world") # we already did this, but we can do it again
+library(cowplot)
+
+perennialmap<-ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
+  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
+  geom_point(GBIF_cleaned %>% filter(life_history=="p"), mapping=aes(x=decimalLongitude, y=decimalLatitude), pch=3, col="dodgerblue3", alpha=0.3) +
+  theme_classic()+
+  theme( text = element_text(size=8))+
+  labs(title="Perennial species", x="Longitude (°)", y="Latitude (°)")+
+  annotate("text", x=32, y=-34, label=paste("n obs. =", nrow(GBIF_cleaned %>% filter(life_history=="p"))), size=2)
+
+annualmap<-ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group=group),  lwd=.5, fill = NA, color = "black") + 
+  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
+  geom_point(GBIF_cleaned %>% filter(life_history=="a"), mapping=aes(x=decimalLongitude, y=decimalLatitude), pch=3, col="orange3", alpha=0.3) +
+  theme_classic()+
+  theme(text = element_text(size=8))+
+  labs(title="Annual species", x="Longitude (°)", y="Latitude (°)")+
+  annotate("text", x=32, y=-34, label=paste("n obs. =", nrow(GBIF_cleaned %>% filter(life_history=="a"))), size=2)
+
+Winter_SA <-crop(Winter, rbind(c(12,37),c(-35, -18)))
+
+wintermap<-gplot(Winter_SA) + 
+  geom_tile(aes(fill=value)) + 
+  scale_fill_gradientn(colors=c( 'white',"gray80", "gray20", 'black'), na.value="white", name=NULL) +
+  geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
+  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
+  theme_classic()+
+  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
+  labs(title="Winter drought freq.", x="Longitude (°)", y="Latitude (°)")
+
+Spring_SA <-crop(Spring, rbind(c(12,37),c(-35, -18)))
+
+springmap<-gplot(Spring_SA) + 
+  geom_tile(aes(fill=value)) + 
+  scale_fill_gradientn(colors=c( 'white',"gray80", "gray20", 'black'), na.value="white", name=NULL) +
+  geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
+  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
+  theme_classic()+
+  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
+  labs(title="Spring drought freq.", x="Longitude (°)", y="Latitude (°)")
+
+Summer_SA <-crop(Summer, rbind(c(12,37),c(-35, -18)))
+
+summermap<-gplot(Summer_SA) + 
+  geom_tile(aes(fill=value)) + 
+  scale_fill_gradientn(colors=c( 'white',"gray80", "gray20", 'black'), na.value="white", name=NULL) +
+  geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
+  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
+  theme_classic()+
+  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
+  labs(title="Summer drought freq.", x="Longitude (°)", y="Latitude (°)")
+
+Fall_SA <-crop(Fall, rbind(c(12,37),c(-35, -18)))
+
+fallmap<-gplot(Fall_SA) + 
+  geom_tile(aes(fill=value)) + 
+  scale_fill_gradientn(colors=c( 'white',"gray80", "gray20", 'black'), na.value="white", name=NULL) +
+  geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
+  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
+  theme_classic()+
+  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
+  labs(title="Fall drought freq.", x="Longitude (°)", y="Latitude (°)")
+
+GBIF_cleaned_long<- GBIF_cleaned %>%
+  dplyr::select(species, life_history, Winter, Spring, Summer,Fall) %>%
+  gather(season, drought_freq, Winter, Spring, Summer,Fall)
+
+GBIF_cleaned_long$season<-factor(GBIF_cleaned_long$season, levels=c("Winter","Spring","Summer","Fall"))
+
+boxplots<-ggplot(GBIF_cleaned_long, aes(x=life_history, col=life_history, y=drought_freq))+
+  facet_wrap(~season, ncol=4 )+
+  geom_jitter(width = .3, alpha=0.1, pch=3)+
+  geom_boxplot(outlier.colour = "white", outlier.fill = "white", fill=NA, width=.3)+
+  scale_color_manual(values=c("orange3","dodgerblue3"))+
+  theme_classic()+
+  theme(legend.position = "none", text = element_text(size=8))+
+  scale_x_discrete(name="Life history", labels=c("Annual","Perennial"))+
+  labs(y="Drought frequency")
+
+maps<-plot_grid(annualmap, 
+          perennialmap, 
+          wintermap, 
+          springmap,
+          summermap,
+          fallmap, 
+          labels = "auto", ncol = 2)
+plot_grid(maps, boxplots, ncol=1,labels=c("","g"), rel_heights = c(.75,.35))
+
+dev.off()
+
+```
+
+```{r mapsboxplots, fig.cap='(ref:mapsboxplots)', out.width = "\\textwidth", fig.pos = "!h"}
+knitr::include_graphics("../figures/maps_boxplots.pdf", dpi = 108)
+```
+
+```{r modelstable, results = 'asis', echo = F}
+
+modelstable<-bind_cols(firthmodelstable,phylomodelstable)
+
+modelstable<- modelstable %>%
+  dplyr::select(Predictor, Estimate, p.value,Estimate1, p.value1)
+
+colnames(modelstable)<-c("Predictor", "Estimate`","P`","Estimate*","P*")
+
+apa_table(modelstable
+  , caption = "Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species the winter, spring, summer, and fall."
+  , note = "` = Firth's penalized logistic regression. * = Phylogenetically constrained logistic regression. Annual species were scored as 0 and perennial species as 1.",
+  midrules=c(2,4,6,8),
+  digits=4
+)
+
+```
+
+### Collection dates
+
+(ref:lineplots) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of Heliophila.
+(B) Collection dates of GBIF records of annual and perennial species of Heliophila.
+
+```{r makelineplots, eval=T, message=F, results="hide"}
+
+pdf("../figures/line_and_dates.pdf", height=5, width=5)
+
+life_history_drought_means<- species_means %>% 
+  dplyr::select(life_history, Winter_mean, Spring_mean, Summer_mean,Fall_mean)  %>%
+  gather(variable, value, -life_history) %>%
+  group_by(life_history, variable) %>%
+  summarise(mean = mean(value), se = sd(value)/sqrt(length(value))) %>%
+  mutate(variable = reorder(variable, c(4,2,3,1)))
+
+lines<-ggplot(life_history_drought_means, aes(x=variable, col=life_history,group=life_history, y=mean))+
+  theme_classic()+
+  geom_point(size=2)+
+  geom_line(lwd=0.5, lty=1)+
+  geom_errorbar(width=0.25,lwd=1, aes(ymin=mean-se, ymax=mean+se))+
+  scale_y_continuous(limits=c(0.3,0.45), labels = c("0.300", "0.350", "0.400", "0.450"))+
+  scale_x_discrete(labels=c("Winter","Spring","Summer","Fall"))+
+  scale_color_manual(values=c( "orange3","dodgerblue3"), name="Life history", labels=c("Annual", "Prennial"))+
+  labs(x="Season", y="Drought Frequency")+
+  annotate("text", x=2, y=0.44, label="*", size=7) +
+  annotate("text", x=3:4, y=0.44, label="**", size=7)
+
+GBIF_cleaned$date_calc<-as.Date(paste(
+  GBIF_cleaned$month,
+  GBIF_cleaned$day, sep="-"),"%m-%d")
+GBIF_cleaned$days<-lubridate::yday(GBIF_cleaned$date_calc)
+GBIF_cleaned$days_adjusted<-sapply(GBIF_cleaned$days, function(x)
+  if(is.na(x)){ return(NA)
+  }else if(x>135) {
+    return (x-135)
+  }else return (x+230))
+
+dates<-ggplot(GBIF_cleaned %>% filter(is.finite(days_adjusted)), aes(x=days_adjusted, fill=life_history)) +
+  geom_density(alpha=.5)+
+  scale_fill_manual(values=c("orange3","dodgerblue3"), name="Life history", labels=c("Annual", "Perennial"))+
+  theme_classic()+
+  scale_x_continuous(name="Season", breaks=c(45,135,225, 315), labels=c("Winter","Spring","Summer","Fall"))+
+  labs(title="Collection date\nof herbaria specimens", y="Density")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+plot_grid(lines, dates, ncol=1, labels="auto")
+
+annuals<-subset(GBIF_cleaned, life_history=="a")
+perennials<-subset(GBIF_cleaned, life_history=="p")
+
+#ks.test(annuals$days,perennials$days)
+t.test(days~life_history, GBIF_cleaned)
+bartlett.test(days~life_history, GBIF_cleaned)
+
+dev.off()
+
+```
+
+```{r lineplots, fig.cap='(ref:lineplots)'}
+knitr::include_graphics("../figures/line_and_dates.pdf", dpi = 108)
+```
 
 # Discussion
-The best paper ever [@monroe2016adaptation]
+## Summary
+We found that the distribution of annual and perennial species of Heliophila is significantly predicted by satellite detected historic drought frequencies.
+Annual species are found in environments that experience more frequent drought during the summer and fall quarters compared to perennials.
+This relationship was consistent while controlling for phylogenetic relatedness among the taxa studied, indicating that these distributions cannot be explained entirely by common ancestry.
+These results support the hypothesis that natural selection has played a role in shaping the contemporary distributions of these alternative life-history strategies.
 
-There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.There is a package installer at pandoc’s download page.
-If you later want to uninstall the package, you can do so by downloading this script and running it with perl uninstall-pandoc.pl.
+## Relationship to previous hypotheses/work
 
-\newpage
+Herbaria records and satellite detected drought provide data with which the distributions of annual and perennial species can be compared with respect to historic drought frequency.
+However, it is necessary to control for the demographic history caused by common ancestry of species if evolutionary processes such as natural selection are to be invoked as explanations for any differences observed.
+If, for example, annual and perennial species show significantly different ranges with respect to historic drought this could be confounded by common ancestry if annuals originated from a common ancestor and vice versa.
+In this case, it would be challenging to distinguish between natural selection and demographic history.
+On the other hand, if annual and perennial life history strategies arose independently in multiple species, controlling for phylogenetic relationships allows us to better account for demography and make stronger assertions about the importance of processes such as natural selection to explain patterns.
+
+These findings support classical theoretical predictions about the adaptive value of annual and perennial life history strategies.
+Taken together, they suggest that in Heliophila, annual species are adapted to environments with increased summer droughts by avoiding these seasons in a dormant seed phase of their life cycle.
+Indeed, we found that very few annuals are collected during this season, supporting the prediction that they are not in a vegetative and/or reproductive phase at this time.
+Traditionally, the focus has been on the evolutionary origins of annual life histories {citations}.
+However, we also find evidence that the transition to perenniality could be explained by historical drought regimes.
+The phylogeny reveals several transitions from annual to perennial life history.
+Perennials may be able to out complete annual relatives in environments where the infrequency of drought favors strategies that allow plants to benefit from growth over many seasons.
+
+## Caveats
+Sampling bias [@daru2018widespread].
+Correlation does not prove causation. But it does indicate predictive power and is consistent with adaptive hypotheses.
+Herbarium collections and their associated data do not represent systematic or random sampling of a species distribution.
+Significant biases in collecting exist, which we have not necessarily controlled for here, and may have some effect on our findings, such as a bias toward collecting near roads or near the locations of natural history collections (Daru et al.
+2018, Heberling, in press).
+Despite these biases, the Cape Floristic region is a biodiversity hotspot and one of the most botanically well sampled regions on Earth (Daru et al.
+2018, Heberling, in press?), suggesting that this may currently be the optimal region for our analyses of life history distribution.
+Future research will benefit from systematic sampling efforts to avoid these noted biases.
+The climate data used here are assessed only from 198X-XXXX and do not reflect conditions at the estimated divergence dates of these species (XXX million of years ago).
+Rather, the results suggest that the current distributions of annual and perennial species reflect a history of environmental filtration and ongoing natural selection.
+That is, their distributions are non-random with respect to historic drought and this is not explained by phylogeny.
+
+## Broader implications 
+These findings suggest that rapidly changing drought regimes threaten species adapted to current environments.
+Studies predict changing drought regimes.
+This could have impacts on ecosystem functioning.
+This should also be considered when thinking about using perennial crops.
+Studies predict changing drought regimes.
+
+Do annuals require drought? We found independent origins of perenniality in environments where droughts are infrequent. Annuals may not be able to compete with perennials in the absence of droughts.
+
+## Conclusions 
+Perenniality appears to be adaptive in environments with less frequent drought. This work demonstrates the power of emerging data to address outstanding classic hypotheses in ecology and evolution.
+
+# Acknowledgments 
+This work was supported by NSF no. USDA grant no. XXX to J.G.M.
 
 # References
+\newpage
 ```{r create_r-references}
-r_refs(file = "r-references.bib")
+#r_refs(file = "r-references.bib")
 ```
-
 \begingroup
@@
 \endgroup
+
+\newpage
+\beginsupplement
+
+# Supplement
+
+### Software used
+We used `r cite_r("r-references.bib")` for all our analyses.
+
+### Supplementary tables and figures
+```{r speciesmeanstable, results = 'asis', echo = F}
+species_means_table<-species_means
+row.names(species_means_table) <- NULL
+species_means_table<-species_means_table %>% dplyr::select(-life_history_num)
+colnames(species_means_table)<-c("Species","LH","n","Winter","Spring","Summer","Fall")
+apa_table(
+  species_means_table
+  , caption = "Heliophila species records and the mean drought frequencies during different seasons at the location of records "
+  , note = "LH = Life history (a = annual, p = perennial). n=sample size of GBIF records. Seasons are mean drought frequencies observed at locations of records.",
+  digits=2,
+  font_size="small",
+  longtable = TRUE
+)
+```

### 2018-11-26 manuscript/manuscript.tex
round 2

@@
 \hypersetup{unicode=true,
-            pdftitle={Plant life history strategies predicted by satellite-detected drought},
-            pdfauthor={J Grey Monroe, Brian Gill, \& Kathryn Turner},
+            pdftitle={Satellite-detected drought frequencies predict plant life history strategies},
+            pdfauthor={J. Grey Monroe, Brian Gill, Kathryn Turner, \& John K. McKay},
             pdfkeywords={drought adaptation, life history evolution, remote sensing,
@@
 
-  \title{Plant life history strategies predicted by satellite-detected drought}
-    \author{J Grey Monroe\textsuperscript{1,2}, Brian Gill\textsuperscript{3}, \&
-Kathryn Turner\textsuperscript{4}}
+  \title{Satellite-detected drought frequencies predict plant life history
+strategies}
+    \author{J. Grey Monroe\textsuperscript{1,2}, Brian Gill\textsuperscript{3},
+Kathryn Turner\textsuperscript{4}, \& John K. McKay\textsuperscript{2}}
     \date{}
@@
 \vspace{0.5cm}
-\textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO\\\textsuperscript{3} Department of Ecology and Evolutionary Biology, Brown University, Provience, RI\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA}
-\keywords{drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records\newline\indent Word count: }
+\textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80523, USA\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO 80523, USA\\\textsuperscript{3} Institute for Environment and Society, Brown University, Providence, RI 02912, USA\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA 16802, USA}
+\keywords{drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records}
 \usepackage{csquotes}
@@
 \linenumbers
+\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
 
@@
 
-Correspondence concerning this article should be addressed to J Grey
+Correspondence concerning this article should be addressed to J. Grey
 Monroe, 307 University Ave, Fort Collins, CO 80523. E-mail:
@@
 \abstract{
-Identifying the environmental factors that predict the evolution of
-annual and perennial life history strategies in plants is important for
-understanding ecosystem functioning, perennial cropping systems, and is
-a long-standing goal of evolutionary ecology. A classic hypothesis is
-that annual and perennial strategies reflect adaptation to environments
-that differ in relative drought frequency. We test this hypothesis in
-Heliophila (Brassicaceae), a genus of flowering plants native to
-Southern Africa using herbarium occurrence records and
-satellite-detected drought histories. We find that perennial Heliophila
-species are observed in environments where droughts are significantly
-less frequent compared to annuals, especially with regards to summer
-drought. These correlations remain predictive while controlling for
-phylogeny, lending support to the hypothesis that drought associated
-natural selection has shaped differences in the distributions of these
-strategies. Additionally, the collection dates of herbarium records are
-consistent with a scenario in which annual species escape drought prone
-seasons. Together these results point to a model of evolution that
-supports classical hypotheses favoring perenniality in environments that
-experience less drought, compared to annuals which escape drought prone
-seasons as seed. Thus, changing drought regimens during the Anthropocene
-may threaten locally adapted species. 
+Explaining variation in life history strategies is a long-standing goal
+of evolutionary biology. For plants, annual and perennial life histories
+are thought to reflect adaptation to environments that differ in the
+frequency of environmental stress such as drought. Here we test this
+hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of
+flowering plants native to Southern Africa using herbarium occurrence
+records and satellite-detected drought histories. We find that perennial
+\textit{Heliophila} species are found in environments where droughts are
+significantly less frequent compared to annuals. These associations are
+predictive while controlling for phylogeny, lending support to the
+hypothesis that drought related natural selection has influenced the
+distributions of these strategies. Additionally, the difference in
+drought frequency between annual and perennial species distributions is
+greatest during the summer and fall, which also appears to be when
+annuals are in the seed phase of their life cycle based on collection
+dates of annual species. Together, these findings support traditional
+hypotheses about the drivers of life history strategy in plants - that
+perenniality is favored in environments with less frequent drought and
+annuality in enviroments with more frequent drought by allowing them to
+escape drought prone seasons as seeds. 
 }
@@
 
-\hypertarget{life-history}{%
-\subsection{Life history}\label{life-history}}
+\hypertarget{life-history-variation-in-plants}{%
+\subsection{Life history variation in
+plants}\label{life-history-variation-in-plants}}
 
-Plants exhibit extraordinary diversity in life history strategies, from
-herbaceous species that complete the entire seed to seed life cycle in a
-number of weeks \{Guo, 2012 \#73\} to trees that live for thousands of
-years \{Brown 1996 \url{http://www.rmtrr.org/oldlist.htm} \}. Along this
-continuum an important division exists, distinguishing annuals which
-complete their seed to seed life cycle within a single calendar year
-from perennials which can persist over multiple years and therefore
-experience the entire range of seasonal conditions. The ecological
-factors that explain the evolution of these alternative strategies
-remains poorly resolved \{Friedman, 2015 \#523\}. Understanding the
-drivers of selection for these alternative strategies is important
-because annuals and perennials have differing impacts on ecosystem
-functioning, such as higher nitrogen concentration in perennials and
-larger specific root length in annuals that affect nutrient cycling
-\{Garnier, 1997 \#280\}\{Roumet, 2006 \#738\}. Furthermore, predicting
-the conditions that favor these strategies in nature is useful for
-developing more sustainable perennial cropping systems \{Lelievre and
-Volaire, 2009\}.
+Understanding the causes and consequences of life history variation is a
+longstanding goal of ecology and evolutionary biology (Cole, 1954). In
+plants, life histories are especially diverse, with herbaceous species
+that complete their life cycle in a number of weeks to trees that live
+for thousands of years (Brown, 1996). Along this continuum an important
+division exists, distinguishing annuals (i.e.~monocarpic or semelparous)
+which complete their seed to seed life cycle within a single calendar
+year from perennials (i.e.~polycarpic or iteroparous{]} which can
+persist over multiple years. Annual plants flower once, set seed,
+senesce, and then die, spending at least some portion of the year as a
+seed. In contrast, perennial plants can continue vegetative growth after
+reproduction and experience all seasons. These represent fundamentally
+different life history strategies, but the ecological factors that
+explain their evolution remain uresolved (Friedman \& Rubin, 2015).
 
-\hypertarget{hypotheses}{%
-\subsection{Hypotheses}\label{hypotheses}}
+Identifying the drivers of selection for these alternative strategies is
+important because annuals and perennials have differing impacts on
+ecosystem functioning. For example, perennials have higher nitrogen
+concentration and larger specific root length in annuals, which both
+affect nutrient cycling (Garnier, Cordonnier, Guillerm, \& Sonié, 1997;
+Roumet, Urcelay, \& Dı'az, 2006). Furthermore, predicting the conditions
+that favor these strategies in nature is useful for developing more
+sustainable perennial cropping systems (Lelièvre \& Volaire, 2009).
 
-Seasonal drought conditions can represent a severe challenge to the
-persistence of individual organisms. Well-established hypotheses suggest
-that an annual life history is adaptive in environments where frequent
-droughts makes an escape from stressful drought conditions in the form
-of seed advantageous \{Stearns 1992\}\{Silvertown and Charlesworth,
-2001\}. These hypotheses are based on models of life history evolution
-in which short life cycles are favored in unpredictable environments or
-those with frequent stressful conditions. Plants are expected to be more
-likely to survive a drought that cooccurs with the seed phase of a
-plant's life cycle in which they are most resistant to desiccation.
-Previous efforts to address this hypothesis have yielded mixed or
-qualitative results. Transitions to annuality in Oenothera were
-associated with warmer summers and drier winters, but not with increased
-drought directly \{Evans, 2005 \#260\}. Annual and perennial species of
-Nemesia were qualitatively associated with winter rather and summer
-rainfall environments respectively \{Datson, 2008\}. Similarly, annual
+Classical theory predict shorter life spans and annuality in
+environments where adult mortality is high (Charnov \& Schaffer, 1973;
+Franco \& Silvertown, 1996; Stearns, 1992). In plants, this has been
+extended to the hypothessis that annuality is adaptive when it allows
+plants to escape drought (Schaffer \& M, 1975). Lack of water is perhaps
+the greatest threat to survival during vegetatitive or reproductive
+growth and annuals can remain dormant (and protected as a seed) during
+drought. Thus, environments with greater seasonal drought frequency may
+select for annual life histories that complete reproduction prior to
+drought prone seasons. Conversely, environments with less frequent
+drought may select for perennial species, which may benefit from
+multiple bouts of reproduction and competitive advantage (Corbin \&
+D'Antonio, 2004). These prediction has been supported by the association
+of annualilty with arid environments in wild rice (Morishima, Sano, \&
+Oka, 1984) and \emph{Oenothera} (Evans, Hearn, Hahn, Spangle, \&
+Venable, 2005). Additionally, annual and perennial species of Nemesia
+were qualitatively associated with winter rather and summer rainfall
+environments respectively (Datson, Murray, \& Steiner, 2008) and annual
 species of Scorzoneroides were associated with environments classified
-as \enquote{unpredictable} \{Cruz-Mazo, 2009 \#598\}. Annual and
-perennial varieties of the wild rice species Oryza rufipogon were
-observed more frequently in dry and humid climates, respectively
-\{Morishima et al.~1984\}. While these reports suggest the prominence of
-annuals in environments considered more arid, a quantitative assessment
-of how historic drought frequency may favor annual or perennial life
-history strategies remains lacking.
+as unpredictable (Cruz-Mazo, Buide, Samuel, \& Narbona, 2009). These
+findings provide qualitative evidence supporting the prediction that
+annual species are found in environments that experience more frequent
+drought, and vice versa, but whether the history frequency of drought
+events indeed predicts the distributions annual or perennial life
+history strategies has yet to be tested.
 
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
+Here we combine a long-term global dataset of satellite detected drought
+events with metadata from natural collections to test these classic
+hypotheses about the evolution of life history strategies within the
+African endemic mustard genus, \emph{Heliophila} L. (Brassicaceae). If
+annuallity is an adaptive strategy allowing plants to escape drought
+prone seasons, then drought frequency should predict the distribution of
+life history strategies across landscapes, and annual species should be
+more commonly associated with drought prone regions than perennial
+species. Furthermore, if annual species have adapted to escape drought
+prone seasons as seeds, observations of annual species should be rare
+during drought prone seasons. Phylogenetic relatedness can have
+significant non-random effects on species distributions and life history
+traits {[}{]}, and therefore we assessed the relationship between life
+history distribution and drought frequency in a phylogenetically
+controlled background.
 
-Here we combine a long-term dataset of global imaging with metadata from
-natural history collections to test classic hypotheses about the
-evolution of life history strategies within the African endemic mustard
-genus, Heliophila L. (Brassicaceae). If annual species temporally escape
-drought as seed, then drought frequency should be an important
-determinant of the distribution of life history strategies across the
-landscape, and annual species should be more commonly associated with
-drought prone regions than perennial species. Furthermore, if annual
-species have adapted to escape drought prone seasons as seeds,
-observations of annual species should be rare during drought prone
-seasons. Phylogenetic relatedness can have significant non-random
-effects on species distributions and life history traits (cite), and
-therefore we assessed the relationship between life history distribution
-and drought frequency in a phylogenetically controlled background.
+\hypertarget{materials-and-methods}{%
+\section{Materials and Methods}\label{materials-and-methods}}
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\hypertarget{availability}{%
+\subsection{Availability}\label{availability}}
+
+All data and the source code to produce this manuscript are available at
+\url{https://github.com/greymonroe/heliophila}.
 
@@
 
-\hypertarget{drought}{%
-\subsubsection{Drought}\label{drought}}
+\hypertarget{satellite-detected-drought-data}{%
+\subsubsection{Satellite-detected drought
+data}\label{satellite-detected-drought-data}}
 
-Remotely sensed data is a valuable tool for characterizing seasonal
-patterns in drought \{AghaKouchak 2015\}. With these data, we can
-exploit the fact that satellite imagery can detect reductions in plant
-cover and moisture at landscape scales that are characteristic of
-drought conditions. Unlike weather station data, which is limited in
-geographic coverage, and manual observation, which is limited in scope,
-satellite detected data provides a long term global historical record of
-drought events at fine temporal and spatial resolutions. One such
-measure, the Vegetative Health Index (VHI), has been collected since
-1981 by NOAA AVHRR satellites and is a combined measure of thermal and
-vegetative parameters. VHI quantifies drought induced vegetative stress
-weekly by detecting reduced plant cover based on the normalized
-difference vegetation index (Vegetative Condition Index) combined with
-reductions in moisture associated with anomalies in thermal spectra
-(Temperature Condition Index). Put most simply, the VHI detects drought
-by identifying environments that appear unusually brown and dry for that
-location and time of the year.
+Remotely sensed data is a powerful tool for characterizing seasonal
+patterns in drought because it is less limited in spatial and temporal
+scope and resolution than weather stations or field observations
+(AghaKouchak et al., 2015). To quantify the frequency of drought during
+different seasons across landscapes, we used the remotely sensed
+Vegetative Health Index (VHI), which measures landscape scale reductions
+in plant cover and temperature conditions characteristic of drought
+(Kogan, 2001). Generated from data collected by NOAA AVHRR satellites
+since 1981, the VHI combines normalized difference vegetation index
+(NDVI) derived measures of vegetative stress (Vegetative Condition Index
+- VCI) with temperature stress indicated by anomalies in thermal spectra
+(Temperature Condition Index - TCI).
 
-This data provides a global, long term, and quantitative perspective on
-drought variability. The VHI database (see methods) presents a valuable
-resource to study seasonal patterns in the frequency of drought across
-environments and to test hypotheses about the effect of drought on
-ecological and evolutionary processes. As such, it has been validated as
-a tool for detecting drought and predicting crop yields \{Kogan 1997\}.
-It has also proven useful for predicting infraspecific variation in
-drought tolerance traits and genes \{Mojica et al 2016\}\{Dittberner et
-al 2018\}\{Monroe et al 2018\}.
+The VHI of year \(y\) during week \(w\) of \([1,52]\) at pixel \(i\) is
+derived from the following equations, where \(n\) is the number of years
+observed.
 
-VHI quantifies drought induced vegetative stress weekly by detecting
-reduced plant cover based on the normalized difference vegetation index
-(Vegetative Condition Index) combined with reductions in moisture
-associated with anomalies in thermal spectra (Temperature Condition
-Index). Put most simply, the VHI detects drought by identifying
-environments that appear unusually brown and dry for that location and
-time of the year.
+\[VCI_{y,w,i} = 100\frac{NDVI_{y,w,i} - NDVI_{min}}{NDVI_{max} - NDVI_{min}}\]
 
-The Vegetative Health Index (VHI) was used to examine seasonal drought
-frequency at the collection locations of Heliophila herbarium specimens.
-The VHI is a satellite-detected drought measurement method based on
-observations of vegetative stress caused by drought \{Kogan, 1995
-\#534\}, combining deviations from historic climatic (Temperature
-Condition Index) and vegetative conditions (Vegetative Condition Index).
-The VHI has been measured weekly at 16km2 spatial resolution since XXXX.
-The frequencies of observing drought conditions (VHI\textless{}40,
-standard recommended by NOAA developers of VHI) during the southern
-hemisphere spring (quarter surrounding spring equinox), summer (quarter
-surrounding summer solstice), fall (quarter surrounding fall equinox),
-and winter (quarter surrounding winter solstice) were calculated
-globally from 1981 to 2015. These values were calculated for the
-collection location of each Heliophila herbarium specimen in our
-filtered dataset.
+where \(NDVI_{min} = min(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})\) and
+\(NDVI_{max} = max(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})\)
 
-NDVI = (Ch2 - Chl)/(Ch2 + Chl)
+\[TCI_{y,w,i} = 100\frac{T_{y,w,i} - T_{min}}{T_{max} - T_{min}}\]
 
-\[NDVI = \frac{NIR - Red}{NIR + Red}\]
+where \(T_{min} = min(T_{1981,w,i}...T_{1981+n,w,i})\) and
+\(T_{max} = max(T_{1981,w,i}...T_{1981+n,w,i})\)
 
-\[VCI = 100\frac{NDVI - NDVI_{min}}{NDVI_{max} - NDVI_{min}}\]
+\[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\]
 
-\[TCI = 100\frac{T - T{min}}{T{max} - T{min}}\]
-
-\[VHI = 0.5(VCI) + 0.5(TCI)\] This data provides a global, long term,
-and quantitative perspective on drought variability. The VHI database
-presents a valuable resource to study seasonal patterns in the frequency
-of drought across environments and to test hypotheses about the effect
-of drought on ecological and evolutionary processes. As such, it has
-been validated as a tool for detecting drought and predicting crop
-yields ({\textbf{???}}). It has also proven useful for predicting
+Thus, VHI measurements are standardized according to conditions
+historically observed at each locations. These measurements have been
+validated and generally used for evaluating drought risk and predicting
+crop yields in agriculture (e.g., Rojas, Vrieling, \& Rembold, 2011;
+Kogan et al., 2016). But they also present a tool to study seasonal
+patterns in the frequency of drought across environments and to test
+hypotheses about the effect of drought on ecological and evolutionary
+processes. Indeed, the VHI has been applied recently to study drought
+related ecology of natural species and proven useful for predicting
 infraspecific variation in drought tolerance traits and genes
 (Dittberner et al., 2018; Mojica et al., 2016; Monroe et al., 2018).
+Here, we accessed VHI data at \(16km^2\) resolution from 1981 to 2015
+(\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php}) to
+characterize the seasonal drought frequencies experienced by annual and
+perennial \emph{Heliophila} sepecies.
 
-\hypertarget{life-history-1}{%
-\subsubsection{Life history}\label{life-history-1}}
+\hypertarget{life-history-data-for-heliophila}{%
+\subsubsection{Life history data for
+Heliophila}\label{life-history-data-for-heliophila}}
 
-Heliophila is a genus of flowering plants endemic to the southern
+\emph{Heliophila} is a genus of flowering plants endemic to the southern
 portion of Africa including the Cape Floristic and Succulent Karoo
-Regions which are among the most botanically diverse environments on
-Earth. Indeed, the estimated \textasciitilde{}50 Heliophila species are
-considered to represent the most diverse genus of the family
-Brassicaceae \{Mummenhoff, 2005 \#494\}. This genus includes both
-perennial and annual species and this change in life history strategy
-has likely arisen multiple independent times \{Appel \& Al-Shehbaz,
-1997\}\{Mummenhoff, 2005 \#494\}. These multiple transitions between
-life history strategy and it's well documented record in global herbaria
-provided the ideal system to address classic hypotheses about the
-evolutionary drivers of annual and perennial life history strategies.
+Regions. These are among the most botanically diverse environments on
+Earth and the estimated \textasciitilde{}50 \emph{Heliophila} species
+are considered to be the most diverse genus of the family Brassicaceae
+(Mummenhoff, Al-Shehbaz, Bakker, Linder, \& Mühlhausen, 2005). This
+genus includes both perennial and annual species and this change in life
+history strategy has likely arisen multiple independent times (Appel \&
+Al-Shehbaz, 1997; Mummenhoff et al., 2005). We used life histories
+reported by Mummenhoff et al. (2005), grouping species with annual or
+perennial life histories. Perenniality was defined based on the
+traditional defition and included any form of perennial life history
+(e.g., herbs, shrubs, mixed, etc).
 
-\hypertarget{phyologeny}{%
-\subsubsection{Phyologeny}\label{phyologeny}}
+\hypertarget{heliophila-herbarium-specimens}{%
+\subsubsection{Heliophila herbarium
+specimens}\label{heliophila-herbarium-specimens}}
 
-Herbaria records and satellite detected drought provide data with which
-the distributions of annual and perennial species can be compared with
-respect to historic drought frequency. However, it is necessary to
-control for the demographic history caused by common ancestry of species
-if evolutionary processes such as natural selection are to be invoked as
-explanations for any differences observed. If, for example, annual and
-perennial species show significantly different ranges with respect to
-historic drought this could be confounded by common ancestry if annuals
-originated from a common ancestor and vice versa. In this case, it would
-be challenging to distinguish between natural selection and demographic
-history. On the other hand, if annual and perennial life history
-strategies arose independently in multiple species, controlling for
-phylogenetic relationships allows us to better account for demography
-and make stronger assertions about the importance of processes such as
-natural selection to explain patterns.
+Botanists have collected and maintained over 350 million botanical
+specimens worldwide over the past 300 years. Herbarium specimens and
+their associated metadata have been used since the 1960s to study
+species' geographical distributions (reviewed by Willis et al. (2017)
+and Lang, Willems, Scheepens, Burbano, and Bossdorf (2018)). And as they
+become digitized (Soltis, 2017), these collections have been used to
+study relationships between trait distributions, geography, and climate
+(Davis, Willis, Connolly, Kelly, \& Ellison, 2015; Stropp et al., 2016;
+Václavı'k, Beckmann, Cord, \& Bindewald, 2017; Wolf, Zimmerman,
+Anderegg, Busby, \& Christensen, 2016). To characterize the diributions
+of annual and perennial \emph{Heliophila} species, all records for the
+genus \emph{Heliophila} were downloaded from the Global Biodiversity
+Information Facility (gbif.org) on July 21, 2018.
 
-Heliophila is a charismatic genus of flowering plants from the
-Brassicaceae provides a valuable model to study the evolution of annual
-and perennial life history strategies because each has arisen
-independently multiple times within the genus. These independent origins
-allow for phylogenetically constrained analyses comparing the climate
-distributions of annual and perennial species, to gain greater insight
-into the role of selection.
+\hypertarget{sequence-data-for-phylogeny}{%
+\subsubsection{Sequence data for
+phylogeny}\label{sequence-data-for-phylogeny}}
 
 Aligned Heliophila ITS sequences were obtained from previous work by
-Mandáková et al.~\{Mandáková, 2012 \#339\}. Aethionema, Alliaria,
-Cardamine, Chamira, and Rorippa ITS records from were downloaded from
-Genbank (link) and aligned together with Heliophila ITS sequences using
-MAFFT as outgroups? (cite). Model selection for construction of
-phylogeny was performed in jModeltest2 with CIPRES (cite). Based on this
-analysis, \(GTR + L\) were selected. Ultrametric phylogeny was estimated
-with branch lengths as relative time (details). Life history character
-states were used based on habit reported by Mummemhoff et
-al.~\{Mummenhoff, 2005 \#494\}. Species reported to have any form of
-perennial life history (such as biennial\ldots{}) were classified for
-analyses here as perennial, where species reported to have strictly
-annual life histories were classified for analyses here as annual.
-
-\hypertarget{herbarium-specimens}{%
-\subsubsection{Herbarium Specimens}\label{herbarium-specimens}}
-
-Botanists have collected and maintained over 350 million botanical
-specimens worldwide over the past 300 years (Lang et al 2018). These
-collections, housed in herbaria, comprise an enormous, yet largely
-untapped, ecological dataset. These specimens are increasingly being
-recognized as an invaluable and underutilized source of data pertaining
-to biological responses to past environmental conditions (Pyke et
-al.~2010). Herbarium specimens and their associated metadata have been
-used since the 1960s to study species' geographical distributions
-(reviewed in Lang et al.~2018). Herbarium specimens have been used to
-track plant responses to climate change using herbarium time-series,
-including the relationships between traits, geography, and climate.
-\{Wolf et al.~2016, Davis et al.~2015, Stropp et al.~2017\}. A
-potentially fruitful approach combines herbarium data with remotely
-sensed data sets that characterize the environmental and climatic
-conditions across landscapes. For example, remotely sensed global
-ultraviolet-B radiation exposure may explain differences in trichome
-phenotypes, observed from herbarium specimens, between the native and
-invasive ranges of two herbaceous plants (Václavík et al.~2017).
-
-Herbarium specimen records for all Heliophila species were downloaded
-from the Global Biodiversity Information Facility (gbif.org) on July 21,
-2018. Several filtering steps were employed to clean this data (see Sup
-Mat. R code). For these analyses, records were restricted to those which
-are associated with a physical herbarium specimen, rather than other
-types of observations. Consequentially, this sample represents
-individuals in a narrow range of vegetative and/or flowering states --
-for example, herbarium specimens cannot be collected for individuals
-existing as seed. Analyses were restricted to Heliophila species with
-previously-reported life history information \{Mummenhoff, 2005 \#494\};
-42 of \textasciitilde{}50 named species in this genus were used. Only
-records including geo-referenced information (latitude and longitude of
-collection site) were used for downstream analyses of historic drought
-frequency. To discard records likely to be erroneous and to restrict
-analyses to plants in their native, non-cultivated, habitats, records
-occurring outside of the native range (roughly, the southern half of
-Africa) were removed. This excluded specimens such as those growing in
-likely cultivation in Australia or terrestrial plants recorded from
-non-terrestrial oceanic locations.. Finally, duplicate records, based on
-species, collection location coordinates, and date, were removed.
-
-We downloaded 8670 records from GBIF.
+Mandáková et al. (2012). Aethionema, Alliaria, Cardamine, Chamira, and
+Rorippa ITS records from were downloaded from Genbank {[}{]} and aligned
+together with Heliophila ITS sequences using MAFFT as outgroups {[}{]}.
 
@@
 
-\hypertarget{drought-frequency}{%
-\subsubsection{Drought frequency}\label{drought-frequency}}
+\hypertarget{drought-frequency-calculations}{%
+\subsubsection{Drought frequency
+calculations}\label{drought-frequency-calculations}}
 
-\hypertarget{herbarium-clean-up}{%
-\subsubsection{Herbarium clean up}\label{herbarium-clean-up}}
+To characterize drought regimens across the distrubtions of annual and
+perennial species of \emph{Heliohpila}, we calculated drought during
+different seasons at the location of observations for \emph{Heliophila}
+records using the VHI. Specifically, we created global maps of the
+frequencies of observing drought conditions (VHI\textless{}40, NOAA)
+during the winter (quarter surrounding winter solstice), spring (quarter
+surrounding spring equinox), summer (quarter surrounding summer
+solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015.
+From these maps, the drought frequency during the winter, spring,
+summer, and fall were extracted for the locations of all GBIF records.
 
-We filtered raw GBIF by restricting our analyses to
+\hypertarget{herbarium-record-quality-control}{%
+\subsubsection{Herbarium record quality
+control}\label{herbarium-record-quality-control}}
 
-\begin{itemize}
-\tightlist
-\item
-  Species with reported life history\\
-\item
-  Records with geospatial data\\
-\item
-  Records from collection sites classified as land pixels
-\item
-  Records for perserved specimens
-\item
-  Records from Africa
-\item
-  Non-duplicate records (ie. identical species, location, collection
+To remove instance with spurious location data, we filtered raw GBIF by
+restricting our analyses to include only:\\
+* Species with reported life history\\
+* Records with geospatial data\\
+* Records without known geospatial issues\\
+* Records from collection sites classified as land pixels\\
+* Records from Africa\\
+* Non-duplicate records (ie. identical species, location, collection
 date)
-\end{itemize}
 
-\hypertarget{phylogeny}{%
-\subsubsection{Phylogeny}\label{phylogeny}}
+\hypertarget{phylogeny-construction}{%
+\subsubsection{Phylogeny construction}\label{phylogeny-construction}}
 
-\hypertarget{contrast}{%
-\subsubsection{Contrast}\label{contrast}}
+Model selection for construction of phylogeny was performed in
+jModeltest2 with CIPRES (cite). Based on this analysis, \(GTR + L\) were
+selected. Ultrametric phylogeny was estimated with branch lengths as
+relative time (details).
+
+\hypertarget{comparison-of-drought-frequency-between-annual-and-perennial}{%
+\subsubsection{Comparison of drought frequency between annual and
+perennial}\label{comparison-of-drought-frequency-between-annual-and-perennial}}
 
@@
 the corresponding prediction that the observed distributions of annual
-and perennial Heliophila species would be significantly associated with
-historic drought frequency. To do so, the average drought frequency for
-each season was calculated across the herbarium specimen collection
-locations for each species. The relationship between drought frequencies
-across each taxon's range and life habitat (annual or perennial) was
-evaluated using Firth's penalized-likelihood logistic regression and
-phylogenetic logistic regression.
+and perennial \emph{Heliophila} species would be significantly
+associated with historic drought frequency. The relationship between
+drought frequencies across each taxon's range and life habitat (annual
+or perennial) was evaluated using Firth's penalized-likelihood logistic
+regression and phylogenetic logistic regression.
 
@@
 distributions were made by Two-sample Kolmogorov-Smirnov test, t-test,
-and Barlett variance test using R (supplemental script 2) \{cite R\}.
-\#\#\# Software We used R (Version 3.5.1; R Core Team, 2018) and the
-R-packages \emph{ape} (Version 5.2; Paradis \& Schliep, 2018; Orme et
-al., 2018; Soetaert, 2018), \emph{bindrcpp} (Version 0.2.2; Müller,
-2018), \emph{caper} (Version 1.0.1; Orme et al., 2018), \emph{coda}
-(Version 0.19.2; Plummer, Best, Cowles, \& Vines, 2006), \emph{diagram}
-(Version 1.6.4; Soetaert, 2017), \emph{dplyr} (Version 0.7.8; Wickham et
-al., 2018), \emph{forcats} (Version 0.3.0; Wickham, 2018a), \emph{gee}
-(Version 4.13.19; R by Thomas Lumley \& author., 2015), \emph{geiger}
-(Version 2.0.6; Alfaro et al., 2009; Eastman, Alfaro, Joyce, Hipp, \&
-Harmon, 2011; Harmon, Weir, Brock, Glor, \& Challenger, 2008; Slater et
-al., 2012), \emph{ggplot2} (Version 3.1.0; Wickham, 2016),
-\emph{logistf} (Version 1.23; Heinze \& Ploner, 2018), \emph{maps}
-(Version 3.3.0; Richard A. Becker, Ray Brownrigg. Enhancements by Thomas
-P Minka, \& Deckmyn., 2018), \emph{MASS} (Version 7.3.51.1; Venables \&
-Ripley, 2002), \emph{Matrix} (Version 1.2.15; Bates \& Maechler, 2018),
-\emph{MCMCglmm} (Version 2.26; Hadfield, 2010), \emph{mvtnorm} (Version
-1.0.8; Genz \& Bretz, 2009), \emph{papaja} (Version 0.1.0.9842; Aust \&
-Barth, 2018), \emph{phylolm} (Version 2.6; Ho \& Ane, 2014),
-\emph{phytools} (Version 0.6.60; Revell, 2012), \emph{purrr} (Version
-0.2.5; Henry \& Wickham, 2018), \emph{raster} (Version 2.8.4; Hijmans,
-2018), \emph{readr} (Version 1.2.1; Wickham et al., 2017), \emph{shape}
-(Version 1.4.4; Soetaert, 2018), \emph{sp} (Version 1.3.1; Pebesma \&
-Bivand, 2005), \emph{stringr} (Version 1.3.1; Wickham, 2018b),
-\emph{tibble} (Version 1.4.2; Müller \& Wickham, 2018), \emph{tidyr}
-(Version 0.8.2; Wickham \& Henry, 2018), and \emph{tidyverse} (Version
-1.2.1; Wickham, 2017) for all our analyses.
+and Barlett variance test.
 
@@
 Out of 8670 \emph{Heliophila} GBIF recrods, 6634 were for species with
-reported life history (Mummenhoff, Al-Shehbaz, Bakker, Linder, \&
-Mühlhausen, 2005), 3653 had geospatial data, 3460 were located on pixels
+reported life history (Mummenhoff et al., 2005), 3653 had geospatial
+data, 3627 did not have geospatial issues, 3460 were located on pixels
 classified as land having drought measurements, 3457 were located in
 Africa, 3162 were not duplicated. After all filtering steps, 2192
-records for 42 species (Figure \ref{tab:gbifrecordsmap}, Table
+records for 42 species (Figure \ref{fig:mapsboxplots}, Table
 \ref{tab:speciesmeanstable}) passed.
+
+\hypertarget{drought-frequency}{%
+\subsubsection{Drought frequency}\label{drought-frequency}}
+
+Drought frequency was signicantly different between annual and perennial
+species.
+
+\begin{figure}[!h]
+\includegraphics[width=\textwidth]{../figures/maps_boxplots} \caption{Locations of annual and perennial Heliophila. Drought
+frequency during the winter, spring, summer and fall. Drought
+frequencies during each season observed at the collection sites of
+Heliophila records.}\label{fig:mapsboxplots}
+\end{figure}
+
+\begin{table}[tbp]
+\begin{center}
+\begin{threeparttable}
+\caption{\label{tab:modelstable}Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species the winter, spring, summer, and fall.}
+\begin{tabular}{lllll}
+\toprule
+Predictor & \multicolumn{1}{c}{Estimate`} & \multicolumn{1}{c}{P`} & \multicolumn{1}{c}{Estimate*} & \multicolumn{1}{c}{P*}\\
+\midrule
+Intercept & 2.2575 & 0.1739 & 0.7231 & 0.6636\\
+Winter drought freq. & -6.7484 & 0.1661 & -1.5452 & 0.7274\\ \midrule
+Intercept & 4.5594 & 0.0443 & 5.0107 & 0.0534\\
+Spring drought freq. & -11.7895 & 0.0423 & -12.9014 & 0.0464\\ \midrule
+Intercept & 7.1742 & 0.0011 & 7.7093 & 0.0054\\
+Summer drought freq. & -18.2999 & 0.0010 & -19.9056 & 0.0042\\ \midrule
+Intercept & 6.4226 & 0.0029 & 7.0162 & 0.0082\\
+Fall drought freq. & -19.0512 & 0.0026 & -20.8174 & 0.0067\\ \midrule
+\bottomrule
+\addlinespace
+\end{tabular}
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} ` = Firth's penalized logistic regression. * = Phylogenetically constrained logistic regression. Annual species were scored as 0 and perennial species as 1.}
+\end{tablenotes}
+\end{threeparttable}
+\end{center}
+\end{table}
+
+\hypertarget{collection-dates-1}{%
+\subsubsection{Collection dates}\label{collection-dates-1}}
 
@@
 \centering
-\includegraphics{../figures/gbifrecords.pdf}
-\caption{\label{fig:gbifrecordsmap}Map of 2192 GBIF records that passed quality
-filtering.}
+\includegraphics{../figures/line_and_dates.pdf}
+\caption{\label{fig:lineplots}Comparison (mean +- SE) of drought frequency across
+seasons measured at the GBIF records of annual and perennial species of
+Heliophila. (B) Collection dates of GBIF records of annual and perennial
+species of Heliophila.}
 \end{figure}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\hypertarget{summary}{%
+\subsection{Summary}\label{summary}}
+
+We found that the distribution of annual and perennial species of
+Heliophila is significantly predicted by satellite detected historic
+drought frequencies. Annual species are found in environments that
+experience more frequent drought during the summer and fall quarters
+compared to perennials. This relationship was consistent while
+controlling for phylogenetic relatedness among the taxa studied,
+indicating that these distributions cannot be explained entirely by
+common ancestry. These results support the hypothesis that natural
+selection has played a role in shaping the contemporary distributions of
+these alternative life-history strategies.
+
+\hypertarget{relationship-to-previous-hypotheseswork}{%
+\subsection{Relationship to previous
+hypotheses/work}\label{relationship-to-previous-hypotheseswork}}
+
+Herbaria records and satellite detected drought provide data with which
+the distributions of annual and perennial species can be compared with
+respect to historic drought frequency. However, it is necessary to
+control for the demographic history caused by common ancestry of species
+if evolutionary processes such as natural selection are to be invoked as
+explanations for any differences observed. If, for example, annual and
+perennial species show significantly different ranges with respect to
+historic drought this could be confounded by common ancestry if annuals
+originated from a common ancestor and vice versa. In this case, it would
+be challenging to distinguish between natural selection and demographic
+history. On the other hand, if annual and perennial life history
+strategies arose independently in multiple species, controlling for
+phylogenetic relationships allows us to better account for demography
+and make stronger assertions about the importance of processes such as
+natural selection to explain patterns.
+
+These findings support classical theoretical predictions about the
+adaptive value of annual and perennial life history strategies. Taken
+together, they suggest that in Heliophila, annual species are adapted to
+environments with increased summer droughts by avoiding these seasons in
+a dormant seed phase of their life cycle. Indeed, we found that very few
+annuals are collected during this season, supporting the prediction that
+they are not in a vegetative and/or reproductive phase at this time.
+Traditionally, the focus has been on the evolutionary origins of annual
+life histories \{citations\}. However, we also find evidence that the
+transition to perenniality could be explained by historical drought
+regimes. The phylogeny reveals several transitions from annual to
+perennial life history. Perennials may be able to out complete annual
+relatives in environments where the infrequency of drought favors
+strategies that allow plants to benefit from growth over many seasons.
+
+\hypertarget{caveats}{%
+\subsection{Caveats}\label{caveats}}
+
+Sampling bias (Daru et al., 2018). Correlation does not prove causation.
+But it does indicate predictive power and is consistent with adaptive
+hypotheses. Herbarium collections and their associated data do not
+represent systematic or random sampling of a species distribution.
+Significant biases in collecting exist, which we have not necessarily
+controlled for here, and may have some effect on our findings, such as a
+bias toward collecting near roads or near the locations of natural
+history collections (Daru et al.~2018, Heberling, in press). Despite
+these biases, the Cape Floristic region is a biodiversity hotspot and
+one of the most botanically well sampled regions on Earth (Daru et
+al.~2018, Heberling, in press?), suggesting that this may currently be
+the optimal region for our analyses of life history distribution. Future
+research will benefit from systematic sampling efforts to avoid these
+noted biases. The climate data used here are assessed only from
+198X-XXXX and do not reflect conditions at the estimated divergence
+dates of these species (XXX million of years ago). Rather, the results
+suggest that the current distributions of annual and perennial species
+reflect a history of environmental filtration and ongoing natural
+selection. That is, their distributions are non-random with respect to
+historic drought and this is not explained by phylogeny.
+
+\hypertarget{broader-implications}{%
+\subsection{Broader implications}\label{broader-implications}}
+
+These findings suggest that rapidly changing drought regimes threaten
+species adapted to current environments. Studies predict changing
+drought regimes. This could have impacts on ecosystem functioning. This
+should also be considered when thinking about using perennial crops.
+Studies predict changing drought regimes.
+
+Do annuals require drought? We found independent origins of perenniality
+in environments where droughts are infrequent. Annuals may not be able
+to compete with perennials in the absence of droughts.
+
+\hypertarget{conclusions}{%
+\subsection{Conclusions}\label{conclusions}}
+
+Perenniality appears to be adaptive in environments with less frequent
+drought. This work demonstrates the power of emerging data to address
+outstanding classic hypotheses in ecology and evolution.
+
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
+
+This work was supported by NSF no. USDA grant no. XXX to J.G.M.
+
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\newpage
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-aghakouchak2015remote}{}%
+AghaKouchak, A., Farahmand, A., Melton, F., Teixeira, J., Anderson, M.,
+Wardlow, B. D., \& Hain, C. (2015). Remote sensing of drought: Progress,
+challenges and opportunities. \emph{Reviews of Geophysics},
+\emph{53}(2), 452--480.
+
+\leavevmode\hypertarget{ref-R-geiger_a}{}%
+Alfaro, M., Santini, F., Brock, C., Alamillo, H., Dornburg, A., Rabosky,
+D., \ldots{} Harmon, L. (2009). Nine exceptional radiations plus high
+turnover explain species diversity in jawed vertebrates.
+\emph{Proceedings of the National Academy of Sciences of the United
+States of America}, \emph{106}, 13410--13414.
+
+\leavevmode\hypertarget{ref-appel1997generic}{}%
+Appel, O., \& Al-Shehbaz, I. A. (1997). Generic limits and taxonomy of
+hornungia, pritzelago, and hymenolobus (brassicaceae). \emph{Novon},
+338--340.
+
+\leavevmode\hypertarget{ref-R-papaja}{}%
+Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with
+R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}
+
+\leavevmode\hypertarget{ref-R-Matrix}{}%
+Bates, D., \& Maechler, M. (2018). \emph{Matrix: Sparse and dense matrix
+classes and methods}. Retrieved from
+\url{https://CRAN.R-project.org/package=Matrix}
+
+\leavevmode\hypertarget{ref-brown1996oldlist}{}%
+Brown, P. M. (1996). OLDLIST: A database of maximum tree ages.
+\emph{Tree Rings, Environment, and Humanity. Radiocarbon}, \emph{1996},
+727--731.
+
+\leavevmode\hypertarget{ref-charnov1973life}{}%
+Charnov, E. L., \& Schaffer, W. M. (1973). Life-history consequences of
+natural selection: Cole's result revisited. \emph{The American
+Naturalist}, \emph{107}(958), 791--793.
+
+\leavevmode\hypertarget{ref-cole1954population}{}%
+Cole, L. C. (1954). The population consequences of life history
+phenomena. \emph{The Quarterly Review of Biology}, \emph{29}(2),
+103--137.
+
+\leavevmode\hypertarget{ref-corbin2004competition}{}%
+Corbin, J. D., \& D'Antonio, C. M. (2004). Competition between native
+perennial and exotic annual grasses: Implications for an historical
+invasion. \emph{Ecology}, \emph{85}(5), 1273--1283.
+
+\leavevmode\hypertarget{ref-cruz2009molecular}{}%
+Cruz-Mazo, G., Buide, M., Samuel, R., \& Narbona, E. (2009). Molecular
+phylogeny of scorzoneroides (asteraceae): Evolution of heterocarpy and
+annual habit in unpredictable environments. \emph{Molecular
+Phylogenetics and Evolution}, \emph{53}(3), 835--847.
+
+\leavevmode\hypertarget{ref-daru2018widespread}{}%
+Daru, B. H., Park, D. S., Primack, R. B., Willis, C. G., Barrington, D.
+S., Whitfeld, T. J., \ldots{} others. (2018). Widespread sampling biases
+in herbaria revealed from large-scale digitization. \emph{New
+Phytologist}, \emph{217}(2), 939--955.
+
+\leavevmode\hypertarget{ref-datson2008climate}{}%
+Datson, P., Murray, B., \& Steiner, K. (2008). Climate and the evolution
+of annual/perennial life-histories in nemesia (scrophulariaceae).
+\emph{Plant Systematics and Evolution}, \emph{270}(1-2), 39--57.
+
+\leavevmode\hypertarget{ref-davis2015herbarium}{}%
+Davis, C. C., Willis, C. G., Connolly, B., Kelly, C., \& Ellison, A. M.
+(2015). Herbarium records are reliable sources of phenological change
+driven by climate and provide novel insights into species' phenological
+cueing mechanisms. \emph{American Journal of Botany}, \emph{102}(10),
+1599--1609.
+
+\leavevmode\hypertarget{ref-dittberner2018natural}{}%
+Dittberner, H., Korte, A., Mettler-Altmann, T., Weber, A., Monroe, G.,
+\& Meaux, J. de. (2018). Natural variation in stomata size contributes
+to the local adaptation of water-use efficiency in arabidopsis thaliana.
+\emph{bioRxiv}, 253021.
+
+\leavevmode\hypertarget{ref-R-geiger_b}{}%
+Eastman, J., Alfaro, M., Joyce, P., Hipp, A., \& Harmon, L. (2011). A
+novel comparative method for identifying shifts in the rate of character
+evolution on trees. \emph{Evolution}, \emph{65}, 3578--3589.
+
+\leavevmode\hypertarget{ref-evans2005climate}{}%
+Evans, M. E., Hearn, D. J., Hahn, W. J., Spangle, J. M., \& Venable, D.
+L. (2005). Climate and life-history evolution in evening primroses
+(oenothera, onagraceae): A phylogenetic comparative analysis.
+\emph{Evolution}, \emph{59}(9), 1914--1927.
+
+\leavevmode\hypertarget{ref-franco1996life}{}%
+Franco, M., \& Silvertown, J. (1996). Life history variation in plants:
+An exploration of the fast-slow continuum hypothesis. \emph{Phil. Trans.
+R. Soc. Lond. B}, \emph{351}(1345), 1341--1348.
+
+\leavevmode\hypertarget{ref-friedman2015all}{}%
+Friedman, J., \& Rubin, M. J. (2015). All in good time: Understanding
+annual and perennial strategies in plants. \emph{American Journal of
+Botany}, \emph{102}(4), 497--499.
+
+\leavevmode\hypertarget{ref-garnier1997specific}{}%
+Garnier, E., Cordonnier, P., Guillerm, J.-L., \& Sonié, L. (1997).
+Specific leaf area and leaf nitrogen concentration in annual and
+perennial grass species growing in mediterranean old-fields.
+\emph{Oecologia}, \emph{111}(4), 490--498.
+
+\leavevmode\hypertarget{ref-R-mvtnorm}{}%
+Genz, A., \& Bretz, F. (2009). \emph{Computation of multivariate normal
+and t probabilities}. Heidelberg: Springer-Verlag.
+
+\leavevmode\hypertarget{ref-R-MCMCglmm}{}%
+Hadfield, J. D. (2010). MCMC methods for multi-response generalized
+linear mixed models: The MCMCglmm R package. \emph{Journal of
+Statistical Software}, \emph{33}(2), 1--22. Retrieved from
+\url{http://www.jstatsoft.org/v33/i02/}
+
+\leavevmode\hypertarget{ref-R-geiger_d}{}%
+Harmon, L., Weir, J., Brock, C., Glor, R., \& Challenger, W. (2008).
+GEIGER: Investigating evolutionary radiations. \emph{Bioinformatics},
+\emph{24}, 129--131.
+
+\leavevmode\hypertarget{ref-R-logistf}{}%
+Heinze, G., \& Ploner, M. (2018). \emph{Logistf: Firth's bias-reduced
+logistic regression}. Retrieved from
+\url{https://CRAN.R-project.org/package=logistf}
+
+\leavevmode\hypertarget{ref-R-purrr}{}%
+Henry, L., \& Wickham, H. (2018). \emph{Purrr: Functional programming
+tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}
+
+\leavevmode\hypertarget{ref-R-raster}{}%
+Hijmans, R. J. (2018). \emph{Raster: Geographic data analysis and
+modeling}. Retrieved from
+\url{https://CRAN.R-project.org/package=raster}
+
+\leavevmode\hypertarget{ref-R-phylolm}{}%
+Ho, L. S. T., \& Ane, C. (2014). A linear-time algorithm for gaussian
+and non-gaussian trait evolution models. \emph{Systematic Biology},
+\emph{63}, 397--408.
+
+\leavevmode\hypertarget{ref-kogan2016modelling}{}%
+Kogan, F., Guo, W., Strashnaia, A., Kleshenko, A., Chub, O., \&
+Virchenko, O. (2016). Modelling and prediction of crop losses from noaa
+polar-orbiting operational satellites. \emph{Geomatics, Natural Hazards
+and Risk}, \emph{7}(3), 886--900.
+
+\leavevmode\hypertarget{ref-kogan2001operational}{}%
+Kogan, F. N. (2001). Operational space technology for global vegetation
+assessment. \emph{Bulletin of the American Meteorological Society},
+\emph{82}(9), 1949--1964.
+
+\leavevmode\hypertarget{ref-lang2018using}{}%
+Lang, P. L., Willems, F. M., Scheepens, J., Burbano, H. A., \& Bossdorf,
+O. (2018). \emph{Using herbaria to study global environmental change}.
+PeerJ Preprints.
+
+\leavevmode\hypertarget{ref-lelievre2009current}{}%
+Lelièvre, F., \& Volaire, F. (2009). Current and potential development
+of perennial grasses in rainfed mediterranean farming systems.
+\emph{Crop Science}, \emph{49}(6), 2371--2378.
+
+\leavevmode\hypertarget{ref-mandakova2012whole}{}%
+Mandáková, T., Mummenhoff, K., Al-Shehbaz, I. A., Mucina, L.,
+Mühlhausen, A., \& Lysak, M. A. (2012). Whole-genome triplication and
+species radiation in the southern african tribe heliophileae
+(brassicaceae). \emph{Taxon}, \emph{61}(5), 989--1000.
+
+\leavevmode\hypertarget{ref-Mojica2016}{}%
+Mojica, J. P., Mullen, J., Lovell, J. T., Monroe, J. G., Paul, J. R.,
+Oakley, C. G., \& McKay, J. K. (2016). Genetics of water use physiology
+in locally adapted Arabidopsis thaliana. \emph{Plant Science}.
+doi:\href{https://doi.org/10.1016/j.plantsci.2016.03.015}{10.1016/j.plantsci.2016.03.015}
+
+\leavevmode\hypertarget{ref-monroe2018drought}{}%
+Monroe, J. G., Powell, T., Price, N., Mullen, J., Howard, A., Evans, K.,
+\ldots{} McKay, J. (2018). Drought adaptation in nature by extensive
+genetic loss-of-function. \emph{bioRxiv}, 372854.
+
+\leavevmode\hypertarget{ref-morishima1984differentiation}{}%
+Morishima, H., Sano, Y., \& Oka, H. (1984). Differentiation of perennial
+and annual types due to habitat conditions in the wild riceOryza
+perennis. \emph{Plant Systematics and Evolution}, \emph{144}(2),
+119--135.
+
+\leavevmode\hypertarget{ref-mummenhoff2005phylogeny}{}%
+Mummenhoff, K., Al-Shehbaz, I. A., Bakker, F. T., Linder, H. P., \&
+Mühlhausen, A. (2005). Phylogeny, morphological evolution, and
+speciation of endemic brassicaceae genera in the cape flora of southern
+africa. \emph{Annals of the Missouri Botanical Garden}, 400--424.
+
+\leavevmode\hypertarget{ref-R-bindrcpp}{}%
+Müller, K. (2018). \emph{Bindrcpp: An 'rcpp' interface to active
+bindings}. Retrieved from
+\url{https://CRAN.R-project.org/package=bindrcpp}
+
+\leavevmode\hypertarget{ref-R-tibble}{}%
+Müller, K., \& Wickham, H. (2018). \emph{Tibble: Simple data frames}.
+Retrieved from \url{https://CRAN.R-project.org/package=tibble}
+
+\leavevmode\hypertarget{ref-R-caper}{}%
+Orme, D., Freckleton, R., Thomas, G., Petzoldt, T., Fritz, S., Isaac,
+N., \& Pearse, W. (2018). \emph{Caper: Comparative analyses of
+phylogenetics and evolution in r}. Retrieved from
+\url{https://CRAN.R-project.org/package=caper}
+
+\leavevmode\hypertarget{ref-R-ape}{}%
+Paradis, E., \& Schliep, K. (2018). Ape 5.0: An environment for modern
+phylogenetics and evolutionary analyses in R. \emph{Bioinformatics},
+\emph{xx}, xxx--xxx.
+
+\leavevmode\hypertarget{ref-R-sp}{}%
+Pebesma, E. J., \& Bivand, R. S. (2005). Classes and methods for spatial
+data in R. \emph{R News}, \emph{5}(2), 9--13. Retrieved from
+\url{https://CRAN.R-project.org/doc/Rnews/}
+
+\leavevmode\hypertarget{ref-R-coda}{}%
+Plummer, M., Best, N., Cowles, K., \& Vines, K. (2006). CODA:
+Convergence diagnosis and output analysis for mcmc. \emph{R News},
+\emph{6}(1), 7--11. Retrieved from
+\url{https://journal.r-project.org/archive/}
+
+\leavevmode\hypertarget{ref-R-gee}{}%
+R by Thomas Lumley, V. J. C. P. to, \& author., B. R. N. that
+maintainers are not available to give advice on using a package they did
+not. (2015). \emph{Gee: Generalized estimation equation solver}.
+Retrieved from \url{https://CRAN.R-project.org/package=gee}
+
+\leavevmode\hypertarget{ref-R-base}{}%
+R Core Team. (2018). \emph{R: A language and environment for statistical
+computing}. Vienna, Austria: R Foundation for Statistical Computing.
+Retrieved from \url{https://www.R-project.org/}
+
+\leavevmode\hypertarget{ref-R-phytools}{}%
+Revell, L. J. (2012). Phytools: An r package for phylogenetic
+comparative biology (and other things). \emph{Methods in Ecology and
+Evolution}, \emph{3}, 217--223.
+
+\leavevmode\hypertarget{ref-R-maps}{}%
+Richard A. Becker, O. S. code by, Ray Brownrigg. Enhancements by Thomas
+P Minka, A. R. W. R. version by, \& Deckmyn., A. (2018). \emph{Maps:
+Draw geographical maps}. Retrieved from
+\url{https://CRAN.R-project.org/package=maps}
+
+\leavevmode\hypertarget{ref-rojas2011assessing}{}%
+Rojas, O., Vrieling, A., \& Rembold, F. (2011). Assessing drought
+probability for agricultural areas in africa with coarse resolution
+remote sensing imagery. \emph{Remote Sensing of Environment},
+\emph{115}(2), 343--352.
+
+\leavevmode\hypertarget{ref-roumet2006suites}{}%
+Roumet, C., Urcelay, C., \& Dı'az, S. (2006). Suites of root traits
+differ between annual and perennial species growing in the field.
+\emph{New Phytologist}, \emph{170}(2), 357--368.
+
+\leavevmode\hypertarget{ref-schaffer1975selection}{}%
+Schaffer, W., \& M, G. (1975). Selection for optimal life histories in
+plants. \emph{Ecology and Evolution of Communities.}, 142--157.
+
+\leavevmode\hypertarget{ref-R-geiger_c}{}%
+Slater, G., Harmon, L., Wegmann, D., Joyce, P., Revell, L., \& Alfaro,
+M. (2012). Fitting models of continuous trait evolution to incompletely
+sampled comparative data using approximate bayesian computation.
+\emph{Evolution}, \emph{66}, 752--762.
+
+\leavevmode\hypertarget{ref-R-diagram}{}%
+Soetaert, K. (2017). \emph{Diagram: Functions for visualising simple
+graphs (networks), plotting flow diagrams}. Retrieved from
+\url{https://CRAN.R-project.org/package=diagram}
+
+\leavevmode\hypertarget{ref-R-shape}{}%
+Soetaert, K. (2018). \emph{Shape: Functions for plotting graphical
+shapes, colors}. Retrieved from
+\url{https://CRAN.R-project.org/package=shape}
+
+\leavevmode\hypertarget{ref-soltis2017digitization}{}%
+Soltis, P. S. (2017). Digitization of herbaria enables novel research.
+\emph{American Journal of Botany}, \emph{104}(9), 1281--1284.
+
+\leavevmode\hypertarget{ref-stearns1992evolution}{}%
+Stearns, S. C. (1992). \emph{The evolution of life histories}.
+
+\leavevmode\hypertarget{ref-stropp2016mapping}{}%
+Stropp, J., Ladle, R. J., M. Malhado, A. C., Hortal, J., Gaffuri, J., H.
+Temperley, W., \ldots{} Mayaux, P. (2016). Mapping ignorance: 300 years
+of collecting flowering plants in africa. \emph{Global Ecology and
+Biogeography}, \emph{25}(9), 1085--1096.
+
+\leavevmode\hypertarget{ref-vaclavik2017effects}{}%
+Václavı'k, T., Beckmann, M., Cord, A. F., \& Bindewald, A. M. (2017).
+Effects of uv-b radiation on leaf hair traits of invasive
+plants---combining historical herbarium records with novel remote
+sensing data. \emph{PloS One}, \emph{12}(4), e0175671.
+
+\leavevmode\hypertarget{ref-R-MASS}{}%
+Venables, W. N., \& Ripley, B. D. (2002). \emph{Modern applied
+statistics with s} (Fourth.). New York: Springer. Retrieved from
+\url{http://www.stats.ox.ac.uk/pub/MASS4}
+
+\leavevmode\hypertarget{ref-R-ggplot2}{}%
+Wickham, H. (2016). \emph{Ggplot2: Elegant graphics for data analysis}.
+Springer-Verlag New York. Retrieved from \url{http://ggplot2.org}
+
+\leavevmode\hypertarget{ref-R-tidyverse}{}%
+Wickham, H. (2017). \emph{Tidyverse: Easily install and load the
+'tidyverse'}. Retrieved from
+\url{https://CRAN.R-project.org/package=tidyverse}
+
+\leavevmode\hypertarget{ref-R-forcats}{}%
+Wickham, H. (2018a). \emph{Forcats: Tools for working with categorical
+variables (factors)}. Retrieved from
+\url{https://CRAN.R-project.org/package=forcats}
+
+\leavevmode\hypertarget{ref-R-stringr}{}%
+Wickham, H. (2018b). \emph{Stringr: Simple, consistent wrappers for
+common string operations}. Retrieved from
+\url{https://CRAN.R-project.org/package=stringr}
+
+\leavevmode\hypertarget{ref-R-dplyr}{}%
+Wickham, H., François, R., Henry, L., \& Müller, K. (2018). \emph{Dplyr:
+A grammar of data manipulation}. Retrieved from
+\url{https://CRAN.R-project.org/package=dplyr}
+
+\leavevmode\hypertarget{ref-R-tidyr}{}%
+Wickham, H., \& Henry, L. (2018). \emph{Tidyr: Easily tidy data with
+'spread()' and 'gather()' functions}. Retrieved from
+\url{https://CRAN.R-project.org/package=tidyr}
+
+\leavevmode\hypertarget{ref-R-readr}{}%
+Wickham, H., Hester, J., \& Francois, R. (2017). \emph{Readr: Read
+rectangular text data}. Retrieved from
+\url{https://CRAN.R-project.org/package=readr}
+
+\leavevmode\hypertarget{ref-willis2017old}{}%
+Willis, C. G., Ellwood, E. R., Primack, R. B., Davis, C. C., Pearson, K.
+D., Gallinat, A. S., \ldots{} others. (2017). Old plants, new tricks:
+Phenological research using herbarium specimens. \emph{Trends in Ecology
+\& Evolution}, \emph{32}(7), 531--546.
+
+\leavevmode\hypertarget{ref-wolf2016altitudinal}{}%
+Wolf, A., Zimmerman, N. B., Anderegg, W. R., Busby, P. E., \&
+Christensen, J. (2016). Altitudinal shifts of the native and introduced
+flora of c alifornia in the context of 20th-century warming.
+\emph{Global Ecology and Biogeography}, \emph{25}(4), 418--429.
+
+\endgroup
+
+\newpage
+\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
+
+\hypertarget{supplement}{%
+\section{Supplement}\label{supplement}}
+
+\hypertarget{software-used}{%
+\subsubsection{Software used}\label{software-used}}
+
+We used R (Version 3.5.1; R Core Team, 2018) and the R-packages
+\emph{ape} (Version 5.2; Paradis \& Schliep, 2018; Orme et al., 2018;
+Soetaert, 2018), \emph{bindrcpp} (Version 0.2.2; Müller, 2018),
+\emph{caper} (Version 1.0.1; Orme et al., 2018), \emph{coda} (Version
+0.19.2; Plummer, Best, Cowles, \& Vines, 2006), \emph{diagram} (Version
+1.6.4; Soetaert, 2017), \emph{dplyr} (Version 0.7.8; Wickham et al.,
+2018), \emph{forcats} (Version 0.3.0; Wickham, 2018a), \emph{gee}
+(Version 4.13.19; R by Thomas Lumley \& author., 2015), \emph{geiger}
+(Version 2.0.6; Alfaro et al., 2009; Eastman, Alfaro, Joyce, Hipp, \&
+Harmon, 2011; Harmon, Weir, Brock, Glor, \& Challenger, 2008; Slater et
+al., 2012), \emph{ggplot2} (Version 3.1.0; Wickham, 2016),
+\emph{logistf} (Version 1.23; Heinze \& Ploner, 2018), \emph{maps}
+(Version 3.3.0; Richard A. Becker, Ray Brownrigg. Enhancements by Thomas
+P Minka, \& Deckmyn., 2018), \emph{MASS} (Version 7.3.51.1; Venables \&
+Ripley, 2002), \emph{Matrix} (Version 1.2.15; Bates \& Maechler, 2018),
+\emph{MCMCglmm} (Version 2.26; Hadfield, 2010), \emph{mvtnorm} (Version
+1.0.8; Genz \& Bretz, 2009), \emph{papaja} (Version 0.1.0.9842; Aust \&
+Barth, 2018), \emph{phylolm} (Version 2.6; Ho \& Ane, 2014),
+\emph{phytools} (Version 0.6.60; Revell, 2012), \emph{purrr} (Version
+0.2.5; Henry \& Wickham, 2018), \emph{raster} (Version 2.8.4; Hijmans,
+2018), \emph{readr} (Version 1.2.1; Wickham et al., 2017), \emph{shape}
+(Version 1.4.4; Soetaert, 2018), \emph{sp} (Version 1.3.1; Pebesma \&
+Bivand, 2005), \emph{stringr} (Version 1.3.1; Wickham, 2018b),
+\emph{tibble} (Version 1.4.2; Müller \& Wickham, 2018), \emph{tidyr}
+(Version 0.8.2; Wickham \& Henry, 2018), and \emph{tidyverse} (Version
+1.2.1; Wickham, 2017) for all our analyses.
+
+\hypertarget{supplementary-tables-and-figures}{%
+\subsubsection{Supplementary tables and
+figures}\label{supplementary-tables-and-figures}}
 
@@
 
-\hypertarget{drought-frequency.}{%
-\subsubsection{Drought frequency.}\label{drought-frequency.}}
-
-\begin{figure}
-\centering
-\includegraphics{../figures/life_history_drought_boxplots.pdf}
-\caption{\label{fig:boxplots}Comparison of drought frequency across seasons measured
-at the GBIF records of annual and perennial species of Heliophila.}
-\end{figure}
-
-\begin{table}[tbp]
-\begin{center}
-\begin{threeparttable}
-\caption{\label{tab:firthmodelstable}Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species. }
-\begin{tabular}{lll}
-\toprule
-Predictor & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{p.value}\\
-\midrule
-Intercept & 2.2575 & 0.1739\\
-Winter drought freq. & -6.7484 & 0.1661\\ \midrule
-Intercept & 4.5594 & 0.0443\\
-Spring drought freq. & -11.7895 & 0.0423\\ \midrule
-Intercept & 7.1742 & 0.0011\\
-Summer drought freq. & -18.2999 & 0.0010\\ \midrule
-Intercept & 6.4226 & 0.0029\\
-Fall drought freq. & -19.0512 & 0.0026\\ \midrule
-\bottomrule
-\addlinespace
-\end{tabular}
-\begin{tablenotes}[para]
-\normalsize{\textit{Note.} Firth's penalized logistic regression. Annual species were scored as 0 and perennial species as 1.}
-\end{tablenotes}
-\end{threeparttable}
-\end{center}
-\end{table}
-
-\begin{table}[tbp]
-\begin{center}
-\begin{threeparttable}
-\caption{\label{tab:phylomodelstable}Phlyogenetically constrained logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species.
-}
-\begin{tabular}{lll}
-\toprule
-Predictor & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{p.value}\\
-\midrule
-Intercept & 0.7231 & 0.6636\\
-Winter drought freq. & -1.5452 & 0.7274\\ \midrule
-Intercept & 5.0107 & 0.0534\\
-Spring drought freq. & -12.9014 & 0.0464\\ \midrule
-Intercept & 7.7093 & 0.0054\\
-Summer drought freq. & -19.9056 & 0.0042\\ \midrule
-Intercept & 7.0162 & 0.0082\\
-Fall drought freq. & -20.8174 & 0.0067\\ \midrule
-\bottomrule
-\addlinespace
-\end{tabular}
-\begin{tablenotes}[para]
-\normalsize{\textit{Note.} Annual species were scored as 0 and perennial species as 1.}
-\end{tablenotes}
-\end{threeparttable}
-\end{center}
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics{../figures/life_history_drought_means_line.pdf}
-\caption{\label{fig:lineplots}Comparison (mean +- SE) of drought frequency across
-seasons measured at the GBIF records of annual and perennial species of
-Heliophila.}
-\end{figure}
-
-\hypertarget{collection-dates-1}{%
-\subsubsection{Collection dates}\label{collection-dates-1}}
-
-\begin{figure}
-\centering
-\includegraphics{../figures/collection_dates.pdf}
-\caption{\label{fig:collectiondates}Collection dates of GBIF records of annual and
-perennial species of Heliophila.}
-\end{figure}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-\hypertarget{summary-1}{%
-\subsection{Summary}\label{summary-1}}
-
-We found that the distribution of annual and perennial species of
-Heliophila is significantly predicted by satellite detected historic
-drought frequencies. Annual species are found in environments that
-experience more frequent drought during the summer and fall quarters
-compared to perennials. This relationship was consistent while
-controlling for phylogenetic relatedness among the taxa studied,
-indicating that these distributions cannot be explained entirely by
-common ancestry., These results support the hypothesis that natural
-selection has played a role in shaping the contemporary distributions of
-these alternative life-history strategies.
-
-\hypertarget{relationship-to-previous-hypotheseswork}{%
-\subsection{Relationship to previous
-hypotheses/work}\label{relationship-to-previous-hypotheseswork}}
-
-These findings support classical theoretical predictions about the
-adaptive value of annual and perennial life history strategies. Taken
-together, they suggest that in Heliophila, annual species are adapted to
-environments with increased summer droughts by avoiding these seasons in
-a dormant seed phase of their life cycle. Indeed, we found that very few
-annuals are collected during this season, supporting the prediction that
-they are not in a vegetative and/or reproductive phase at this time.
-Traditionally, the focus has been on the evolutionary origins of annual
-life histories \{citations\}. However, we also find evidence that the
-transition to perenniality could be explained by historical drought
-regimes. The phylogeny reveals several transitions from annual to
-perennial life history. Perennials may be able to out complete annual
-relatives in environments where the infrequency of drought favors
-strategies that allow plants to benefit from growth over many seasons.
-
-\hypertarget{caveats}{%
-\subsection{Caveats}\label{caveats}}
-
-Correlation does not prove causation. But it does indicate predictive
-power and is consistent with adaptive hypotheses. Herbarium collections
-and their associated data do not represent systematic or random sampling
-of a species distribution. Significant biases in collecting exist, which
-we have not necessarily controlled for here, and may have some effect on
-our findings, such as a bias toward collecting near roads or near the
-locations of natural history collections (Daru et al.~2018, Heberling,
-in press). Despite these biases, the Cape Floristic region is a
-biodiversity hotspot and one of the most botanically well sampled
-regions on Earth (Daru et al.~2018, Heberling, in press?), suggesting
-that this may currently be the optimal region for our analyses of life
-history distribution. Future research will benefit from systematic
-sampling efforts to avoid these noted biases. The climate data used here
-are assessed only from 198X-XXXX and do not reflect conditions at the
-estimated divergence dates of these species (XXX million of years ago).
-Rather, the results suggest that the current distributions of annual and
-perennial species reflect a history of environmental filtration and
-ongoing natural selection. That is, their distributions are non-random
-with respect to historic drought and this is not explained by phylogeny.
-
-\hypertarget{broader-implications}{%
-\subsection{Broader implications}\label{broader-implications}}
-
-These findings suggest that rapidly changing drought regimes threaten
-species adapted to current environments. Studies predict changing
-drought regimes. This could have impacts on ecosystem functioning. This
-should also be considered when thinking about using perennial crops.
-Studies predict changing drought regimes.
-
-\hypertarget{conclusions}{%
-\subsection{Conclusions}\label{conclusions}}
-
-Perenniality appears to be adaptive in environments with less frequent
-drought. This work demonstrates the power of emerging data to address
-outstanding classic hypotheses in ecology and evolution.
-
-\hypertarget{acknowledgments}{%
-\section{Acknowledgments}\label{acknowledgments}}
-
-This work was supported by NSF no. USDA grant no. XXX to J.G.M.
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\newpage
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-R-geiger_a}{}%
-Alfaro, M., Santini, F., Brock, C., Alamillo, H., Dornburg, A., Rabosky,
-D., \ldots{} Harmon, L. (2009). Nine exceptional radiations plus high
-turnover explain species diversity in jawed vertebrates.
-\emph{Proceedings of the National Academy of Sciences of the United
-States of America}, \emph{106}, 13410--13414.
-
-\leavevmode\hypertarget{ref-R-papaja}{}%
-Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with
-R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}
-
-\leavevmode\hypertarget{ref-R-Matrix}{}%
-Bates, D., \& Maechler, M. (2018). \emph{Matrix: Sparse and dense matrix
-classes and methods}. Retrieved from
-\url{https://CRAN.R-project.org/package=Matrix}
-
-\leavevmode\hypertarget{ref-dittberner2018natural}{}%
-Dittberner, H., Korte, A., Mettler-Altmann, T., Weber, A., Monroe, G.,
-\& Meaux, J. de. (2018). Natural variation in stomata size contributes
-to the local adaptation of water-use efficiency in arabidopsis thaliana.
-\emph{bioRxiv}, 253021.
-
-\leavevmode\hypertarget{ref-R-geiger_b}{}%
-Eastman, J., Alfaro, M., Joyce, P., Hipp, A., \& Harmon, L. (2011). A
-novel comparative method for identifying shifts in the rate of character
-evolution on trees. \emph{Evolution}, \emph{65}, 3578--3589.
-
-\leavevmode\hypertarget{ref-R-mvtnorm}{}%
-Genz, A., \& Bretz, F. (2009). \emph{Computation of multivariate normal
-and t probabilities}. Heidelberg: Springer-Verlag.
-
-\leavevmode\hypertarget{ref-R-MCMCglmm}{}%
-Hadfield, J. D. (2010). MCMC methods for multi-response generalized
-linear mixed models: The MCMCglmm R package. \emph{Journal of
-Statistical Software}, \emph{33}(2), 1--22. Retrieved from
-\url{http://www.jstatsoft.org/v33/i02/}
-
-\leavevmode\hypertarget{ref-R-geiger_d}{}%
-Harmon, L., Weir, J., Brock, C., Glor, R., \& Challenger, W. (2008).
-GEIGER: Investigating evolutionary radiations. \emph{Bioinformatics},
-\emph{24}, 129--131.
-
-\leavevmode\hypertarget{ref-R-logistf}{}%
-Heinze, G., \& Ploner, M. (2018). \emph{Logistf: Firth's bias-reduced
-logistic regression}. Retrieved from
-\url{https://CRAN.R-project.org/package=logistf}
-
-\leavevmode\hypertarget{ref-R-purrr}{}%
-Henry, L., \& Wickham, H. (2018). \emph{Purrr: Functional programming
-tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}
-
-\leavevmode\hypertarget{ref-R-raster}{}%
-Hijmans, R. J. (2018). \emph{Raster: Geographic data analysis and
-modeling}. Retrieved from
-\url{https://CRAN.R-project.org/package=raster}
-
-\leavevmode\hypertarget{ref-R-phylolm}{}%
-Ho, L. S. T., \& Ane, C. (2014). A linear-time algorithm for gaussian
-and non-gaussian trait evolution models. \emph{Systematic Biology},
-\emph{63}, 397--408.
-
-\leavevmode\hypertarget{ref-Mojica2016}{}%
-Mojica, J. P., Mullen, J., Lovell, J. T., Monroe, J. G., Paul, J. R.,
-Oakley, C. G., \& McKay, J. K. (2016). Genetics of water use physiology
-in locally adapted Arabidopsis thaliana. \emph{Plant Science}.
-doi:\href{https://doi.org/10.1016/j.plantsci.2016.03.015}{10.1016/j.plantsci.2016.03.015}
-
-\leavevmode\hypertarget{ref-monroe2018drought}{}%
-Monroe, J. G., Powell, T., Price, N., Mullen, J., Howard, A., Evans, K.,
-\ldots{} McKay, J. (2018). Drought adaptation in nature by extensive
-genetic loss-of-function. \emph{bioRxiv}, 372854.
-
-\leavevmode\hypertarget{ref-mummenhoff2005phylogeny}{}%
-Mummenhoff, K., Al-Shehbaz, I. A., Bakker, F. T., Linder, H. P., \&
-Mühlhausen, A. (2005). Phylogeny, morphological evolution, and
-speciation of endemic brassicaceae genera in the cape flora of southern
-africa. \emph{Annals of the Missouri Botanical Garden}, 400--424.
-
-\leavevmode\hypertarget{ref-R-bindrcpp}{}%
-Müller, K. (2018). \emph{Bindrcpp: An 'rcpp' interface to active
-bindings}. Retrieved from
-\url{https://CRAN.R-project.org/package=bindrcpp}
-
-\leavevmode\hypertarget{ref-R-tibble}{}%
-Müller, K., \& Wickham, H. (2018). \emph{Tibble: Simple data frames}.
-Retrieved from \url{https://CRAN.R-project.org/package=tibble}
-
-\leavevmode\hypertarget{ref-R-caper}{}%
-Orme, D., Freckleton, R., Thomas, G., Petzoldt, T., Fritz, S., Isaac,
-N., \& Pearse, W. (2018). \emph{Caper: Comparative analyses of
-phylogenetics and evolution in r}. Retrieved from
-\url{https://CRAN.R-project.org/package=caper}
-
-\leavevmode\hypertarget{ref-R-ape}{}%
-Paradis, E., \& Schliep, K. (2018). Ape 5.0: An environment for modern
-phylogenetics and evolutionary analyses in R. \emph{Bioinformatics},
-\emph{xx}, xxx--xxx.
-
-\leavevmode\hypertarget{ref-R-sp}{}%
-Pebesma, E. J., \& Bivand, R. S. (2005). Classes and methods for spatial
-data in R. \emph{R News}, \emph{5}(2), 9--13. Retrieved from
-\url{https://CRAN.R-project.org/doc/Rnews/}
-
-\leavevmode\hypertarget{ref-R-coda}{}%
-Plummer, M., Best, N., Cowles, K., \& Vines, K. (2006). CODA:
-Convergence diagnosis and output analysis for mcmc. \emph{R News},
-\emph{6}(1), 7--11. Retrieved from
-\url{https://journal.r-project.org/archive/}
-
-\leavevmode\hypertarget{ref-R-gee}{}%
-R by Thomas Lumley, V. J. C. P. to, \& author., B. R. N. that
-maintainers are not available to give advice on using a package they did
-not. (2015). \emph{Gee: Generalized estimation equation solver}.
-Retrieved from \url{https://CRAN.R-project.org/package=gee}
-
-\leavevmode\hypertarget{ref-R-base}{}%
-R Core Team. (2018). \emph{R: A language and environment for statistical
-computing}. Vienna, Austria: R Foundation for Statistical Computing.
-Retrieved from \url{https://www.R-project.org/}
-
-\leavevmode\hypertarget{ref-R-phytools}{}%
-Revell, L. J. (2012). Phytools: An r package for phylogenetic
-comparative biology (and other things). \emph{Methods in Ecology and
-Evolution}, \emph{3}, 217--223.
-
-\leavevmode\hypertarget{ref-R-maps}{}%
-Richard A. Becker, O. S. code by, Ray Brownrigg. Enhancements by Thomas
-P Minka, A. R. W. R. version by, \& Deckmyn., A. (2018). \emph{Maps:
-Draw geographical maps}. Retrieved from
-\url{https://CRAN.R-project.org/package=maps}
-
-\leavevmode\hypertarget{ref-R-geiger_c}{}%
-Slater, G., Harmon, L., Wegmann, D., Joyce, P., Revell, L., \& Alfaro,
-M. (2012). Fitting models of continuous trait evolution to incompletely
-sampled comparative data using approximate bayesian computation.
-\emph{Evolution}, \emph{66}, 752--762.
-
-\leavevmode\hypertarget{ref-R-diagram}{}%
-Soetaert, K. (2017). \emph{Diagram: Functions for visualising simple
-graphs (networks), plotting flow diagrams}. Retrieved from
-\url{https://CRAN.R-project.org/package=diagram}
-
-\leavevmode\hypertarget{ref-R-shape}{}%
-Soetaert, K. (2018). \emph{Shape: Functions for plotting graphical
-shapes, colors}. Retrieved from
-\url{https://CRAN.R-project.org/package=shape}
-
-\leavevmode\hypertarget{ref-R-MASS}{}%
-Venables, W. N., \& Ripley, B. D. (2002). \emph{Modern applied
-statistics with s} (Fourth.). New York: Springer. Retrieved from
-\url{http://www.stats.ox.ac.uk/pub/MASS4}
-
-\leavevmode\hypertarget{ref-R-ggplot2}{}%
-Wickham, H. (2016). \emph{Ggplot2: Elegant graphics for data analysis}.
-Springer-Verlag New York. Retrieved from \url{http://ggplot2.org}
-
-\leavevmode\hypertarget{ref-R-tidyverse}{}%
-Wickham, H. (2017). \emph{Tidyverse: Easily install and load the
-'tidyverse'}. Retrieved from
-\url{https://CRAN.R-project.org/package=tidyverse}
-
-\leavevmode\hypertarget{ref-R-forcats}{}%
-Wickham, H. (2018a). \emph{Forcats: Tools for working with categorical
-variables (factors)}. Retrieved from
-\url{https://CRAN.R-project.org/package=forcats}
-
-\leavevmode\hypertarget{ref-R-stringr}{}%
-Wickham, H. (2018b). \emph{Stringr: Simple, consistent wrappers for
-common string operations}. Retrieved from
-\url{https://CRAN.R-project.org/package=stringr}
-
-\leavevmode\hypertarget{ref-R-dplyr}{}%
-Wickham, H., François, R., Henry, L., \& Müller, K. (2018). \emph{Dplyr:
-A grammar of data manipulation}. Retrieved from
-\url{https://CRAN.R-project.org/package=dplyr}
-
-\leavevmode\hypertarget{ref-R-tidyr}{}%
-Wickham, H., \& Henry, L. (2018). \emph{Tidyr: Easily tidy data with
-'spread()' and 'gather()' functions}. Retrieved from
-\url{https://CRAN.R-project.org/package=tidyr}
-
-\leavevmode\hypertarget{ref-R-readr}{}%
-Wickham, H., Hester, J., \& Francois, R. (2017). \emph{Readr: Read
-rectangular text data}. Retrieved from
-\url{https://CRAN.R-project.org/package=readr}
-
-\endgroup
-
 \end{document}

### 2018-11-26 src/GBIF_records_drought_frequency.R

@@ -382,7 +382,6 @@ p<-ggplot(GBIF_cleaned, aes(x=days_adjusted, fill=life_history)) +
 p
 dev.off()
 
-
 # TEST ANNUAL AND PERENNIAL DATES -----------------------------------------
 annuals<-subset(GBIF_cleaned, life_history=="a")
 perennials<-subset(GBIF_cleaned, life_history=="p")

### 2018-11-26 src/sandbox.R

@@ -1,0 +1,32 @@
+
+pdf("../figures/phylogeny.pdf", height=6, width=5)
+
+treeplot<-ggtree(tree, layout = "rectangular")+
+  xlim(c(0,2))+
+  geom_tippoint(size=3,col=ifelse(species_life_history$life_history[match(tree$tip.label,species_life_history$species)]=="a", "orange3", "dodgerblue3"))+
+  geom_tiplab(offset = .01 )
+
+minima <- readPNG("../pictures/minima.png")
+minima <- rasterGrob(minima, interpolate=TRUE)
+#https://www.gbif.org/occurrence/1099023487
+
+deserticola <- readPNG("../pictures/deserticola.png")
+deserticola <- rasterGrob(deserticola, interpolate=TRUE)
+#https://www.gbif.org/occurrence/1057389408
+
+coronopifolia<-readPNG("../pictures/coronopifolia.png")
+coronopifolia <- rasterGrob(coronopifolia, interpolate=TRUE)
+#https://www.gbif.org/occurrence/1099023562
+
+ephemera<-readPNG("../pictures/ephemera.png")
+ephemera <- rasterGrob(ephemera, interpolate=TRUE)
+
+#https://www.gbif.org/occurrence/1099023490
+
+pics<-plot_grid(minima, deserticola, coronopifolia, ephemera,
+                labels = c("b","c","d","e"), ncol = 1)
+
+plot_grid(treeplot, pics, ncol=2,labels=c("a",""), rel_widths = c(1,0.35))
+
+
+dev.off()

### 2018-11-27 manuscript/manuscript.Rmd
version 3

@@
 ---
-title             : "Satellite-detected drought frequencies predict plant life history strategies"
+title             : "Satellite-detected droughts predict plant life history"
 shorttitle        : "Drought and life history"
@@
 For plants, annual and perennial life histories are thought to reflect adaptation to environments that differ in the frequency of environmental stress such as drought.
-Here we test this hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Southern Africa using herbarium occurrence records and satellite-detected drought histories.
-We find that perennial \textit{Heliophila} species are found in environments where droughts are significantly less frequent compared to annuals.
+Here we test this hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa by integrating 2192 herbaria occurrence records with 34 years of satellite-based drought detection.
+Consistent with predictions from classic life history theory, we find that perennial \textit{Heliophila} species occur in environments where droughts are significantly less frequent compared to annuals.
 These associations are predictive while controlling for phylogeny, lending support to the hypothesis that drought related natural selection has influenced the distributions of these strategies.
 Additionally, the difference in drought frequency between annual and perennial species distributions is greatest during the summer and fall, which also appears to be when annuals are in the seed phase of their life cycle based on collection dates of annual species.
-Together, these findings support traditional hypotheses about the drivers of life history strategy in plants - that perenniality is favored in environments with less frequent drought and annuality in enviroments with more frequent drought by allowing them to escape drought prone seasons as seeds.
+Together, these finding provide empirical support for classic hypotheses about the drivers of life history strategy in plants - that perennials out compete annuals in environments with less frequent drought and that annuals are adpated to enviroments with more frequent drought by escaping drought prone seasons as seeds.
 
@@
 library(logistf)
-#library(gee) 
-#library(phytools) 
 library(geiger) 
 library(ape) 
-#library(caper) 
-#library(MCMCglmm) 
 library(phylolm)
@@
 
-## Life history variation in plants
-
 Understanding the causes and consequences of life history variation is a longstanding goal of ecology and evolutionary biology  [@cole1954population].
@@
 monocarpic or semelparous) which complete their seed to seed life cycle within a single calendar year from perennials (i.e.
-polycarpic or iteroparous] which can persist over multiple years.
-Annual plants flower once, set seed, senesce, and then die, spending at least some portion of the year as a seed.
-In contrast, perennial plants can continue vegetative growth after reproduction and experience all seasons.
-These represent fundamentally different life history strategies, but the ecological factors that explain their evolution remain uresolved [@friedman2015all].
-
-Identifying the drivers of selection for these alternative strategies is important because annuals and perennials have differing impacts on ecosystem functioning.
-For example, perennials have higher nitrogen concentration and larger specific root length in annuals, which both affect nutrient cycling [@garnier1997specific; @roumet2006suites].
-Furthermore, predicting the conditions that favor these strategies in nature is useful for developing more sustainable perennial cropping systems [@lelievre2009current].
-
-Classical theory predict shorter life spans and annuality in environments where adult mortality is high [@charnov1973life; @stearns1992evolution;@franco1996life].
+polycarpic or iteroparous) which can persist over multiple years.
+Annual plants flower once, set seed, senesce, and then die, spending at least some portion of the year as a seed, where they are relatively protected from environmental stress.
+In contrast, perennial plants can continue vegetative growth after reproduction and must survive conditions experienced during all seasons.
+These represent fundamentally different life history strategies, but the ecological factors that explain their evolution and distributions remain empirically uresolved [@friedman2015all].
+Classical theory predict shorter life spans in environments where adult mortality is high [@charnov1973life; @stearns1992evolution;@franco1996life].
 In plants, this has been extended to the hypothessis that annuality is adaptive when it allows plants to escape drought [@schaffer1975selection].
@@
 Thus, environments with greater seasonal drought frequency may select for annual life histories that complete reproduction prior to drought prone seasons.
-Conversely, environments with less frequent drought may select for perennial species, which may benefit from multiple bouts of reproduction and competitive advantage [@corbin2004competition].
-These prediction has been supported by the association of annualilty with arid environments in wild rice [@morishima1984differentiation] and _Oenothera_ [@evans2005climate].
-Additionally, annual and perennial species of Nemesia were qualitatively associated with winter rather and summer rainfall environments respectively [@datson2008climate] and annual species of Scorzoneroides were associated with environments classified as unpredictable [@cruz2009molecular].
-These findings provide qualitative evidence supporting the prediction that annual species are found in environments that experience more frequent drought, and vice versa, but whether the history frequency of drought events indeed predicts the distributions annual or perennial life history strategies has yet to be tested.
-
+Conversely, environments with less frequent drought may select for perennial species, which may benefit from multiple bouts of reproduction and competitive advantage by preventing recruitment of annual species [@corbin2004competition].
+These predictions have been supported by the association of annualilty with arid environments in wild rice [@morishima1984differentiation] and _Oenothera_ [@evans2005climate].
+Additionally, annual and perennial species of _Nemesia_ were qualitatively associated with winter rather and summer rainfall environments respectively [@datson2008climate] and annual species of _Scorzoneroides_ were associated with environments classified as unpredictable [@cruz2009molecular].
+However, whether the history frequency of drought events indeed predicts the distributions annual or perennial life history strategies has yet to be tested in a phylogenetic context with suitable sample sizes of distrubution and climate data.
 Here we combine a long-term global dataset of satellite detected drought events with metadata from natural collections to test these classic hypotheses about the evolution of life history strategies within the African endemic mustard genus, _Heliophila_ L.
@@
 Furthermore, if annual species have adapted to escape drought prone seasons as seeds, observations of annual species should be rare during drought prone seasons.
-Phylogenetic relatedness can have significant non-random effects on species distributions and life history traits [], and therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetically controlled background.
+Phylogenetic relatedness can have significant non-random effects on species distributions and life history traits [@barrett1996comparative], and therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetically controlled background.
 
 # Materials and Methods
+## Data  
+### Availability
+All analyses were peformed using R. All data and the source code to produce this manuscript are available at https://github.com/greymonroe/heliophila.
 
-## Availability
-All data and the source code to produce this manuscript are available at https://github.com/greymonroe/heliophila.
-
-## Data
 ### Satellite-detected drought data
@@
 ```
-
 Remotely sensed data is a powerful tool for characterizing seasonal patterns in drought because it is less limited in spatial and temporal scope and resolution than weather stations or field observations [@aghakouchak2015remote].
@@
 Generated from data collected by NOAA AVHRR satellites since 1981, the VHI combines normalized difference vegetation index (NDVI) derived measures of vegetative stress (Vegetative Condition Index - VCI) with temperature stress indicated by anomalies in thermal spectra (Temperature Condition Index - TCI).
-
 The VHI of year $y$ during week $w$ of $[1,52]$ at pixel $i$ is derived from the following equations, where $n$ is the number of years observed.
@@
 $$VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})$$
-
 Thus, VHI measurements are standardized according to conditions historically observed at each locations.
 These measurements have been validated and generally used for evaluating drought risk and predicting crop yields in agriculture [e.g., @rojas2011assessing; @kogan2016modelling].
-But they also present a tool to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes.
-Indeed, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting infraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought].
-Here, we accessed VHI data at $16km^2$ resolution from 1981 to 2015 (https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php) to characterize the seasonal drought frequencies experienced by annual and perennial _Heliophila_ sepecies.
+But they also present a new tool to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes [@kerr2003space].
+As such, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting infraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought].
+Here, we accessed VHI data at $16km^2$ resolution from 1981 to 2015 (https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php) to characterize the seasonal drought frequencies experienced by annual and perennial _Heliophila_ species.
 
@@
 This genus includes both perennial and annual species and this change in life history strategy has likely arisen multiple independent times [@appel1997generic;@mummenhoff2005phylogeny].
+Furthermore, the fine scale climatic heterogeneity of Southern Africa is ideal for studying the distribution of traits in relation to environmental parameters [@sayre2013new].
 We used life histories reported by @mummenhoff2005phylogeny, grouping species with annual or perennial life histories.
-Perenniality was defined based on the traditional defition and included any form of perennial life history (e.g., herbs, shrubs, mixed, etc).
+Perenniality was defined based any form of perennial life history (e.g., herbs, shrubs, mixed, etc).
+We also repeated these analyses without _H.
+subulata_ and _H.
+minima_ which have mixed life histories and found similar results.
+Because the original authors didn't report the nature of the mixed trait [@mummenhoff2005phylogeny], (ie.
+plasticity vs.
+genetic variation), we classified these species here as perennial since they have the capacity to maintain vegetative growth after reproduction at least to some capacity.
 
-### Heliophila herbarium specimens
+### Heliophila occurance records
 ```{r raw_GBIF}
@@
 And as they become digitized [@soltis2017digitization], these collections have been used to study relationships between trait distributions, geography, and climate [@davis2015herbarium;@wolf2016altitudinal;@stropp2016mapping;@vaclavik2017effects].
-To characterize the diributions of annual and perennial _Heliophila_ species, all records for the genus _Heliophila_ were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018.
+To characterize the diributions of annual and perennial _Heliophila_ species, all records for the genus _Heliophila_ were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018 [@gbifdownload].
 
 ### Sequence data for phylogeny
-Aligned Heliophila ITS sequences were obtained from previous work by @mandakova2012whole.
-Aethionema, Alliaria, Cardamine, Chamira, and Rorippa ITS records from were downloaded from Genbank [] and aligned together with Heliophila ITS sequences using MAFFT as outgroups [].
+Aligned Heliophila ITS sequences were obtained from previous work by @mandakova2012whole.  Aethionema, Alliaria, Cardamine, Chamira, and Rorippa ITS records from were downloaded from Genbank.
 
@@
 ### Drought frequency calculations
-
 To characterize drought regimens across the distrubtions of annual and perennial species of _Heliohpila_, we calculated drought during different seasons at the location of observations for _Heliophila_ records using the VHI.
@@
 
-### Herbarium record quality control
+### Filtering of occurance records
 ```{r cleanup}
@@
 
-To remove instance with spurious location data, we filtered raw GBIF by restricting our analyses to include only:  
-* Species with reported life history  
+To avoid instances with spurious location data, we filtered raw GBIF by restricting our analyses to include only:
+
+* Records for species with reported life history  
 * Records with geospatial data  
-* Records without known geospatial issues  
+* Records without known geospatial issues (i.e., coordinates reported are those of herbarium)  
 * Records from collection sites classified as land pixels   
 * Records from Africa  
-* Non-duplicate records (ie. identical species, location, collection date)
+* Records without duplicates (i.e., identical species, location, collection date)
 
 ### Phylogeny construction
-Model selection for construction of phylogeny was performed in jModeltest2 with CIPRES (cite).
+Out group species were aligned together with _Heliophila_ ITS sequences using MAFFT.
+Model selection for construction of phylogeny was performed in jModeltest2 with CIPRES.
 Based on this analysis, $GTR + L$ were selected.
-Ultrametric phylogeny was estimated with branch lengths as relative time (details).
+Ultrametric phylogeny was estimated with branch lengths as relative time.
 
-### Comparison of drought frequency between annual and perennial
+### Comparison of drought frequency between annual and perennial species
 ```{r speciesmeans}
@@
 tree<- rescale(tree, "depth", 1)
-
+tree<-keep.tip(tree, tip = gsub(" ","_", as.character(species_means$species)))
 phylomodelstable<-data.frame()
@@
 To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimes, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency.
-The relationship between drought frequencies across each taxon’s range and life habitat (annual or perennial) was evaluated using Firth's penalized-likelihood logistic regression and phylogenetic logistic regression.
+First, we compared the frequency of drought during the winter, spring, summer, and fall between raw occurance records of annual and perennial species by t-tests.
+To account for variation in the number of occurance records per species, we next calculated the mean drought frequency during the winter, spring, summer and fall for each species.
+The relationships between species mean values of drought frequency during each season and life habitat (annual or perennial) were tested using Firth's penalized-likelihood logistic regressions.
+Because demographic histories caused by ancestry can confound trait - environment associations, we then tested for the relationships between drought frequeny and life history while controlling for relationships between species using phylogenetic logistic regressions.
 
@@
 # Results
-
-To test the hypothesis that annual and perennial plants reflect adaptation to alternative drought environments we examined the landscape distribution of life history strategies in the large and diverse mustard genus, _Heliophila_ Figure \@ref(fig:phylogeny).
-Using both herbarium specimen metadata and a 30 year dataset of satellite generated climate information, we tested the prediction that annual species  are more often observed in drought-prone locations than perennial species, when controlling for phylogenetic relatedness.
-We found that drought frequency is significantly different between the distributions of annual and perennial species, with annuals being found in environments with significantly more frequent drought, and that this signal is strongest during the summer.
-These results remain significant while controlling for the phylogenetic relationships of Heliophila species, yielding support for the role that natural selection has played in driving contemporary distributions of these alternatives strategies in relation to drought regimes.
-
-(ref:phylogeny) Phylogeny of Heliophila.
+(ref:phylogeny) Species and examples of herbaria specimens of _Heliophila_ (a) Phylogeny and life history strategies of species studied.
+Orange circles at branch tips mark annual species and blue circles mark perennial species.
+Example herbaria specimens accessed via GBIF of (a) _H.
+minima_, (b) _H.
+deserticola_, (c) _H.
+coronopifolia_ and (d) _H.
+ephemera_.
+Images (a,c,d) courtesy of The Bavarian Natural History Collections (CC BY-SA 4.0) and (b) The London  Natural History Museum (CC BY 4.0).
+Links to images are found in the supplement.
 
@@
 tree$tip.label<-gsub("_"," ", tree$tip.label)
+tree<-keep.tip(tree, tip =as.character(species_means$species))
 
@@
   geom_tippoint(size=3,col=ifelse(species_life_history$life_history[match(tree$tip.label,species_life_history$species)]=="a", "orange3", "dodgerblue3"))+
-  geom_tiplab(offset = .05 , size=3)
+  geom_tiplab(offset = .05 , size=3)+
+  theme(plot.background = element_rect(fill=alpha("white",0)))
 
@@
 pics<-plot_grid(minima, deserticola, coronopifolia, ephemera,
-                labels = c("b","c","d","e"), ncol = 1)
+                labels = c("b","c","d","e"), ncol = 1, hjust = -1)
 
-plot_grid(treeplot, pics, ncol=2,labels=c("a",""), rel_widths = c(1,0.3))
+plot_grid(treeplot, pics, ncol=2,labels=c("a",""), rel_widths = c(1,0.5))
 
@@
 ```
-
-### GBIF records
-
-Out of `r nrow(GBIF)` _Heliophila_ GBIF recrods,  `r sum(GBIF$species %in% species_life_history$species)` were for species with reported life history [@mummenhoff2005phylogeny], `r sum(!is.na(GBIF$decimalLatitude))` had geospatial data, `r sum(!(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)))` did not have geospatial issues, `r sum(!(is.na(GBIF$Summer)) & !(is.na(GBIF$decimalLatitude)))` were located on pixels classified as land having drought measurements, `r sum(GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude))` were located in Africa, `r sum(!duplicated(GBIF$ymd_lat_lon_sp) & !is.na(GBIF$decimalLatitude)) ` were not duplicated.
-After all filtering steps, `r nrow(GBIF_cleaned)` records for `r length(unique(GBIF_cleaned$species))` species (Figure \@ref(fig:mapsboxplots), Table \@ref(tab:speciesmeanstable)) passed.
-
-### Drought frequency
-
-Drought frequency was signicantly different between annual and perennial species.
-
-(ref:mapsboxplots) Locations of annual and perennial Heliophila.
-Drought frequency during the winter, spring, summer and fall.
-Drought frequencies during each season observed at the collection sites of Heliophila records.
+(ref:mapsboxplots) Locations of (a) annual and (b) perennial Heliophila.
+Drought frequency during the (c) winter, (d) spring, (e) summer and (f) fall.
+(g) Drought frequencies during each season observed at the collection sites of _Heliophila_ records.
 
@@
 world <- map_data("world") # we already did this, but we can do it again
-library(cowplot)
 
@@
   geom_jitter(width = .3, alpha=0.1, pch=3)+
-  geom_boxplot(outlier.colour = "white", outlier.fill = "white", fill=NA, width=.3)+
+  geom_boxplot(outlier.colour = "white", outlier.fill = NA, fill=NA, width=.3)+
   scale_color_manual(values=c("orange3","dodgerblue3"))+
@@
 ```
+Out of `r nrow(GBIF)` _Heliophila_ GBIF recrods,  `r sum(GBIF$species %in% species_life_history$species)` were for species with reported life history [@mummenhoff2005phylogeny], `r sum(!is.na(GBIF$decimalLatitude))` had geospatial data, `r sum(!(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)))` did not have geospatial issues, `r sum(!(is.na(GBIF$Summer)) & !(is.na(GBIF$decimalLatitude)))` were located on pixels classified as land having drought measurements, `r sum(GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude))` were located in Africa, `r sum(!duplicated(GBIF$ymd_lat_lon_sp) & !is.na(GBIF$decimalLatitude)) ` were not duplicated.
+After all filtering steps, `r nrow(GBIF_cleaned)` records for `r length(unique(GBIF_cleaned$species))` species (Figure \@ref(fig:phylogeny), Table \@ref(tab:speciesmeanstable)) passed for further analyses.
+The number of samples varied between species, with a mean of `r mean(species_means$n)` samples per species.
+`r as.character(species_means$species[species_means$n==max(species_means$n)])` had the most records, `r max(species_means$n)`,  and  `r as.character(species_means$species[species_means$n==min(species_means$n)])` the fewest,  `r min(species_means$n)` (Table \@ref(tab:speciesmeanstable)).
+
+There were clear visual differences between the distributions of the `r sum(GBIF_cleaned$life_history=="a")` annual and the `r sum(GBIF_cleaned$life_history=="p")` perennial _Heliophila_ observation records (see Figure \@ref(fig:speciesmaps) for maps of individual species).
+While annual species were generally found in the western regions of South Africa and the Namibia, primarily in the Cape Floristic Region and Succulent Karoo (Figure \@ref(fig:mapsboxplots)a), the occurances of perennials extended to the east coast of South Africa (Figure \@ref(fig:mapsboxplots)b).
+
+```{r ttests, eval=T, message=F, warning="hide", results="hide"}
+twin<-t.test(Winter~life_history, GBIF_cleaned)
+tspr<-t.test(Spring~life_history, GBIF_cleaned)
+tsum<-t.test(Summer~life_history, GBIF_cleaned)
+tfal<-t.test(Fall~life_history, GBIF_cleaned)
+```
+
+The frequency of drought varied considerably across the ranges of _Heliophila_ species (Figure \@ref(fig:mapsboxplots)c-f).
+This heterogeneity is expected, given that this is one of the most climatically diverse regions of the Earth [@sayre2013new].
+It is worth noting the east to west cline in drought frequency observed during the summer, which distinguishes the high drought frequency of the Kalahari Sands and Namid Destert phyotogeographic regions from the low drought frequency of the Drakensberg Mountains and Coastal Zambesian phytogeographic regions.
+In the Cape phytogeographic region there was finer scale heterogeneity in drought frequency during the summer.
+
+Theory has predicted that annuality should be adaptive in places where stresses such as drought are more common.
+Conversely, perenniality should be adaptive in places where such stresses are less frequent.
+We found that the frequency of drought was significantly higher at the locations of occurance records for annual species.
+In terms of raw observation records
+(Figure @\ref(fig:mapsboxplots)g), the frequency of drought was signficantly higher at the location of annuals during the winter (t = `r twin$statistic`, p = `r twin$p.value`), spring (t = `r tspr$statistic`, p = `r tspr$p.value`), summer (t = `r tsum$statistic`, p = `r tsum$p.value`), and fall (t = `r tfal$statistic`, p = `r tfal$p.value`).
+Because raw occurance records do not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable))), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials by a Firths-penalized logistic regression.
+We found that the mean drought frequencies were signficantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a).
+We further tested whether annual species are found in places where droughts occur more frequently while controlling for the phylogentic relatedness of _Heliophila_ species.
+This is important, because environmental differences in species distributions can be confounded with demographic history caused by acestry.
+We found that while controlling for phylogeny, the mean drought frequencies were signficantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a).
+These findings indicate that common acestry alone does not explain differences the drought frequencies experienced in the environemnts of annual and perennial _Heliophila_.
 
@@
 
-### Collection dates
-
 (ref:lineplots) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of Heliophila.
@@
 
-```{r makelineplots, eval=T, message=F, results="hide"}
+```{r makelineplots, eval=T, message=F, results="hide", warning=F}
 
@@
 
-#ks.test(annuals$days,perennials$days)
-t.test(days~life_history, GBIF_cleaned)
-bartlett.test(days~life_history, GBIF_cleaned)
+kout<-ks.test(annuals$days,perennials$days)
+tout<-t.test(days~life_history, GBIF_cleaned)
+bout<-bartlett.test(days~life_history, GBIF_cleaned)
 
@@
 
+The preceding results indicated that annual speices are found in environments where droughts are signifacantly more frequent, especially in the summer and fall.
+Classic life history theory hypothesizes that annuality reflects adaptation to such environments because it allows species to escape stressful conditions.
+If this is the case, we would expect that annuals spend the drought prone seasons of summer and fall as seeds.
+To test this hypothesis, we compared the dates of occurance records between annual and perennial _Heliophila_ species.
+The distributions reveal a considerable difference in the timing of observation of these two life histories.
+In comparison to perennials, which appear to be collected throughout the year, annuals are almost exclusively observed during the spring (Figure \@ref(fig:lineplots)b).
+The differences between the distrubution of collection dates were signicant by all tests (ks.test D = `r  kout$statistic`, p = `r  kout$p.value`; bartlett.test K2 = `r bout$statistic`, p = `r bout$p.value`) This is consistent with a model of life history in which annual species flower in the spring, set seed, senesce, and die before the summer.
+Thus, these annual species are likely to remain dormant during the summer and fall, when drought is the strongest predictor of the distrubutions of annual and perennial life histories (Figure \@ref(fig:lineplots)a).
+
 # Discussion
-## Summary
-We found that the distribution of annual and perennial species of Heliophila is significantly predicted by satellite detected historic drought frequencies.
-Annual species are found in environments that experience more frequent drought during the summer and fall quarters compared to perennials.
-This relationship was consistent while controlling for phylogenetic relatedness among the taxa studied, indicating that these distributions cannot be explained entirely by common ancestry.
-These results support the hypothesis that natural selection has played a role in shaping the contemporary distributions of these alternative life-history strategies.
 
-## Relationship to previous hypotheses/work
+To test the hypothesis that annual and perennial plants reflect adaptation to alternative drought environments we examined the landscape distribution of life history strategies in the large and diverse mustard genus, _Heliophila_.
+Using metadata of `r nrow(GBIF_cleaned)` occurance records and a 34 year dataset of satellite-detected droughts, we tested the prediction that annual species are more often observed in drought-prone locations than perennial species, when controlling for phylogenetic relatedness.
+We found that drought frequency is significantly different between the distributions of annual and perennial species, with annuals being found in environments with more frequent drought, and that this signal is strongest during the summer.
+These results remain significant while controlling for the phylogenetic relationships of _Heliophila_ species, yielding support for the role that natural selection has played in driving contemporary distributions of these alternatives strategies in relation to drought regimes.
 
-Herbaria records and satellite detected drought provide data with which the distributions of annual and perennial species can be compared with respect to historic drought frequency.
-However, it is necessary to control for the demographic history caused by common ancestry of species if evolutionary processes such as natural selection are to be invoked as explanations for any differences observed.
-If, for example, annual and perennial species show significantly different ranges with respect to historic drought this could be confounded by common ancestry if annuals originated from a common ancestor and vice versa.
-In this case, it would be challenging to distinguish between natural selection and demographic history.
-On the other hand, if annual and perennial life history strategies arose independently in multiple species, controlling for phylogenetic relationships allows us to better account for demography and make stronger assertions about the importance of processes such as natural selection to explain patterns.
+We cannot elimiate the possibility that counfounding traits or environmental variables are the causative factors explaining variation in the distributions of annual and perennial species.
+Nevertheless, these results reveal quantitative support for the classic prediction that annual species are found in environemnts that experience more frequent drought than perennial species.
+These findings complement previous reports of qualitative associations between annaulity with enviroments characterized as having increased aridity [@evans2005climate], alternative precipitation defined habitats [@morishima1984differentiation;@datson2008climate], or greater unpredictability [@cruz2009molecular].
+However, to our knowledge this is the first study to demonstrate a signficant association between life history and drought in a phylogenetic context informed by large scale species distribution data and long term drought measures.
+
+Unfortunately, herbarium collections and their associated data do not represent systematic or random sampling of a species distribution.
+Significant biases in collecting exist, which we have not necessarily controlled for here, and may have some effect on our findings, such as a bias toward collecting near roads or near the locations of natural history collections [@daru2018widespread].
+Future research will benefit from systematic sampling efforts to avoid these noted biases.
+However, the ecosystems of southern Africa include several biodiversity hotspots and are among the most botanically well sampled regions on Earth [@daru2018widespread], suggesting that this may currently be the optimal region for our analyses of life history distribution.
+Indeed, we were able to use `r nrow(GBIF_cleaned)` occurance records to study 42 species, which represents a significant advance over relying on personal observations to characterize species distrubtuions.
 
 These findings support classical theoretical predictions about the adaptive value of annual and perennial life history strategies.
-Taken together, they suggest that in Heliophila, annual species are adapted to environments with increased summer droughts by avoiding these seasons in a dormant seed phase of their life cycle.
-Indeed, we found that very few annuals are collected during this season, supporting the prediction that they are not in a vegetative and/or reproductive phase at this time.
-Traditionally, the focus has been on the evolutionary origins of annual life histories {citations}.
-However, we also find evidence that the transition to perenniality could be explained by historical drought regimes.
-The phylogeny reveals several transitions from annual to perennial life history.
+Taken together, they suggest that in _Heliophila_, annual species are adapted to environments with increased summer droughts by avoiding these seasons in a dormant seed phase of their life cycle.
+They also suggest that perenniality is adaptive in environemnts where droughts are less frequent.
+While most previous work has focused on describing the evolutionary origins of annuality [@barrett1996comparative;@conti1999phylogenetic;@andreasen2001unequal;@verboom2004testing;@friedman2015all] there are at least a few other cases where perenniality appears to have arisen from an annual acestor [@bena1998molecular;@tank2008annuals].
+And while early theory predicted selection for annuality when adult morality is high [@stearns1992evolution],  we also find evidence that the transition to perenniality could be explained by historical drought regimes.
+The phylogeny reveals several transitions from annual to perennial life history (Figure \@ref(fig:phylogeny)a) and that perennials distributions extend into regions where drought frequency is low (Figure \@ref(fig:mapsboxplots)b,Figure \@ref(fig:speciesmaps)).
 Perennials may be able to out complete annual relatives in environments where the infrequency of drought favors strategies that allow plants to benefit from growth over many seasons.
+It may also indicate that annuals rely on drought as a source of disturbance for seedling recruitment when competing with perennials [@corbin2004competition].
+Indeed, no annual species were observed in the low drought regions of east South Africa \@ref(fig:mapsboxplots)a,Figure \@ref(fig:speciesmaps)).
 
-## Caveats
-Sampling bias [@daru2018widespread].
-Correlation does not prove causation. But it does indicate predictive power and is consistent with adaptive hypotheses.
-Herbarium collections and their associated data do not represent systematic or random sampling of a species distribution.
-Significant biases in collecting exist, which we have not necessarily controlled for here, and may have some effect on our findings, such as a bias toward collecting near roads or near the locations of natural history collections (Daru et al.
-2018, Heberling, in press).
-Despite these biases, the Cape Floristic region is a biodiversity hotspot and one of the most botanically well sampled regions on Earth (Daru et al.
-2018, Heberling, in press?), suggesting that this may currently be the optimal region for our analyses of life history distribution.
-Future research will benefit from systematic sampling efforts to avoid these noted biases.
-The climate data used here are assessed only from 198X-XXXX and do not reflect conditions at the estimated divergence dates of these species (XXX million of years ago).
-Rather, the results suggest that the current distributions of annual and perennial species reflect a history of environmental filtration and ongoing natural selection.
-That is, their distributions are non-random with respect to historic drought and this is not explained by phylogeny.
+These findings suggest that species with locally adapted life history strategies could be threatened by rapidly changing drought regimes [@dai2011drought].
+This could have impacts on ecosystem functioning if the composition of annual and perennial species changes as a response [@garnier1997specific; @roumet2006suites].
+Furthermore, the frequency of drought may be an important factor when considering the use of perennial cropping sysyems [@parry2005prospects;@lelievre2009current].
 
-## Broader implications 
-These findings suggest that rapidly changing drought regimes threaten species adapted to current environments.
-Studies predict changing drought regimes.
-This could have impacts on ecosystem functioning.
-This should also be considered when thinking about using perennial crops.
-Studies predict changing drought regimes.
-
-Do annuals require drought? We found independent origins of perenniality in environments where droughts are infrequent. Annuals may not be able to compete with perennials in the absence of droughts.
-
-## Conclusions 
-Perenniality appears to be adaptive in environments with less frequent drought. This work demonstrates the power of emerging data to address outstanding classic hypotheses in ecology and evolution.
+In conclusion, we find strong support for classic life history theory which predicts that annuality is adaptive environments where droughts occur more frequently.
+Additionally, we report evidence consistent with a life history model in annuals in which they escape drought prone seasons during the seed phase of their life cycle.
+Finally, we find evidence that the distributions of perennial lineages may indicate a competitive advantage in areas where droughts are infrequent.
+More broadly, this work demonstrates the power of emerging data to address outstanding classic hypotheses in ecology and evolution.
 
 # Acknowledgments 
-This work was supported by NSF no. USDA grant no. XXX to J.G.M.
+We thank __ for generous feedback that improved the quality of this work. This research was supported by NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
 
@@
 # Supplement
+
+### Images used
+https://www.gbif.org/occurrence/1099023487
+https://www.gbif.org/occurrence/1057389408
+https://www.gbif.org/occurrence/1099023562
+https://www.gbif.org/occurrence/1099023490
 
@@
 ```
+
+(ref:speciesmaps) Maps of occurance records for individual species. Orange points indicate annual species. Blue points indicate perennial species.
+
+```{r make_spceiesmaps, results = 'hide', echo = F, warnings="hide"}
+pdf("../figures/speciesmaps.pdf", height=8, width=8)
+
+world <- map_data("world") # we already did this, but we can do it again
+
+plot_grid(plotlist=lapply(unique(GBIF_cleaned$species), function(x){
+
+  lh<-unique((GBIF_cleaned %>% filter(species==x))$life_history)
+  lh_col=ifelse(lh=="a", "orange3", "dodgerblue3")
+
+  ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
+    coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
+    geom_point(GBIF_cleaned %>% filter(species==x), mapping=aes(x=decimalLongitude, y=decimalLatitude), pch=3, col=lh_col, alpha=0.3) +
+    theme_classic()+
+    theme( text = element_text(size=5))+
+    labs(title=x, x=NULL, y=NULL)
+
+}), ncol=6)
+
+dev.off()
+
+```
+
+```{r speciesmaps, fig.cap='(ref:speciesmaps)'}
+knitr::include_graphics("../figures/speciesmaps.pdf", dpi = 108)
+```

### 2018-11-27 manuscript/manuscript.tex
version 3

@@
 \hypersetup{unicode=true,
-            pdftitle={Satellite-detected drought frequencies predict plant life history strategies},
+            pdftitle={Satellite-detected droughts predict plant life history},
             pdfauthor={J. Grey Monroe, Brian Gill, Kathryn Turner, \& John K. McKay},
@@
 
-  \title{Satellite-detected drought frequencies predict plant life history
-strategies}
+  \title{Satellite-detected droughts predict plant life history}
     \author{J. Grey Monroe\textsuperscript{1,2}, Brian Gill\textsuperscript{3},
@@
 hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of
-flowering plants native to Southern Africa using herbarium occurrence
-records and satellite-detected drought histories. We find that perennial
-\textit{Heliophila} species are found in environments where droughts are
-significantly less frequent compared to annuals. These associations are
-predictive while controlling for phylogeny, lending support to the
-hypothesis that drought related natural selection has influenced the
-distributions of these strategies. Additionally, the difference in
-drought frequency between annual and perennial species distributions is
-greatest during the summer and fall, which also appears to be when
-annuals are in the seed phase of their life cycle based on collection
-dates of annual species. Together, these findings support traditional
-hypotheses about the drivers of life history strategy in plants - that
-perenniality is favored in environments with less frequent drought and
-annuality in enviroments with more frequent drought by allowing them to
-escape drought prone seasons as seeds. 
+flowering plants native to Africa by integrating 2192 herbaria
+occurrence records with 34 years of satellite-based drought detection.
+Consistent with predictions from classic life history theory, we find
+that perennial \textit{Heliophila} species occur in environments where
+droughts are significantly less frequent compared to annuals. These
+associations are predictive while controlling for phylogeny, lending
+support to the hypothesis that drought related natural selection has
+influenced the distributions of these strategies. Additionally, the
+difference in drought frequency between annual and perennial species
+distributions is greatest during the summer and fall, which also appears
+to be when annuals are in the seed phase of their life cycle based on
+collection dates of annual species. Together, these finding provide
+empirical support for classic hypotheses about the drivers of life
+history strategy in plants - that perennials out compete annuals in
+environments with less frequent drought and that annuals are adpated to
+enviroments with more frequent drought by escaping drought prone seasons
+as seeds. 
 }
@@
 
-\hypertarget{life-history-variation-in-plants}{%
-\subsection{Life history variation in
-plants}\label{life-history-variation-in-plants}}
-
 Understanding the causes and consequences of life history variation is a
@@
 which complete their seed to seed life cycle within a single calendar
-year from perennials (i.e.~polycarpic or iteroparous{]} which can
-persist over multiple years. Annual plants flower once, set seed,
-senesce, and then die, spending at least some portion of the year as a
-seed. In contrast, perennial plants can continue vegetative growth after
-reproduction and experience all seasons. These represent fundamentally
-different life history strategies, but the ecological factors that
-explain their evolution remain uresolved (Friedman \& Rubin, 2015).
-
-Identifying the drivers of selection for these alternative strategies is
-important because annuals and perennials have differing impacts on
-ecosystem functioning. For example, perennials have higher nitrogen
-concentration and larger specific root length in annuals, which both
-affect nutrient cycling (Garnier, Cordonnier, Guillerm, \& Sonié, 1997;
-Roumet, Urcelay, \& Dı'az, 2006). Furthermore, predicting the conditions
-that favor these strategies in nature is useful for developing more
-sustainable perennial cropping systems (Lelièvre \& Volaire, 2009).
-
-Classical theory predict shorter life spans and annuality in
-environments where adult mortality is high (Charnov \& Schaffer, 1973;
-Franco \& Silvertown, 1996; Stearns, 1992). In plants, this has been
-extended to the hypothessis that annuality is adaptive when it allows
-plants to escape drought (Schaffer \& M, 1975). Lack of water is perhaps
-the greatest threat to survival during vegetatitive or reproductive
-growth and annuals can remain dormant (and protected as a seed) during
-drought. Thus, environments with greater seasonal drought frequency may
-select for annual life histories that complete reproduction prior to
-drought prone seasons. Conversely, environments with less frequent
-drought may select for perennial species, which may benefit from
-multiple bouts of reproduction and competitive advantage (Corbin \&
-D'Antonio, 2004). These prediction has been supported by the association
-of annualilty with arid environments in wild rice (Morishima, Sano, \&
-Oka, 1984) and \emph{Oenothera} (Evans, Hearn, Hahn, Spangle, \&
-Venable, 2005). Additionally, annual and perennial species of Nemesia
-were qualitatively associated with winter rather and summer rainfall
-environments respectively (Datson, Murray, \& Steiner, 2008) and annual
-species of Scorzoneroides were associated with environments classified
-as unpredictable (Cruz-Mazo, Buide, Samuel, \& Narbona, 2009). These
-findings provide qualitative evidence supporting the prediction that
-annual species are found in environments that experience more frequent
-drought, and vice versa, but whether the history frequency of drought
-events indeed predicts the distributions annual or perennial life
-history strategies has yet to be tested.
-
+year from perennials (i.e.~polycarpic or iteroparous) which can persist
+over multiple years. Annual plants flower once, set seed, senesce, and
+then die, spending at least some portion of the year as a seed, where
+they are relatively protected from environmental stress. In contrast,
+perennial plants can continue vegetative growth after reproduction and
+must survive conditions experienced during all seasons. These represent
+fundamentally different life history strategies, but the ecological
+factors that explain their evolution and distributions remain
+empirically uresolved (Friedman \& Rubin, 2015).\\
+Classical theory predict shorter life spans in environments where adult
+mortality is high (Charnov \& Schaffer, 1973; Franco \& Silvertown,
+1996; Stearns, 1992). In plants, this has been extended to the
+hypothessis that annuality is adaptive when it allows plants to escape
+drought (Schaffer \& M, 1975). Lack of water is perhaps the greatest
+threat to survival during vegetatitive or reproductive growth and
+annuals can remain dormant (and protected as a seed) during drought.
+Thus, environments with greater seasonal drought frequency may select
+for annual life histories that complete reproduction prior to drought
+prone seasons. Conversely, environments with less frequent drought may
+select for perennial species, which may benefit from multiple bouts of
+reproduction and competitive advantage by preventing recruitment of
+annual species (Corbin \& D'Antonio, 2004). These predictions have been
+supported by the association of annualilty with arid environments in
+wild rice (Morishima, Sano, \& Oka, 1984) and \emph{Oenothera} (Evans,
+Hearn, Hahn, Spangle, \& Venable, 2005). Additionally, annual and
+perennial species of \emph{Nemesia} were qualitatively associated with
+winter rather and summer rainfall environments respectively (Datson,
+Murray, \& Steiner, 2008) and annual species of \emph{Scorzoneroides}
+were associated with environments classified as unpredictable
+(Cruz-Mazo, Buide, Samuel, \& Narbona, 2009). However, whether the
+history frequency of drought events indeed predicts the distributions
+annual or perennial life history strategies has yet to be tested in a
+phylogenetic context with suitable sample sizes of distrubution and
+climate data.\\
 Here we combine a long-term global dataset of satellite detected drought
@@
 significant non-random effects on species distributions and life history
-traits {[}{]}, and therefore we assessed the relationship between life
-history distribution and drought frequency in a phylogenetically
-controlled background.
+traits (Barrett, Harder, \& Worley, 1996), and therefore we assessed the
+relationship between life history distribution and drought frequency in
+a phylogenetically controlled background.
 
@@
 
-\hypertarget{availability}{%
-\subsection{Availability}\label{availability}}
-
-All data and the source code to produce this manuscript are available at
-\url{https://github.com/greymonroe/heliophila}.
-
 \hypertarget{data}{%
 \subsection{Data}\label{data}}
+
+\hypertarget{availability}{%
+\subsubsection{Availability}\label{availability}}
+
+All analyses were peformed using R. All data and the source code to
+produce this manuscript are available at
+\url{https://github.com/greymonroe/heliophila}.
 
@@
 - VCI) with temperature stress indicated by anomalies in thermal spectra
-(Temperature Condition Index - TCI).
-
-The VHI of year \(y\) during week \(w\) of \([1,52]\) at pixel \(i\) is
-derived from the following equations, where \(n\) is the number of years
-observed.
+(Temperature Condition Index - TCI). The VHI of year \(y\) during week
+\(w\) of \([1,52]\) at pixel \(i\) is derived from the following
+equations, where \(n\) is the number of years observed.
 
@@
 
-\[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\]
-
-Thus, VHI measurements are standardized according to conditions
-historically observed at each locations. These measurements have been
-validated and generally used for evaluating drought risk and predicting
-crop yields in agriculture (e.g., Rojas, Vrieling, \& Rembold, 2011;
-Kogan et al., 2016). But they also present a tool to study seasonal
-patterns in the frequency of drought across environments and to test
-hypotheses about the effect of drought on ecological and evolutionary
-processes. Indeed, the VHI has been applied recently to study drought
-related ecology of natural species and proven useful for predicting
-infraspecific variation in drought tolerance traits and genes
+\[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\] Thus, VHI
+measurements are standardized according to conditions historically
+observed at each locations. These measurements have been validated and
+generally used for evaluating drought risk and predicting crop yields in
+agriculture (e.g., Rojas, Vrieling, \& Rembold, 2011; Kogan et al.,
+2016). But they also present a new tool to study seasonal patterns in
+the frequency of drought across environments and to test hypotheses
+about the effect of drought on ecological and evolutionary processes
+(Kerr \& Ostrovsky, 2003). As such, the VHI has been applied recently to
+study drought related ecology of natural species and proven useful for
+predicting infraspecific variation in drought tolerance traits and genes
 (Dittberner et al., 2018; Mojica et al., 2016; Monroe et al., 2018).
@@
 characterize the seasonal drought frequencies experienced by annual and
-perennial \emph{Heliophila} sepecies.
+perennial \emph{Heliophila} species.
 
@@
 history strategy has likely arisen multiple independent times (Appel \&
-Al-Shehbaz, 1997; Mummenhoff et al., 2005). We used life histories
-reported by Mummenhoff et al. (2005), grouping species with annual or
-perennial life histories. Perenniality was defined based on the
-traditional defition and included any form of perennial life history
-(e.g., herbs, shrubs, mixed, etc).
-
-\hypertarget{heliophila-herbarium-specimens}{%
-\subsubsection{Heliophila herbarium
-specimens}\label{heliophila-herbarium-specimens}}
+Al-Shehbaz, 1997; Mummenhoff et al., 2005). Furthermore, the fine scale
+climatic heterogeneity of Southern Africa is ideal for studying the
+distribution of traits in relation to environmental parameters (Sayre et
+al., 2013). We used life histories reported by Mummenhoff et al. (2005),
+grouping species with annual or perennial life histories. Perenniality
+was defined based any form of perennial life history (e.g., herbs,
+shrubs, mixed, etc). We also repeated these analyses without \emph{H.
+subulata} and \emph{H. minima} which have mixed life histories and found
+similar results. Because the original authors didn't report the nature
+of the mixed trait (Mummenhoff et al., 2005), (ie. plasticity
+vs.~genetic variation), we classified these species here as perennial
+since they have the capacity to maintain vegetative growth after
+reproduction at least to some capacity.
+
+\hypertarget{heliophila-occurance-records}{%
+\subsubsection{Heliophila occurance
+records}\label{heliophila-occurance-records}}
 
@@
 genus \emph{Heliophila} were downloaded from the Global Biodiversity
-Information Facility (gbif.org) on July 21, 2018.
+Information Facility (gbif.org) on July 21, 2018 (``GBIF occurrence
+download,'' 2018).
 
@@
 Mandáková et al. (2012). Aethionema, Alliaria, Cardamine, Chamira, and
-Rorippa ITS records from were downloaded from Genbank {[}{]} and aligned
-together with Heliophila ITS sequences using MAFFT as outgroups {[}{]}.
+Rorippa ITS records from were downloaded from Genbank.
 
@@
 
-\hypertarget{herbarium-record-quality-control}{%
-\subsubsection{Herbarium record quality
-control}\label{herbarium-record-quality-control}}
-
-To remove instance with spurious location data, we filtered raw GBIF by
-restricting our analyses to include only:\\
-* Species with reported life history\\
-* Records with geospatial data\\
-* Records without known geospatial issues\\
-* Records from collection sites classified as land pixels\\
-* Records from Africa\\
-* Non-duplicate records (ie. identical species, location, collection
-date)
+\hypertarget{filtering-of-occurance-records}{%
+\subsubsection{Filtering of occurance
+records}\label{filtering-of-occurance-records}}
+
+To avoid instances with spurious location data, we filtered raw GBIF by
+restricting our analyses to include only:
+
+\begin{itemize}
+\tightlist
+\item
+  Records for species with reported life history\\
+\item
+  Records with geospatial data\\
+\item
+  Records without known geospatial issues (i.e., coordinates reported
+  are those of herbarium)\\
+\item
+  Records from collection sites classified as land pixels\\
+\item
+  Records from Africa\\
+\item
+  Records without duplicates (i.e., identical species, location,
+  collection date)
+\end{itemize}
 
@@
 
-Model selection for construction of phylogeny was performed in
-jModeltest2 with CIPRES (cite). Based on this analysis, \(GTR + L\) were
-selected. Ultrametric phylogeny was estimated with branch lengths as
-relative time (details).
-
-\hypertarget{comparison-of-drought-frequency-between-annual-and-perennial}{%
+Out group species were aligned together with \emph{Heliophila} ITS
+sequences using MAFFT. Model selection for construction of phylogeny was
+performed in jModeltest2 with CIPRES. Based on this analysis,
+\(GTR + L\) were selected. Ultrametric phylogeny was estimated with
+branch lengths as relative time.
+
+\hypertarget{comparison-of-drought-frequency-between-annual-and-perennial-species}{%
 \subsubsection{Comparison of drought frequency between annual and
-perennial}\label{comparison-of-drought-frequency-between-annual-and-perennial}}
+perennial
+species}\label{comparison-of-drought-frequency-between-annual-and-perennial-species}}
 
@@
 and perennial \emph{Heliophila} species would be significantly
-associated with historic drought frequency. The relationship between
-drought frequencies across each taxon's range and life habitat (annual
-or perennial) was evaluated using Firth's penalized-likelihood logistic
-regression and phylogenetic logistic regression.
+associated with historic drought frequency. First, we compared the
+frequency of drought during the winter, spring, summer, and fall between
+raw occurance records of annual and perennial species by t-tests. To
+account for variation in the number of occurance records per species, we
+next calculated the mean drought frequency during the winter, spring,
+summer and fall for each species. The relationships between species mean
+values of drought frequency during each season and life habitat (annual
+or perennial) were tested using Firth's penalized-likelihood logistic
+regressions. Because demographic histories caused by ancestry can
+confound trait - environment associations, we then tested for the
+relationships between drought frequeny and life history while
+controlling for relationships between species using phylogenetic
+logistic regressions.
 
@@
 
-To test the hypothesis that annual and perennial plants reflect
-adaptation to alternative drought environments we examined the landscape
-distribution of life history strategies in the large and diverse mustard
-genus, \emph{Heliophila} Figure \ref{fig:phylogeny}. Using both
-herbarium specimen metadata and a 30 year dataset of satellite generated
-climate information, we tested the prediction that annual species are
-more often observed in drought-prone locations than perennial species,
-when controlling for phylogenetic relatedness. We found that drought
-frequency is significantly different between the distributions of annual
-and perennial species, with annuals being found in environments with
-significantly more frequent drought, and that this signal is strongest
-during the summer. These results remain significant while controlling
-for the phylogenetic relationships of Heliophila species, yielding
-support for the role that natural selection has played in driving
-contemporary distributions of these alternatives strategies in relation
-to drought regimes.
-
 \begin{figure}[!h]
-\includegraphics[width=\textwidth]{../figures/phylogeny} \caption{Phylogeny of Heliophila.}\label{fig:phylogeny}
+\includegraphics[width=\textwidth]{../figures/phylogeny} \caption{Species and examples of herbaria specimens of
+\emph{Heliophila} (a) Phylogeny and life history strategies of species
+studied. Orange circles at branch tips mark annual species and blue
+circles mark perennial species. Example herbaria specimens accessed via
+GBIF of (a) \emph{H. minima}, (b) \emph{H. deserticola}, (c) \emph{H.
+coronopifolia} and (d) \emph{H. ephemera}. Images (a,c,d) courtesy of
+The Bavarian Natural History Collections (CC BY-SA 4.0) and (b) The
+London Natural History Museum (CC BY 4.0). Links to images are found in
+the supplement.}\label{fig:phylogeny}
 \end{figure}
 
-\hypertarget{gbif-records}{%
-\subsubsection{GBIF records}\label{gbif-records}}
+\begin{figure}[!h]
+\includegraphics[width=\textwidth]{../figures/maps_boxplots} \caption{Locations of (a) annual and (b) perennial Heliophila.
+Drought frequency during the (c) winter, (d) spring, (e) summer and (f)
+fall. (g) Drought frequencies during each season observed at the
+collection sites of \emph{Heliophila} records.}\label{fig:mapsboxplots}
+\end{figure}
 
@@
 Africa, 3162 were not duplicated. After all filtering steps, 2192
-records for 42 species (Figure \ref{fig:mapsboxplots}, Table
-\ref{tab:speciesmeanstable}) passed.
-
-\hypertarget{drought-frequency}{%
-\subsubsection{Drought frequency}\label{drought-frequency}}
-
-Drought frequency was signicantly different between annual and perennial
-species.
-
-\begin{figure}[!h]
-\includegraphics[width=\textwidth]{../figures/maps_boxplots} \caption{Locations of annual and perennial Heliophila. Drought
-frequency during the winter, spring, summer and fall. Drought
-frequencies during each season observed at the collection sites of
-Heliophila records.}\label{fig:mapsboxplots}
-\end{figure}
+records for 42 species (Figure \ref{fig:phylogeny}, Table
+\ref{tab:speciesmeanstable}) passed for further analyses. The number of
+samples varied between species, with a mean of 52.19 samples per
+species. Heliophila rigidiuscula had the most records, 201, and
+Heliophila cornellsbergia the fewest, 2 (Table
+\ref{tab:speciesmeanstable}).
+
+There were clear visual differences between the distributions of the 960
+annual and the 1232 perennial \emph{Heliophila} observation records (see
+Figure \ref{fig:speciesmaps} for maps of individual species). While
+annual species were generally found in the western regions of South
+Africa and the Namibia, primarily in the Cape Floristic Region and
+Succulent Karoo (Figure \ref{fig:mapsboxplots}a), the occurances of
+perennials extended to the east coast of South Africa (Figure
+\ref{fig:mapsboxplots}b).
+
+The frequency of drought varied considerably across the ranges of
+\emph{Heliophila} species (Figure \ref{fig:mapsboxplots}c-f). This
+heterogeneity is expected, given that this is one of the most
+climatically diverse regions of the Earth (Sayre et al., 2013). It is
+worth noting the east to west cline in drought frequency observed during
+the summer, which distinguishes the high drought frequency of the
+Kalahari Sands and Namid Destert phyotogeographic regions from the low
+drought frequency of the Drakensberg Mountains and Coastal Zambesian
+phytogeographic regions. In the Cape phytogeographic region there was
+finer scale heterogeneity in drought frequency during the summer.
+
+Theory has predicted that annuality should be adaptive in places where
+stresses such as drought are more common. Conversely, perenniality
+should be adaptive in places where such stresses are less frequent. We
+found that the frequency of drought was significantly higher at the
+locations of occurance records for annual species. In terms of raw
+observation records (Figure @\ref(fig:mapsboxplots)g), the frequency of
+drought was signficantly higher at the location of annuals during the
+winter (t = 10.65, p = 0.00), spring (t = 10.73, p = 0.00), summer (t =
+12.67, p = 0.00), and fall (t = 15.26, p = 0.00). Because raw occurance
+records do not account for variation in the number of records per
+species (Table \ref{tab:speciesmeanstable})), we also tested whether
+mean drought frequency values of each species were significantly
+different between annuals and perennials by a Firths-penalized logistic
+regression. We found that the mean drought frequencies were signficantly
+higher (\(\alpha = 0.05\)) in annual species during the spring, summer,
+and fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a). We
+further tested whether annual species are found in places where droughts
+occur more frequently while controlling for the phylogentic relatedness
+of \emph{Heliophila} species. This is important, because environmental
+differences in species distributions can be confounded with demographic
+history caused by acestry. We found that while controlling for
+phylogeny, the mean drought frequencies were signficantly higher
+(\(\alpha = 0.05\)) in annual species during the spring, summer, and
+fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a). These
+findings indicate that common acestry alone does not explain differences
+the drought frequencies experienced in the environemnts of annual and
+perennial \emph{Heliophila}.
 
@@
 
-\hypertarget{collection-dates-1}{%
-\subsubsection{Collection dates}\label{collection-dates-1}}
-
 \begin{figure}
@@
 
+The preceding results indicated that annual speices are found in
+environments where droughts are signifacantly more frequent, especially
+in the summer and fall. Classic life history theory hypothesizes that
+annuality reflects adaptation to such environments because it allows
+species to escape stressful conditions. If this is the case, we would
+expect that annuals spend the drought prone seasons of summer and fall
+as seeds. To test this hypothesis, we compared the dates of occurance
+records between annual and perennial \emph{Heliophila} species. The
+distributions reveal a considerable difference in the timing of
+observation of these two life histories. In comparison to perennials,
+which appear to be collected throughout the year, annuals are almost
+exclusively observed during the spring (Figure \ref{fig:lineplots}b).
+The differences between the distrubution of collection dates were
+signicant by all tests (ks.test D = 0.25, p = 0; bartlett.test K2 =
+503.18, p = 0.00) This is consistent with a model of life history in
+which annual species flower in the spring, set seed, senesce, and die
+before the summer. Thus, these annual species are likely to remain
+dormant during the summer and fall, when drought is the strongest
+predictor of the distrubutions of annual and perennial life histories
+(Figure \ref{fig:lineplots}a).
+
 \hypertarget{discussion}{%
@@
 
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
-
-We found that the distribution of annual and perennial species of
-Heliophila is significantly predicted by satellite detected historic
-drought frequencies. Annual species are found in environments that
-experience more frequent drought during the summer and fall quarters
-compared to perennials. This relationship was consistent while
-controlling for phylogenetic relatedness among the taxa studied,
-indicating that these distributions cannot be explained entirely by
-common ancestry. These results support the hypothesis that natural
-selection has played a role in shaping the contemporary distributions of
-these alternative life-history strategies.
-
-\hypertarget{relationship-to-previous-hypotheseswork}{%
-\subsection{Relationship to previous
-hypotheses/work}\label{relationship-to-previous-hypotheseswork}}
-
-Herbaria records and satellite detected drought provide data with which
-the distributions of annual and perennial species can be compared with
-respect to historic drought frequency. However, it is necessary to
-control for the demographic history caused by common ancestry of species
-if evolutionary processes such as natural selection are to be invoked as
-explanations for any differences observed. If, for example, annual and
-perennial species show significantly different ranges with respect to
-historic drought this could be confounded by common ancestry if annuals
-originated from a common ancestor and vice versa. In this case, it would
-be challenging to distinguish between natural selection and demographic
-history. On the other hand, if annual and perennial life history
-strategies arose independently in multiple species, controlling for
-phylogenetic relationships allows us to better account for demography
-and make stronger assertions about the importance of processes such as
-natural selection to explain patterns.
-
-These findings support classical theoretical predictions about the
-adaptive value of annual and perennial life history strategies. Taken
-together, they suggest that in Heliophila, annual species are adapted to
-environments with increased summer droughts by avoiding these seasons in
-a dormant seed phase of their life cycle. Indeed, we found that very few
-annuals are collected during this season, supporting the prediction that
-they are not in a vegetative and/or reproductive phase at this time.
-Traditionally, the focus has been on the evolutionary origins of annual
-life histories \{citations\}. However, we also find evidence that the
-transition to perenniality could be explained by historical drought
-regimes. The phylogeny reveals several transitions from annual to
-perennial life history. Perennials may be able to out complete annual
-relatives in environments where the infrequency of drought favors
-strategies that allow plants to benefit from growth over many seasons.
-
-\hypertarget{caveats}{%
-\subsection{Caveats}\label{caveats}}
-
-Sampling bias (Daru et al., 2018). Correlation does not prove causation.
-But it does indicate predictive power and is consistent with adaptive
-hypotheses. Herbarium collections and their associated data do not
+To test the hypothesis that annual and perennial plants reflect
+adaptation to alternative drought environments we examined the landscape
+distribution of life history strategies in the large and diverse mustard
+genus, \emph{Heliophila}. Using metadata of 2192 occurance records and a
+34 year dataset of satellite-detected droughts, we tested the prediction
+that annual species are more often observed in drought-prone locations
+than perennial species, when controlling for phylogenetic relatedness.
+We found that drought frequency is significantly different between the
+distributions of annual and perennial species, with annuals being found
+in environments with more frequent drought, and that this signal is
+strongest during the summer. These results remain significant while
+controlling for the phylogenetic relationships of \emph{Heliophila}
+species, yielding support for the role that natural selection has played
+in driving contemporary distributions of these alternatives strategies
+in relation to drought regimes.
+
+We cannot elimiate the possibility that counfounding traits or
+environmental variables are the causative factors explaining variation
+in the distributions of annual and perennial species. Nevertheless,
+these results reveal quantitative support for the classic prediction
+that annual species are found in environemnts that experience more
+frequent drought than perennial species. These findings complement
+previous reports of qualitative associations between annaulity with
+enviroments characterized as having increased aridity (Evans et al.,
+2005), alternative precipitation defined habitats (Datson et al., 2008;
+Morishima et al., 1984), or greater unpredictability (Cruz-Mazo et al.,
+2009). However, to our knowledge this is the first study to demonstrate
+a signficant association between life history and drought in a
+phylogenetic context informed by large scale species distribution data
+and long term drought measures.
+
+Unfortunately, herbarium collections and their associated data do not
 represent systematic or random sampling of a species distribution.
@@
 bias toward collecting near roads or near the locations of natural
-history collections (Daru et al.~2018, Heberling, in press). Despite
-these biases, the Cape Floristic region is a biodiversity hotspot and
-one of the most botanically well sampled regions on Earth (Daru et
-al.~2018, Heberling, in press?), suggesting that this may currently be
-the optimal region for our analyses of life history distribution. Future
-research will benefit from systematic sampling efforts to avoid these
-noted biases. The climate data used here are assessed only from
-198X-XXXX and do not reflect conditions at the estimated divergence
-dates of these species (XXX million of years ago). Rather, the results
-suggest that the current distributions of annual and perennial species
-reflect a history of environmental filtration and ongoing natural
-selection. That is, their distributions are non-random with respect to
-historic drought and this is not explained by phylogeny.
-
-\hypertarget{broader-implications}{%
-\subsection{Broader implications}\label{broader-implications}}
-
-These findings suggest that rapidly changing drought regimes threaten
-species adapted to current environments. Studies predict changing
-drought regimes. This could have impacts on ecosystem functioning. This
-should also be considered when thinking about using perennial crops.
-Studies predict changing drought regimes.
-
-Do annuals require drought? We found independent origins of perenniality
-in environments where droughts are infrequent. Annuals may not be able
-to compete with perennials in the absence of droughts.
-
-\hypertarget{conclusions}{%
-\subsection{Conclusions}\label{conclusions}}
-
-Perenniality appears to be adaptive in environments with less frequent
-drought. This work demonstrates the power of emerging data to address
+history collections (Daru et al., 2018). Future research will benefit
+from systematic sampling efforts to avoid these noted biases. However,
+the ecosystems of southern Africa include several biodiversity hotspots
+and are among the most botanically well sampled regions on Earth (Daru
+et al., 2018), suggesting that this may currently be the optimal region
+for our analyses of life history distribution. Indeed, we were able to
+use 2192 occurance records to study 42 species, which represents a
+significant advance over relying on personal observations to
+characterize species distrubtuions.
+
+These findings support classical theoretical predictions about the
+adaptive value of annual and perennial life history strategies. Taken
+together, they suggest that in \emph{Heliophila}, annual species are
+adapted to environments with increased summer droughts by avoiding these
+seasons in a dormant seed phase of their life cycle. They also suggest
+that perenniality is adaptive in environemnts where droughts are less
+frequent. While most previous work has focused on describing the
+evolutionary origins of annuality (Andreasen \& Baldwin, 2001; Barrett
+et al., 1996; Conti, Soltis, Hardig, \& Schneider, 1999; Friedman \&
+Rubin, 2015; Verboom, Linder, \& Stock, 2004) there are at least a few
+other cases where perenniality appears to have arisen from an annual
+acestor (Bena, Lejeune, Prosperi, \& Olivieri, 1998; Tank \& Olmstead,
+2008). And while early theory predicted selection for annuality when
+adult morality is high (Stearns, 1992), we also find evidence that the
+transition to perenniality could be explained by historical drought
+regimes. The phylogeny reveals several transitions from annual to
+perennial life history (Figure \ref{fig:phylogeny}a) and that perennials
+distributions extend into regions where drought frequency is low (Figure
+\ref{fig:mapsboxplots}b,Figure \ref{fig:speciesmaps}). Perennials may be
+able to out complete annual relatives in environments where the
+infrequency of drought favors strategies that allow plants to benefit
+from growth over many seasons. It may also indicate that annuals rely on
+drought as a source of disturbance for seedling recruitment when
+competing with perennials (Corbin \& D'Antonio, 2004). Indeed, no annual
+species were observed in the low drought regions of east South Africa
+\ref{fig:mapsboxplots}a,Figure \ref{fig:speciesmaps}).
+
+These findings suggest that species with locally adapted life history
+strategies could be threatened by rapidly changing drought regimes (Dai,
+2011). This could have impacts on ecosystem functioning if the
+composition of annual and perennial species changes as a response
+(Garnier, Cordonnier, Guillerm, \& Sonié, 1997; Roumet, Urcelay, \&
+Dı'az, 2006). Furthermore, the frequency of drought may be an important
+factor when considering the use of perennial cropping sysyems (Lelièvre
+\& Volaire, 2009; Parry, Flexas, \& Medrano, 2005).
+
+In conclusion, we find strong support for classic life history theory
+which predicts that annuality is adaptive environments where droughts
+occur more frequently. Additionally, we report evidence consistent with
+a life history model in annuals in which they escape drought prone
+seasons during the seed phase of their life cycle. Finally, we find
+evidence that the distributions of perennial lineages may indicate a
+competitive advantage in areas where droughts are infrequent. More
+broadly, this work demonstrates the power of emerging data to address
 outstanding classic hypotheses in ecology and evolution.
@@
 
-This work was supported by NSF no. USDA grant no. XXX to J.G.M.
+We thank \_\_ for generous feedback that improved the quality of this
+work. This research was supported by NSF Award 1701918 and USDA-NIFA
+Award 2014-38420-21801 to JGM.
 
@@
 
+\leavevmode\hypertarget{ref-andreasen2001unequal}{}%
+Andreasen, K., \& Baldwin, B. G. (2001). Unequal evolutionary rates
+between annual and perennial lineages of checker mallows (sidalcea,
+malvaceae): Evidence from 18S--26S rDNA internal and external
+transcribed spacers. \emph{Molecular Biology and Evolution},
+\emph{18}(6), 936--944.
+
 \leavevmode\hypertarget{ref-appel1997generic}{}%
@@
 
+\leavevmode\hypertarget{ref-barrett1996comparative}{}%
+Barrett, S. C. H., Harder, L. D., \& Worley, A. C. (1996). The
+comparative biology of pollination and mating in flowering plants.
+\emph{Phil. Trans. R. Soc. Lond. B}, \emph{351}(1345), 1271--1280.
+
 \leavevmode\hypertarget{ref-R-Matrix}{}%
@@
 
+\leavevmode\hypertarget{ref-bena1998molecular}{}%
+Bena, G., Lejeune, B., Prosperi, J.-M., \& Olivieri, I. (1998).
+Molecular phylogenetic approach for studying life-history evolution: The
+ambiguous example of the genus medicago l. \emph{Proceedings of the
+Royal Society of London B: Biological Sciences}, \emph{265}(1401),
+1141--1151.
+
 \leavevmode\hypertarget{ref-brown1996oldlist}{}%
@@
 103--137.
+
+\leavevmode\hypertarget{ref-conti1999phylogenetic}{}%
+Conti, E., Soltis, D. E., Hardig, T. M., \& Schneider, J. (1999).
+Phylogenetic relationships of the silver saxifrages (saxifraga, sect.
+Ligulatae haworth): Implications for the evolution of substrate
+specificity, life histories, and biogeography. \emph{Molecular
+Phylogenetics and Evolution}, \emph{13}(3), 536--555.
 
@@
 Phylogenetics and Evolution}, \emph{53}(3), 835--847.
+
+\leavevmode\hypertarget{ref-dai2011drought}{}%
+Dai, A. (2011). Drought under global warming: A review. \emph{Wiley
+Interdisciplinary Reviews: Climate Change}, \emph{2}(1), 45--65.
 
@@
 
+\leavevmode\hypertarget{ref-gbifdownload}{}%
+GBIF occurrence download. (2018, July). Retrieved from
+\url{https://doi.org/10.15468/dl.xmwqxm}
+
 \leavevmode\hypertarget{ref-R-mvtnorm}{}%
@@
 
+\leavevmode\hypertarget{ref-kerr2003space}{}%
+Kerr, J. T., \& Ostrovsky, M. (2003). From space to species: Ecological
+applications for remote sensing. \emph{Trends in Ecology \& Evolution},
+\emph{18}(6), 299--305.
+
 \leavevmode\hypertarget{ref-kogan2016modelling}{}%
@@
 
+\leavevmode\hypertarget{ref-parry2005prospects}{}%
+Parry, M., Flexas, J., \& Medrano, H. (2005). Prospects for crop
+production under drought: Research priorities and future directions.
+\emph{Annals of Applied Biology}, \emph{147}(3), 211--226.
+
 \leavevmode\hypertarget{ref-R-sp}{}%
@@
 
+\leavevmode\hypertarget{ref-sayre2013new}{}%
+Sayre, R. G., Comer, P., Hak, J., Josse, C., Bow, J., Warner, H.,
+\ldots{} others. (2013). A new map of standardized terrestrial
+ecosystems of africa. \emph{African Geographical Review}.
+
 \leavevmode\hypertarget{ref-schaffer1975selection}{}%
@@
 
+\leavevmode\hypertarget{ref-tank2008annuals}{}%
+Tank, D. C., \& Olmstead, R. G. (2008). From annuals to perennials:
+Phylogeny of subtribe castillejinae (orobanchaceae). \emph{American
+Journal of Botany}, \emph{95}(5), 608--625.
+
 \leavevmode\hypertarget{ref-vaclavik2017effects}{}%
@@
 
+\leavevmode\hypertarget{ref-verboom2004testing}{}%
+Verboom, G. A., Linder, H. P., \& Stock, W. D. (2004). Testing the
+adaptive nature of radiation: Growth form and life history divergence in
+the african grass genus ehrharta (poaceae: Ehrhartoideae).
+\emph{American Journal of Botany}, \emph{91}(9), 1364--1370.
+
 \leavevmode\hypertarget{ref-R-ggplot2}{}%
@@
 \section{Supplement}\label{supplement}}
+
+\hypertarget{images-used}{%
+\subsubsection{Images used}\label{images-used}}
+
+\url{https://www.gbif.org/occurrence/1099023487}
+\url{https://www.gbif.org/occurrence/1057389408}
+\url{https://www.gbif.org/occurrence/1099023562}
+\url{https://www.gbif.org/occurrence/1099023490}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{../figures/speciesmaps.pdf}
+\caption{\label{fig:speciesmaps}Maps of occurance records for individual species.
+Orange points indicate annual species. Blue points indicate perennial
+species.}
+\end{figure}
+
 \end{document}

### 2018-11-27 src/sandbox.R

@@ -1,32 +1,0 @@
-
-pdf("../figures/phylogeny.pdf", height=6, width=5)
-
-treeplot<-ggtree(tree, layout = "rectangular")+
-  xlim(c(0,2))+
-  geom_tippoint(size=3,col=ifelse(species_life_history$life_history[match(tree$tip.label,species_life_history$species)]=="a", "orange3", "dodgerblue3"))+
-  geom_tiplab(offset = .01 )
-
-minima <- readPNG("../pictures/minima.png")
-minima <- rasterGrob(minima, interpolate=TRUE)
-#https://www.gbif.org/occurrence/1099023487
-
-deserticola <- readPNG("../pictures/deserticola.png")
-deserticola <- rasterGrob(deserticola, interpolate=TRUE)
-#https://www.gbif.org/occurrence/1057389408
-
-coronopifolia<-readPNG("../pictures/coronopifolia.png")
-coronopifolia <- rasterGrob(coronopifolia, interpolate=TRUE)
-#https://www.gbif.org/occurrence/1099023562
-
-ephemera<-readPNG("../pictures/ephemera.png")
-ephemera <- rasterGrob(ephemera, interpolate=TRUE)
-
-#https://www.gbif.org/occurrence/1099023490
-
-pics<-plot_grid(minima, deserticola, coronopifolia, ephemera,
-                labels = c("b","c","d","e"), ncol = 1)
-
-plot_grid(treeplot, pics, ncol=2,labels=c("a",""), rel_widths = c(1,0.35))
-
-
-dev.off()

### 2018-11-28 manuscript/manuscript.Rmd
edits from Kathryn

@@
 ---
-title             : "Satellite-detected droughts predict plant life history"
+title             : "Drought frequency predicts plant life history strategies"
 shorttitle        : "Drought and life history"
@@
     affiliation   : "3"
-  - name          : "Kathryn Turner"
+  - name          : "Kathryn G. Turner"
     affiliation   : "4"
@@
 In plants, life histories are especially diverse, with herbaceous species that complete their life cycle in a number of weeks to trees that live for thousands of years [@brown1996oldlist].
-Along this continuum an important division exists, distinguishing annuals (i.e.
-monocarpic or semelparous) which complete their seed to seed life cycle within a single calendar year from perennials (i.e.
-polycarpic or iteroparous) which can persist over multiple years.
+Along this continuum an important division exists, distinguishing annuals which complete their seed to seed life cycle within a single calendar year from perennials which can persist over multiple years.
 Annual plants flower once, set seed, senesce, and then die, spending at least some portion of the year as a seed, where they are relatively protected from environmental stress.
@@
 Conversely, environments with less frequent drought may select for perennial species, which may benefit from multiple bouts of reproduction and competitive advantage by preventing recruitment of annual species [@corbin2004competition].
-These predictions have been supported by the association of annualilty with arid environments in wild rice [@morishima1984differentiation] and _Oenothera_ [@evans2005climate].
+These predictions have been supported by the association of annualilty with arid environments in _Oryza perennis_ [@morishima1984differentiation] and _Oenothera_ [@evans2005climate].
 Additionally, annual and perennial species of _Nemesia_ were qualitatively associated with winter rather and summer rainfall environments respectively [@datson2008climate] and annual species of _Scorzoneroides_ were associated with environments classified as unpredictable [@cruz2009molecular].
 However, whether the history frequency of drought events indeed predicts the distributions annual or perennial life history strategies has yet to be tested in a phylogenetic context with suitable sample sizes of distrubution and climate data.
-Here we combine a long-term global dataset of satellite detected drought events with metadata from natural collections to test these classic hypotheses about the evolution of life history strategies within the African endemic mustard genus, _Heliophila_ L.
+Here we combine a long-term global dataset of satellite detected drought events with metadata from natural history collections to test these classic hypotheses about the evolution of life history strategies within the African endemic mustard genus, _Heliophila_ L.
 (Brassicaceae).
 If annuallity is an adaptive strategy allowing plants to escape drought prone seasons, then drought frequency should predict the distribution of life history strategies across landscapes, and annual species should be more commonly associated with drought prone regions than perennial species.
-Furthermore, if annual species have adapted to escape drought prone seasons as seeds, observations of annual species should be rare during drought prone seasons.
+Furthermore, if annual species have adapted to escape drought prone seasons, observations of growing annual species (i.e.
+occurring in forms other than seed) should be rare during drought prone seasons.
 Phylogenetic relatedness can have significant non-random effects on species distributions and life history traits [@barrett1996comparative], and therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetically controlled background.
@@
 ### Availability
-All analyses were peformed using R. All data and the source code to produce this manuscript are available at https://github.com/greymonroe/heliophila.
+All analyses were peformed using R. All data and the source code to produce this manuscript are available at https://github.com/greymonroe/heliophila. Softare used is listed in the supplement.
 
@@
 To quantify the frequency of drought during different seasons across landscapes, we used the remotely sensed Vegetative Health Index (VHI), which measures landscape scale reductions in plant cover and temperature conditions characteristic of drought [@kogan2001operational].
-Generated from data collected by NOAA AVHRR satellites since 1981, the VHI combines normalized difference vegetation index (NDVI) derived measures of vegetative stress (Vegetative Condition Index - VCI) with temperature stress indicated by anomalies in thermal spectra (Temperature Condition Index - TCI).
+Generated from data collected by NOAA AVHRR satellites since 1981, the VHI combines Normalized Difference Vegetation Index (NDVI) derived measures of vegetative stress (Vegetative Condition Index - VCI) with temperature stress indicated by anomalies in thermal spectra (Temperature Condition Index - TCI).
 The VHI of year $y$ during week $w$ of $[1,52]$ at pixel $i$ is derived from the following equations, where $n$ is the number of years observed.
@@
 
-where $NDVI_{min} = min(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})$ and $NDVI_{max} = max(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})$
-
 $$TCI_{y,w,i} = 100\frac{T_{y,w,i} - T_{min}}{T_{max} - T_{min}}$$
 
-where $T_{min} = min(T_{1981,w,i}...T_{1981+n,w,i})$ and $T_{max} = max(T_{1981,w,i}...T_{1981+n,w,i})$
+$$VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})$$
+where $NDVI_{min} = min(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})$ and $NDVI_{max} = max(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})$ and $T_{min} = min(T_{1981,w,i}...T_{1981+n,w,i})$ and $T_{max} = max(T_{1981,w,i}...T_{1981+n,w,i})$
 
-$$VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})$$
 Thus, VHI measurements are standardized according to conditions historically observed at each locations.
@@
 _Heliophila_ is a genus of flowering plants endemic to the southern portion of Africa including the Cape Floristic and Succulent Karoo Regions.
-These are among the most botanically diverse environments on Earth and the estimated ~50 _Heliophila_ species are considered to be the most diverse genus of the family Brassicaceae [@mummenhoff2005phylogeny].
+These are among the most botanically diverse environments on Earth and the estimated ~50 _Heliophila_ species occurring there are considered to make up the most diverse genus of the family Brassicaceae [@mummenhoff2005phylogeny;@mandakova2012whole].
 This genus includes both perennial and annual species and this change in life history strategy has likely arisen multiple independent times [@appel1997generic;@mummenhoff2005phylogeny].
@@
 
-* Records for species with reported life history  
-* Records with geospatial data  
-* Records without known geospatial issues (i.e., coordinates reported are those of herbarium)  
-* Records from collection sites classified as land pixels   
-* Records from Africa  
-* Records without duplicates (i.e., identical species, location, collection date)
+* records for species with reported life history  
+* records with geospatial data  
+* records without known geospatial coodinate issues (i.e., coordinates reported are those of herbarium)  
+* records from collection sites classified as land pixels   
+* records from Africa (to exclude locations of cultivation)
+* records without duplicates (i.e., identical species, location, collection date)
 
@@
 There were clear visual differences between the distributions of the `r sum(GBIF_cleaned$life_history=="a")` annual and the `r sum(GBIF_cleaned$life_history=="p")` perennial _Heliophila_ observation records (see Figure \@ref(fig:speciesmaps) for maps of individual species).
-While annual species were generally found in the western regions of South Africa and the Namibia, primarily in the Cape Floristic Region and Succulent Karoo (Figure \@ref(fig:mapsboxplots)a), the occurances of perennials extended to the east coast of South Africa (Figure \@ref(fig:mapsboxplots)b).
+While annual species were generally found in the western regions of South Africa and Namibia, primarily in the Cape Floristic Region and Succulent Karoo (Figure \@ref(fig:mapsboxplots)a), the occurances of perennials extended to the east coast of South Africa (Figure \@ref(fig:mapsboxplots)b).
 
@@
 
-Theory has predicted that annuality should be adaptive in places where stresses such as drought are more common.
+Theory predicts that annuality should be adaptive in places where stresses such as drought are more common.
 Conversely, perenniality should be adaptive in places where such stresses are less frequent.
@@
 In terms of raw observation records
-(Figure @\ref(fig:mapsboxplots)g), the frequency of drought was signficantly higher at the location of annuals during the winter (t = `r twin$statistic`, p = `r twin$p.value`), spring (t = `r tspr$statistic`, p = `r tspr$p.value`), summer (t = `r tsum$statistic`, p = `r tsum$p.value`), and fall (t = `r tfal$statistic`, p = `r tfal$p.value`).
-Because raw occurance records do not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable))), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials by a Firths-penalized logistic regression.
+(Figure \@ref(fig:mapsboxplots)g), the frequency of drought was signficantly higher at the location of annuals during the winter (t = `r twin$statistic`, p = `r twin$p.value`), spring (t = `r tspr$statistic`, p = `r tspr$p.value`), summer (t = `r tsum$statistic`, p = `r tsum$p.value`), and fall (t = `r tfal$statistic`, p = `r tfal$p.value`).
+Because raw occurance records do not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable)), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials by a Firths-penalized logistic regression.
 We found that the mean drought frequencies were signficantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a).
@@
 
-(ref:lineplots) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of Heliophila.
-(B) Collection dates of GBIF records of annual and perennial species of Heliophila.
+(ref:lineplots) (a) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of Heliophila.
+(b) Collection dates of GBIF records of annual and perennial species of Heliophila.
 
@@
   scale_x_continuous(name="Season", breaks=c(45,135,225, 315), labels=c("Winter","Spring","Summer","Fall"))+
-  labs(title="Collection date\nof herbaria specimens", y="Density")+
-  theme(plot.title = element_text(hjust = 0.5))
+  labs(title="Date of observation for occurance records", y="Density")+
+  theme(plot.title = element_text(hjust = 0.5, size=4))
 
@@
 The distributions reveal a considerable difference in the timing of observation of these two life histories.
-In comparison to perennials, which appear to be collected throughout the year, annuals are almost exclusively observed during the spring (Figure \@ref(fig:lineplots)b).
+In comparison to perennials, which appear to be collected throughout the year, annuals are almost exclusively observed during the winter and spring (Figure \@ref(fig:lineplots)b).
 The differences between the distrubution of collection dates were signicant by all tests (ks.test D = `r  kout$statistic`, p = `r  kout$p.value`; bartlett.test K2 = `r bout$statistic`, p = `r bout$p.value`) This is consistent with a model of life history in which annual species flower in the spring, set seed, senesce, and die before the summer.
@@
 Using metadata of `r nrow(GBIF_cleaned)` occurance records and a 34 year dataset of satellite-detected droughts, we tested the prediction that annual species are more often observed in drought-prone locations than perennial species, when controlling for phylogenetic relatedness.
-We found that drought frequency is significantly different between the distributions of annual and perennial species, with annuals being found in environments with more frequent drought, and that this signal is strongest during the summer.
-These results remain significant while controlling for the phylogenetic relationships of _Heliophila_ species, yielding support for the role that natural selection has played in driving contemporary distributions of these alternatives strategies in relation to drought regimes.
+We found that drought frequency is significantly different between the distributions of annual and perennial species, with annuals being found in environments with more frequent drought, and that this signal is strongest during the seasons when annuals are likely escaping via seed dormancy These results remain significant while controlling for the phylogenetic relationships of _Heliophila_ species, yielding support for the role that natural selection has played in driving contemporary distributions of these alternatives strategies in relation to drought regimens.
 
@@
 While most previous work has focused on describing the evolutionary origins of annuality [@barrett1996comparative;@conti1999phylogenetic;@andreasen2001unequal;@verboom2004testing;@friedman2015all] there are at least a few other cases where perenniality appears to have arisen from an annual acestor [@bena1998molecular;@tank2008annuals].
-And while early theory predicted selection for annuality when adult morality is high [@stearns1992evolution],  we also find evidence that the transition to perenniality could be explained by historical drought regimes.
-The phylogeny reveals several transitions from annual to perennial life history (Figure \@ref(fig:phylogeny)a) and that perennials distributions extend into regions where drought frequency is low (Figure \@ref(fig:mapsboxplots)b,Figure \@ref(fig:speciesmaps)).
+And while early theory predicted selection for annuality when adult morality is high [@stearns1992evolution],  we also find evidence that the transition to perenniality could be explained by historical drought regimens.
+The phylogeny reveals several transitions from annual to perennial life history (Figure \@ref(fig:phylogeny)a) and that the distributions of perennial _Heliophila_ extend into regions where drought frequency is low (Figure \@ref(fig:mapsboxplots)b,Figure \@ref(fig:speciesmaps)).
 Perennials may be able to out complete annual relatives in environments where the infrequency of drought favors strategies that allow plants to benefit from growth over many seasons.
 It may also indicate that annuals rely on drought as a source of disturbance for seedling recruitment when competing with perennials [@corbin2004competition].
-Indeed, no annual species were observed in the low drought regions of east South Africa \@ref(fig:mapsboxplots)a,Figure \@ref(fig:speciesmaps)).
+Indeed, no annual species were observed in the low drought regions of eastern South Africa \@ref(fig:mapsboxplots)a,Figure \@ref(fig:speciesmaps)).
 
-These findings suggest that species with locally adapted life history strategies could be threatened by rapidly changing drought regimes [@dai2011drought].
-This could have impacts on ecosystem functioning if the composition of annual and perennial species changes as a response [@garnier1997specific; @roumet2006suites].
+These findings suggest that species with locally adaptive life history strategies could be threatened by rapidly changing drought regimens [@dai2011drought].
+This could have impacts on ecosystem functioning and processes such as carbon cycling if the composition of annual and perennial species changes as a response [@garnier1997specific; @roumet2006suites;monroe2018ecoevolutionary].
 Furthermore, the frequency of drought may be an important factor when considering the use of perennial cropping sysyems [@parry2005prospects;@lelievre2009current].
 
-In conclusion, we find strong support for classic life history theory which predicts that annuality is adaptive environments where droughts occur more frequently.
+In conclusion, we find strong support for classic life history theory which predicts that annuality is adaptive in environments where droughts occur more frequently.
 Additionally, we report evidence consistent with a life history model in annuals in which they escape drought prone seasons during the seed phase of their life cycle.
 Finally, we find evidence that the distributions of perennial lineages may indicate a competitive advantage in areas where droughts are infrequent.
-More broadly, this work demonstrates the power of emerging data to address outstanding classic hypotheses in ecology and evolution.
+More broadly, this work highlights the irreplaceable value of natural history collections and demonstrates the power of combining such information with large scale remote sensing data to address outstanding classic hypotheses in ecology and evolution.
 
 # Acknowledgments 
-We thank __ for generous feedback that improved the quality of this work. This research was supported by NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
+We thank Jesse Lasky and members of the Sloan lab for generous feedback that improved the quality of this work.
+This research was supported by NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
 

### 2018-11-28 manuscript/manuscript.tex
edits from Kathryn

@@
 \hypersetup{unicode=true,
-            pdftitle={Satellite-detected droughts predict plant life history},
-            pdfauthor={J. Grey Monroe, Brian Gill, Kathryn Turner, \& John K. McKay},
+            pdftitle={Drought frequency predicts plant life history strategies},
+            pdfauthor={J. Grey Monroe, Brian Gill, Kathryn G. Turner, \& John K. McKay},
             pdfkeywords={drought adaptation, life history evolution, remote sensing,
@@
 
-  \title{Satellite-detected droughts predict plant life history}
+  \title{Drought frequency predicts plant life history strategies}
     \author{J. Grey Monroe\textsuperscript{1,2}, Brian Gill\textsuperscript{3},
-Kathryn Turner\textsuperscript{4}, \& John K. McKay\textsuperscript{2}}
+Kathryn G. Turner\textsuperscript{4}, \& John K.
+McKay\textsuperscript{2}}
     \date{}
@@
 for thousands of years (Brown, 1996). Along this continuum an important
-division exists, distinguishing annuals (i.e.~monocarpic or semelparous)
-which complete their seed to seed life cycle within a single calendar
-year from perennials (i.e.~polycarpic or iteroparous) which can persist
-over multiple years. Annual plants flower once, set seed, senesce, and
-then die, spending at least some portion of the year as a seed, where
-they are relatively protected from environmental stress. In contrast,
-perennial plants can continue vegetative growth after reproduction and
-must survive conditions experienced during all seasons. These represent
-fundamentally different life history strategies, but the ecological
-factors that explain their evolution and distributions remain
+division exists, distinguishing annuals which complete their seed to
+seed life cycle within a single calendar year from perennials which can
+persist over multiple years. Annual plants flower once, set seed,
+senesce, and then die, spending at least some portion of the year as a
+seed, where they are relatively protected from environmental stress. In
+contrast, perennial plants can continue vegetative growth after
+reproduction and must survive conditions experienced during all seasons.
+These represent fundamentally different life history strategies, but the
+ecological factors that explain their evolution and distributions remain
 empirically uresolved (Friedman \& Rubin, 2015).\\
@@
 hypothessis that annuality is adaptive when it allows plants to escape
-drought (Schaffer \& M, 1975). Lack of water is perhaps the greatest
-threat to survival during vegetatitive or reproductive growth and
-annuals can remain dormant (and protected as a seed) during drought.
+drought (Schaffer \& Gadgil, 1975). Lack of water is perhaps the
+greatest threat to survival during vegetatitive or reproductive growth
+and annuals can remain dormant (and protected as a seed) during drought.
 Thus, environments with greater seasonal drought frequency may select
@@
 supported by the association of annualilty with arid environments in
-wild rice (Morishima, Sano, \& Oka, 1984) and \emph{Oenothera} (Evans,
-Hearn, Hahn, Spangle, \& Venable, 2005). Additionally, annual and
-perennial species of \emph{Nemesia} were qualitatively associated with
-winter rather and summer rainfall environments respectively (Datson,
-Murray, \& Steiner, 2008) and annual species of \emph{Scorzoneroides}
-were associated with environments classified as unpredictable
-(Cruz-Mazo, Buide, Samuel, \& Narbona, 2009). However, whether the
-history frequency of drought events indeed predicts the distributions
-annual or perennial life history strategies has yet to be tested in a
-phylogenetic context with suitable sample sizes of distrubution and
-climate data.\\
+\emph{Oryza perennis} (Morishima, Sano, \& Oka, 1984) and
+\emph{Oenothera} (Evans, Hearn, Hahn, Spangle, \& Venable, 2005).
+Additionally, annual and perennial species of \emph{Nemesia} were
+qualitatively associated with winter rather and summer rainfall
+environments respectively (Datson, Murray, \& Steiner, 2008) and annual
+species of \emph{Scorzoneroides} were associated with environments
+classified as unpredictable (Cruz-Mazo, Buide, Samuel, \& Narbona,
+2009). However, whether the history frequency of drought events indeed
+predicts the distributions annual or perennial life history strategies
+has yet to be tested in a phylogenetic context with suitable sample
+sizes of distrubution and climate data.\\
 Here we combine a long-term global dataset of satellite detected drought
-events with metadata from natural collections to test these classic
-hypotheses about the evolution of life history strategies within the
-African endemic mustard genus, \emph{Heliophila} L. (Brassicaceae). If
-annuallity is an adaptive strategy allowing plants to escape drought
+events with metadata from natural history collections to test these
+classic hypotheses about the evolution of life history strategies within
+the African endemic mustard genus, \emph{Heliophila} L. (Brassicaceae).
+If annuallity is an adaptive strategy allowing plants to escape drought
 prone seasons, then drought frequency should predict the distribution of
@@
 species. Furthermore, if annual species have adapted to escape drought
-prone seasons as seeds, observations of annual species should be rare
-during drought prone seasons. Phylogenetic relatedness can have
-significant non-random effects on species distributions and life history
-traits (Barrett, Harder, \& Worley, 1996), and therefore we assessed the
-relationship between life history distribution and drought frequency in
-a phylogenetically controlled background.
+prone seasons, observations of growing annual species (i.e.~occurring in
+forms other than seed) should be rare during drought prone seasons.
+Phylogenetic relatedness can have significant non-random effects on
+species distributions and life history traits (Barrett, Harder, \&
+Worley, 1996), and therefore we assessed the relationship between life
+history distribution and drought frequency in a phylogenetically
+controlled background.
 
@@
 produce this manuscript are available at
-\url{https://github.com/greymonroe/heliophila}.
+\url{https://github.com/greymonroe/heliophila}. Softare used is listed
+in the supplement.
 
@@
 (Kogan, 2001). Generated from data collected by NOAA AVHRR satellites
-since 1981, the VHI combines normalized difference vegetation index
+since 1981, the VHI combines Normalized Difference Vegetation Index
 (NDVI) derived measures of vegetative stress (Vegetative Condition Index
@@
 
-where \(NDVI_{min} = min(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})\) and
-\(NDVI_{max} = max(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})\)
-
 \[TCI_{y,w,i} = 100\frac{T_{y,w,i} - T_{min}}{T_{max} - T_{min}}\]
 
-where \(T_{min} = min(T_{1981,w,i}...T_{1981+n,w,i})\) and
+\[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\] where
+\(NDVI_{min} = min(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})\) and
+\(NDVI_{max} = max(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})\) and
+\(T_{min} = min(T_{1981,w,i}...T_{1981+n,w,i})\) and
 \(T_{max} = max(T_{1981,w,i}...T_{1981+n,w,i})\)
 
-\[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\] Thus, VHI
-measurements are standardized according to conditions historically
-observed at each locations. These measurements have been validated and
-generally used for evaluating drought risk and predicting crop yields in
-agriculture (e.g., Rojas, Vrieling, \& Rembold, 2011; Kogan et al.,
-2016). But they also present a new tool to study seasonal patterns in
-the frequency of drought across environments and to test hypotheses
-about the effect of drought on ecological and evolutionary processes
-(Kerr \& Ostrovsky, 2003). As such, the VHI has been applied recently to
-study drought related ecology of natural species and proven useful for
-predicting infraspecific variation in drought tolerance traits and genes
-(Dittberner et al., 2018; Mojica et al., 2016; Monroe et al., 2018).
-Here, we accessed VHI data at \(16km^2\) resolution from 1981 to 2015
+Thus, VHI measurements are standardized according to conditions
+historically observed at each locations. These measurements have been
+validated and generally used for evaluating drought risk and predicting
+crop yields in agriculture (e.g., Rojas, Vrieling, \& Rembold, 2011;
+Kogan et al., 2016). But they also present a new tool to study seasonal
+patterns in the frequency of drought across environments and to test
+hypotheses about the effect of drought on ecological and evolutionary
+processes (Kerr \& Ostrovsky, 2003). As such, the VHI has been applied
+recently to study drought related ecology of natural species and proven
+useful for predicting infraspecific variation in drought tolerance
+traits and genes (Dittberner et al., 2018; Mojica et al., 2016; Monroe
+et al., 2018). Here, we accessed VHI data at \(16km^2\) resolution from
+1981 to 2015
 (\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php}) to
@@
 Earth and the estimated \textasciitilde{}50 \emph{Heliophila} species
-are considered to be the most diverse genus of the family Brassicaceae
-(Mummenhoff, Al-Shehbaz, Bakker, Linder, \& Mühlhausen, 2005). This
-genus includes both perennial and annual species and this change in life
-history strategy has likely arisen multiple independent times (Appel \&
-Al-Shehbaz, 1997; Mummenhoff et al., 2005). Furthermore, the fine scale
-climatic heterogeneity of Southern Africa is ideal for studying the
-distribution of traits in relation to environmental parameters (Sayre et
-al., 2013). We used life histories reported by Mummenhoff et al. (2005),
-grouping species with annual or perennial life histories. Perenniality
-was defined based any form of perennial life history (e.g., herbs,
-shrubs, mixed, etc). We also repeated these analyses without \emph{H.
-subulata} and \emph{H. minima} which have mixed life histories and found
-similar results. Because the original authors didn't report the nature
-of the mixed trait (Mummenhoff et al., 2005), (ie. plasticity
-vs.~genetic variation), we classified these species here as perennial
-since they have the capacity to maintain vegetative growth after
-reproduction at least to some capacity.
+occurring there are considered to make up the most diverse genus of the
+family Brassicaceae (Mandáková et al., 2012; Mummenhoff, Al-Shehbaz,
+Bakker, Linder, \& Mühlhausen, 2005). This genus includes both perennial
+and annual species and this change in life history strategy has likely
+arisen multiple independent times (Appel \& Al-Shehbaz, 1997; Mummenhoff
+et al., 2005). Furthermore, the fine scale climatic heterogeneity of
+Southern Africa is ideal for studying the distribution of traits in
+relation to environmental parameters (Sayre et al., 2013). We used life
+histories reported by Mummenhoff et al. (2005), grouping species with
+annual or perennial life histories. Perenniality was defined based any
+form of perennial life history (e.g., herbs, shrubs, mixed, etc). We
+also repeated these analyses without \emph{H. subulata} and \emph{H.
+minima} which have mixed life histories and found similar results.
+Because the original authors didn't report the nature of the mixed trait
+(Mummenhoff et al., 2005), (ie. plasticity vs.~genetic variation), we
+classified these species here as perennial since they have the capacity
+to maintain vegetative growth after reproduction at least to some
+capacity.
 
@@
 genus \emph{Heliophila} were downloaded from the Global Biodiversity
-Information Facility (gbif.org) on July 21, 2018 (``GBIF occurrence
-download,'' 2018).
+Information Facility (gbif.org) on July 21, 2018 (GBIF, 2018).
 
@@
 \item
-  Records for species with reported life history\\
+  records for species with reported life history\\
 \item
-  Records with geospatial data\\
+  records with geospatial data\\
 \item
-  Records without known geospatial issues (i.e., coordinates reported
-  are those of herbarium)\\
+  records without known geospatial coodinate issues (i.e., coordinates
+  reported are those of herbarium)\\
 \item
-  Records from collection sites classified as land pixels\\
+  records from collection sites classified as land pixels\\
 \item
-  Records from Africa\\
+  records from Africa (to exclude locations of cultivation)
 \item
-  Records without duplicates (i.e., identical species, location,
+  records without duplicates (i.e., identical species, location,
   collection date)
@@
 annual species were generally found in the western regions of South
-Africa and the Namibia, primarily in the Cape Floristic Region and
-Succulent Karoo (Figure \ref{fig:mapsboxplots}a), the occurances of
-perennials extended to the east coast of South Africa (Figure
+Africa and Namibia, primarily in the Cape Floristic Region and Succulent
+Karoo (Figure \ref{fig:mapsboxplots}a), the occurances of perennials
+extended to the east coast of South Africa (Figure
 \ref{fig:mapsboxplots}b).
@@
 
-Theory has predicted that annuality should be adaptive in places where
+Theory predicts that annuality should be adaptive in places where
 stresses such as drought are more common. Conversely, perenniality
@@
 locations of occurance records for annual species. In terms of raw
-observation records (Figure @\ref(fig:mapsboxplots)g), the frequency of
+observation records (Figure \ref{fig:mapsboxplots}g), the frequency of
 drought was signficantly higher at the location of annuals during the
@@
 records do not account for variation in the number of records per
-species (Table \ref{tab:speciesmeanstable})), we also tested whether
-mean drought frequency values of each species were significantly
-different between annuals and perennials by a Firths-penalized logistic
+species (Table \ref{tab:speciesmeanstable}), we also tested whether mean
+drought frequency values of each species were significantly different
+between annuals and perennials by a Firths-penalized logistic
 regression. We found that the mean drought frequencies were signficantly
@@
 \includegraphics{../figures/line_and_dates.pdf}
-\caption{\label{fig:lineplots}Comparison (mean +- SE) of drought frequency across
+\caption{\label{fig:lineplots}(a) Comparison (mean +- SE) of drought frequency across
 seasons measured at the GBIF records of annual and perennial species of
-Heliophila. (B) Collection dates of GBIF records of annual and perennial
+Heliophila. (b) Collection dates of GBIF records of annual and perennial
 species of Heliophila.}
@@
 which appear to be collected throughout the year, annuals are almost
-exclusively observed during the spring (Figure \ref{fig:lineplots}b).
-The differences between the distrubution of collection dates were
-signicant by all tests (ks.test D = 0.25, p = 0; bartlett.test K2 =
-503.18, p = 0.00) This is consistent with a model of life history in
-which annual species flower in the spring, set seed, senesce, and die
-before the summer. Thus, these annual species are likely to remain
-dormant during the summer and fall, when drought is the strongest
-predictor of the distrubutions of annual and perennial life histories
-(Figure \ref{fig:lineplots}a).
+exclusively observed during the winter and spring (Figure
+\ref{fig:lineplots}b). The differences between the distrubution of
+collection dates were signicant by all tests (ks.test D = 0.25, p = 0;
+bartlett.test K2 = 503.18, p = 0.00) This is consistent with a model of
+life history in which annual species flower in the spring, set seed,
+senesce, and die before the summer. Thus, these annual species are
+likely to remain dormant during the summer and fall, when drought is the
+strongest predictor of the distrubutions of annual and perennial life
+histories (Figure \ref{fig:lineplots}a).
 
@@
 in environments with more frequent drought, and that this signal is
-strongest during the summer. These results remain significant while
-controlling for the phylogenetic relationships of \emph{Heliophila}
-species, yielding support for the role that natural selection has played
-in driving contemporary distributions of these alternatives strategies
-in relation to drought regimes.
+strongest during the seasons when annuals are likely escaping via seed
+dormancy These results remain significant while controlling for the
+phylogenetic relationships of \emph{Heliophila} species, yielding
+support for the role that natural selection has played in driving
+contemporary distributions of these alternatives strategies in relation
+to drought regimens.
 
@@
 transition to perenniality could be explained by historical drought
-regimes. The phylogeny reveals several transitions from annual to
-perennial life history (Figure \ref{fig:phylogeny}a) and that perennials
-distributions extend into regions where drought frequency is low (Figure
-\ref{fig:mapsboxplots}b,Figure \ref{fig:speciesmaps}). Perennials may be
-able to out complete annual relatives in environments where the
-infrequency of drought favors strategies that allow plants to benefit
-from growth over many seasons. It may also indicate that annuals rely on
-drought as a source of disturbance for seedling recruitment when
-competing with perennials (Corbin \& D'Antonio, 2004). Indeed, no annual
-species were observed in the low drought regions of east South Africa
+regimens. The phylogeny reveals several transitions from annual to
+perennial life history (Figure \ref{fig:phylogeny}a) and that the
+distributions of perennial \emph{Heliophila} extend into regions where
+drought frequency is low (Figure \ref{fig:mapsboxplots}b,Figure
+\ref{fig:speciesmaps}). Perennials may be able to out complete annual
+relatives in environments where the infrequency of drought favors
+strategies that allow plants to benefit from growth over many seasons.
+It may also indicate that annuals rely on drought as a source of
+disturbance for seedling recruitment when competing with perennials
+(Corbin \& D'Antonio, 2004). Indeed, no annual species were observed in
+the low drought regions of eastern South Africa
 \ref{fig:mapsboxplots}a,Figure \ref{fig:speciesmaps}).
 
-These findings suggest that species with locally adapted life history
-strategies could be threatened by rapidly changing drought regimes (Dai,
-2011). This could have impacts on ecosystem functioning if the
-composition of annual and perennial species changes as a response
-(Garnier, Cordonnier, Guillerm, \& Sonié, 1997; Roumet, Urcelay, \&
-Dı'az, 2006). Furthermore, the frequency of drought may be an important
-factor when considering the use of perennial cropping sysyems (Lelièvre
-\& Volaire, 2009; Parry, Flexas, \& Medrano, 2005).
+These findings suggest that species with locally adaptive life history
+strategies could be threatened by rapidly changing drought regimens
+(Dai, 2011). This could have impacts on ecosystem functioning and
+processes such as carbon cycling if the composition of annual and
+perennial species changes as a response {[}@garnier1997specific; Roumet,
+Urcelay, and Dı'az (2006);monroe2018ecoevolutionary{]}. Furthermore, the
+frequency of drought may be an important factor when considering the use
+of perennial cropping sysyems (Lelièvre \& Volaire, 2009; Parry, Flexas,
+\& Medrano, 2005).
 
 In conclusion, we find strong support for classic life history theory
-which predicts that annuality is adaptive environments where droughts
+which predicts that annuality is adaptive in environments where droughts
 occur more frequently. Additionally, we report evidence consistent with
@@
 competitive advantage in areas where droughts are infrequent. More
-broadly, this work demonstrates the power of emerging data to address
-outstanding classic hypotheses in ecology and evolution.
+broadly, this work highlights the irreplaceable value of natural history
+collections and demonstrates the power of combining such information
+with large scale remote sensing data to address outstanding classic
+hypotheses in ecology and evolution.
 
@@
 
-We thank \_\_ for generous feedback that improved the quality of this
-work. This research was supported by NSF Award 1701918 and USDA-NIFA
-Award 2014-38420-21801 to JGM.
+We thank Jesse Lasky and members of the Sloan lab for generous feedback
+that improved the quality of this work. This research was supported by
+NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
 
@@
 
-\leavevmode\hypertarget{ref-garnier1997specific}{}%
-Garnier, E., Cordonnier, P., Guillerm, J.-L., \& Sonié, L. (1997).
-Specific leaf area and leaf nitrogen concentration in annual and
-perennial grass species growing in mediterranean old-fields.
-\emph{Oecologia}, \emph{111}(4), 490--498.
-
 \leavevmode\hypertarget{ref-gbifdownload}{}%
-GBIF occurrence download. (2018, July). Retrieved from
+GBIF. (2018, July). GBIF occurrence download. Retrieved from
 \url{https://doi.org/10.15468/dl.xmwqxm}
@@
 \leavevmode\hypertarget{ref-schaffer1975selection}{}%
-Schaffer, W., \& M, G. (1975). Selection for optimal life histories in
-plants. \emph{Ecology and Evolution of Communities.}, 142--157.
+Schaffer, W., \& Gadgil, M. (1975). Selection for optimal life histories
+in plants. \emph{Ecology and Evolution of Communities.}, 142--157.
 

### 2018-11-28 manuscript/manuscript.Rmd
change csl

@@
 bibliography      : ["r-references.bib"]
+csl               : mycsl.csl
 
@@
 Perenniality was defined based any form of perennial life history (e.g., herbs, shrubs, mixed, etc).
-We also repeated these analyses without _H.
-subulata_ and _H.
-minima_ which have mixed life histories and found similar results.
-Because the original authors didn't report the nature of the mixed trait [@mummenhoff2005phylogeny], (ie.
+Because the nature of species reported with mixed traits were unknown (ie.
 plasticity vs.
-genetic variation), we classified these species here as perennial since they have the capacity to maintain vegetative growth after reproduction at least to some capacity.
+genetic variation), we classified these species here as perennial since they can maintain vegetative growth after reproduction at least to some capacity.
 

### 2018-11-28 manuscript/manuscript.tex
change csl

@@
 ecological factors that explain their evolution and distributions remain
-empirically uresolved (Friedman \& Rubin, 2015).\\
+empirically uresolved (Friedman and Rubin, 2015).\\
 Classical theory predict shorter life spans in environments where adult
-mortality is high (Charnov \& Schaffer, 1973; Franco \& Silvertown,
+mortality is high (Charnov and Schaffer, 1973; Franco and Silvertown,
 1996; Stearns, 1992). In plants, this has been extended to the
 hypothessis that annuality is adaptive when it allows plants to escape
-drought (Schaffer \& Gadgil, 1975). Lack of water is perhaps the
+drought (Schaffer and Gadgil, 1975). Lack of water is perhaps the
 greatest threat to survival during vegetatitive or reproductive growth
@@
 reproduction and competitive advantage by preventing recruitment of
-annual species (Corbin \& D'Antonio, 2004). These predictions have been
+annual species (Corbin and D'Antonio, 2004). These predictions have been
 supported by the association of annualilty with arid environments in
-\emph{Oryza perennis} (Morishima, Sano, \& Oka, 1984) and
-\emph{Oenothera} (Evans, Hearn, Hahn, Spangle, \& Venable, 2005).
-Additionally, annual and perennial species of \emph{Nemesia} were
-qualitatively associated with winter rather and summer rainfall
-environments respectively (Datson, Murray, \& Steiner, 2008) and annual
-species of \emph{Scorzoneroides} were associated with environments
-classified as unpredictable (Cruz-Mazo, Buide, Samuel, \& Narbona,
-2009). However, whether the history frequency of drought events indeed
-predicts the distributions annual or perennial life history strategies
-has yet to be tested in a phylogenetic context with suitable sample
-sizes of distrubution and climate data.\\
+\emph{Oryza perennis} (Morishima et al., 1984) and \emph{Oenothera}
+(Evans et al., 2005). Additionally, annual and perennial species of
+\emph{Nemesia} were qualitatively associated with winter rather and
+summer rainfall environments respectively (Datson et al., 2008) and
+annual species of \emph{Scorzoneroides} were associated with
+environments classified as unpredictable (Cruz-Mazo et al., 2009).
+However, whether the history frequency of drought events indeed predicts
+the distributions annual or perennial life history strategies has yet to
+be tested in a phylogenetic context with suitable sample sizes of
+distrubution and climate data.\\
 Here we combine a long-term global dataset of satellite detected drought
@@
 Phylogenetic relatedness can have significant non-random effects on
-species distributions and life history traits (Barrett, Harder, \&
-Worley, 1996), and therefore we assessed the relationship between life
-history distribution and drought frequency in a phylogenetically
-controlled background.
+species distributions and life history traits (Barrett et al., 1996),
+and therefore we assessed the relationship between life history
+distribution and drought frequency in a phylogenetically controlled
+background.
 
@@
 validated and generally used for evaluating drought risk and predicting
-crop yields in agriculture (e.g., Rojas, Vrieling, \& Rembold, 2011;
-Kogan et al., 2016). But they also present a new tool to study seasonal
-patterns in the frequency of drought across environments and to test
-hypotheses about the effect of drought on ecological and evolutionary
-processes (Kerr \& Ostrovsky, 2003). As such, the VHI has been applied
-recently to study drought related ecology of natural species and proven
-useful for predicting infraspecific variation in drought tolerance
-traits and genes (Dittberner et al., 2018; Mojica et al., 2016; Monroe
-et al., 2018). Here, we accessed VHI data at \(16km^2\) resolution from
-1981 to 2015
-(\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php}) to
-characterize the seasonal drought frequencies experienced by annual and
-perennial \emph{Heliophila} species.
+crop yields in agriculture (e.g., Rojas et al., 2011; Kogan et al.,
+2016). But they also present a new tool to study seasonal patterns in
+the frequency of drought across environments and to test hypotheses
+about the effect of drought on ecological and evolutionary processes
+(Kerr and Ostrovsky, 2003). As such, the VHI has been applied recently
+to study drought related ecology of natural species and proven useful
+for predicting infraspecific variation in drought tolerance traits and
+genes (Dittberner et al., 2018; Mojica et al., 2016; Monroe et al.,
+2018). Here, we accessed VHI data at \(16km^2\) resolution from 1981 to
+2015 (\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php})
+to characterize the seasonal drought frequencies experienced by annual
+and perennial \emph{Heliophila} species.
 
@@
 occurring there are considered to make up the most diverse genus of the
-family Brassicaceae (Mandáková et al., 2012; Mummenhoff, Al-Shehbaz,
-Bakker, Linder, \& Mühlhausen, 2005). This genus includes both perennial
-and annual species and this change in life history strategy has likely
-arisen multiple independent times (Appel \& Al-Shehbaz, 1997; Mummenhoff
-et al., 2005). Furthermore, the fine scale climatic heterogeneity of
-Southern Africa is ideal for studying the distribution of traits in
-relation to environmental parameters (Sayre et al., 2013). We used life
-histories reported by Mummenhoff et al. (2005), grouping species with
-annual or perennial life histories. Perenniality was defined based any
-form of perennial life history (e.g., herbs, shrubs, mixed, etc). We
-also repeated these analyses without \emph{H. subulata} and \emph{H.
-minima} which have mixed life histories and found similar results.
-Because the original authors didn't report the nature of the mixed trait
-(Mummenhoff et al., 2005), (ie. plasticity vs.~genetic variation), we
-classified these species here as perennial since they have the capacity
-to maintain vegetative growth after reproduction at least to some
-capacity.
+family Brassicaceae (Mandáková et al., 2012; Mummenhoff et al., 2005).
+This genus includes both perennial and annual species and this change in
+life history strategy has likely arisen multiple independent times
+(Appel and Al-Shehbaz, 1997; Mummenhoff et al., 2005). Furthermore, the
+fine scale climatic heterogeneity of Southern Africa is ideal for
+studying the distribution of traits in relation to environmental
+parameters (Sayre et al., 2013). We used life histories reported by
+Mummenhoff et al. (2005), grouping species with annual or perennial life
+histories. Perenniality was defined based any form of perennial life
+history (e.g., herbs, shrubs, mixed, etc). We also repeated these
+analyses without \emph{H. subulata} and \emph{H. minima} which have
+mixed life histories and found similar results. Because the original
+authors didn't report the nature of the mixed trait (Mummenhoff et al.,
+2005), (ie. plasticity vs.~genetic variation), we classified these
+species here as perennial since they have the capacity to maintain
+vegetative growth after reproduction at least to some capacity.
 
@@
 species' geographical distributions (reviewed by Willis et al. (2017)
-and Lang, Willems, Scheepens, Burbano, and Bossdorf (2018)). And as they
-become digitized (Soltis, 2017), these collections have been used to
-study relationships between trait distributions, geography, and climate
-(Davis, Willis, Connolly, Kelly, \& Ellison, 2015; Stropp et al., 2016;
-Václavı'k, Beckmann, Cord, \& Bindewald, 2017; Wolf, Zimmerman,
-Anderegg, Busby, \& Christensen, 2016). To characterize the diributions
-of annual and perennial \emph{Heliophila} species, all records for the
-genus \emph{Heliophila} were downloaded from the Global Biodiversity
-Information Facility (gbif.org) on July 21, 2018 (GBIF, 2018).
+and Lang et al. (2018)). And as they become digitized (Soltis, 2017),
+these collections have been used to study relationships between trait
+distributions, geography, and climate (Davis et al., 2015; Stropp et
+al., 2016; Václavı'k et al., 2017; Wolf et al., 2016). To characterize
+the diributions of annual and perennial \emph{Heliophila} species, all
+records for the genus \emph{Heliophila} were downloaded from the Global
+Biodiversity Information Facility (gbif.org) on July 21, 2018 (GBIF,
+2018).
 
@@
 frequent. While most previous work has focused on describing the
-evolutionary origins of annuality (Andreasen \& Baldwin, 2001; Barrett
-et al., 1996; Conti, Soltis, Hardig, \& Schneider, 1999; Friedman \&
-Rubin, 2015; Verboom, Linder, \& Stock, 2004) there are at least a few
-other cases where perenniality appears to have arisen from an annual
-acestor (Bena, Lejeune, Prosperi, \& Olivieri, 1998; Tank \& Olmstead,
-2008). And while early theory predicted selection for annuality when
-adult morality is high (Stearns, 1992), we also find evidence that the
-transition to perenniality could be explained by historical drought
-regimens. The phylogeny reveals several transitions from annual to
-perennial life history (Figure \ref{fig:phylogeny}a) and that the
-distributions of perennial \emph{Heliophila} extend into regions where
-drought frequency is low (Figure \ref{fig:mapsboxplots}b,Figure
-\ref{fig:speciesmaps}). Perennials may be able to out complete annual
-relatives in environments where the infrequency of drought favors
-strategies that allow plants to benefit from growth over many seasons.
-It may also indicate that annuals rely on drought as a source of
-disturbance for seedling recruitment when competing with perennials
-(Corbin \& D'Antonio, 2004). Indeed, no annual species were observed in
-the low drought regions of eastern South Africa
-\ref{fig:mapsboxplots}a,Figure \ref{fig:speciesmaps}).
+evolutionary origins of annuality (Andreasen and Baldwin, 2001; Barrett
+et al., 1996; Conti et al., 1999; Friedman and Rubin, 2015; Verboom et
+al., 2004) there are at least a few other cases where perenniality
+appears to have arisen from an annual acestor (Bena et al., 1998; Tank
+and Olmstead, 2008). And while early theory predicted selection for
+annuality when adult morality is high (Stearns, 1992), we also find
+evidence that the transition to perenniality could be explained by
+historical drought regimens. The phylogeny reveals several transitions
+from annual to perennial life history (Figure \ref{fig:phylogeny}a) and
+that the distributions of perennial \emph{Heliophila} extend into
+regions where drought frequency is low (Figure
+\ref{fig:mapsboxplots}b,Figure \ref{fig:speciesmaps}). Perennials may be
+able to out complete annual relatives in environments where the
+infrequency of drought favors strategies that allow plants to benefit
+from growth over many seasons. It may also indicate that annuals rely on
+drought as a source of disturbance for seedling recruitment when
+competing with perennials (Corbin and D'Antonio, 2004). Indeed, no
+annual species were observed in the low drought regions of eastern South
+Africa \ref{fig:mapsboxplots}a,Figure \ref{fig:speciesmaps}).
 
@@
 processes such as carbon cycling if the composition of annual and
-perennial species changes as a response {[}@garnier1997specific; Roumet,
-Urcelay, and Dı'az (2006);monroe2018ecoevolutionary{]}. Furthermore, the
-frequency of drought may be an important factor when considering the use
-of perennial cropping sysyems (Lelièvre \& Volaire, 2009; Parry, Flexas,
-\& Medrano, 2005).
+perennial species changes as a response {[}@garnier1997specific; Roumet
+et al. (2006);monroe2018ecoevolutionary{]}. Furthermore, the frequency
+of drought may be an important factor when considering the use of
+perennial cropping sysyems (Lelièvre and Volaire, 2009; Parry et al.,
+2005).
 
@@
 \leavevmode\hypertarget{ref-aghakouchak2015remote}{}%
-AghaKouchak, A., Farahmand, A., Melton, F., Teixeira, J., Anderson, M.,
-Wardlow, B. D., \& Hain, C. (2015). Remote sensing of drought: Progress,
-challenges and opportunities. \emph{Reviews of Geophysics},
-\emph{53}(2), 452--480.
+AghaKouchak A, Farahmand A, Melton F, Teixeira J, Anderson M, Wardlow
+BD, Hain C. 2015. Remote sensing of drought: Progress, challenges and
+opportunities. \emph{Reviews of Geophysics} \textbf{53}:452--480.
 
 \leavevmode\hypertarget{ref-R-geiger_a}{}%
-Alfaro, M., Santini, F., Brock, C., Alamillo, H., Dornburg, A., Rabosky,
-D., \ldots{} Harmon, L. (2009). Nine exceptional radiations plus high
+Alfaro M, Santini F, Brock C, Alamillo H, Dornburg A, Rabosky D,
+Carnevale G, Harmon L. 2009. Nine exceptional radiations plus high
 turnover explain species diversity in jawed vertebrates.
 \emph{Proceedings of the National Academy of Sciences of the United
-States of America}, \emph{106}, 13410--13414.
+States of America} \textbf{106}:13410--13414.
 
 \leavevmode\hypertarget{ref-andreasen2001unequal}{}%
-Andreasen, K., \& Baldwin, B. G. (2001). Unequal evolutionary rates
-between annual and perennial lineages of checker mallows (sidalcea,
-malvaceae): Evidence from 18S--26S rDNA internal and external
-transcribed spacers. \emph{Molecular Biology and Evolution},
-\emph{18}(6), 936--944.
+Andreasen K, Baldwin BG. 2001. Unequal evolutionary rates between annual
+and perennial lineages of checker mallows (sidalcea, malvaceae):
+Evidence from 18S--26S rDNA internal and external transcribed spacers.
+\emph{Molecular Biology and Evolution} \textbf{18}:936--944.
 
 \leavevmode\hypertarget{ref-appel1997generic}{}%
-Appel, O., \& Al-Shehbaz, I. A. (1997). Generic limits and taxonomy of
-hornungia, pritzelago, and hymenolobus (brassicaceae). \emph{Novon},
-338--340.
+Appel O, Al-Shehbaz IA. 1997. Generic limits and taxonomy of hornungia,
+pritzelago, and hymenolobus (brassicaceae). \emph{Novon} 338--340.
 
 \leavevmode\hypertarget{ref-R-papaja}{}%
-Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with
-R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}
+Aust F, Barth M. 2018. papaja: Create APA manuscripts with R Markdown.
 
 \leavevmode\hypertarget{ref-barrett1996comparative}{}%
-Barrett, S. C. H., Harder, L. D., \& Worley, A. C. (1996). The
-comparative biology of pollination and mating in flowering plants.
-\emph{Phil. Trans. R. Soc. Lond. B}, \emph{351}(1345), 1271--1280.
+Barrett SCH, Harder LD, Worley AC. 1996. The comparative biology of
+pollination and mating in flowering plants. \emph{Phil Trans R Soc Lond
+B} \textbf{351}:1271--1280.
 
 \leavevmode\hypertarget{ref-R-Matrix}{}%
-Bates, D., \& Maechler, M. (2018). \emph{Matrix: Sparse and dense matrix
-classes and methods}. Retrieved from
-\url{https://CRAN.R-project.org/package=Matrix}
+Bates D, Maechler M. 2018. Matrix: Sparse and dense matrix classes and
+methods.
 
 \leavevmode\hypertarget{ref-bena1998molecular}{}%
-Bena, G., Lejeune, B., Prosperi, J.-M., \& Olivieri, I. (1998).
-Molecular phylogenetic approach for studying life-history evolution: The
-ambiguous example of the genus medicago l. \emph{Proceedings of the
-Royal Society of London B: Biological Sciences}, \emph{265}(1401),
-1141--1151.
+Bena G, Lejeune B, Prosperi J-M, Olivieri I. 1998. Molecular
+phylogenetic approach for studying life-history evolution: The ambiguous
+example of the genus medicago l. \emph{Proceedings of the Royal Society
+of London B: Biological Sciences} \textbf{265}:1141--1151.
 
 \leavevmode\hypertarget{ref-brown1996oldlist}{}%
-Brown, P. M. (1996). OLDLIST: A database of maximum tree ages.
-\emph{Tree Rings, Environment, and Humanity. Radiocarbon}, \emph{1996},
-727--731.
+Brown PM. 1996. OLDLIST: A database of maximum tree ages. \emph{Tree
+rings, environment, and humanity Radiocarbon} \textbf{1996}:727--731.
 
 \leavevmode\hypertarget{ref-charnov1973life}{}%
-Charnov, E. L., \& Schaffer, W. M. (1973). Life-history consequences of
-natural selection: Cole's result revisited. \emph{The American
-Naturalist}, \emph{107}(958), 791--793.
+Charnov EL, Schaffer WM. 1973. Life-history consequences of natural
+selection: Cole's result revisited. \emph{The American Naturalist}
+\textbf{107}:791--793.
 
 \leavevmode\hypertarget{ref-cole1954population}{}%
-Cole, L. C. (1954). The population consequences of life history
-phenomena. \emph{The Quarterly Review of Biology}, \emph{29}(2),
-103--137.
+Cole LC. 1954. The population consequences of life history phenomena.
+\emph{The Quarterly Review of Biology} \textbf{29}:103--137.
 
 \leavevmode\hypertarget{ref-conti1999phylogenetic}{}%
-Conti, E., Soltis, D. E., Hardig, T. M., \& Schneider, J. (1999).
-Phylogenetic relationships of the silver saxifrages (saxifraga, sect.
-Ligulatae haworth): Implications for the evolution of substrate
-specificity, life histories, and biogeography. \emph{Molecular
-Phylogenetics and Evolution}, \emph{13}(3), 536--555.
+Conti E, Soltis DE, Hardig TM, Schneider J. 1999. Phylogenetic
+relationships of the silver saxifrages (saxifraga, sect. Ligulatae
+haworth): Implications for the evolution of substrate specificity, life
+histories, and biogeography. \emph{Molecular Phylogenetics and
+Evolution} \textbf{13}:536--555.
 
 \leavevmode\hypertarget{ref-corbin2004competition}{}%
-Corbin, J. D., \& D'Antonio, C. M. (2004). Competition between native
-perennial and exotic annual grasses: Implications for an historical
-invasion. \emph{Ecology}, \emph{85}(5), 1273--1283.
+Corbin JD, D'Antonio CM. 2004. Competition between native perennial and
+exotic annual grasses: Implications for an historical invasion.
+\emph{Ecology} \textbf{85}:1273--1283.
 
 \leavevmode\hypertarget{ref-cruz2009molecular}{}%
-Cruz-Mazo, G., Buide, M., Samuel, R., \& Narbona, E. (2009). Molecular
-phylogeny of scorzoneroides (asteraceae): Evolution of heterocarpy and
-annual habit in unpredictable environments. \emph{Molecular
-Phylogenetics and Evolution}, \emph{53}(3), 835--847.
+Cruz-Mazo G, Buide M, Samuel R, Narbona E. 2009. Molecular phylogeny of
+scorzoneroides (asteraceae): Evolution of heterocarpy and annual habit
+in unpredictable environments. \emph{Molecular phylogenetics and
+evolution} \textbf{53}:835--847.
 
 \leavevmode\hypertarget{ref-dai2011drought}{}%
-Dai, A. (2011). Drought under global warming: A review. \emph{Wiley
-Interdisciplinary Reviews: Climate Change}, \emph{2}(1), 45--65.
+Dai A. 2011. Drought under global warming: A review. \emph{Wiley
+Interdisciplinary Reviews: Climate Change} \textbf{2}:45--65.
 
 \leavevmode\hypertarget{ref-daru2018widespread}{}%
-Daru, B. H., Park, D. S., Primack, R. B., Willis, C. G., Barrington, D.
-S., Whitfeld, T. J., \ldots{} others. (2018). Widespread sampling biases
-in herbaria revealed from large-scale digitization. \emph{New
-Phytologist}, \emph{217}(2), 939--955.
+Daru BH, Park DS, Primack RB, Willis CG, Barrington DS, Whitfeld TJ,
+Seidler TG, Sweeney PW, Foster DR, Ellison AM, others. 2018. Widespread
+sampling biases in herbaria revealed from large-scale digitization.
+\emph{New Phytologist} \textbf{217}:939--955.
 
 \leavevmode\hypertarget{ref-datson2008climate}{}%
-Datson, P., Murray, B., \& Steiner, K. (2008). Climate and the evolution
-of annual/perennial life-histories in nemesia (scrophulariaceae).
-\emph{Plant Systematics and Evolution}, \emph{270}(1-2), 39--57.
+Datson P, Murray B, Steiner K. 2008. Climate and the evolution of
+annual/perennial life-histories in nemesia (scrophulariaceae).
+\emph{Plant Systematics and Evolution} \textbf{270}:39--57.
 
 \leavevmode\hypertarget{ref-davis2015herbarium}{}%
-Davis, C. C., Willis, C. G., Connolly, B., Kelly, C., \& Ellison, A. M.
-(2015). Herbarium records are reliable sources of phenological change
-driven by climate and provide novel insights into species' phenological
-cueing mechanisms. \emph{American Journal of Botany}, \emph{102}(10),
-1599--1609.
+Davis CC, Willis CG, Connolly B, Kelly C, Ellison AM. 2015. Herbarium
+records are reliable sources of phenological change driven by climate
+and provide novel insights into species' phenological cueing mechanisms.
+\emph{American Journal of Botany} \textbf{102}:1599--1609.
 
 \leavevmode\hypertarget{ref-dittberner2018natural}{}%
-Dittberner, H., Korte, A., Mettler-Altmann, T., Weber, A., Monroe, G.,
-\& Meaux, J. de. (2018). Natural variation in stomata size contributes
-to the local adaptation of water-use efficiency in arabidopsis thaliana.
-\emph{bioRxiv}, 253021.
+Dittberner H, Korte A, Mettler-Altmann T, Weber A, Monroe G, Meaux J de.
+2018. Natural variation in stomata size contributes to the local
+adaptation of water-use efficiency in arabidopsis thaliana.
+\emph{bioRxiv} 253021.
 
 \leavevmode\hypertarget{ref-R-geiger_b}{}%
-Eastman, J., Alfaro, M., Joyce, P., Hipp, A., \& Harmon, L. (2011). A
-novel comparative method for identifying shifts in the rate of character
-evolution on trees. \emph{Evolution}, \emph{65}, 3578--3589.
+Eastman J, Alfaro M, Joyce P, Hipp A, Harmon L. 2011. A novel
+comparative method for identifying shifts in the rate of character
+evolution on trees. \emph{Evolution} \textbf{65}:3578--3589.
 
 \leavevmode\hypertarget{ref-evans2005climate}{}%
-Evans, M. E., Hearn, D. J., Hahn, W. J., Spangle, J. M., \& Venable, D.
-L. (2005). Climate and life-history evolution in evening primroses
-(oenothera, onagraceae): A phylogenetic comparative analysis.
-\emph{Evolution}, \emph{59}(9), 1914--1927.
+Evans ME, Hearn DJ, Hahn WJ, Spangle JM, Venable DL. 2005. Climate and
+life-history evolution in evening primroses (oenothera, onagraceae): A
+phylogenetic comparative analysis. \emph{Evolution}
+\textbf{59}:1914--1927.
 
 \leavevmode\hypertarget{ref-franco1996life}{}%
-Franco, M., \& Silvertown, J. (1996). Life history variation in plants:
-An exploration of the fast-slow continuum hypothesis. \emph{Phil. Trans.
-R. Soc. Lond. B}, \emph{351}(1345), 1341--1348.
+Franco M, Silvertown J. 1996. Life history variation in plants: An
+exploration of the fast-slow continuum hypothesis. \emph{Phil Trans R
+Soc Lond B} \textbf{351}:1341--1348.
 
 \leavevmode\hypertarget{ref-friedman2015all}{}%
-Friedman, J., \& Rubin, M. J. (2015). All in good time: Understanding
-annual and perennial strategies in plants. \emph{American Journal of
-Botany}, \emph{102}(4), 497--499.
+Friedman J, Rubin MJ. 2015. All in good time: Understanding annual and
+perennial strategies in plants. \emph{American journal of botany}
+\textbf{102}:497--499.
 
 \leavevmode\hypertarget{ref-gbifdownload}{}%
-GBIF. (2018, July). GBIF occurrence download. Retrieved from
-\url{https://doi.org/10.15468/dl.xmwqxm}
+GBIF. 2018. GBIF occurrence download.
 
 \leavevmode\hypertarget{ref-R-mvtnorm}{}%
-Genz, A., \& Bretz, F. (2009). \emph{Computation of multivariate normal
-and t probabilities}. Heidelberg: Springer-Verlag.
+Genz A, Bretz F. 2009. Computation of multivariate normal and t
+probabilities, Lecture notes in statistics. Heidelberg: Springer-Verlag.
 
 \leavevmode\hypertarget{ref-R-MCMCglmm}{}%
-Hadfield, J. D. (2010). MCMC methods for multi-response generalized
-linear mixed models: The MCMCglmm R package. \emph{Journal of
-Statistical Software}, \emph{33}(2), 1--22. Retrieved from
-\url{http://www.jstatsoft.org/v33/i02/}
+Hadfield JD. 2010. MCMC methods for multi-response generalized linear
+mixed models: The MCMCglmm R package. \emph{Journal of Statistical
+Software} \textbf{33}:1--22.
 
 \leavevmode\hypertarget{ref-R-geiger_d}{}%
-Harmon, L., Weir, J., Brock, C., Glor, R., \& Challenger, W. (2008).
-GEIGER: Investigating evolutionary radiations. \emph{Bioinformatics},
-\emph{24}, 129--131.
+Harmon L, Weir J, Brock C, Glor R, Challenger W. 2008. GEIGER:
+Investigating evolutionary radiations. \emph{Bioinformatics}
+\textbf{24}:129--131.
 
 \leavevmode\hypertarget{ref-R-logistf}{}%
-Heinze, G., \& Ploner, M. (2018). \emph{Logistf: Firth's bias-reduced
-logistic regression}. Retrieved from
-\url{https://CRAN.R-project.org/package=logistf}
+Heinze G, Ploner M. 2018. Logistf: Firth's bias-reduced logistic
+regression.
 
 \leavevmode\hypertarget{ref-R-purrr}{}%
-Henry, L., \& Wickham, H. (2018). \emph{Purrr: Functional programming
-tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}
+Henry L, Wickham H. 2018. Purrr: Functional programming tools.
 
 \leavevmode\hypertarget{ref-R-raster}{}%
-Hijmans, R. J. (2018). \emph{Raster: Geographic data analysis and
-modeling}. Retrieved from
-\url{https://CRAN.R-project.org/package=raster}
+Hijmans RJ. 2018. Raster: Geographic data analysis and modeling.
 
 \leavevmode\hypertarget{ref-R-phylolm}{}%
-Ho, L. S. T., \& Ane, C. (2014). A linear-time algorithm for gaussian
-and non-gaussian trait evolution models. \emph{Systematic Biology},
-\emph{63}, 397--408.
+Ho LST, Ane C. 2014. A linear-time algorithm for gaussian and
+non-gaussian trait evolution models. \emph{Systematic Biology}
+\textbf{63}:397--408.
 
 \leavevmode\hypertarget{ref-kerr2003space}{}%
-Kerr, J. T., \& Ostrovsky, M. (2003). From space to species: Ecological
-applications for remote sensing. \emph{Trends in Ecology \& Evolution},
-\emph{18}(6), 299--305.
+Kerr JT, Ostrovsky M. 2003. From space to species: Ecological
+applications for remote sensing. \emph{Trends in ecology \& evolution}
+\textbf{18}:299--305.
 
 \leavevmode\hypertarget{ref-kogan2016modelling}{}%
-Kogan, F., Guo, W., Strashnaia, A., Kleshenko, A., Chub, O., \&
-Virchenko, O. (2016). Modelling and prediction of crop losses from noaa
-polar-orbiting operational satellites. \emph{Geomatics, Natural Hazards
-and Risk}, \emph{7}(3), 886--900.
+Kogan F, Guo W, Strashnaia A, Kleshenko A, Chub O, Virchenko O. 2016.
+Modelling and prediction of crop losses from noaa polar-orbiting
+operational satellites. \emph{Geomatics, Natural Hazards and Risk}
+\textbf{7}:886--900.
 
 \leavevmode\hypertarget{ref-kogan2001operational}{}%
-Kogan, F. N. (2001). Operational space technology for global vegetation
-assessment. \emph{Bulletin of the American Meteorological Society},
-\emph{82}(9), 1949--1964.
+Kogan FN. 2001. Operational space technology for global vegetation
+assessment. \emph{Bulletin of the American Meteorological Society}
+\textbf{82}:1949--1964.
 
 \leavevmode\hypertarget{ref-lang2018using}{}%
-Lang, P. L., Willems, F. M., Scheepens, J., Burbano, H. A., \& Bossdorf,
-O. (2018). \emph{Using herbaria to study global environmental change}.
-PeerJ Preprints.
+Lang PL, Willems FM, Scheepens J, Burbano HA, Bossdorf O. 2018. Using
+herbaria to study global environmental change. PeerJ Preprints.
 
 \leavevmode\hypertarget{ref-lelievre2009current}{}%
-Lelièvre, F., \& Volaire, F. (2009). Current and potential development
-of perennial grasses in rainfed mediterranean farming systems.
-\emph{Crop Science}, \emph{49}(6), 2371--2378.
+Lelièvre F, Volaire F. 2009. Current and potential development of
+perennial grasses in rainfed mediterranean farming systems. \emph{Crop
+Science} \textbf{49}:2371--2378.
 
 \leavevmode\hypertarget{ref-mandakova2012whole}{}%
-Mandáková, T., Mummenhoff, K., Al-Shehbaz, I. A., Mucina, L.,
-Mühlhausen, A., \& Lysak, M. A. (2012). Whole-genome triplication and
-species radiation in the southern african tribe heliophileae
-(brassicaceae). \emph{Taxon}, \emph{61}(5), 989--1000.
+Mandáková T, Mummenhoff K, Al-Shehbaz IA, Mucina L, Mühlhausen A, Lysak
+MA. 2012. Whole-genome triplication and species radiation in the
+southern african tribe heliophileae (brassicaceae). \emph{Taxon}
+\textbf{61}:989--1000.
 
 \leavevmode\hypertarget{ref-Mojica2016}{}%
-Mojica, J. P., Mullen, J., Lovell, J. T., Monroe, J. G., Paul, J. R.,
-Oakley, C. G., \& McKay, J. K. (2016). Genetics of water use physiology
-in locally adapted Arabidopsis thaliana. \emph{Plant Science}.
+Mojica JP, Mullen J, Lovell JT, Monroe JG, Paul JR, Oakley CG, McKay JK.
+2016. Genetics of water use physiology in locally adapted Arabidopsis
+thaliana. \emph{Plant Science}.
 doi:\href{https://doi.org/10.1016/j.plantsci.2016.03.015}{10.1016/j.plantsci.2016.03.015}
@@
 \leavevmode\hypertarget{ref-monroe2018drought}{}%
-Monroe, J. G., Powell, T., Price, N., Mullen, J., Howard, A., Evans, K.,
-\ldots{} McKay, J. (2018). Drought adaptation in nature by extensive
-genetic loss-of-function. \emph{bioRxiv}, 372854.
+Monroe JG, Powell T, Price N, Mullen J, Howard A, Evans K, Lovell J,
+McKay J. 2018. Drought adaptation in nature by extensive genetic
+loss-of-function. \emph{bioRxiv} 372854.
 
 \leavevmode\hypertarget{ref-morishima1984differentiation}{}%
-Morishima, H., Sano, Y., \& Oka, H. (1984). Differentiation of perennial
-and annual types due to habitat conditions in the wild riceOryza
-perennis. \emph{Plant Systematics and Evolution}, \emph{144}(2),
-119--135.
+Morishima H, Sano Y, Oka H. 1984. Differentiation of perennial and
+annual types due to habitat conditions in the wild riceOryza perennis.
+\emph{Plant Systematics and Evolution} \textbf{144}:119--135.
 
 \leavevmode\hypertarget{ref-mummenhoff2005phylogeny}{}%
-Mummenhoff, K., Al-Shehbaz, I. A., Bakker, F. T., Linder, H. P., \&
-Mühlhausen, A. (2005). Phylogeny, morphological evolution, and
-speciation of endemic brassicaceae genera in the cape flora of southern
-africa. \emph{Annals of the Missouri Botanical Garden}, 400--424.
+Mummenhoff K, Al-Shehbaz IA, Bakker FT, Linder HP, Mühlhausen A. 2005.
+Phylogeny, morphological evolution, and speciation of endemic
+brassicaceae genera in the cape flora of southern africa. \emph{Annals
+of the Missouri Botanical Garden} 400--424.
 
 \leavevmode\hypertarget{ref-R-bindrcpp}{}%
-Müller, K. (2018). \emph{Bindrcpp: An 'rcpp' interface to active
-bindings}. Retrieved from
-\url{https://CRAN.R-project.org/package=bindrcpp}
+Müller K. 2018. Bindrcpp: An 'rcpp' interface to active bindings.
 
 \leavevmode\hypertarget{ref-R-tibble}{}%
-Müller, K., \& Wickham, H. (2018). \emph{Tibble: Simple data frames}.
-Retrieved from \url{https://CRAN.R-project.org/package=tibble}
+Müller K, Wickham H. 2018. Tibble: Simple data frames.
 
 \leavevmode\hypertarget{ref-R-caper}{}%
-Orme, D., Freckleton, R., Thomas, G., Petzoldt, T., Fritz, S., Isaac,
-N., \& Pearse, W. (2018). \emph{Caper: Comparative analyses of
-phylogenetics and evolution in r}. Retrieved from
-\url{https://CRAN.R-project.org/package=caper}
+Orme D, Freckleton R, Thomas G, Petzoldt T, Fritz S, Isaac N, Pearse W.
+2018. Caper: Comparative analyses of phylogenetics and evolution in r.
 
 \leavevmode\hypertarget{ref-R-ape}{}%
-Paradis, E., \& Schliep, K. (2018). Ape 5.0: An environment for modern
-phylogenetics and evolutionary analyses in R. \emph{Bioinformatics},
-\emph{xx}, xxx--xxx.
+Paradis E, Schliep K. 2018. Ape 5.0: An environment for modern
+phylogenetics and evolutionary analyses in R. \emph{Bioinformatics}
+\textbf{xx}:xxx--xxx.
 
 \leavevmode\hypertarget{ref-parry2005prospects}{}%
-Parry, M., Flexas, J., \& Medrano, H. (2005). Prospects for crop
-production under drought: Research priorities and future directions.
-\emph{Annals of Applied Biology}, \emph{147}(3), 211--226.
+Parry M, Flexas J, Medrano H. 2005. Prospects for crop production under
+drought: Research priorities and future directions. \emph{Annals of
+Applied Biology} \textbf{147}:211--226.
 
 \leavevmode\hypertarget{ref-R-sp}{}%
-Pebesma, E. J., \& Bivand, R. S. (2005). Classes and methods for spatial
-data in R. \emph{R News}, \emph{5}(2), 9--13. Retrieved from
-\url{https://CRAN.R-project.org/doc/Rnews/}
+Pebesma EJ, Bivand RS. 2005. Classes and methods for spatial data in R.
+\emph{R News} \textbf{5}:9--13.
 
 \leavevmode\hypertarget{ref-R-coda}{}%
-Plummer, M., Best, N., Cowles, K., \& Vines, K. (2006). CODA:
-Convergence diagnosis and output analysis for mcmc. \emph{R News},
-\emph{6}(1), 7--11. Retrieved from
-\url{https://journal.r-project.org/archive/}
+Plummer M, Best N, Cowles K, Vines K. 2006. CODA: Convergence diagnosis
+and output analysis for mcmc. \emph{R News} \textbf{6}:7--11.
 
 \leavevmode\hypertarget{ref-R-gee}{}%
-R by Thomas Lumley, V. J. C. P. to, \& author., B. R. N. that
-maintainers are not available to give advice on using a package they did
-not. (2015). \emph{Gee: Generalized estimation equation solver}.
-Retrieved from \url{https://CRAN.R-project.org/package=gee}
+R by Thomas Lumley VJCP to, author. BRN that maintainers are not
+available to give advice on using a package they did not. 2015. Gee:
+Generalized estimation equation solver.
 
 \leavevmode\hypertarget{ref-R-base}{}%
-R Core Team. (2018). \emph{R: A language and environment for statistical
-computing}. Vienna, Austria: R Foundation for Statistical Computing.
-Retrieved from \url{https://www.R-project.org/}
+R Core Team. 2018. R: A language and environment for statistical
+computing. Vienna, Austria: R Foundation for Statistical Computing.
 
 \leavevmode\hypertarget{ref-R-phytools}{}%
-Revell, L. J. (2012). Phytools: An r package for phylogenetic
-comparative biology (and other things). \emph{Methods in Ecology and
-Evolution}, \emph{3}, 217--223.
+Revell LJ. 2012. Phytools: An r package for phylogenetic comparative
+biology (and other things). \emph{Methods in Ecology and Evolution}
+\textbf{3}:217--223.
 
 \leavevmode\hypertarget{ref-R-maps}{}%
-Richard A. Becker, O. S. code by, Ray Brownrigg. Enhancements by Thomas
-P Minka, A. R. W. R. version by, \& Deckmyn., A. (2018). \emph{Maps:
-Draw geographical maps}. Retrieved from
-\url{https://CRAN.R-project.org/package=maps}
+Richard A. Becker OS code by, Ray Brownrigg. Enhancements by Thomas P
+Minka ARWR version by, Deckmyn. A. 2018. Maps: Draw geographical maps.
 
 \leavevmode\hypertarget{ref-rojas2011assessing}{}%
-Rojas, O., Vrieling, A., \& Rembold, F. (2011). Assessing drought
-probability for agricultural areas in africa with coarse resolution
-remote sensing imagery. \emph{Remote Sensing of Environment},
-\emph{115}(2), 343--352.
+Rojas O, Vrieling A, Rembold F. 2011. Assessing drought probability for
+agricultural areas in africa with coarse resolution remote sensing
+imagery. \emph{Remote sensing of Environment} \textbf{115}:343--352.
 
 \leavevmode\hypertarget{ref-roumet2006suites}{}%
-Roumet, C., Urcelay, C., \& Dı'az, S. (2006). Suites of root traits
-differ between annual and perennial species growing in the field.
-\emph{New Phytologist}, \emph{170}(2), 357--368.
+Roumet C, Urcelay C, Dı'az S. 2006. Suites of root traits differ between
+annual and perennial species growing in the field. \emph{New
+phytologist} \textbf{170}:357--368.
 
 \leavevmode\hypertarget{ref-sayre2013new}{}%
-Sayre, R. G., Comer, P., Hak, J., Josse, C., Bow, J., Warner, H.,
-\ldots{} others. (2013). A new map of standardized terrestrial
+Sayre RG, Comer P, Hak J, Josse C, Bow J, Warner H, Larwanou M, Kelbessa
+E, Bekele T, Kehl H, others. 2013. A new map of standardized terrestrial
 ecosystems of africa. \emph{African Geographical Review}.
@@
 \leavevmode\hypertarget{ref-schaffer1975selection}{}%
-Schaffer, W., \& Gadgil, M. (1975). Selection for optimal life histories
-in plants. \emph{Ecology and Evolution of Communities.}, 142--157.
+Schaffer W, Gadgil M. 1975. Selection for optimal life histories in
+plants. \emph{Ecology and evolution of communities} 142--157.
 
 \leavevmode\hypertarget{ref-R-geiger_c}{}%
-Slater, G., Harmon, L., Wegmann, D., Joyce, P., Revell, L., \& Alfaro,
-M. (2012). Fitting models of continuous trait evolution to incompletely
-sampled comparative data using approximate bayesian computation.
-\emph{Evolution}, \emph{66}, 752--762.
+Slater G, Harmon L, Wegmann D, Joyce P, Revell L, Alfaro M. 2012.
+Fitting models of continuous trait evolution to incompletely sampled
+comparative data using approximate bayesian computation.
+\emph{Evolution} \textbf{66}:752--762.
+
+\leavevmode\hypertarget{ref-R-shape}{}%
+Soetaert K. 2018. Shape: Functions for plotting graphical shapes,
+colors.
 
 \leavevmode\hypertarget{ref-R-diagram}{}%
-Soetaert, K. (2017). \emph{Diagram: Functions for visualising simple
-graphs (networks), plotting flow diagrams}. Retrieved from
-\url{https://CRAN.R-project.org/package=diagram}
-
-\leavevmode\hypertarget{ref-R-shape}{}%
-Soetaert, K. (2018). \emph{Shape: Functions for plotting graphical
-shapes, colors}. Retrieved from
-\url{https://CRAN.R-project.org/package=shape}
+Soetaert K. 2017. Diagram: Functions for visualising simple graphs
+(networks), plotting flow diagrams.
 
 \leavevmode\hypertarget{ref-soltis2017digitization}{}%
-Soltis, P. S. (2017). Digitization of herbaria enables novel research.
-\emph{American Journal of Botany}, \emph{104}(9), 1281--1284.
+Soltis PS. 2017. Digitization of herbaria enables novel research.
+\emph{American journal of botany} \textbf{104}:1281--1284.
 
 \leavevmode\hypertarget{ref-stearns1992evolution}{}%
-Stearns, S. C. (1992). \emph{The evolution of life histories}.
+Stearns SC. 1992. The evolution of life histories.
 
 \leavevmode\hypertarget{ref-stropp2016mapping}{}%
-Stropp, J., Ladle, R. J., M. Malhado, A. C., Hortal, J., Gaffuri, J., H.
-Temperley, W., \ldots{} Mayaux, P. (2016). Mapping ignorance: 300 years
-of collecting flowering plants in africa. \emph{Global Ecology and
-Biogeography}, \emph{25}(9), 1085--1096.
+Stropp J, Ladle RJ, M. Malhado AC, Hortal J, Gaffuri J, H. Temperley W,
+Olav Skøien J, Mayaux P. 2016. Mapping ignorance: 300 years of
+collecting flowering plants in africa. \emph{Global Ecology and
+Biogeography} \textbf{25}:1085--1096.
 
 \leavevmode\hypertarget{ref-tank2008annuals}{}%
-Tank, D. C., \& Olmstead, R. G. (2008). From annuals to perennials:
-Phylogeny of subtribe castillejinae (orobanchaceae). \emph{American
-Journal of Botany}, \emph{95}(5), 608--625.
+Tank DC, Olmstead RG. 2008. From annuals to perennials: Phylogeny of
+subtribe castillejinae (orobanchaceae). \emph{American Journal of
+Botany} \textbf{95}:608--625.
 
 \leavevmode\hypertarget{ref-vaclavik2017effects}{}%
-Václavı'k, T., Beckmann, M., Cord, A. F., \& Bindewald, A. M. (2017).
-Effects of uv-b radiation on leaf hair traits of invasive
-plants---combining historical herbarium records with novel remote
-sensing data. \emph{PloS One}, \emph{12}(4), e0175671.
+Václavı'k T, Beckmann M, Cord AF, Bindewald AM. 2017. Effects of uv-b
+radiation on leaf hair traits of invasive plants---combining historical
+herbarium records with novel remote sensing data. \emph{PloS one}
+\textbf{12}:e0175671.
 
 \leavevmode\hypertarget{ref-R-MASS}{}%
-Venables, W. N., \& Ripley, B. D. (2002). \emph{Modern applied
-statistics with s} (Fourth.). New York: Springer. Retrieved from
-\url{http://www.stats.ox.ac.uk/pub/MASS4}
+Venables WN, Ripley BD. 2002. Modern applied statistics with s, Fourth.
+ed. New York: Springer.
 
 \leavevmode\hypertarget{ref-verboom2004testing}{}%
-Verboom, G. A., Linder, H. P., \& Stock, W. D. (2004). Testing the
-adaptive nature of radiation: Growth form and life history divergence in
-the african grass genus ehrharta (poaceae: Ehrhartoideae).
-\emph{American Journal of Botany}, \emph{91}(9), 1364--1370.
+Verboom GA, Linder HP, Stock WD. 2004. Testing the adaptive nature of
+radiation: Growth form and life history divergence in the african grass
+genus ehrharta (poaceae: Ehrhartoideae). \emph{American Journal of
+Botany} \textbf{91}:1364--1370.
+
+\leavevmode\hypertarget{ref-R-forcats}{}%
+Wickham H. 2018a. Forcats: Tools for working with categorical variables
+(factors).
+
+\leavevmode\hypertarget{ref-R-stringr}{}%
+Wickham H. 2018b. Stringr: Simple, consistent wrappers for common string
+operations.
+
+\leavevmode\hypertarget{ref-R-tidyverse}{}%
+Wickham H. 2017. Tidyverse: Easily install and load the 'tidyverse'.
 
 \leavevmode\hypertarget{ref-R-ggplot2}{}%
-Wickham, H. (2016). \emph{Ggplot2: Elegant graphics for data analysis}.
-Springer-Verlag New York. Retrieved from \url{http://ggplot2.org}
-
-\leavevmode\hypertarget{ref-R-tidyverse}{}%
-Wickham, H. (2017). \emph{Tidyverse: Easily install and load the
-'tidyverse'}. Retrieved from
-\url{https://CRAN.R-project.org/package=tidyverse}
-
-\leavevmode\hypertarget{ref-R-forcats}{}%
-Wickham, H. (2018a). \emph{Forcats: Tools for working with categorical
-variables (factors)}. Retrieved from
-\url{https://CRAN.R-project.org/package=forcats}
-
-\leavevmode\hypertarget{ref-R-stringr}{}%
-Wickham, H. (2018b). \emph{Stringr: Simple, consistent wrappers for
-common string operations}. Retrieved from
-\url{https://CRAN.R-project.org/package=stringr}
+Wickham H. 2016. Ggplot2: Elegant graphics for data analysis.
+Springer-Verlag New York.
 
 \leavevmode\hypertarget{ref-R-dplyr}{}%
-Wickham, H., François, R., Henry, L., \& Müller, K. (2018). \emph{Dplyr:
-A grammar of data manipulation}. Retrieved from
-\url{https://CRAN.R-project.org/package=dplyr}
+Wickham H, François R, Henry L, Müller K. 2018. Dplyr: A grammar of data
+manipulation.
 
 \leavevmode\hypertarget{ref-R-tidyr}{}%
-Wickham, H., \& Henry, L. (2018). \emph{Tidyr: Easily tidy data with
-'spread()' and 'gather()' functions}. Retrieved from
-\url{https://CRAN.R-project.org/package=tidyr}
+Wickham H, Henry L. 2018. Tidyr: Easily tidy data with 'spread()' and
+'gather()' functions.
 
 \leavevmode\hypertarget{ref-R-readr}{}%
-Wickham, H., Hester, J., \& Francois, R. (2017). \emph{Readr: Read
-rectangular text data}. Retrieved from
-\url{https://CRAN.R-project.org/package=readr}
+Wickham H, Hester J, Francois R. 2017. Readr: Read rectangular text
+data.
 
 \leavevmode\hypertarget{ref-willis2017old}{}%
-Willis, C. G., Ellwood, E. R., Primack, R. B., Davis, C. C., Pearson, K.
-D., Gallinat, A. S., \ldots{} others. (2017). Old plants, new tricks:
-Phenological research using herbarium specimens. \emph{Trends in Ecology
-\& Evolution}, \emph{32}(7), 531--546.
+Willis CG, Ellwood ER, Primack RB, Davis CC, Pearson KD, Gallinat AS,
+Yost JM, Nelson G, Mazer SJ, Rossington NL, others. 2017. Old plants,
+new tricks: Phenological research using herbarium specimens.
+\emph{Trends in ecology \& evolution} \textbf{32}:531--546.
 
 \leavevmode\hypertarget{ref-wolf2016altitudinal}{}%
-Wolf, A., Zimmerman, N. B., Anderegg, W. R., Busby, P. E., \&
-Christensen, J. (2016). Altitudinal shifts of the native and introduced
-flora of c alifornia in the context of 20th-century warming.
-\emph{Global Ecology and Biogeography}, \emph{25}(4), 418--429.
+Wolf A, Zimmerman NB, Anderegg WR, Busby PE, Christensen J. 2016.
+Altitudinal shifts of the native and introduced flora of c alifornia in
+the context of 20th-century warming. \emph{Global ecology and
+biogeography} \textbf{25}:418--429.
 
@@
 We used R (Version 3.5.1; R Core Team, 2018) and the R-packages
-\emph{ape} (Version 5.2; Paradis \& Schliep, 2018; Orme et al., 2018;
+\emph{ape} (Version 5.2; Paradis and Schliep, 2018; Orme et al., 2018;
 Soetaert, 2018), \emph{bindrcpp} (Version 0.2.2; Müller, 2018),
 \emph{caper} (Version 1.0.1; Orme et al., 2018), \emph{coda} (Version
-0.19.2; Plummer, Best, Cowles, \& Vines, 2006), \emph{diagram} (Version
-1.6.4; Soetaert, 2017), \emph{dplyr} (Version 0.7.8; Wickham et al.,
-2018), \emph{forcats} (Version 0.3.0; Wickham, 2018a), \emph{gee}
-(Version 4.13.19; R by Thomas Lumley \& author., 2015), \emph{geiger}
-(Version 2.0.6; Alfaro et al., 2009; Eastman, Alfaro, Joyce, Hipp, \&
-Harmon, 2011; Harmon, Weir, Brock, Glor, \& Challenger, 2008; Slater et
-al., 2012), \emph{ggplot2} (Version 3.1.0; Wickham, 2016),
-\emph{logistf} (Version 1.23; Heinze \& Ploner, 2018), \emph{maps}
-(Version 3.3.0; Richard A. Becker, Ray Brownrigg. Enhancements by Thomas
-P Minka, \& Deckmyn., 2018), \emph{MASS} (Version 7.3.51.1; Venables \&
-Ripley, 2002), \emph{Matrix} (Version 1.2.15; Bates \& Maechler, 2018),
-\emph{MCMCglmm} (Version 2.26; Hadfield, 2010), \emph{mvtnorm} (Version
-1.0.8; Genz \& Bretz, 2009), \emph{papaja} (Version 0.1.0.9842; Aust \&
-Barth, 2018), \emph{phylolm} (Version 2.6; Ho \& Ane, 2014),
-\emph{phytools} (Version 0.6.60; Revell, 2012), \emph{purrr} (Version
-0.2.5; Henry \& Wickham, 2018), \emph{raster} (Version 2.8.4; Hijmans,
-2018), \emph{readr} (Version 1.2.1; Wickham et al., 2017), \emph{shape}
-(Version 1.4.4; Soetaert, 2018), \emph{sp} (Version 1.3.1; Pebesma \&
-Bivand, 2005), \emph{stringr} (Version 1.3.1; Wickham, 2018b),
-\emph{tibble} (Version 1.4.2; Müller \& Wickham, 2018), \emph{tidyr}
-(Version 0.8.2; Wickham \& Henry, 2018), and \emph{tidyverse} (Version
-1.2.1; Wickham, 2017) for all our analyses.
+0.19.2; Plummer et al., 2006), \emph{diagram} (Version 1.6.4; Soetaert,
+2017), \emph{dplyr} (Version 0.7.8; Wickham et al., 2018),
+\emph{forcats} (Version 0.3.0; Wickham, 2018a), \emph{gee} (Version
+4.13.19; R by Thomas Lumley and author., 2015), \emph{geiger} (Version
+2.0.6; Alfaro et al., 2009; Eastman et al., 2011; Harmon et al., 2008;
+Slater et al., 2012), \emph{ggplot2} (Version 3.1.0; Wickham, 2016),
+\emph{logistf} (Version 1.23; Heinze and Ploner, 2018), \emph{maps}
+(Version 3.3.0; Richard A. Becker et al., 2018), \emph{MASS} (Version
+7.3.51.1; Venables and Ripley, 2002), \emph{Matrix} (Version 1.2.15;
+Bates and Maechler, 2018), \emph{MCMCglmm} (Version 2.26; Hadfield,
+2010), \emph{mvtnorm} (Version 1.0.8; Genz and Bretz, 2009),
+\emph{papaja} (Version 0.1.0.9842; Aust and Barth, 2018), \emph{phylolm}
+(Version 2.6; Ho and Ane, 2014), \emph{phytools} (Version 0.6.60;
+Revell, 2012), \emph{purrr} (Version 0.2.5; Henry and Wickham, 2018),
+\emph{raster} (Version 2.8.4; Hijmans, 2018), \emph{readr} (Version
+1.2.1; Wickham et al., 2017), \emph{shape} (Version 1.4.4; Soetaert,
+2018), \emph{sp} (Version 1.3.1; Pebesma and Bivand, 2005),
+\emph{stringr} (Version 1.3.1; Wickham, 2018b), \emph{tibble} (Version
+1.4.2; Müller and Wickham, 2018), \emph{tidyr} (Version 0.8.2; Wickham
+and Henry, 2018), and \emph{tidyverse} (Version 1.2.1; Wickham, 2017)
+for all our analyses.
 

### 2018-11-30 manuscript/manuscript.Rmd
updates, edits, spell checked

@@
 ---
-title             : "Drought frequency predicts plant life history strategies"
+title             : "Drought frequency predicts life history strategies in African _Heliophila_"
 shorttitle        : "Drought and life history"
@@
     corresponding : yes    # Define only one corresponding author
-    address       : "307 University Ave, Fort Collins, CO 80523"
+    address       : "307 University Ave, Fort Collins, CO 80521"
     email         : "monroejg@colostate.edu"
@@
   - id            : "1"
-    institution   : "Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80523, USA"
+    institution   : "Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80521, USA"
   - id            : "2"
-    institution   : "College of Agriculture, Colorado State University, Fort Collins, CO 80523, USA"
+    institution   : "College of Agriculture, Colorado State University, Fort Collins, CO 80521, USA"
   - id            : "3"
@@
 For plants, annual and perennial life histories are thought to reflect adaptation to environments that differ in the frequency of environmental stress such as drought.
-Here we test this hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa by integrating 2192 herbaria occurrence records with 34 years of satellite-based drought detection.
+Here we test this hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa, by integrating 34 years of satellite-based drought measurements with 2192 herbaria occurrence records.
 Consistent with predictions from classic life history theory, we find that perennial \textit{Heliophila} species occur in environments where droughts are significantly less frequent compared to annuals.
@@
 Additionally, the difference in drought frequency between annual and perennial species distributions is greatest during the summer and fall, which also appears to be when annuals are in the seed phase of their life cycle based on collection dates of annual species.
-Together, these finding provide empirical support for classic hypotheses about the drivers of life history strategy in plants - that perennials out compete annuals in environments with less frequent drought and that annuals are adpated to enviroments with more frequent drought by escaping drought prone seasons as seeds.
+Together, these finding provide empirical support for classic hypotheses about the drivers of life history strategy in plants - that perennials out compete annuals in environments with less frequent drought and that annuals are adapted to environments with more frequent drought by escaping drought prone seasons as seeds.
 
@@
 Classical theory predict shorter life spans in environments where adult mortality is high [@charnov1973life; @stearns1992evolution;@franco1996life].
-In plants, this has been extended to the hypothessis that annuality is adaptive when it allows plants to escape drought [@schaffer1975selection].
-Lack of water is perhaps the greatest threat to survival during vegetatitive or reproductive growth and annuals can remain dormant (and protected as a seed) during drought.
+In plants, this has been extended to the hypothesis that annuality is adaptive when it allows plants to escape drought [@schaffer1975selection].
+Lack of water is perhaps the greatest threat to survival during vegetative or reproductive growth and annuals can remain dormant (and protected as a seed) during drought.
 Thus, environments with greater seasonal drought frequency may select for annual life histories that complete reproduction prior to drought prone seasons.
 Conversely, environments with less frequent drought may select for perennial species, which may benefit from multiple bouts of reproduction and competitive advantage by preventing recruitment of annual species [@corbin2004competition].
-These predictions have been supported by the association of annualilty with arid environments in _Oryza perennis_ [@morishima1984differentiation] and _Oenothera_ [@evans2005climate].
+These predictions have been supported by the observation of annuals in arid environments in _Oryza perennis_ [@morishima1984differentiation] and _Oenothera_ [@evans2005climate].
 Additionally, annual and perennial species of _Nemesia_ were qualitatively associated with winter rather and summer rainfall environments respectively [@datson2008climate] and annual species of _Scorzoneroides_ were associated with environments classified as unpredictable [@cruz2009molecular].
-However, whether the history frequency of drought events indeed predicts the distributions annual or perennial life history strategies has yet to be tested in a phylogenetic context with suitable sample sizes of distrubution and climate data.
+However, whether the history frequency of drought events indeed predicts the distributions annual or perennial life history strategies has yet to be tested in a phylogenetic context with suitable sample sizes of distribution and climate data.
 Here we combine a long-term global dataset of satellite detected drought events with metadata from natural history collections to test these classic hypotheses about the evolution of life history strategies within the African endemic mustard genus, _Heliophila_ L.
 (Brassicaceae).
-If annuallity is an adaptive strategy allowing plants to escape drought prone seasons, then drought frequency should predict the distribution of life history strategies across landscapes, and annual species should be more commonly associated with drought prone regions than perennial species.
+If annuality is an adaptive strategy allowing plants to escape drought prone seasons, then drought frequency should predict the distribution of life history strategies across landscapes, and annual species should be more commonly associated with drought prone regions than perennial species.
 Furthermore, if annual species have adapted to escape drought prone seasons, observations of growing annual species (i.e.
@@
 ### Availability
-All analyses were peformed using R. All data and the source code to produce this manuscript are available at https://github.com/greymonroe/heliophila. Softare used is listed in the supplement.
+All analyses were performed using R. All data and the source code to produce this manuscript are available at https://github.com/greymonroe/heliophila. Software used is listed in the supplement.
 
@@
 But they also present a new tool to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes [@kerr2003space].
-As such, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting infraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought].
+As such, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting intraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought].
 Here, we accessed VHI data at $16km^2$ resolution from 1981 to 2015 (https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php) to characterize the seasonal drought frequencies experienced by annual and perennial _Heliophila_ species.
@@
 Perenniality was defined based any form of perennial life history (e.g., herbs, shrubs, mixed, etc).
-Because the nature of species reported with mixed traits were unknown (ie.
+Because the nature of species reported with mixed traits were unknown (i.e.
 plasticity vs.
@@
 
-### Heliophila occurance records
+### Heliophila occurrence records
 ```{r raw_GBIF}
@@
 And as they become digitized [@soltis2017digitization], these collections have been used to study relationships between trait distributions, geography, and climate [@davis2015herbarium;@wolf2016altitudinal;@stropp2016mapping;@vaclavik2017effects].
-To characterize the diributions of annual and perennial _Heliophila_ species, all records for the genus _Heliophila_ were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018 [@gbifdownload].
+To characterize the distributions of annual and perennial _Heliophila_ species, all records for the genus _Heliophila_ were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018 [@gbifdownload].
 
@@
 ### Drought frequency calculations
-To characterize drought regimens across the distrubtions of annual and perennial species of _Heliohpila_, we calculated drought during different seasons at the location of observations for _Heliophila_ records using the VHI.
+To characterize drought regimens across the distributions of annual and perennial species of _Heliophila_, we calculated drought during different seasons at the location of observations for _Heliophila_ records using the VHI.
 Specifically, we created global maps of the frequencies of observing drought conditions (VHI<40, NOAA) during the winter (quarter surrounding winter solstice), spring (quarter surrounding spring equinox), summer (quarter surrounding summer solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015.
@@
 
-### Filtering of occurance records
+### Filtering of occurrence records
 ```{r cleanup}
@@
 * records with geospatial data  
-* records without known geospatial coodinate issues (i.e., coordinates reported are those of herbarium)  
+* records without known geospatial coordinate issues (i.e., coordinates reported are those of herbarium)  
 * records from collection sites classified as land pixels   
@@
 To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimes, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency.
-First, we compared the frequency of drought during the winter, spring, summer, and fall between raw occurance records of annual and perennial species by t-tests.
-To account for variation in the number of occurance records per species, we next calculated the mean drought frequency during the winter, spring, summer and fall for each species.
+First, we compared the frequency of drought during the winter, spring, summer, and fall between raw occurrence records of annual and perennial species by t-tests.
+To account for variation in the number of occurrence records per species, we next calculated the mean drought frequency during the winter, spring, summer and fall for each species.
 The relationships between species mean values of drought frequency during each season and life habitat (annual or perennial) were tested using Firth's penalized-likelihood logistic regressions.
-Because demographic histories caused by ancestry can confound trait - environment associations, we then tested for the relationships between drought frequeny and life history while controlling for relationships between species using phylogenetic logistic regressions.
+Because demographic histories caused by ancestry can confound trait - environment associations, we then tested for the relationships between drought frequency and life history while controlling for relationships between species using phylogenetic logistic regressions.
 
@@
 ```
-(ref:mapsboxplots) Locations of (a) annual and (b) perennial Heliophila.
-Drought frequency during the (c) winter, (d) spring, (e) summer and (f) fall.
-(g) Drought frequencies during each season observed at the collection sites of _Heliophila_ records.
+(ref:mapsboxplots) Locations of occurrence records of (a) annual and (b) perennial _Heliophila._ Drought frequency during the (c) winter, (d) spring, (e) summer and (f) fall measured using the VHI.
+(g) Drought frequencies during each season at the observation locations of annual and perennial _Heliophila_ (t tests, \*\* = p < 0.01).
 
@@
 
-pdf("../figures/maps_boxplots.pdf", width=5, height=7.5)
+pdf("../figures/maps_boxplots.pdf", width=4.5, height=7.2)
 world <- map_data("world") # we already did this, but we can do it again
@@
 
+col_scale<-c( 'white',"gray90", "gray10", 'black')
+#col_scale<-c( "green4","green", "white","purple4","black")
+
 Winter_SA <-crop(Winter, rbind(c(12,37),c(-35, -18)))
@@
   geom_tile(aes(fill=value)) + 
-  scale_fill_gradientn(colors=c( 'white',"gray80", "gray20", 'black'), na.value="white", name=NULL) +
+  scale_fill_gradientn(colors=col_scale, na.value="white", name=NULL) +
   geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
@@
   theme_classic()+
-  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
+  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.4,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
   labs(title="Winter drought freq.", x="Longitude (°)", y="Latitude (°)")
@@
   theme_classic()+
-  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
+  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.4,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
   labs(title="Spring drought freq.", x="Longitude (°)", y="Latitude (°)")
@@
   theme_classic()+
-  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
+  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.4,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
   labs(title="Summer drought freq.", x="Longitude (°)", y="Latitude (°)")
@@
   theme_classic()+
-  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.5,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
+  theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.4,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
   labs(title="Fall drought freq.", x="Longitude (°)", y="Latitude (°)")
@@
   facet_wrap(~season, ncol=4 )+
-  geom_jitter(width = .3, alpha=0.1, pch=3)+
+  geom_jitter(width = .3, alpha=0.05, pch=3)+
   geom_boxplot(outlier.colour = "white", outlier.fill = NA, fill=NA, width=.3)+
@@
   scale_x_discrete(name="Life history", labels=c("Annual","Perennial"))+
-  labs(y="Drought frequency")
+  labs(y="Drought frequency")+
+  annotate("text", x=1.5, y=0.8, label= "**")
 
@@
 ```
-Out of `r nrow(GBIF)` _Heliophila_ GBIF recrods,  `r sum(GBIF$species %in% species_life_history$species)` were for species with reported life history [@mummenhoff2005phylogeny], `r sum(!is.na(GBIF$decimalLatitude))` had geospatial data, `r sum(!(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)))` did not have geospatial issues, `r sum(!(is.na(GBIF$Summer)) & !(is.na(GBIF$decimalLatitude)))` were located on pixels classified as land having drought measurements, `r sum(GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude))` were located in Africa, `r sum(!duplicated(GBIF$ymd_lat_lon_sp) & !is.na(GBIF$decimalLatitude)) ` were not duplicated.
-After all filtering steps, `r nrow(GBIF_cleaned)` records for `r length(unique(GBIF_cleaned$species))` species (Figure \@ref(fig:phylogeny), Table \@ref(tab:speciesmeanstable)) passed for further analyses.
+Out of `r nrow(GBIF)` _Heliophila_ GBIF records,  `r sum(GBIF$species %in% species_life_history$species)` were for species with reported life history [@mummenhoff2005phylogeny], `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude))` had geospatial data, `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)))` did not have geospatial issues, `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)) & !(is.na(GBIF$Summer)) & !(is.na(GBIF$decimalLatitude)))` were located on pixels classified as land having drought measurements, `r sum(GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & (GBIF$species %in% species_life_history$species) & !(is.na(GBIF$Summer)))` were located in Africa, `r sum(!duplicated(GBIF$ymd_lat_lon_sp) & GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & (GBIF$species %in% species_life_history$species) & !(is.na(GBIF$Summer))) ` were not duplicated.
+Thus, after all filtering steps, `r nrow(GBIF_cleaned)` records for `r length(unique(GBIF_cleaned$species))` species (Figure \@ref(fig:phylogeny), Table \@ref(tab:speciesmeanstable)) passed for further analyses.
 The number of samples varied between species, with a mean of `r mean(species_means$n)` samples per species.
@@
 There were clear visual differences between the distributions of the `r sum(GBIF_cleaned$life_history=="a")` annual and the `r sum(GBIF_cleaned$life_history=="p")` perennial _Heliophila_ observation records (see Figure \@ref(fig:speciesmaps) for maps of individual species).
-While annual species were generally found in the western regions of South Africa and Namibia, primarily in the Cape Floristic Region and Succulent Karoo (Figure \@ref(fig:mapsboxplots)a), the occurances of perennials extended to the east coast of South Africa (Figure \@ref(fig:mapsboxplots)b).
+While annual species were generally found in the western regions of South Africa and Namibia, primarily in the Cape Floristic Region and Succulent Karoo (Figure \@ref(fig:mapsboxplots)a), the occurrence of perennials extended to the east coast of South Africa (Figure \@ref(fig:mapsboxplots)b).
 
@@
 This heterogeneity is expected, given that this is one of the most climatically diverse regions of the Earth [@sayre2013new].
-It is worth noting the east to west cline in drought frequency observed during the summer, which distinguishes the high drought frequency of the Kalahari Sands and Namid Destert phyotogeographic regions from the low drought frequency of the Drakensberg Mountains and Coastal Zambesian phytogeographic regions.
+It is worth noting the east to west cline in drought frequency observed during the summer, which distinguishes the high drought frequency of the Kalahari Sands and Namid Desert phytogeographic regions from the low drought frequency of the Drakensberg Mountains and Coastal Zambesian phytogeographic regions.
 In the Cape phytogeographic region there was finer scale heterogeneity in drought frequency during the summer.
@@
 Conversely, perenniality should be adaptive in places where such stresses are less frequent.
-We found that the frequency of drought was significantly higher at the locations of occurance records for annual species.
+We found that the frequency of drought was significantly higher at the locations of occurrence records for annual species.
 In terms of raw observation records
-(Figure \@ref(fig:mapsboxplots)g), the frequency of drought was signficantly higher at the location of annuals during the winter (t = `r twin$statistic`, p = `r twin$p.value`), spring (t = `r tspr$statistic`, p = `r tspr$p.value`), summer (t = `r tsum$statistic`, p = `r tsum$p.value`), and fall (t = `r tfal$statistic`, p = `r tfal$p.value`).
-Because raw occurance records do not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable)), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials by a Firths-penalized logistic regression.
-We found that the mean drought frequencies were signficantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a).
+(Figure \@ref(fig:mapsboxplots)g), the frequency of drought was significantly higher at the location of annuals during the winter (t = `r twin$statistic`, p = `r twin$p.value`), spring (t = `r tspr$statistic`, p = `r tspr$p.value`), summer (t = `r tsum$statistic`, p = `r tsum$p.value`), and fall (t = `r tfal$statistic`, p = `r tfal$p.value`).
+Because raw occurrence records do not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable)), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials by a Firths-penalized logistic regression.
+We found that the mean drought frequencies were significantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a).
 We further tested whether annual species are found in places where droughts occur more frequently while controlling for the phylogentic relatedness of _Heliophila_ species.
-This is important, because environmental differences in species distributions can be confounded with demographic history caused by acestry.
-We found that while controlling for phylogeny, the mean drought frequencies were signficantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a).
-These findings indicate that common acestry alone does not explain differences the drought frequencies experienced in the environemnts of annual and perennial _Heliophila_.
+This is important, because environmental differences in species distributions can be confounded with demographic history caused by ancestry.
+We found that while controlling for phylogeny, the mean drought frequencies were significantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a).
+These findings indicate that common acestry alone does not explain differences the drought frequencies experienced in the environments of annual and perennial _Heliophila_.
 
@@
 (ref:lineplots) (a) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of Heliophila.
-(b) Collection dates of GBIF records of annual and perennial species of Heliophila.
+(phylogenetic logistic regression, \* = p < 0.05, \*\* = p < 0.01) (b) Collection dates of GBIF records of annual and perennial species of Heliophila.
 
@@
   scale_x_continuous(name="Season", breaks=c(45,135,225, 315), labels=c("Winter","Spring","Summer","Fall"))+
-  labs(title="Date of observation for occurance records", y="Density")+
-  theme(plot.title = element_text(hjust = 0.5, size=4))
+  labs(title="Date of observation for occurrence records", y="Density")+
+  theme(plot.title = element_text(hjust = 0.5, size=10))
 
@@
 
-```{r lineplots, fig.cap='(ref:lineplots)'}
+```{r lineplots, fig.cap='(ref:lineplots)', out.width = "\\textwidth", fig.pos = "!h"}
 knitr::include_graphics("../figures/line_and_dates.pdf", dpi = 108)
@@
 
-The preceding results indicated that annual speices are found in environments where droughts are signifacantly more frequent, especially in the summer and fall.
+The preceding results indicated that annual species are found in environments where droughts are significantly more frequent, especially in the summer and fall.
 Classic life history theory hypothesizes that annuality reflects adaptation to such environments because it allows species to escape stressful conditions.
 If this is the case, we would expect that annuals spend the drought prone seasons of summer and fall as seeds.
-To test this hypothesis, we compared the dates of occurance records between annual and perennial _Heliophila_ species.
+To test this hypothesis, we compared the dates of occurrence records between annual and perennial _Heliophila_ species.
 The distributions reveal a considerable difference in the timing of observation of these two life histories.
 In comparison to perennials, which appear to be collected throughout the year, annuals are almost exclusively observed during the winter and spring (Figure \@ref(fig:lineplots)b).
-The differences between the distrubution of collection dates were signicant by all tests (ks.test D = `r  kout$statistic`, p = `r  kout$p.value`; bartlett.test K2 = `r bout$statistic`, p = `r bout$p.value`) This is consistent with a model of life history in which annual species flower in the spring, set seed, senesce, and die before the summer.
-Thus, these annual species are likely to remain dormant during the summer and fall, when drought is the strongest predictor of the distrubutions of annual and perennial life histories (Figure \@ref(fig:lineplots)a).
+The differences between the distribution of collection dates were significant by all tests (ks.test D = `r  kout$statistic`, p = `r  kout$p.value`; bartlett.test K2 = `r bout$statistic`, p = `r bout$p.value`) This is consistent with a model of life history in which annual species flower in the spring, set seed, senesce, and die before the summer.
+Thus, these annual species are likely to remain dormant during the summer and fall, when drought is the strongest predictor of the distributions of annual and perennial life histories (Figure \@ref(fig:lineplots)a).
 
@@
 To test the hypothesis that annual and perennial plants reflect adaptation to alternative drought environments we examined the landscape distribution of life history strategies in the large and diverse mustard genus, _Heliophila_.
-Using metadata of `r nrow(GBIF_cleaned)` occurance records and a 34 year dataset of satellite-detected droughts, we tested the prediction that annual species are more often observed in drought-prone locations than perennial species, when controlling for phylogenetic relatedness.
+Using metadata of `r nrow(GBIF_cleaned)` occurrence records and a 34 year dataset of satellite-detected droughts, we tested the prediction that annual species are more often observed in drought-prone locations than perennial species, when controlling for phylogenetic relatedness.
 We found that drought frequency is significantly different between the distributions of annual and perennial species, with annuals being found in environments with more frequent drought, and that this signal is strongest during the seasons when annuals are likely escaping via seed dormancy These results remain significant while controlling for the phylogenetic relationships of _Heliophila_ species, yielding support for the role that natural selection has played in driving contemporary distributions of these alternatives strategies in relation to drought regimens.
 
-We cannot elimiate the possibility that counfounding traits or environmental variables are the causative factors explaining variation in the distributions of annual and perennial species.
+We cannot eliminate the possibility that confounding traits or environmental variables are the causative factors explaining variation in the distributions of annual and perennial species.
 Nevertheless, these results reveal quantitative support for the classic prediction that annual species are found in environemnts that experience more frequent drought than perennial species.
-These findings complement previous reports of qualitative associations between annaulity with enviroments characterized as having increased aridity [@evans2005climate], alternative precipitation defined habitats [@morishima1984differentiation;@datson2008climate], or greater unpredictability [@cruz2009molecular].
-However, to our knowledge this is the first study to demonstrate a signficant association between life history and drought in a phylogenetic context informed by large scale species distribution data and long term drought measures.
+These findings complement previous reports of qualitative associations between annuality with enviroments characterized as having increased aridity [@evans2005climate], alternative precipitation defined habitats [@morishima1984differentiation;@datson2008climate], or greater unpredictability [@cruz2009molecular].
+However, to our knowledge this is the first study to demonstrate a significant association between life history and drought in a phylogenetic context informed by large scale species distribution data and long term drought measures.
 
@@
 However, the ecosystems of southern Africa include several biodiversity hotspots and are among the most botanically well sampled regions on Earth [@daru2018widespread], suggesting that this may currently be the optimal region for our analyses of life history distribution.
-Indeed, we were able to use `r nrow(GBIF_cleaned)` occurance records to study 42 species, which represents a significant advance over relying on personal observations to characterize species distrubtuions.
+Indeed, we were able to use `r nrow(GBIF_cleaned)` occurrence records to study 42 species, which represents a significant advance over relying on personal observations to characterize species distributions.
 
@@
 They also suggest that perenniality is adaptive in environemnts where droughts are less frequent.
-While most previous work has focused on describing the evolutionary origins of annuality [@barrett1996comparative;@conti1999phylogenetic;@andreasen2001unequal;@verboom2004testing;@friedman2015all] there are at least a few other cases where perenniality appears to have arisen from an annual acestor [@bena1998molecular;@tank2008annuals].
-And while early theory predicted selection for annuality when adult morality is high [@stearns1992evolution],  we also find evidence that the transition to perenniality could be explained by historical drought regimens.
+While most previous work has focused on describing the evolutionary origins of annuality [@barrett1996comparative;@conti1999phylogenetic;@andreasen2001unequal;@verboom2004testing;@friedman2015all] there are at least a few other cases where perenniality appears to have arisen from an annual ancestor [@bena1998molecular;@tank2008annuals].
+And while early theory predicted selection for annuality when adult morality is high [@stearns1992evolution],  we also find evidence that the transition to perenniality could be explained by the historical frequency of drought, a likely cause of mortality in plants.
 The phylogeny reveals several transitions from annual to perennial life history (Figure \@ref(fig:phylogeny)a) and that the distributions of perennial _Heliophila_ extend into regions where drought frequency is low (Figure \@ref(fig:mapsboxplots)b,Figure \@ref(fig:speciesmaps)).
@@
 These findings suggest that species with locally adaptive life history strategies could be threatened by rapidly changing drought regimens [@dai2011drought].
-This could have impacts on ecosystem functioning and processes such as carbon cycling if the composition of annual and perennial species changes as a response [@garnier1997specific; @roumet2006suites;monroe2018ecoevolutionary].
-Furthermore, the frequency of drought may be an important factor when considering the use of perennial cropping sysyems [@parry2005prospects;@lelievre2009current].
+This could have impacts on ecosystem functioning and processes such as carbon cycling if the composition of annual and perennial species changes as a response [@garnier1997specific; @roumet2006suites;@monroe2018ecoevolutionary].
+Furthermore, the frequency of drought may be an important factor when considering the use of perennial cropping systems [@parry2005prospects;@lelievre2009current].
 
@@
 This research was supported by NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
+
+# Author contributions
+
+JGM, BG, KGT and JKM contributed to the design of the research, interpretation, and writing the manuscript. JGM, BG, and KGT contributed to the performance of the research and data analysis.
 
@@
 
-(ref:speciesmaps) Maps of occurance records for individual species. Orange points indicate annual species. Blue points indicate perennial species.
+(ref:speciesmaps) Maps of occurrence records for individual species. Orange points indicate annual species. Blue points indicate perennial species.
 

### 2018-11-30 manuscript/manuscript.tex
updates, edits, spell checked

@@
 \hypersetup{unicode=true,
-            pdftitle={Drought frequency predicts plant life history strategies},
+            pdftitle={Drought frequency predicts life history strategies in African Heliophila},
             pdfauthor={J. Grey Monroe, Brian Gill, Kathryn G. Turner, \& John K. McKay},
@@
 
-  \title{Drought frequency predicts plant life history strategies}
+  \title{Drought frequency predicts life history strategies in African
+\emph{Heliophila}}
     \author{J. Grey Monroe\textsuperscript{1,2}, Brian Gill\textsuperscript{3},
@@
 \vspace{0.5cm}
-\textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80523, USA\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO 80523, USA\\\textsuperscript{3} Institute for Environment and Society, Brown University, Providence, RI 02912, USA\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA 16802, USA}
+\textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{3} Institute for Environment and Society, Brown University, Providence, RI 02912, USA\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA 16802, USA}
 \keywords{drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records}
@@
 Correspondence concerning this article should be addressed to J. Grey
-Monroe, 307 University Ave, Fort Collins, CO 80523. E-mail:
+Monroe, 307 University Ave, Fort Collins, CO 80521. E-mail:
 \href{mailto:monroejg@colostate.edu}{\nolinkurl{monroejg@colostate.edu}}}
@@
 hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of
-flowering plants native to Africa by integrating 2192 herbaria
-occurrence records with 34 years of satellite-based drought detection.
-Consistent with predictions from classic life history theory, we find
-that perennial \textit{Heliophila} species occur in environments where
-droughts are significantly less frequent compared to annuals. These
-associations are predictive while controlling for phylogeny, lending
-support to the hypothesis that drought related natural selection has
-influenced the distributions of these strategies. Additionally, the
+flowering plants native to Africa, by integrating 34 years of
+satellite-based drought measurements with 2192 herbaria occurrence
+records. Consistent with predictions from classic life history theory,
+we find that perennial \textit{Heliophila} species occur in environments
+where droughts are significantly less frequent compared to annuals.
+These associations are predictive while controlling for phylogeny,
+lending support to the hypothesis that drought related natural selection
+has influenced the distributions of these strategies. Additionally, the
 difference in drought frequency between annual and perennial species
@@
 history strategy in plants - that perennials out compete annuals in
-environments with less frequent drought and that annuals are adpated to
-enviroments with more frequent drought by escaping drought prone seasons
-as seeds. 
+environments with less frequent drought and that annuals are adapted to
+environments with more frequent drought by escaping drought prone
+seasons as seeds. 
 }
@@
 1996; Stearns, 1992). In plants, this has been extended to the
-hypothessis that annuality is adaptive when it allows plants to escape
+hypothesis that annuality is adaptive when it allows plants to escape
 drought (Schaffer and Gadgil, 1975). Lack of water is perhaps the
-greatest threat to survival during vegetatitive or reproductive growth
-and annuals can remain dormant (and protected as a seed) during drought.
+greatest threat to survival during vegetative or reproductive growth and
+annuals can remain dormant (and protected as a seed) during drought.
 Thus, environments with greater seasonal drought frequency may select
@@
 annual species (Corbin and D'Antonio, 2004). These predictions have been
-supported by the association of annualilty with arid environments in
+supported by the observation of annuals in arid environments in
 \emph{Oryza perennis} (Morishima et al., 1984) and \emph{Oenothera}
@@
 be tested in a phylogenetic context with suitable sample sizes of
-distrubution and climate data.\\
+distribution and climate data.\\
 Here we combine a long-term global dataset of satellite detected drought
@@
 the African endemic mustard genus, \emph{Heliophila} L. (Brassicaceae).
-If annuallity is an adaptive strategy allowing plants to escape drought
+If annuality is an adaptive strategy allowing plants to escape drought
 prone seasons, then drought frequency should predict the distribution of
@@
 
-All analyses were peformed using R. All data and the source code to
+All analyses were performed using R. All data and the source code to
 produce this manuscript are available at
-\url{https://github.com/greymonroe/heliophila}. Softare used is listed
+\url{https://github.com/greymonroe/heliophila}. Software used is listed
 in the supplement.
@@
 to study drought related ecology of natural species and proven useful
-for predicting infraspecific variation in drought tolerance traits and
-genes (Dittberner et al., 2018; Mojica et al., 2016; Monroe et al.,
-2018). Here, we accessed VHI data at \(16km^2\) resolution from 1981 to
-2015 (\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php})
-to characterize the seasonal drought frequencies experienced by annual
-and perennial \emph{Heliophila} species.
+for predicting intraspecific variation in drought tolerance traits and
+genes (Dittberner et al., 2018; Mojica et al., 2016; John G Monroe et
+al., 2018). Here, we accessed VHI data at \(16km^2\) resolution from
+1981 to 2015
+(\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php}) to
+characterize the seasonal drought frequencies experienced by annual and
+perennial \emph{Heliophila} species.
 
@@
 histories. Perenniality was defined based any form of perennial life
-history (e.g., herbs, shrubs, mixed, etc). We also repeated these
-analyses without \emph{H. subulata} and \emph{H. minima} which have
-mixed life histories and found similar results. Because the original
-authors didn't report the nature of the mixed trait (Mummenhoff et al.,
-2005), (ie. plasticity vs.~genetic variation), we classified these
-species here as perennial since they have the capacity to maintain
-vegetative growth after reproduction at least to some capacity.
-
-\hypertarget{heliophila-occurance-records}{%
-\subsubsection{Heliophila occurance
-records}\label{heliophila-occurance-records}}
+history (e.g., herbs, shrubs, mixed, etc). Because the nature of species
+reported with mixed traits were unknown (i.e.~plasticity vs.~genetic
+variation), we classified these species here as perennial since they can
+maintain vegetative growth after reproduction at least to some capacity.
+
+\hypertarget{heliophila-occurrence-records}{%
+\subsubsection{Heliophila occurrence
+records}\label{heliophila-occurrence-records}}
 
@@
 al., 2016; Václavı'k et al., 2017; Wolf et al., 2016). To characterize
-the diributions of annual and perennial \emph{Heliophila} species, all
+the distributions of annual and perennial \emph{Heliophila} species, all
 records for the genus \emph{Heliophila} were downloaded from the Global
@@
 
-To characterize drought regimens across the distrubtions of annual and
-perennial species of \emph{Heliohpila}, we calculated drought during
+To characterize drought regimens across the distributions of annual and
+perennial species of \emph{Heliophila}, we calculated drought during
 different seasons at the location of observations for \emph{Heliophila}
@@
 
-\hypertarget{filtering-of-occurance-records}{%
-\subsubsection{Filtering of occurance
-records}\label{filtering-of-occurance-records}}
+\hypertarget{filtering-of-occurrence-records}{%
+\subsubsection{Filtering of occurrence
+records}\label{filtering-of-occurrence-records}}
 
@@
 \item
-  records without known geospatial coodinate issues (i.e., coordinates
+  records without known geospatial coordinate issues (i.e., coordinates
   reported are those of herbarium)\\
@@
 frequency of drought during the winter, spring, summer, and fall between
-raw occurance records of annual and perennial species by t-tests. To
-account for variation in the number of occurance records per species, we
-next calculated the mean drought frequency during the winter, spring,
+raw occurrence records of annual and perennial species by t-tests. To
+account for variation in the number of occurrence records per species,
+we next calculated the mean drought frequency during the winter, spring,
 summer and fall for each species. The relationships between species mean
@@
 confound trait - environment associations, we then tested for the
-relationships between drought frequeny and life history while
+relationships between drought frequency and life history while
 controlling for relationships between species using phylogenetic
@@
 \begin{figure}[!h]
-\includegraphics[width=\textwidth]{../figures/maps_boxplots} \caption{Locations of (a) annual and (b) perennial Heliophila.
-Drought frequency during the (c) winter, (d) spring, (e) summer and (f)
-fall. (g) Drought frequencies during each season observed at the
-collection sites of \emph{Heliophila} records.}\label{fig:mapsboxplots}
+\includegraphics[width=\textwidth]{../figures/maps_boxplots} \caption{Locations of occurrence records of (a) annual and (b)
+perennial \emph{Heliophila.} Drought frequency during the (c) winter,
+(d) spring, (e) summer and (f) fall measured using the VHI. (g) Drought
+frequencies during each season at the observation locations of annual
+and perennial \emph{Heliophila} (t tests, ** = p \textless{} 0.01).}\label{fig:mapsboxplots}
 \end{figure}
 
-Out of 8670 \emph{Heliophila} GBIF recrods, 6634 were for species with
-reported life history (Mummenhoff et al., 2005), 3653 had geospatial
-data, 3627 did not have geospatial issues, 3460 were located on pixels
-classified as land having drought measurements, 3457 were located in
-Africa, 3162 were not duplicated. After all filtering steps, 2192
+Out of 8670 \emph{Heliophila} GBIF records, 6634 were for species with
+reported life history (Mummenhoff et al., 2005), 2856 had geospatial
+data, 2833 did not have geospatial issues, 2684 were located on pixels
+classified as land having drought measurements, 2543 were located in
+Africa, 2192 were not duplicated. Thus, after all filtering steps, 2192
 records for 42 species (Figure \ref{fig:phylogeny}, Table
@@
 Africa and Namibia, primarily in the Cape Floristic Region and Succulent
-Karoo (Figure \ref{fig:mapsboxplots}a), the occurances of perennials
+Karoo (Figure \ref{fig:mapsboxplots}a), the occurrence of perennials
 extended to the east coast of South Africa (Figure
@@
 the summer, which distinguishes the high drought frequency of the
-Kalahari Sands and Namid Destert phyotogeographic regions from the low
+Kalahari Sands and Namid Desert phytogeographic regions from the low
 drought frequency of the Drakensberg Mountains and Coastal Zambesian
@@
 found that the frequency of drought was significantly higher at the
-locations of occurance records for annual species. In terms of raw
+locations of occurrence records for annual species. In terms of raw
 observation records (Figure \ref{fig:mapsboxplots}g), the frequency of
-drought was signficantly higher at the location of annuals during the
+drought was significantly higher at the location of annuals during the
 winter (t = 10.65, p = 0.00), spring (t = 10.73, p = 0.00), summer (t =
-12.67, p = 0.00), and fall (t = 15.26, p = 0.00). Because raw occurance
+12.67, p = 0.00), and fall (t = 15.26, p = 0.00). Because raw occurrence
 records do not account for variation in the number of records per
@@
 between annuals and perennials by a Firths-penalized logistic
-regression. We found that the mean drought frequencies were signficantly
-higher (\(\alpha = 0.05\)) in annual species during the spring, summer,
-and fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a). We
-further tested whether annual species are found in places where droughts
-occur more frequently while controlling for the phylogentic relatedness
-of \emph{Heliophila} species. This is important, because environmental
-differences in species distributions can be confounded with demographic
-history caused by acestry. We found that while controlling for
-phylogeny, the mean drought frequencies were signficantly higher
-(\(\alpha = 0.05\)) in annual species during the spring, summer, and
-fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a). These
-findings indicate that common acestry alone does not explain differences
-the drought frequencies experienced in the environemnts of annual and
-perennial \emph{Heliophila}.
+regression. We found that the mean drought frequencies were
+significantly higher (\(\alpha = 0.05\)) in annual species during the
+spring, summer, and fall (Table \ref{tab:modelstable}, Figure
+\ref{fig:lineplots}a). We further tested whether annual species are
+found in places where droughts occur more frequently while controlling
+for the phylogentic relatedness of \emph{Heliophila} species. This is
+important, because environmental differences in species distributions
+can be confounded with demographic history caused by ancestry. We found
+that while controlling for phylogeny, the mean drought frequencies were
+significantly higher (\(\alpha = 0.05\)) in annual species during the
+spring, summer, and fall (Table \ref{tab:modelstable}, Figure
+\ref{fig:lineplots}a). These findings indicate that common acestry alone
+does not explain differences the drought frequencies experienced in the
+environments of annual and perennial \emph{Heliophila}.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{../figures/line_and_dates.pdf}
-\caption{\label{fig:lineplots}(a) Comparison (mean +- SE) of drought frequency across
+\begin{figure}[!h]
+\includegraphics[width=\textwidth]{../figures/line_and_dates} \caption{(a) Comparison (mean +- SE) of drought frequency across
 seasons measured at the GBIF records of annual and perennial species of
-Heliophila. (b) Collection dates of GBIF records of annual and perennial
-species of Heliophila.}
+Heliophila. (phylogenetic logistic regression, * = p \textless{} 0.05,
+** = p \textless{} 0.01) (b) Collection dates of GBIF records of annual
+and perennial species of Heliophila.}\label{fig:lineplots}
 \end{figure}
 
-The preceding results indicated that annual speices are found in
-environments where droughts are signifacantly more frequent, especially
+The preceding results indicated that annual species are found in
+environments where droughts are significantly more frequent, especially
 in the summer and fall. Classic life history theory hypothesizes that
@@
 expect that annuals spend the drought prone seasons of summer and fall
-as seeds. To test this hypothesis, we compared the dates of occurance
+as seeds. To test this hypothesis, we compared the dates of occurrence
 records between annual and perennial \emph{Heliophila} species. The
@@
 exclusively observed during the winter and spring (Figure
-\ref{fig:lineplots}b). The differences between the distrubution of
-collection dates were signicant by all tests (ks.test D = 0.25, p = 0;
+\ref{fig:lineplots}b). The differences between the distribution of
+collection dates were significant by all tests (ks.test D = 0.25, p = 0;
 bartlett.test K2 = 503.18, p = 0.00) This is consistent with a model of
@@
 likely to remain dormant during the summer and fall, when drought is the
-strongest predictor of the distrubutions of annual and perennial life
+strongest predictor of the distributions of annual and perennial life
 histories (Figure \ref{fig:lineplots}a).
@@
 distribution of life history strategies in the large and diverse mustard
-genus, \emph{Heliophila}. Using metadata of 2192 occurance records and a
-34 year dataset of satellite-detected droughts, we tested the prediction
-that annual species are more often observed in drought-prone locations
-than perennial species, when controlling for phylogenetic relatedness.
-We found that drought frequency is significantly different between the
-distributions of annual and perennial species, with annuals being found
-in environments with more frequent drought, and that this signal is
-strongest during the seasons when annuals are likely escaping via seed
-dormancy These results remain significant while controlling for the
-phylogenetic relationships of \emph{Heliophila} species, yielding
+genus, \emph{Heliophila}. Using metadata of 2192 occurrence records and
+a 34 year dataset of satellite-detected droughts, we tested the
+prediction that annual species are more often observed in drought-prone
+locations than perennial species, when controlling for phylogenetic
+relatedness. We found that drought frequency is significantly different
+between the distributions of annual and perennial species, with annuals
+being found in environments with more frequent drought, and that this
+signal is strongest during the seasons when annuals are likely escaping
+via seed dormancy These results remain significant while controlling for
+the phylogenetic relationships of \emph{Heliophila} species, yielding
 support for the role that natural selection has played in driving
@@
 
-We cannot elimiate the possibility that counfounding traits or
+We cannot eliminate the possibility that confounding traits or
 environmental variables are the causative factors explaining variation
@@
 frequent drought than perennial species. These findings complement
-previous reports of qualitative associations between annaulity with
+previous reports of qualitative associations between annuality with
 enviroments characterized as having increased aridity (Evans et al.,
@@
 2009). However, to our knowledge this is the first study to demonstrate
-a signficant association between life history and drought in a
+a significant association between life history and drought in a
 phylogenetic context informed by large scale species distribution data
@@
 for our analyses of life history distribution. Indeed, we were able to
-use 2192 occurance records to study 42 species, which represents a
+use 2192 occurrence records to study 42 species, which represents a
 significant advance over relying on personal observations to
-characterize species distrubtuions.
+characterize species distributions.
 
@@
 al., 2004) there are at least a few other cases where perenniality
-appears to have arisen from an annual acestor (Bena et al., 1998; Tank
+appears to have arisen from an annual ancestor (Bena et al., 1998; Tank
 and Olmstead, 2008). And while early theory predicted selection for
 annuality when adult morality is high (Stearns, 1992), we also find
-evidence that the transition to perenniality could be explained by
-historical drought regimens. The phylogeny reveals several transitions
-from annual to perennial life history (Figure \ref{fig:phylogeny}a) and
-that the distributions of perennial \emph{Heliophila} extend into
-regions where drought frequency is low (Figure
-\ref{fig:mapsboxplots}b,Figure \ref{fig:speciesmaps}). Perennials may be
-able to out complete annual relatives in environments where the
-infrequency of drought favors strategies that allow plants to benefit
-from growth over many seasons. It may also indicate that annuals rely on
-drought as a source of disturbance for seedling recruitment when
+evidence that the transition to perenniality could be explained by the
+historical frequency of drought, a likely cause of mortality in plants.
+The phylogeny reveals several transitions from annual to perennial life
+history (Figure \ref{fig:phylogeny}a) and that the distributions of
+perennial \emph{Heliophila} extend into regions where drought frequency
+is low (Figure \ref{fig:mapsboxplots}b,Figure \ref{fig:speciesmaps}).
+Perennials may be able to out complete annual relatives in environments
+where the infrequency of drought favors strategies that allow plants to
+benefit from growth over many seasons. It may also indicate that annuals
+rely on drought as a source of disturbance for seedling recruitment when
 competing with perennials (Corbin and D'Antonio, 2004). Indeed, no
@@
 processes such as carbon cycling if the composition of annual and
-perennial species changes as a response {[}@garnier1997specific; Roumet
-et al. (2006);monroe2018ecoevolutionary{]}. Furthermore, the frequency
-of drought may be an important factor when considering the use of
-perennial cropping sysyems (Lelièvre and Volaire, 2009; Parry et al.,
-2005).
+perennial species changes as a response (Garnier et al., 1997; J Grey
+Monroe et al., 2018; Roumet et al., 2006). Furthermore, the frequency of
+drought may be an important factor when considering the use of perennial
+cropping systems (Lelièvre and Volaire, 2009; Parry et al., 2005).
 
@@
 
+\hypertarget{author-contributions}{%
+\section{Author contributions}\label{author-contributions}}
+
+JGM, BG, KGT and JKM contributed to the design of the research,
+interpretation, and writing the manuscript. JGM, BG, and KGT contributed
+to the performance of the research and data analysis.
+
 \hypertarget{references}{%
@@
 
+\leavevmode\hypertarget{ref-garnier1997specific}{}%
+Garnier E, Cordonnier P, Guillerm J-L, Sonié L. 1997. Specific leaf area
+and leaf nitrogen concentration in annual and perennial grass species
+growing in mediterranean old-fields. \emph{Oecologia}
+\textbf{111}:490--498.
+
 \leavevmode\hypertarget{ref-gbifdownload}{}%
@@
 doi:\href{https://doi.org/10.1016/j.plantsci.2016.03.015}{10.1016/j.plantsci.2016.03.015}
+
+\leavevmode\hypertarget{ref-monroe2018ecoevolutionary}{}%
+Monroe JG, Markman DW, Beck WS, Felton AJ, Vahsen ML, Pressler Y. 2018.
+Ecoevolutionary dynamics of carbon cycling in the anthropocene.
+\emph{Trends in ecology \& evolution} \textbf{33}:213--225.
 
@@
 \includegraphics{../figures/speciesmaps.pdf}
-\caption{\label{fig:speciesmaps}Maps of occurance records for individual species.
+\caption{\label{fig:speciesmaps}Maps of occurrence records for individual species.
 Orange points indicate annual species. Blue points indicate perennial

### 2018-12-01 manuscript/manuscript.Rmd
formatting

@@
 ---
-title             : "Drought frequency predicts life history strategies in African _Heliophila_"
+title             : "Drought frequency predicts life history strategies in _Heliophila_"
 shorttitle        : "Drought and life history"
@@
 Additionally, annual and perennial species of _Nemesia_ were qualitatively associated with winter rather and summer rainfall environments respectively [@datson2008climate] and annual species of _Scorzoneroides_ were associated with environments classified as unpredictable [@cruz2009molecular].
-However, whether the history frequency of drought events indeed predicts the distributions annual or perennial life history strategies has yet to be tested in a phylogenetic context with suitable sample sizes of distribution and climate data.
-Here we combine a long-term global dataset of satellite detected drought events with metadata from natural history collections to test these classic hypotheses about the evolution of life history strategies within the African endemic mustard genus, _Heliophila_ L.
+However, whether the history frequency of drought events indeed predicts the distributions annual or perennial life history strategies has yet to be tested.
+Here we combine a long-term global dataset of satellite detected drought events with metadata from natural history collections to test these classic hypotheses within the African endemic mustard genus, _Heliophila_ L.
 (Brassicaceae).
@@
 
-### Life history data for Heliophila
+### Life history data for_ Heliophila_
 _Heliophila_ is a genus of flowering plants endemic to the southern portion of Africa including the Cape Floristic and Succulent Karoo Regions.
@@
 
-### Heliophila occurrence records
+###_ Heliophila_ occurrence records
 ```{r raw_GBIF}
@@
 ### Sequence data for phylogeny
-Aligned Heliophila ITS sequences were obtained from previous work by @mandakova2012whole.  Aethionema, Alliaria, Cardamine, Chamira, and Rorippa ITS records from were downloaded from Genbank.
+Aligned_ Heliophila_ ITS sequences were obtained from previous work by @mandakova2012whole.  Aethionema, Alliaria, Cardamine, Chamira, and Rorippa ITS records from were downloaded from Genbank.
 
@@
 apa_table(modelstable
-  , caption = "Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species the winter, spring, summer, and fall."
+  , caption = "Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of_ Heliophila_ species the winter, spring, summer, and fall."
   , note = "` = Firth's penalized logistic regression. * = Phylogenetically constrained logistic regression. Annual species were scored as 0 and perennial species as 1.",
@@
 
-(ref:lineplots) (a) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of Heliophila.
-(phylogenetic logistic regression, \* = p < 0.05, \*\* = p < 0.01) (b) Collection dates of GBIF records of annual and perennial species of Heliophila.
+(ref:lineplots) (a) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of_ Heliophila_.
+(phylogenetic logistic regression, \* = p < 0.05, \*\* = p < 0.01) (b) Collection dates of GBIF records of annual and perennial species of_ Heliophila_.
 

### 2018-12-01 manuscript/manuscript.tex
formatting

@@
 \hypersetup{unicode=true,
-            pdftitle={Drought frequency predicts life history strategies in African Heliophila},
+            pdftitle={Drought frequency predicts life history strategies in Heliophila},
             pdfauthor={J. Grey Monroe, Brian Gill, Kathryn G. Turner, \& John K. McKay},
@@
 
-  \title{Drought frequency predicts life history strategies in African
-\emph{Heliophila}}
+  \title{Drought frequency predicts life history strategies in \emph{Heliophila}}
     \author{J. Grey Monroe\textsuperscript{1,2}, Brian Gill\textsuperscript{3},
@@
 the distributions annual or perennial life history strategies has yet to
-be tested in a phylogenetic context with suitable sample sizes of
-distribution and climate data.\\
+be tested.\\
 Here we combine a long-term global dataset of satellite detected drought
 events with metadata from natural history collections to test these
-classic hypotheses about the evolution of life history strategies within
-the African endemic mustard genus, \emph{Heliophila} L. (Brassicaceae).
-If annuality is an adaptive strategy allowing plants to escape drought
-prone seasons, then drought frequency should predict the distribution of
-life history strategies across landscapes, and annual species should be
-more commonly associated with drought prone regions than perennial
-species. Furthermore, if annual species have adapted to escape drought
-prone seasons, observations of growing annual species (i.e.~occurring in
-forms other than seed) should be rare during drought prone seasons.
-Phylogenetic relatedness can have significant non-random effects on
-species distributions and life history traits (Barrett et al., 1996),
-and therefore we assessed the relationship between life history
-distribution and drought frequency in a phylogenetically controlled
-background.
+classic hypotheses within the African endemic mustard genus,
+\emph{Heliophila} L. (Brassicaceae). If annuality is an adaptive
+strategy allowing plants to escape drought prone seasons, then drought
+frequency should predict the distribution of life history strategies
+across landscapes, and annual species should be more commonly associated
+with drought prone regions than perennial species. Furthermore, if
+annual species have adapted to escape drought prone seasons,
+observations of growing annual species (i.e.~occurring in forms other
+than seed) should be rare during drought prone seasons. Phylogenetic
+relatedness can have significant non-random effects on species
+distributions and life history traits (Barrett et al., 1996), and
+therefore we assessed the relationship between life history distribution
+and drought frequency in a phylogenetically controlled background.
 
@@
 
-\hypertarget{life-history-data-for-heliophila}{%
-\subsubsection{Life history data for
-Heliophila}\label{life-history-data-for-heliophila}}
+\hypertarget{life-history-data-for_-heliophila_}{%
+\subsubsection{Life history data for\_
+Heliophila\_}\label{life-history-data-for_-heliophila_}}
 
@@
 
-\hypertarget{heliophila-occurrence-records}{%
-\subsubsection{Heliophila occurrence
-records}\label{heliophila-occurrence-records}}
+\#\#\#\_ Heliophila\_ occurrence records
 
@@
 
-Aligned Heliophila ITS sequences were obtained from previous work by
+Aligned\_ Heliophila\_ ITS sequences were obtained from previous work by
 Mandáková et al. (2012). Aethionema, Alliaria, Cardamine, Chamira, and
@@
 \begin{threeparttable}
-\caption{\label{tab:modelstable}Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species the winter, spring, summer, and fall.}
+\caption{\label{tab:modelstable}Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of\_ Heliophila\_ species the winter, spring, summer, and fall.}
 \begin{tabular}{lllll}
@@
 \includegraphics[width=\textwidth]{../figures/line_and_dates} \caption{(a) Comparison (mean +- SE) of drought frequency across
-seasons measured at the GBIF records of annual and perennial species of
-Heliophila. (phylogenetic logistic regression, * = p \textless{} 0.05,
-** = p \textless{} 0.01) (b) Collection dates of GBIF records of annual
-and perennial species of Heliophila.}\label{fig:lineplots}
+seasons measured at the GBIF records of annual and perennial species
+of\_ Heliophila\_. (phylogenetic logistic regression, * = p \textless{}
+0.05, ** = p \textless{} 0.01) (b) Collection dates of GBIF records of
+annual and perennial species of\_ Heliophila\_.}\label{fig:lineplots}
 \end{figure}

### 2018-12-01 manuscript/manuscript.Rmd
species #

@@
 _Heliophila_ is a genus of flowering plants endemic to the southern portion of Africa including the Cape Floristic and Succulent Karoo Regions.
-These are among the most botanically diverse environments on Earth and the estimated ~50 _Heliophila_ species occurring there are considered to make up the most diverse genus of the family Brassicaceae [@mummenhoff2005phylogeny;@mandakova2012whole].
+These are among the most botanically diverse environments on Earth and the _Heliophila_ species occurring there are considered to make up the most diverse genus of the family Brassicaceae [@mummenhoff2005phylogeny;@mandakova2012whole].
 This genus includes both perennial and annual species and this change in life history strategy has likely arisen multiple independent times [@appel1997generic;@mummenhoff2005phylogeny].

### 2018-12-01 manuscript/manuscript.tex
species #

@@
 Regions. These are among the most botanically diverse environments on
-Earth and the estimated \textasciitilde{}50 \emph{Heliophila} species
-occurring there are considered to make up the most diverse genus of the
-family Brassicaceae (Mandáková et al., 2012; Mummenhoff et al., 2005).
-This genus includes both perennial and annual species and this change in
-life history strategy has likely arisen multiple independent times
-(Appel and Al-Shehbaz, 1997; Mummenhoff et al., 2005). Furthermore, the
-fine scale climatic heterogeneity of Southern Africa is ideal for
-studying the distribution of traits in relation to environmental
-parameters (Sayre et al., 2013). We used life histories reported by
-Mummenhoff et al. (2005), grouping species with annual or perennial life
-histories. Perenniality was defined based any form of perennial life
-history (e.g., herbs, shrubs, mixed, etc). Because the nature of species
-reported with mixed traits were unknown (i.e.~plasticity vs.~genetic
-variation), we classified these species here as perennial since they can
-maintain vegetative growth after reproduction at least to some capacity.
+Earth and the \emph{Heliophila} species occurring there are considered
+to make up the most diverse genus of the family Brassicaceae (Mandáková
+et al., 2012; Mummenhoff et al., 2005). This genus includes both
+perennial and annual species and this change in life history strategy
+has likely arisen multiple independent times (Appel and Al-Shehbaz,
+1997; Mummenhoff et al., 2005). Furthermore, the fine scale climatic
+heterogeneity of Southern Africa is ideal for studying the distribution
+of traits in relation to environmental parameters (Sayre et al., 2013).
+We used life histories reported by Mummenhoff et al. (2005), grouping
+species with annual or perennial life histories. Perenniality was
+defined based any form of perennial life history (e.g., herbs, shrubs,
+mixed, etc). Because the nature of species reported with mixed traits
+were unknown (i.e.~plasticity vs.~genetic variation), we classified
+these species here as perennial since they can maintain vegetative
+growth after reproduction at least to some capacity.
 

### 2018-12-10 manuscript/manuscript.Rmd
remove firth

@@
 occurring in forms other than seed) should be rare during drought prone seasons.
-Phylogenetic relatedness can have significant non-random effects on species distributions and life history traits [@barrett1996comparative], and therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetically controlled background.
+Phylogenetic relatedness can influence tests of associations between species’ traits and their environments [@barrett1996comparative;@felsenstein1985phylogenies], and therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetic context.
 
@@
 
-### Life history data for_ Heliophila_
+### Life history data for _Heliophila_
 _Heliophila_ is a genus of flowering plants endemic to the southern portion of Africa including the Cape Floristic and Succulent Karoo Regions.
@@
 
-###_ Heliophila_ occurrence records
+### _Heliophila_ occurrence records
 ```{r raw_GBIF}
@@
 ### Sequence data for phylogeny
-Aligned_ Heliophila_ ITS sequences were obtained from previous work by @mandakova2012whole.  Aethionema, Alliaria, Cardamine, Chamira, and Rorippa ITS records from were downloaded from Genbank.
+An alignment of ITS I and II sequences for _Heliophila_ species was obtained from the authors of @mandakova2012whole.
+Individual ITS I and II sequences for _Aethionema grandiflorum, Alliaria petiolata, Cardamine matthioli, Chamira circaeoides_, and _Rorippa amphibia_ were downloaded from Genbank.
 
@@
 ### Phylogeny construction
-Out group species were aligned together with _Heliophila_ ITS sequences using MAFFT.
-Model selection for construction of phylogeny was performed in jModeltest2 with CIPRES.
-Based on this analysis, $GTR + L$ were selected.
-Ultrametric phylogeny was estimated with branch lengths as relative time.
+Out group ( _Aethionema grandiflorum, Alliaria petiolata, Cardamine matthioli, Chamira circaeoides_, and _Rorippa amphibia_) and ingroup _Heliophila_ ITS I and II sequences were aaligned using MAFFT [@katoh2002mafft] with strategy G-INS-I, offset value 0.1, and all other options set as default.
+The $GTR + L$ model of nucleotide substitution was determined to best fit the data based on AIC using jModelTest2 [@darriba2012jmodeltest;@guindon2003a].
+A maximum clade credibility tree with branch lengths as relative time was estimated by summarizing data from six runs of 100,000,000 generations of Bayesian Markov chain Monte Carlo conducted in BEAST 2 [@bouckaert2014beast].
+Model selection and phylogenetic analyses were conducted through the CIPRES Science Gateway [@miller2010creating].
 
@@
 ```
-```{r firthmodels, echo=F,  results="hide"}
 
-firthmodelstable<-data.frame()
-
-# non-phylogenetic logistic regression firth penalized
-species_means$life_history_num<-as.numeric(species_means$life_history)-1
-row.names(species_means)<-gsub(" ", "_", species_means$species)
-seasons<-c("Winter_mean", "Spring_mean", "Summer_mean","Fall_mean")
-for(s in 1:length(seasons)){
-  fit1 <- logistf(species_means$life_history_num~(species_means[,seasons[s]]), firth = TRUE)
-  sum<-summary(fit1)
-  row.names(sum$coefficients)<-NULL
-  sum_df<-data.frame(Predictor=c("Intercept", gsub("_mean", " drought freq.", seasons[s])), 
-                     Estimate=sum$coefficients, 
-                     p.value=sum$prob)
-  row.names(sum_df)<-NULL
-  firthmodelstable<-rbind(firthmodelstable, sum_df)
-}
-
-```
 ```{r phylomodels, results="hide"}
@@
 tree<-keep.tip(tree, tip = gsub(" ","_", as.character(species_means$species)))
+
+species_means$life_history_num<-as.numeric(species_means$life_history)-1
+row.names(species_means)<-gsub(" ", "_", species_means$species)
+seasons<-c("Winter_mean", "Spring_mean", "Summer_mean","Fall_mean")
+
 phylomodelstable<-data.frame()
@@
 To account for variation in the number of occurrence records per species, we next calculated the mean drought frequency during the winter, spring, summer and fall for each species.
-The relationships between species mean values of drought frequency during each season and life habitat (annual or perennial) were tested using Firth's penalized-likelihood logistic regressions.
-Because demographic histories caused by ancestry can confound trait - environment associations, we then tested for the relationships between drought frequency and life history while controlling for relationships between species using phylogenetic logistic regressions.
+Because shared evolutionary history of closely related species can lead to spurious associations between traits and environments [@felsenstein1985phylogenies], we tested for a relationship between life history strategy and drought frequency while controlling for phylogeny using phylogenetic logistic regression [@ives2010phylogenetic].
 
@@
 (Figure \@ref(fig:mapsboxplots)g), the frequency of drought was significantly higher at the location of annuals during the winter (t = `r twin$statistic`, p = `r twin$p.value`), spring (t = `r tspr$statistic`, p = `r tspr$p.value`), summer (t = `r tsum$statistic`, p = `r tsum$p.value`), and fall (t = `r tfal$statistic`, p = `r tfal$p.value`).
-Because raw occurrence records do not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable)), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials by a Firths-penalized logistic regression.
+Because raw occurrence records do not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable)) or species relatedness (Figure \@ref(fig:phylogeny)a), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials using phylogenetic logistic regression.
 We found that the mean drought frequencies were significantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a).
-We further tested whether annual species are found in places where droughts occur more frequently while controlling for the phylogentic relatedness of _Heliophila_ species.
-This is important, because environmental differences in species distributions can be confounded with demographic history caused by ancestry.
-We found that while controlling for phylogeny, the mean drought frequencies were significantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a).
 These findings indicate that common acestry alone does not explain differences the drought frequencies experienced in the environments of annual and perennial _Heliophila_.
@@
 
-modelstable<-bind_cols(firthmodelstable,phylomodelstable)
+modelstable<-bind_cols(phylomodelstable)
 
-modelstable<- modelstable %>%
-  dplyr::select(Predictor, Estimate, p.value,Estimate1, p.value1)
+modelstable<- phylomodelstable %>%
+  dplyr::select(Predictor, Estimate, p.value)
 
-colnames(modelstable)<-c("Predictor", "Estimate`","P`","Estimate*","P*")
+colnames(modelstable)<-c("Predictor", "Estimate","P")
 
 apa_table(modelstable
-  , caption = "Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of_ Heliophila_ species the winter, spring, summer, and fall."
-  , note = "` = Firth's penalized logistic regression. * = Phylogenetically constrained logistic regression. Annual species were scored as 0 and perennial species as 1.",
+  , caption = "Phylogenetic logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species the winter, spring, summer, and fall."
+  , note = "Annual species were scored as 0 and perennial species as 1.",
   midrules=c(2,4,6,8),
@@
 
-(ref:lineplots) (a) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of_ Heliophila_.
-(phylogenetic logistic regression, \* = p < 0.05, \*\* = p < 0.01) (b) Collection dates of GBIF records of annual and perennial species of_ Heliophila_.
+(ref:lineplots) (a) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of _Heliophila_.
+(phylogenetic logistic regression, \* = p < 0.05, \*\* = p < 0.01) (b) Collection dates of GBIF records of annual and perennial species of _Heliophila_.
 
@@
 ```
-
 The preceding results indicated that annual species are found in environments where droughts are significantly more frequent, especially in the summer and fall.
@@
 Using metadata of `r nrow(GBIF_cleaned)` occurrence records and a 34 year dataset of satellite-detected droughts, we tested the prediction that annual species are more often observed in drought-prone locations than perennial species, when controlling for phylogenetic relatedness.
-We found that drought frequency is significantly different between the distributions of annual and perennial species, with annuals being found in environments with more frequent drought, and that this signal is strongest during the seasons when annuals are likely escaping via seed dormancy These results remain significant while controlling for the phylogenetic relationships of _Heliophila_ species, yielding support for the role that natural selection has played in driving contemporary distributions of these alternatives strategies in relation to drought regimens.
+We found that drought frequency is significantly different between the distributions of annual and perennial species, with annuals being found in environments with more frequent drought, and that this signal is strongest during the seasons when annuals are likely escaping via seed dormancy.
+These results remain significant while controlling for the phylogenetic relationships of _Heliophila_ species, yielding support for the role that natural selection has played in driving contemporary distributions of these alternatives strategies in relation to drought regimens.
 
 We cannot eliminate the possibility that confounding traits or environmental variables are the causative factors explaining variation in the distributions of annual and perennial species.
-Nevertheless, these results reveal quantitative support for the classic prediction that annual species are found in environemnts that experience more frequent drought than perennial species.
+Nevertheless, these results provide quantitative support for the classic prediction that annual species are found in environemnts that experience more frequent drought than perennial species.
 These findings complement previous reports of qualitative associations between annuality with enviroments characterized as having increased aridity [@evans2005climate], alternative precipitation defined habitats [@morishima1984differentiation;@datson2008climate], or greater unpredictability [@cruz2009molecular].
@@
 And while early theory predicted selection for annuality when adult morality is high [@stearns1992evolution],  we also find evidence that the transition to perenniality could be explained by the historical frequency of drought, a likely cause of mortality in plants.
-The phylogeny reveals several transitions from annual to perennial life history (Figure \@ref(fig:phylogeny)a) and that the distributions of perennial _Heliophila_ extend into regions where drought frequency is low (Figure \@ref(fig:mapsboxplots)b,Figure \@ref(fig:speciesmaps)).
+The phylogeny reveals several transitions from annual to perennial life history (Figure \@ref(fig:phylogeny)a) and that the distributions of perennial _Heliophila_ extend into regions where drought frequency is low (Figure \@ref(fig:mapsboxplots)b, Figure \@ref(fig:speciesmaps)).
 Perennials may be able to out complete annual relatives in environments where the infrequency of drought favors strategies that allow plants to benefit from growth over many seasons.
 It may also indicate that annuals rely on drought as a source of disturbance for seedling recruitment when competing with perennials [@corbin2004competition].
-Indeed, no annual species were observed in the low drought regions of eastern South Africa \@ref(fig:mapsboxplots)a,Figure \@ref(fig:speciesmaps)).
+Indeed, no annual species were observed in the low drought regions of eastern South Africa (Figure \@ref(fig:mapsboxplots), Figure \@ref(fig:speciesmaps)).
 
 These findings suggest that species with locally adaptive life history strategies could be threatened by rapidly changing drought regimens [@dai2011drought].
-This could have impacts on ecosystem functioning and processes such as carbon cycling if the composition of annual and perennial species changes as a response [@garnier1997specific; @roumet2006suites;@monroe2018ecoevolutionary].
+This could have impacts on ecosystem functioning and processes such as carbon cycling if life history traits evolve or the composition of annual and perennial species changes in response [@garnier1997specific; @roumet2006suites;@monroe2018ecoevolutionary].
 Furthermore, the frequency of drought may be an important factor when considering the use of perennial cropping systems [@parry2005prospects;@lelievre2009current].
@@
 We used `r cite_r("r-references.bib")` for all our analyses.
+\newpage
 

### 2018-12-10 manuscript/manuscript.tex
remove firth

@@
 than seed) should be rare during drought prone seasons. Phylogenetic
-relatedness can have significant non-random effects on species
-distributions and life history traits (Barrett et al., 1996), and
+relatedness can influence tests of associations between species' traits
+and their environments (Barrett et al., 1996; Felsenstein, 1985), and
 therefore we assessed the relationship between life history distribution
-and drought frequency in a phylogenetically controlled background.
+and drought frequency in a phylogenetic context.
 
@@
 for predicting intraspecific variation in drought tolerance traits and
-genes (Dittberner et al., 2018; Mojica et al., 2016; John G Monroe et
-al., 2018). Here, we accessed VHI data at \(16km^2\) resolution from
+genes (Dittberner et al., 2018; Mojica et al., 2016; J. Monroe, Powell,
+et al., 2018). Here, we accessed VHI data at \(16km^2\) resolution from
 1981 to 2015
@@
 
-\hypertarget{life-history-data-for_-heliophila_}{%
-\subsubsection{Life history data for\_
-Heliophila\_}\label{life-history-data-for_-heliophila_}}
+\hypertarget{life-history-data-for-heliophila}{%
+\subsubsection{\texorpdfstring{Life history data for
+\emph{Heliophila}}{Life history data for Heliophila}}\label{life-history-data-for-heliophila}}
 
@@
 
-\#\#\#\_ Heliophila\_ occurrence records
+\hypertarget{heliophila-occurrence-records}{%
+\subsubsection{\texorpdfstring{\emph{Heliophila} occurrence
+records}{Heliophila occurrence records}}\label{heliophila-occurrence-records}}
 
@@
 
-Aligned\_ Heliophila\_ ITS sequences were obtained from previous work by
-Mandáková et al. (2012). Aethionema, Alliaria, Cardamine, Chamira, and
-Rorippa ITS records from were downloaded from Genbank.
+An alignment of ITS I and II sequences for \emph{Heliophila} species was
+obtained from the authors of Mandáková et al. (2012). Individual ITS I
+and II sequences for \emph{Aethionema grandiflorum, Alliaria petiolata,
+Cardamine matthioli, Chamira circaeoides}, and \emph{Rorippa amphibia}
+were downloaded from Genbank.
 
@@
 
-Out group species were aligned together with \emph{Heliophila} ITS
-sequences using MAFFT. Model selection for construction of phylogeny was
-performed in jModeltest2 with CIPRES. Based on this analysis,
-\(GTR + L\) were selected. Ultrametric phylogeny was estimated with
-branch lengths as relative time.
+Out group ( \emph{Aethionema grandiflorum, Alliaria petiolata, Cardamine
+matthioli, Chamira circaeoides}, and \emph{Rorippa amphibia}) and
+ingroup \emph{Heliophila} ITS I and II sequences were aaligned using
+MAFFT (Katoh et al., 2002) with strategy G-INS-I, offset value 0.1, and
+all other options set as default. The \(GTR + L\) model of nucleotide
+substitution was determined to best fit the data based on AIC using
+jModelTest2 (Darriba et al., 2012; Guindon and Gascuel, 2003). A maximum
+clade credibility tree with branch lengths as relative time was
+estimated by summarizing data from six runs of 100,000,000 generations
+of Bayesian Markov chain Monte Carlo conducted in BEAST 2 (Bouckaert et
+al., 2014). Model selection and phylogenetic analyses were conducted
+through the CIPRES Science Gateway (Miller et al., 2010).
 
@@
 we next calculated the mean drought frequency during the winter, spring,
-summer and fall for each species. The relationships between species mean
-values of drought frequency during each season and life habitat (annual
-or perennial) were tested using Firth's penalized-likelihood logistic
-regressions. Because demographic histories caused by ancestry can
-confound trait - environment associations, we then tested for the
-relationships between drought frequency and life history while
-controlling for relationships between species using phylogenetic
-logistic regressions.
+summer and fall for each species. Because shared evolutionary history of
+closely related species can lead to spurious associations between traits
+and environments (Felsenstein, 1985), we tested for a relationship
+between life history strategy and drought frequency while controlling
+for phylogeny using phylogenetic logistic regression (Ives and Garland,
+2010).
 
@@
 records do not account for variation in the number of records per
-species (Table \ref{tab:speciesmeanstable}), we also tested whether mean
-drought frequency values of each species were significantly different
-between annuals and perennials by a Firths-penalized logistic
-regression. We found that the mean drought frequencies were
-significantly higher (\(\alpha = 0.05\)) in annual species during the
-spring, summer, and fall (Table \ref{tab:modelstable}, Figure
-\ref{fig:lineplots}a). We further tested whether annual species are
-found in places where droughts occur more frequently while controlling
-for the phylogentic relatedness of \emph{Heliophila} species. This is
-important, because environmental differences in species distributions
-can be confounded with demographic history caused by ancestry. We found
-that while controlling for phylogeny, the mean drought frequencies were
-significantly higher (\(\alpha = 0.05\)) in annual species during the
-spring, summer, and fall (Table \ref{tab:modelstable}, Figure
-\ref{fig:lineplots}a). These findings indicate that common acestry alone
-does not explain differences the drought frequencies experienced in the
-environments of annual and perennial \emph{Heliophila}.
+species (Table \ref{tab:speciesmeanstable}) or species relatedness
+(Figure \ref{fig:phylogeny}a), we also tested whether mean drought
+frequency values of each species were significantly different between
+annuals and perennials using phylogenetic logistic regression. We found
+that the mean drought frequencies were significantly higher
+(\(\alpha = 0.05\)) in annual species during the spring, summer, and
+fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a). These
+findings indicate that common acestry alone does not explain differences
+the drought frequencies experienced in the environments of annual and
+perennial \emph{Heliophila}.
 
@@
 \begin{threeparttable}
-\caption{\label{tab:modelstable}Logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of\_ Heliophila\_ species the winter, spring, summer, and fall.}
-\begin{tabular}{lllll}
+\caption{\label{tab:modelstable}Phylogenetic logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species the winter, spring, summer, and fall.}
+\begin{tabular}{lll}
 \toprule
-Predictor & \multicolumn{1}{c}{Estimate`} & \multicolumn{1}{c}{P`} & \multicolumn{1}{c}{Estimate*} & \multicolumn{1}{c}{P*}\\
+Predictor & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{P}\\
 \midrule
-Intercept & 2.2575 & 0.1739 & 0.7231 & 0.6636\\
-Winter drought freq. & -6.7484 & 0.1661 & -1.5452 & 0.7274\\ \midrule
-Intercept & 4.5594 & 0.0443 & 5.0107 & 0.0534\\
-Spring drought freq. & -11.7895 & 0.0423 & -12.9014 & 0.0464\\ \midrule
-Intercept & 7.1742 & 0.0011 & 7.7093 & 0.0054\\
-Summer drought freq. & -18.2999 & 0.0010 & -19.9056 & 0.0042\\ \midrule
-Intercept & 6.4226 & 0.0029 & 7.0162 & 0.0082\\
-Fall drought freq. & -19.0512 & 0.0026 & -20.8174 & 0.0067\\ \midrule
+Intercept & 0.7231 & 0.6636\\
+Winter drought freq. & -1.5452 & 0.7274\\ \midrule
+Intercept & 5.0107 & 0.0534\\
+Spring drought freq. & -12.9014 & 0.0464\\ \midrule
+Intercept & 7.7093 & 0.0054\\
+Summer drought freq. & -19.9056 & 0.0042\\ \midrule
+Intercept & 7.0162 & 0.0082\\
+Fall drought freq. & -20.8174 & 0.0067\\ \midrule
 \bottomrule
@@
 \begin{tablenotes}[para]
-\normalsize{\textit{Note.} ` = Firth's penalized logistic regression. * = Phylogenetically constrained logistic regression. Annual species were scored as 0 and perennial species as 1.}
+\normalsize{\textit{Note.} Annual species were scored as 0 and perennial species as 1.}
 \end{tablenotes}
@@
 \includegraphics[width=\textwidth]{../figures/line_and_dates} \caption{(a) Comparison (mean +- SE) of drought frequency across
-seasons measured at the GBIF records of annual and perennial species
-of\_ Heliophila\_. (phylogenetic logistic regression, * = p \textless{}
+seasons measured at the GBIF records of annual and perennial species of
+\emph{Heliophila}. (phylogenetic logistic regression, * = p \textless{}
 0.05, ** = p \textless{} 0.01) (b) Collection dates of GBIF records of
-annual and perennial species of\_ Heliophila\_.}\label{fig:lineplots}
+annual and perennial species of \emph{Heliophila}.}\label{fig:lineplots}
 \end{figure}
@@
 signal is strongest during the seasons when annuals are likely escaping
-via seed dormancy These results remain significant while controlling for
-the phylogenetic relationships of \emph{Heliophila} species, yielding
-support for the role that natural selection has played in driving
-contemporary distributions of these alternatives strategies in relation
-to drought regimens.
+via seed dormancy. These results remain significant while controlling
+for the phylogenetic relationships of \emph{Heliophila} species,
+yielding support for the role that natural selection has played in
+driving contemporary distributions of these alternatives strategies in
+relation to drought regimens.
 
@@
 in the distributions of annual and perennial species. Nevertheless,
-these results reveal quantitative support for the classic prediction
+these results provide quantitative support for the classic prediction
 that annual species are found in environemnts that experience more
@@
 perennial \emph{Heliophila} extend into regions where drought frequency
-is low (Figure \ref{fig:mapsboxplots}b,Figure \ref{fig:speciesmaps}).
+is low (Figure \ref{fig:mapsboxplots}b, Figure \ref{fig:speciesmaps}).
 Perennials may be able to out complete annual relatives in environments
@@
 annual species were observed in the low drought regions of eastern South
-Africa \ref{fig:mapsboxplots}a,Figure \ref{fig:speciesmaps}).
+Africa (Figure \ref{fig:mapsboxplots}, Figure \ref{fig:speciesmaps}).
 
@@
 (Dai, 2011). This could have impacts on ecosystem functioning and
-processes such as carbon cycling if the composition of annual and
-perennial species changes as a response (Garnier et al., 1997; J Grey
-Monroe et al., 2018; Roumet et al., 2006). Furthermore, the frequency of
-drought may be an important factor when considering the use of perennial
-cropping systems (Lelièvre and Volaire, 2009; Parry et al., 2005).
+processes such as carbon cycling if life history traits evolve or the
+composition of annual and perennial species changes in response (Garnier
+et al., 1997; J. Monroe, Markman, et al., 2018; Roumet et al., 2006).
+Furthermore, the frequency of drought may be an important factor when
+considering the use of perennial cropping systems (Lelièvre and Volaire,
+2009; Parry et al., 2005).
 
@@
 
+\leavevmode\hypertarget{ref-bouckaert2014beast}{}%
+Bouckaert R, Heled J, Kühnert D, Vaughan T, Wu C-H, Xie D, Suchard M,
+Rambaut A, Drummond A. 2014. BEAST 2: A software platform for bayesian
+evolutionary analysis. \emph{PLoS Computational Biology}
+\textbf{10}:doi:10.1371/journal.pcbi.1003537.
+
 \leavevmode\hypertarget{ref-brown1996oldlist}{}%
@@
 
+\leavevmode\hypertarget{ref-darriba2012jmodeltest}{}%
+Darriba D, Taboada G, Doallo R, Posada D. 2012. JModelTest 2: More
+models, new heuristics and parallel computing. \emph{Nature Methods}
+\textbf{9}:772.
+
 \leavevmode\hypertarget{ref-daru2018widespread}{}%
@@
 
+\leavevmode\hypertarget{ref-felsenstein1985phylogenies}{}%
+Felsenstein J. 1985. Phylogenies and the comparative method.
+\emph{American Naturalist} \textbf{125}:1--15.
+
 \leavevmode\hypertarget{ref-franco1996life}{}%
@@
 
+\leavevmode\hypertarget{ref-guindon2003a}{}%
+Guindon S, Gascuel O. 2003. A simple, fast and accurate method to
+estimate large phylogenies by maximum-likelihood. \emph{Systematic
+Biology} \textbf{52}:696--704.
+
 \leavevmode\hypertarget{ref-R-MCMCglmm}{}%
@@
 \textbf{63}:397--408.
+
+\leavevmode\hypertarget{ref-ives2010phylogenetic}{}%
+Ives A, Garland T. 2010. CPhylogenetic logistic regression for binary
+dependent variables. \emph{Systematic Biology} \textbf{59}:9--26.
+
+\leavevmode\hypertarget{ref-katoh2002mafft}{}%
+Katoh, Misawa, Kuma, Miyata. 2002. MAFFT: A novel method for rapid
+multiple sequence alignment based on fast fourier transform.
+\emph{Nucleic Acids Research} \textbf{30}:3059--3066.
 
@@
 
+\leavevmode\hypertarget{ref-miller2010creating}{}%
+Miller M, Pfeiffer W, Schwartz T and. 2010. Creating the cipres science
+gateway for inference of large phylogenetic trees. \emph{Proceedings of
+the Gateway Computing Environments Workshop} 1--8.
+
 \leavevmode\hypertarget{ref-Mojica2016}{}%
@@
 \leavevmode\hypertarget{ref-monroe2018ecoevolutionary}{}%
-Monroe JG, Markman DW, Beck WS, Felton AJ, Vahsen ML, Pressler Y. 2018.
+Monroe J, Markman D, Beck W, Felton A, Vahsen M, Pressler Y. 2018.
 Ecoevolutionary dynamics of carbon cycling in the anthropocene.
@@
 \leavevmode\hypertarget{ref-monroe2018drought}{}%
-Monroe JG, Powell T, Price N, Mullen J, Howard A, Evans K, Lovell J,
+Monroe J, Powell T, Price N, Mullen J, Howard A, Evans K, Lovell J,
 McKay J. 2018. Drought adaptation in nature by extensive genetic
-loss-of-function. \emph{bioRxiv} 372854.
+loss-of-function. \emph{eLife} DOI: 10.7554/eLife.41038.
 
@@
 and Henry, 2018), and \emph{tidyverse} (Version 1.2.1; Wickham, 2017)
-for all our analyses.
+for all our analyses. \newpage
 

### 2018-12-11 manuscript/manuscript.Rmd
update citation style

@@
 abstract: Explaining variation in life history strategies is a long-standing goal of evolutionary biology.
-For plants, annual and perennial life histories are thought to reflect adaptation to environments that differ in the frequency of environmental stress such as drought.
+For plants, annual and perennial life histories are thought to reflect adaptation to environments that differ in the frequency of stress events such as drought.
 Here we test this hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa, by integrating 34 years of satellite-based drought measurements with 2192 herbaria occurrence records.
@@
 These associations are predictive while controlling for phylogeny, lending support to the hypothesis that drought related natural selection has influenced the distributions of these strategies.
-Additionally, the difference in drought frequency between annual and perennial species distributions is greatest during the summer and fall, which also appears to be when annuals are in the seed phase of their life cycle based on collection dates of annual species.
+Additionally, the collection dates of annual and perennial species indicate that annuals escape drought prone seasons during the seed phase of their life cycle.
 Together, these finding provide empirical support for classic hypotheses about the drivers of life history strategy in plants - that perennials out compete annuals in environments with less frequent drought and that annuals are adapted to environments with more frequent drought by escaping drought prone seasons as seeds.
@@
 ```
-
 # Introduction
@@
 Out group ( _Aethionema grandiflorum, Alliaria petiolata, Cardamine matthioli, Chamira circaeoides_, and _Rorippa amphibia_) and ingroup _Heliophila_ ITS I and II sequences were aaligned using MAFFT [@katoh2002mafft] with strategy G-INS-I, offset value 0.1, and all other options set as default.
-The $GTR + L$ model of nucleotide substitution was determined to best fit the data based on AIC using jModelTest2 [@darriba2012jmodeltest;@guindon2003a].
+The $GTR + \Gamma$ model of nucleotide substitution was determined to best fit the data based on AIC using jModelTest2 [@darriba2012jmodeltest;@guindon2003a].
 A maximum clade credibility tree with branch lengths as relative time was estimated by summarizing data from six runs of 100,000,000 generations of Bayesian Markov chain Monte Carlo conducted in BEAST 2 [@bouckaert2014beast].
@@
 To test the hypothesis that annual species have adapted to escape drought prone seasons as seeds, collection dates for herbarium specimens were compared between annual and perennial species.
-Comparisons of distributions were made by Two-sample Kolmogorov-Smirnov test, t-test, and Barlett variance test.
+Comparisons of distributions were made by Two-sample Kolmogorov-Smirnov test and Barlett variance test.
 

### 2018-12-11 manuscript/manuscript.tex
update citation style

@@
 are thought to reflect adaptation to environments that differ in the
-frequency of environmental stress such as drought. Here we test this
-hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of
-flowering plants native to Africa, by integrating 34 years of
-satellite-based drought measurements with 2192 herbaria occurrence
-records. Consistent with predictions from classic life history theory,
-we find that perennial \textit{Heliophila} species occur in environments
-where droughts are significantly less frequent compared to annuals.
-These associations are predictive while controlling for phylogeny,
-lending support to the hypothesis that drought related natural selection
-has influenced the distributions of these strategies. Additionally, the
-difference in drought frequency between annual and perennial species
-distributions is greatest during the summer and fall, which also appears
-to be when annuals are in the seed phase of their life cycle based on
-collection dates of annual species. Together, these finding provide
-empirical support for classic hypotheses about the drivers of life
-history strategy in plants - that perennials out compete annuals in
-environments with less frequent drought and that annuals are adapted to
-environments with more frequent drought by escaping drought prone
-seasons as seeds. 
+frequency of stress events such as drought. Here we test this hypothesis
+in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering
+plants native to Africa, by integrating 34 years of satellite-based
+drought measurements with 2192 herbaria occurrence records. Consistent
+with predictions from classic life history theory, we find that
+perennial \textit{Heliophila} species occur in environments where
+droughts are significantly less frequent compared to annuals. These
+associations are predictive while controlling for phylogeny, lending
+support to the hypothesis that drought related natural selection has
+influenced the distributions of these strategies. Additionally, the
+collection dates of annual and perennial species indicate that annuals
+escape drought prone seasons during the seed phase of their life cycle.
+Together, these finding provide empirical support for classic hypotheses
+about the drivers of life history strategy in plants - that perennials
+out compete annuals in environments with less frequent drought and that
+annuals are adapted to environments with more frequent drought by
+escaping drought prone seasons as seeds. 
 }
@@
 ecological factors that explain their evolution and distributions remain
-empirically uresolved (Friedman and Rubin, 2015).\\
+empirically uresolved (Friedman \& Rubin, 2015).\\
 Classical theory predict shorter life spans in environments where adult
-mortality is high (Charnov and Schaffer, 1973; Franco and Silvertown,
-1996; Stearns, 1992). In plants, this has been extended to the
-hypothesis that annuality is adaptive when it allows plants to escape
-drought (Schaffer and Gadgil, 1975). Lack of water is perhaps the
-greatest threat to survival during vegetative or reproductive growth and
-annuals can remain dormant (and protected as a seed) during drought.
-Thus, environments with greater seasonal drought frequency may select
-for annual life histories that complete reproduction prior to drought
-prone seasons. Conversely, environments with less frequent drought may
-select for perennial species, which may benefit from multiple bouts of
+mortality is high (Charnov \& Schaffer, 1973; Stearns, 1992; Franco \&
+Silvertown, 1996). In plants, this has been extended to the hypothesis
+that annuality is adaptive when it allows plants to escape drought
+(Schaffer \& Gadgil, 1975). Lack of water is perhaps the greatest threat
+to survival during vegetative or reproductive growth and annuals can
+remain dormant (and protected as a seed) during drought. Thus,
+environments with greater seasonal drought frequency may select for
+annual life histories that complete reproduction prior to drought prone
+seasons. Conversely, environments with less frequent drought may select
+for perennial species, which may benefit from multiple bouts of
 reproduction and competitive advantage by preventing recruitment of
-annual species (Corbin and D'Antonio, 2004). These predictions have been
+annual species (Corbin \& D'Antonio, 2004). These predictions have been
 supported by the observation of annuals in arid environments in
-\emph{Oryza perennis} (Morishima et al., 1984) and \emph{Oenothera}
-(Evans et al., 2005). Additionally, annual and perennial species of
-\emph{Nemesia} were qualitatively associated with winter rather and
-summer rainfall environments respectively (Datson et al., 2008) and
-annual species of \emph{Scorzoneroides} were associated with
-environments classified as unpredictable (Cruz-Mazo et al., 2009).
-However, whether the history frequency of drought events indeed predicts
-the distributions annual or perennial life history strategies has yet to
-be tested.\\
+\emph{Oryza perennis} (Morishima \emph{et al.}, 1984) and
+\emph{Oenothera} (Evans \emph{et al.}, 2005). Additionally, annual and
+perennial species of \emph{Nemesia} were qualitatively associated with
+winter rather and summer rainfall environments respectively (Datson
+\emph{et al.}, 2008) and annual species of \emph{Scorzoneroides} were
+associated with environments classified as unpredictable (Cruz-Mazo
+\emph{et al.}, 2009). However, whether the history frequency of drought
+events indeed predicts the distributions annual or perennial life
+history strategies has yet to be tested.\\
 Here we combine a long-term global dataset of satellite detected drought
@@
 relatedness can influence tests of associations between species' traits
-and their environments (Barrett et al., 1996; Felsenstein, 1985), and
-therefore we assessed the relationship between life history distribution
-and drought frequency in a phylogenetic context.
+and their environments (Felsenstein, 1985; Barrett \emph{et al.}, 1996),
+and therefore we assessed the relationship between life history
+distribution and drought frequency in a phylogenetic context.
 
@@
 scope and resolution than weather stations or field observations
-(AghaKouchak et al., 2015). To quantify the frequency of drought during
-different seasons across landscapes, we used the remotely sensed
+(AghaKouchak \emph{et al.}, 2015). To quantify the frequency of drought
+during different seasons across landscapes, we used the remotely sensed
 Vegetative Health Index (VHI), which measures landscape scale reductions
@@
 validated and generally used for evaluating drought risk and predicting
-crop yields in agriculture (e.g., Rojas et al., 2011; Kogan et al.,
-2016). But they also present a new tool to study seasonal patterns in
-the frequency of drought across environments and to test hypotheses
-about the effect of drought on ecological and evolutionary processes
-(Kerr and Ostrovsky, 2003). As such, the VHI has been applied recently
-to study drought related ecology of natural species and proven useful
-for predicting intraspecific variation in drought tolerance traits and
-genes (Dittberner et al., 2018; Mojica et al., 2016; J. Monroe, Powell,
-et al., 2018). Here, we accessed VHI data at \(16km^2\) resolution from
-1981 to 2015
+crop yields in agriculture (e.g., Rojas \emph{et al.}, 2011; Kogan
+\emph{et al.}, 2016). But they also present a new tool to study seasonal
+patterns in the frequency of drought across environments and to test
+hypotheses about the effect of drought on ecological and evolutionary
+processes (Kerr \& Ostrovsky, 2003). As such, the VHI has been applied
+recently to study drought related ecology of natural species and proven
+useful for predicting intraspecific variation in drought tolerance
+traits and genes (Mojica \emph{et al.}, 2016; Dittberner \emph{et al.},
+2018; Monroe \emph{et al.}, 2018b). Here, we accessed VHI data at
+\(16km^2\) resolution from 1981 to 2015
 (\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php}) to
@@
 Earth and the \emph{Heliophila} species occurring there are considered
-to make up the most diverse genus of the family Brassicaceae (Mandáková
-et al., 2012; Mummenhoff et al., 2005). This genus includes both
-perennial and annual species and this change in life history strategy
-has likely arisen multiple independent times (Appel and Al-Shehbaz,
-1997; Mummenhoff et al., 2005). Furthermore, the fine scale climatic
-heterogeneity of Southern Africa is ideal for studying the distribution
-of traits in relation to environmental parameters (Sayre et al., 2013).
-We used life histories reported by Mummenhoff et al. (2005), grouping
-species with annual or perennial life histories. Perenniality was
-defined based any form of perennial life history (e.g., herbs, shrubs,
-mixed, etc). Because the nature of species reported with mixed traits
-were unknown (i.e.~plasticity vs.~genetic variation), we classified
-these species here as perennial since they can maintain vegetative
-growth after reproduction at least to some capacity.
+to make up the most diverse genus of the family Brassicaceae (Mummenhoff
+\emph{et al.}, 2005; Mandáková \emph{et al.}, 2012). This genus includes
+both perennial and annual species and this change in life history
+strategy has likely arisen multiple independent times (Appel \&
+Al-Shehbaz, 1997; Mummenhoff \emph{et al.}, 2005). Furthermore, the fine
+scale climatic heterogeneity of Southern Africa is ideal for studying
+the distribution of traits in relation to environmental parameters
+(Sayre \emph{et al.}, 2013). We used life histories reported by
+Mummenhoff \emph{et al.} (2005), grouping species with annual or
+perennial life histories. Perenniality was defined based any form of
+perennial life history (e.g., herbs, shrubs, mixed, etc). Because the
+nature of species reported with mixed traits were unknown
+(i.e.~plasticity vs.~genetic variation), we classified these species
+here as perennial since they can maintain vegetative growth after
+reproduction at least to some capacity.
 
@@
 their associated metadata have been used since the 1960s to study
-species' geographical distributions (reviewed by Willis et al. (2017)
-and Lang et al. (2018)). And as they become digitized (Soltis, 2017),
-these collections have been used to study relationships between trait
-distributions, geography, and climate (Davis et al., 2015; Stropp et
-al., 2016; Václavı'k et al., 2017; Wolf et al., 2016). To characterize
-the distributions of annual and perennial \emph{Heliophila} species, all
-records for the genus \emph{Heliophila} were downloaded from the Global
-Biodiversity Information Facility (gbif.org) on July 21, 2018 (GBIF,
-2018).
+species' geographical distributions (reviewed by Willis \emph{et al.}
+(2017) and Lang \emph{et al.} (2018)). And as they become digitized
+(Soltis, 2017), these collections have been used to study relationships
+between trait distributions, geography, and climate (Davis \emph{et
+al.}, 2015; Stropp \emph{et al.}, 2016; Wolf \emph{et al.}, 2016;
+Václavı'k \emph{et al.}, 2017). To characterize the distributions of
+annual and perennial \emph{Heliophila} species, all records for the
+genus \emph{Heliophila} were downloaded from the Global Biodiversity
+Information Facility (gbif.org) on July 21, 2018 (GBIF, 2018).
 
@@
 An alignment of ITS I and II sequences for \emph{Heliophila} species was
-obtained from the authors of Mandáková et al. (2012). Individual ITS I
-and II sequences for \emph{Aethionema grandiflorum, Alliaria petiolata,
-Cardamine matthioli, Chamira circaeoides}, and \emph{Rorippa amphibia}
-were downloaded from Genbank.
+obtained from the authors of Mandáková \emph{et al.} (2012). Individual
+ITS I and II sequences for \emph{Aethionema grandiflorum, Alliaria
+petiolata, Cardamine matthioli, Chamira circaeoides}, and \emph{Rorippa
+amphibia} were downloaded from Genbank.
 
@@
 ingroup \emph{Heliophila} ITS I and II sequences were aaligned using
-MAFFT (Katoh et al., 2002) with strategy G-INS-I, offset value 0.1, and
-all other options set as default. The \(GTR + L\) model of nucleotide
-substitution was determined to best fit the data based on AIC using
-jModelTest2 (Darriba et al., 2012; Guindon and Gascuel, 2003). A maximum
-clade credibility tree with branch lengths as relative time was
-estimated by summarizing data from six runs of 100,000,000 generations
-of Bayesian Markov chain Monte Carlo conducted in BEAST 2 (Bouckaert et
-al., 2014). Model selection and phylogenetic analyses were conducted
-through the CIPRES Science Gateway (Miller et al., 2010).
+MAFFT (Katoh \emph{et al.}, 2002) with strategy G-INS-I, offset value
+0.1, and all other options set as default. The \(GTR + \Gamma\) model of
+nucleotide substitution was determined to best fit the data based on AIC
+using jModelTest2 (Guindon \& Gascuel, 2003; Darriba \emph{et al.},
+2012). A maximum clade credibility tree with branch lengths as relative
+time was estimated by summarizing data from six runs of 100,000,000
+generations of Bayesian Markov chain Monte Carlo conducted in BEAST 2
+(Bouckaert \emph{et al.}, 2014). Model selection and phylogenetic
+analyses were conducted through the CIPRES Science Gateway (Miller
+\emph{et al.}, 2010).
 
@@
 between life history strategy and drought frequency while controlling
-for phylogeny using phylogenetic logistic regression (Ives and Garland,
+for phylogeny using phylogenetic logistic regression (Ives \& Garland,
 2010).
@@
 were compared between annual and perennial species. Comparisons of
-distributions were made by Two-sample Kolmogorov-Smirnov test, t-test,
-and Barlett variance test.
+distributions were made by Two-sample Kolmogorov-Smirnov test and
+Barlett variance test.
 
@@
 Out of 8670 \emph{Heliophila} GBIF records, 6634 were for species with
-reported life history (Mummenhoff et al., 2005), 2856 had geospatial
-data, 2833 did not have geospatial issues, 2684 were located on pixels
-classified as land having drought measurements, 2543 were located in
-Africa, 2192 were not duplicated. Thus, after all filtering steps, 2192
-records for 42 species (Figure \ref{fig:phylogeny}, Table
+reported life history (Mummenhoff \emph{et al.}, 2005), 2856 had
+geospatial data, 2833 did not have geospatial issues, 2684 were located
+on pixels classified as land having drought measurements, 2543 were
+located in Africa, 2192 were not duplicated. Thus, after all filtering
+steps, 2192 records for 42 species (Figure \ref{fig:phylogeny}, Table
 \ref{tab:speciesmeanstable}) passed for further analyses. The number of
@@
 heterogeneity is expected, given that this is one of the most
-climatically diverse regions of the Earth (Sayre et al., 2013). It is
-worth noting the east to west cline in drought frequency observed during
-the summer, which distinguishes the high drought frequency of the
+climatically diverse regions of the Earth (Sayre \emph{et al.}, 2013).
+It is worth noting the east to west cline in drought frequency observed
+during the summer, which distinguishes the high drought frequency of the
 Kalahari Sands and Namid Desert phytogeographic regions from the low
@@
 previous reports of qualitative associations between annuality with
-enviroments characterized as having increased aridity (Evans et al.,
-2005), alternative precipitation defined habitats (Datson et al., 2008;
-Morishima et al., 1984), or greater unpredictability (Cruz-Mazo et al.,
-2009). However, to our knowledge this is the first study to demonstrate
-a significant association between life history and drought in a
-phylogenetic context informed by large scale species distribution data
-and long term drought measures.
+enviroments characterized as having increased aridity (Evans \emph{et
+al.}, 2005), alternative precipitation defined habitats (Morishima
+\emph{et al.}, 1984; Datson \emph{et al.}, 2008), or greater
+unpredictability (Cruz-Mazo \emph{et al.}, 2009). However, to our
+knowledge this is the first study to demonstrate a significant
+association between life history and drought in a phylogenetic context
+informed by large scale species distribution data and long term drought
+measures.
 
@@
 bias toward collecting near roads or near the locations of natural
-history collections (Daru et al., 2018). Future research will benefit
-from systematic sampling efforts to avoid these noted biases. However,
-the ecosystems of southern Africa include several biodiversity hotspots
-and are among the most botanically well sampled regions on Earth (Daru
-et al., 2018), suggesting that this may currently be the optimal region
-for our analyses of life history distribution. Indeed, we were able to
-use 2192 occurrence records to study 42 species, which represents a
-significant advance over relying on personal observations to
-characterize species distributions.
+history collections (Daru \emph{et al.}, 2018). Future research will
+benefit from systematic sampling efforts to avoid these noted biases.
+However, the ecosystems of southern Africa include several biodiversity
+hotspots and are among the most botanically well sampled regions on
+Earth (Daru \emph{et al.}, 2018), suggesting that this may currently be
+the optimal region for our analyses of life history distribution.
+Indeed, we were able to use 2192 occurrence records to study 42 species,
+which represents a significant advance over relying on personal
+observations to characterize species distributions.
 
@@
 frequent. While most previous work has focused on describing the
-evolutionary origins of annuality (Andreasen and Baldwin, 2001; Barrett
-et al., 1996; Conti et al., 1999; Friedman and Rubin, 2015; Verboom et
-al., 2004) there are at least a few other cases where perenniality
-appears to have arisen from an annual ancestor (Bena et al., 1998; Tank
-and Olmstead, 2008). And while early theory predicted selection for
-annuality when adult morality is high (Stearns, 1992), we also find
-evidence that the transition to perenniality could be explained by the
-historical frequency of drought, a likely cause of mortality in plants.
-The phylogeny reveals several transitions from annual to perennial life
-history (Figure \ref{fig:phylogeny}a) and that the distributions of
-perennial \emph{Heliophila} extend into regions where drought frequency
-is low (Figure \ref{fig:mapsboxplots}b, Figure \ref{fig:speciesmaps}).
-Perennials may be able to out complete annual relatives in environments
-where the infrequency of drought favors strategies that allow plants to
-benefit from growth over many seasons. It may also indicate that annuals
-rely on drought as a source of disturbance for seedling recruitment when
-competing with perennials (Corbin and D'Antonio, 2004). Indeed, no
-annual species were observed in the low drought regions of eastern South
-Africa (Figure \ref{fig:mapsboxplots}, Figure \ref{fig:speciesmaps}).
+evolutionary origins of annuality (Barrett \emph{et al.}, 1996; Conti
+\emph{et al.}, 1999; Andreasen \& Baldwin, 2001; Verboom \emph{et al.},
+2004; Friedman \& Rubin, 2015) there are at least a few other cases
+where perenniality appears to have arisen from an annual ancestor (Bena
+\emph{et al.}, 1998; Tank \& Olmstead, 2008). And while early theory
+predicted selection for annuality when adult morality is high (Stearns,
+1992), we also find evidence that the transition to perenniality could
+be explained by the historical frequency of drought, a likely cause of
+mortality in plants. The phylogeny reveals several transitions from
+annual to perennial life history (Figure \ref{fig:phylogeny}a) and that
+the distributions of perennial \emph{Heliophila} extend into regions
+where drought frequency is low (Figure \ref{fig:mapsboxplots}b, Figure
+\ref{fig:speciesmaps}). Perennials may be able to out complete annual
+relatives in environments where the infrequency of drought favors
+strategies that allow plants to benefit from growth over many seasons.
+It may also indicate that annuals rely on drought as a source of
+disturbance for seedling recruitment when competing with perennials
+(Corbin \& D'Antonio, 2004). Indeed, no annual species were observed in
+the low drought regions of eastern South Africa (Figure
+\ref{fig:mapsboxplots}, Figure \ref{fig:speciesmaps}).
 
@@
 composition of annual and perennial species changes in response (Garnier
-et al., 1997; J. Monroe, Markman, et al., 2018; Roumet et al., 2006).
-Furthermore, the frequency of drought may be an important factor when
-considering the use of perennial cropping systems (Lelièvre and Volaire,
-2009; Parry et al., 2005).
+\emph{et al.}, 1997; Roumet \emph{et al.}, 2006; Monroe \emph{et al.},
+2018a). Furthermore, the frequency of drought may be an important factor
+when considering the use of perennial cropping systems (Parry \emph{et
+al.}, 2005; Lelièvre \& Volaire, 2009).
 
@@
 \leavevmode\hypertarget{ref-aghakouchak2015remote}{}%
-AghaKouchak A, Farahmand A, Melton F, Teixeira J, Anderson M, Wardlow
-BD, Hain C. 2015. Remote sensing of drought: Progress, challenges and
-opportunities. \emph{Reviews of Geophysics} \textbf{53}:452--480.
+\textbf{\textnormal{AghaKouchak A}, \textnormal{Farahmand A},
+\textnormal{Melton F}, \textnormal{Teixeira J}, \textnormal{Anderson M},
+\textnormal{Wardlow BD}, \textnormal{Hain C}}. \textbf{2015}. Remote
+sensing of drought: Progress, challenges and opportunities.
+\emph{Reviews of Geophysics} \textbf{53}: 452--480.
 
 \leavevmode\hypertarget{ref-R-geiger_a}{}%
-Alfaro M, Santini F, Brock C, Alamillo H, Dornburg A, Rabosky D,
-Carnevale G, Harmon L. 2009. Nine exceptional radiations plus high
-turnover explain species diversity in jawed vertebrates.
-\emph{Proceedings of the National Academy of Sciences of the United
-States of America} \textbf{106}:13410--13414.
+\textbf{\textnormal{Alfaro M}, \textnormal{Santini F}, \textnormal{Brock
+C}, \textnormal{Alamillo H}, \textnormal{Dornburg A},
+\textnormal{Rabosky D}, \textnormal{Carnevale G}, \textnormal{Harmon
+L}}. \textbf{2009}. Nine exceptional radiations plus high turnover
+explain species diversity in jawed vertebrates. \emph{Proceedings of the
+National Academy of Sciences of the United States of America}
+\textbf{106}: 13410--13414.
 
 \leavevmode\hypertarget{ref-andreasen2001unequal}{}%
-Andreasen K, Baldwin BG. 2001. Unequal evolutionary rates between annual
-and perennial lineages of checker mallows (sidalcea, malvaceae):
-Evidence from 18S--26S rDNA internal and external transcribed spacers.
-\emph{Molecular Biology and Evolution} \textbf{18}:936--944.
+\textbf{\textnormal{Andreasen K}, \textnormal{Baldwin BG}}.
+\textbf{2001}. Unequal evolutionary rates between annual and perennial
+lineages of checker mallows (sidalcea, malvaceae): Evidence from
+18S--26S rDNA internal and external transcribed spacers. \emph{Molecular
+Biology and Evolution} \textbf{18}: 936--944.
 
 \leavevmode\hypertarget{ref-appel1997generic}{}%
-Appel O, Al-Shehbaz IA. 1997. Generic limits and taxonomy of hornungia,
-pritzelago, and hymenolobus (brassicaceae). \emph{Novon} 338--340.
+\textbf{\textnormal{Appel O}, \textnormal{Al-Shehbaz IA}}.
+\textbf{1997}. Generic limits and taxonomy of hornungia, pritzelago, and
+hymenolobus (brassicaceae). \emph{Novon}: 338--340.
 
 \leavevmode\hypertarget{ref-R-papaja}{}%
-Aust F, Barth M. 2018. papaja: Create APA manuscripts with R Markdown.
+\textbf{\textnormal{Aust F}, \textnormal{Barth M}}. \textbf{2018}.
+\emph{papaja: Create APA manuscripts with R Markdown}.
 
 \leavevmode\hypertarget{ref-barrett1996comparative}{}%
-Barrett SCH, Harder LD, Worley AC. 1996. The comparative biology of
-pollination and mating in flowering plants. \emph{Phil Trans R Soc Lond
-B} \textbf{351}:1271--1280.
+\textbf{\textnormal{Barrett SCH}, \textnormal{Harder LD},
+\textnormal{Worley AC}}. \textbf{1996}. The comparative biology of
+pollination and mating in flowering plants. \emph{Phil. Trans. R. Soc.
+Lond. B} \textbf{351}: 1271--1280.
 
 \leavevmode\hypertarget{ref-R-Matrix}{}%
-Bates D, Maechler M. 2018. Matrix: Sparse and dense matrix classes and
-methods.
+\textbf{\textnormal{Bates D}, \textnormal{Maechler M}}. \textbf{2018}.
+\emph{Matrix: Sparse and dense matrix classes and methods}.
 
 \leavevmode\hypertarget{ref-bena1998molecular}{}%
-Bena G, Lejeune B, Prosperi J-M, Olivieri I. 1998. Molecular
-phylogenetic approach for studying life-history evolution: The ambiguous
-example of the genus medicago l. \emph{Proceedings of the Royal Society
-of London B: Biological Sciences} \textbf{265}:1141--1151.
+\textbf{\textnormal{Bena G}, \textnormal{Lejeune B},
+\textnormal{Prosperi J-M}, \textnormal{Olivieri I}}. \textbf{1998}.
+Molecular phylogenetic approach for studying life-history evolution: The
+ambiguous example of the genus medicago l. \emph{Proceedings of the
+Royal Society of London B: Biological Sciences} \textbf{265}:
+1141--1151.
 
 \leavevmode\hypertarget{ref-bouckaert2014beast}{}%
-Bouckaert R, Heled J, Kühnert D, Vaughan T, Wu C-H, Xie D, Suchard M,
-Rambaut A, Drummond A. 2014. BEAST 2: A software platform for bayesian
-evolutionary analysis. \emph{PLoS Computational Biology}
-\textbf{10}:doi:10.1371/journal.pcbi.1003537.
+\textbf{\textnormal{Bouckaert R}, \textnormal{Heled J},
+\textnormal{Kühnert D}, \textnormal{Vaughan T}, \textnormal{Wu C-H},
+\textnormal{Xie D}, \textnormal{Suchard M}, \textnormal{Rambaut A},
+\textnormal{Drummond A}}. \textbf{2014}. BEAST 2: A software platform
+for bayesian evolutionary analysis. \emph{PLoS Computational Biology}
+\textbf{10}: doi:10.1371/journal.pcbi.1003537.
 
 \leavevmode\hypertarget{ref-brown1996oldlist}{}%
-Brown PM. 1996. OLDLIST: A database of maximum tree ages. \emph{Tree
-rings, environment, and humanity Radiocarbon} \textbf{1996}:727--731.
+\textbf{\textnormal{Brown PM}}. \textbf{1996}. OLDLIST: A database of
+maximum tree ages. \emph{Tree rings, environment, and humanity.
+Radiocarbon} \textbf{1996}: 727--731.
 
 \leavevmode\hypertarget{ref-charnov1973life}{}%
-Charnov EL, Schaffer WM. 1973. Life-history consequences of natural
-selection: Cole's result revisited. \emph{The American Naturalist}
-\textbf{107}:791--793.
+\textbf{\textnormal{Charnov EL}, \textnormal{Schaffer WM}}.
+\textbf{1973}. Life-history consequences of natural selection: Cole's
+result revisited. \emph{The American Naturalist} \textbf{107}: 791--793.
 
 \leavevmode\hypertarget{ref-cole1954population}{}%
-Cole LC. 1954. The population consequences of life history phenomena.
-\emph{The Quarterly Review of Biology} \textbf{29}:103--137.
+\textbf{\textnormal{Cole LC}}. \textbf{1954}. The population
+consequences of life history phenomena. \emph{The Quarterly Review of
+Biology} \textbf{29}: 103--137.
 
 \leavevmode\hypertarget{ref-conti1999phylogenetic}{}%
-Conti E, Soltis DE, Hardig TM, Schneider J. 1999. Phylogenetic
+\textbf{\textnormal{Conti E}, \textnormal{Soltis DE}, \textnormal{Hardig
+TM}, \textnormal{Schneider J}}. \textbf{1999}. Phylogenetic
 relationships of the silver saxifrages (saxifraga, sect. Ligulatae
@@
 histories, and biogeography. \emph{Molecular Phylogenetics and
-Evolution} \textbf{13}:536--555.
+Evolution} \textbf{13}: 536--555.
 
 \leavevmode\hypertarget{ref-corbin2004competition}{}%
-Corbin JD, D'Antonio CM. 2004. Competition between native perennial and
-exotic annual grasses: Implications for an historical invasion.
-\emph{Ecology} \textbf{85}:1273--1283.
+\textbf{\textnormal{Corbin JD}, \textnormal{D'Antonio CM}}.
+\textbf{2004}. Competition between native perennial and exotic annual
+grasses: Implications for an historical invasion. \emph{Ecology}
+\textbf{85}: 1273--1283.
 
 \leavevmode\hypertarget{ref-cruz2009molecular}{}%
-Cruz-Mazo G, Buide M, Samuel R, Narbona E. 2009. Molecular phylogeny of
-scorzoneroides (asteraceae): Evolution of heterocarpy and annual habit
-in unpredictable environments. \emph{Molecular phylogenetics and
-evolution} \textbf{53}:835--847.
+\textbf{\textnormal{Cruz-Mazo G}, \textnormal{Buide M},
+\textnormal{Samuel R}, \textnormal{Narbona E}}. \textbf{2009}. Molecular
+phylogeny of scorzoneroides (asteraceae): Evolution of heterocarpy and
+annual habit in unpredictable environments. \emph{Molecular
+phylogenetics and evolution} \textbf{53}: 835--847.
 
 \leavevmode\hypertarget{ref-dai2011drought}{}%
-Dai A. 2011. Drought under global warming: A review. \emph{Wiley
-Interdisciplinary Reviews: Climate Change} \textbf{2}:45--65.
+\textbf{\textnormal{Dai A}}. \textbf{2011}. Drought under global
+warming: A review. \emph{Wiley Interdisciplinary Reviews: Climate
+Change} \textbf{2}: 45--65.
 
 \leavevmode\hypertarget{ref-darriba2012jmodeltest}{}%
-Darriba D, Taboada G, Doallo R, Posada D. 2012. JModelTest 2: More
-models, new heuristics and parallel computing. \emph{Nature Methods}
-\textbf{9}:772.
+\textbf{\textnormal{Darriba D}, \textnormal{Taboada G},
+\textnormal{Doallo R}, \textnormal{Posada D}}. \textbf{2012}. JModelTest
+2: More models, new heuristics and parallel computing. \emph{Nature
+Methods} \textbf{9}: 772.
 
 \leavevmode\hypertarget{ref-daru2018widespread}{}%
-Daru BH, Park DS, Primack RB, Willis CG, Barrington DS, Whitfeld TJ,
-Seidler TG, Sweeney PW, Foster DR, Ellison AM, others. 2018. Widespread
-sampling biases in herbaria revealed from large-scale digitization.
-\emph{New Phytologist} \textbf{217}:939--955.
+\textbf{\textnormal{Daru BH}, \textnormal{Park DS}, \textnormal{Primack
+RB}, \textnormal{Willis CG}, \textnormal{Barrington DS},
+\textnormal{Whitfeld TJ}, \textnormal{Seidler TG}, \textnormal{Sweeney
+PW}, \textnormal{Foster DR}, \textnormal{Ellison AM} \emph{et al.}}
+\textbf{2018}. Widespread sampling biases in herbaria revealed from
+large-scale digitization. \emph{New Phytologist} \textbf{217}: 939--955.
 
 \leavevmode\hypertarget{ref-datson2008climate}{}%
-Datson P, Murray B, Steiner K. 2008. Climate and the evolution of
+\textbf{\textnormal{Datson P}, \textnormal{Murray B},
+\textnormal{Steiner K}}. \textbf{2008}. Climate and the evolution of
 annual/perennial life-histories in nemesia (scrophulariaceae).
-\emph{Plant Systematics and Evolution} \textbf{270}:39--57.
+\emph{Plant Systematics and Evolution} \textbf{270}: 39--57.
 
 \leavevmode\hypertarget{ref-davis2015herbarium}{}%
-Davis CC, Willis CG, Connolly B, Kelly C, Ellison AM. 2015. Herbarium
-records are reliable sources of phenological change driven by climate
-and provide novel insights into species' phenological cueing mechanisms.
-\emph{American Journal of Botany} \textbf{102}:1599--1609.
+\textbf{\textnormal{Davis CC}, \textnormal{Willis CG},
+\textnormal{Connolly B}, \textnormal{Kelly C}, \textnormal{Ellison AM}}.
+\textbf{2015}. Herbarium records are reliable sources of phenological
+change driven by climate and provide novel insights into species'
+phenological cueing mechanisms. \emph{American Journal of Botany}
+\textbf{102}: 1599--1609.
 
 \leavevmode\hypertarget{ref-dittberner2018natural}{}%
-Dittberner H, Korte A, Mettler-Altmann T, Weber A, Monroe G, Meaux J de.
-2018. Natural variation in stomata size contributes to the local
-adaptation of water-use efficiency in arabidopsis thaliana.
-\emph{bioRxiv} 253021.
+\textbf{\textnormal{Dittberner H}, \textnormal{Korte A},
+\textnormal{Mettler-Altmann T}, \textnormal{Weber A}, \textnormal{Monroe
+G}, \textnormal{Meaux J de}}. \textbf{2018}. Natural variation in
+stomata size contributes to the local adaptation of water-use efficiency
+in arabidopsis thaliana. \emph{bioRxiv}: 253021.
 
 \leavevmode\hypertarget{ref-R-geiger_b}{}%
-Eastman J, Alfaro M, Joyce P, Hipp A, Harmon L. 2011. A novel
+\textbf{\textnormal{Eastman J}, \textnormal{Alfaro M}, \textnormal{Joyce
+P}, \textnormal{Hipp A}, \textnormal{Harmon L}}. \textbf{2011}. A novel
 comparative method for identifying shifts in the rate of character
-evolution on trees. \emph{Evolution} \textbf{65}:3578--3589.
+evolution on trees. \emph{Evolution} \textbf{65}: 3578--3589.
 
 \leavevmode\hypertarget{ref-evans2005climate}{}%
-Evans ME, Hearn DJ, Hahn WJ, Spangle JM, Venable DL. 2005. Climate and
-life-history evolution in evening primroses (oenothera, onagraceae): A
-phylogenetic comparative analysis. \emph{Evolution}
-\textbf{59}:1914--1927.
+\textbf{\textnormal{Evans ME}, \textnormal{Hearn DJ}, \textnormal{Hahn
+WJ}, \textnormal{Spangle JM}, \textnormal{Venable DL}}. \textbf{2005}.
+Climate and life-history evolution in evening primroses (oenothera,
+onagraceae): A phylogenetic comparative analysis. \emph{Evolution}
+\textbf{59}: 1914--1927.
 
 \leavevmode\hypertarget{ref-felsenstein1985phylogenies}{}%
-Felsenstein J. 1985. Phylogenies and the comparative method.
-\emph{American Naturalist} \textbf{125}:1--15.
+\textbf{\textnormal{Felsenstein J}}. \textbf{1985}. Phylogenies and the
+comparative method. \emph{American Naturalist} \textbf{125}: 1--15.
 
 \leavevmode\hypertarget{ref-franco1996life}{}%
-Franco M, Silvertown J. 1996. Life history variation in plants: An
-exploration of the fast-slow continuum hypothesis. \emph{Phil Trans R
-Soc Lond B} \textbf{351}:1341--1348.
+\textbf{\textnormal{Franco M}, \textnormal{Silvertown J}}.
+\textbf{1996}. Life history variation in plants: An exploration of the
+fast-slow continuum hypothesis. \emph{Phil. Trans. R. Soc. Lond. B}
+\textbf{351}: 1341--1348.
 
 \leavevmode\hypertarget{ref-friedman2015all}{}%
-Friedman J, Rubin MJ. 2015. All in good time: Understanding annual and
-perennial strategies in plants. \emph{American journal of botany}
-\textbf{102}:497--499.
+\textbf{\textnormal{Friedman J}, \textnormal{Rubin MJ}}. \textbf{2015}.
+All in good time: Understanding annual and perennial strategies in
+plants. \emph{American journal of botany} \textbf{102}: 497--499.
 
 \leavevmode\hypertarget{ref-garnier1997specific}{}%
-Garnier E, Cordonnier P, Guillerm J-L, Sonié L. 1997. Specific leaf area
-and leaf nitrogen concentration in annual and perennial grass species
-growing in mediterranean old-fields. \emph{Oecologia}
-\textbf{111}:490--498.
+\textbf{\textnormal{Garnier E}, \textnormal{Cordonnier P},
+\textnormal{Guillerm J-L}, \textnormal{Sonié L}}. \textbf{1997}.
+Specific leaf area and leaf nitrogen concentration in annual and
+perennial grass species growing in mediterranean old-fields.
+\emph{Oecologia} \textbf{111}: 490--498.
 
 \leavevmode\hypertarget{ref-gbifdownload}{}%
-GBIF. 2018. GBIF occurrence download.
+\textbf{\textnormal{GBIF}}. \textbf{2018}. GBIF occurrence download.
 
 \leavevmode\hypertarget{ref-R-mvtnorm}{}%
-Genz A, Bretz F. 2009. Computation of multivariate normal and t
-probabilities, Lecture notes in statistics. Heidelberg: Springer-Verlag.
+\textbf{\textnormal{Genz A}, \textnormal{Bretz F}}. \textbf{2009}.
+\emph{Computation of multivariate normal and t probabilities}.
+Heidelberg: Springer-Verlag.
 
 \leavevmode\hypertarget{ref-guindon2003a}{}%
-Guindon S, Gascuel O. 2003. A simple, fast and accurate method to
-estimate large phylogenies by maximum-likelihood. \emph{Systematic
-Biology} \textbf{52}:696--704.
+\textbf{\textnormal{Guindon S}, \textnormal{Gascuel O}}. \textbf{2003}.
+A simple, fast and accurate method to estimate large phylogenies by
+maximum-likelihood. \emph{Systematic Biology} \textbf{52}: 696--704.
 
 \leavevmode\hypertarget{ref-R-MCMCglmm}{}%
-Hadfield JD. 2010. MCMC methods for multi-response generalized linear
-mixed models: The MCMCglmm R package. \emph{Journal of Statistical
-Software} \textbf{33}:1--22.
+\textbf{\textnormal{Hadfield JD}}. \textbf{2010}. MCMC methods for
+multi-response generalized linear mixed models: The MCMCglmm R package.
+\emph{Journal of Statistical Software} \textbf{33}: 1--22.
 
 \leavevmode\hypertarget{ref-R-geiger_d}{}%
-Harmon L, Weir J, Brock C, Glor R, Challenger W. 2008. GEIGER:
-Investigating evolutionary radiations. \emph{Bioinformatics}
-\textbf{24}:129--131.
+\textbf{\textnormal{Harmon L}, \textnormal{Weir J}, \textnormal{Brock
+C}, \textnormal{Glor R}, \textnormal{Challenger W}}. \textbf{2008}.
+GEIGER: Investigating evolutionary radiations. \emph{Bioinformatics}
+\textbf{24}: 129--131.
 
 \leavevmode\hypertarget{ref-R-logistf}{}%
-Heinze G, Ploner M. 2018. Logistf: Firth's bias-reduced logistic
-regression.
+\textbf{\textnormal{Heinze G}, \textnormal{Ploner M}}. \textbf{2018}.
+\emph{Logistf: Firth's bias-reduced logistic regression}.
 
 \leavevmode\hypertarget{ref-R-purrr}{}%
-Henry L, Wickham H. 2018. Purrr: Functional programming tools.
+\textbf{\textnormal{Henry L}, \textnormal{Wickham H}}. \textbf{2018}.
+\emph{Purrr: Functional programming tools}.
 
 \leavevmode\hypertarget{ref-R-raster}{}%
-Hijmans RJ. 2018. Raster: Geographic data analysis and modeling.
+\textbf{\textnormal{Hijmans RJ}}. \textbf{2018}. \emph{Raster:
+Geographic data analysis and modeling}.
 
 \leavevmode\hypertarget{ref-R-phylolm}{}%
-Ho LST, Ane C. 2014. A linear-time algorithm for gaussian and
-non-gaussian trait evolution models. \emph{Systematic Biology}
-\textbf{63}:397--408.
+\textbf{\textnormal{Ho LST}, \textnormal{Ane C}}. \textbf{2014}. A
+linear-time algorithm for gaussian and non-gaussian trait evolution
+models. \emph{Systematic Biology} \textbf{63}: 397--408.
 
 \leavevmode\hypertarget{ref-ives2010phylogenetic}{}%
-Ives A, Garland T. 2010. CPhylogenetic logistic regression for binary
-dependent variables. \emph{Systematic Biology} \textbf{59}:9--26.
+\textbf{\textnormal{Ives A}, \textnormal{Garland T}}. \textbf{2010}.
+CPhylogenetic logistic regression for binary dependent variables.
+\emph{Systematic Biology} \textbf{59}: 9--26.
 
 \leavevmode\hypertarget{ref-katoh2002mafft}{}%
-Katoh, Misawa, Kuma, Miyata. 2002. MAFFT: A novel method for rapid
+\textbf{\textnormal{Katoh}, \textnormal{Misawa}, \textnormal{Kuma},
+\textnormal{Miyata}}. \textbf{2002}. MAFFT: A novel method for rapid
 multiple sequence alignment based on fast fourier transform.
-\emph{Nucleic Acids Research} \textbf{30}:3059--3066.
+\emph{Nucleic Acids Research} \textbf{30}: 3059--3066.
 
 \leavevmode\hypertarget{ref-kerr2003space}{}%
-Kerr JT, Ostrovsky M. 2003. From space to species: Ecological
-applications for remote sensing. \emph{Trends in ecology \& evolution}
-\textbf{18}:299--305.
+\textbf{\textnormal{Kerr JT}, \textnormal{Ostrovsky M}}. \textbf{2003}.
+From space to species: Ecological applications for remote sensing.
+\emph{Trends in ecology \& evolution} \textbf{18}: 299--305.
+
+\leavevmode\hypertarget{ref-kogan2001operational}{}%
+\textbf{\textnormal{Kogan FN}}. \textbf{2001}. Operational space
+technology for global vegetation assessment. \emph{Bulletin of the
+American Meteorological Society} \textbf{82}: 1949--1964.
 
 \leavevmode\hypertarget{ref-kogan2016modelling}{}%
-Kogan F, Guo W, Strashnaia A, Kleshenko A, Chub O, Virchenko O. 2016.
-Modelling and prediction of crop losses from noaa polar-orbiting
-operational satellites. \emph{Geomatics, Natural Hazards and Risk}
-\textbf{7}:886--900.
-
-\leavevmode\hypertarget{ref-kogan2001operational}{}%
-Kogan FN. 2001. Operational space technology for global vegetation
-assessment. \emph{Bulletin of the American Meteorological Society}
-\textbf{82}:1949--1964.
+\textbf{\textnormal{Kogan F}, \textnormal{Guo W}, \textnormal{Strashnaia
+A}, \textnormal{Kleshenko A}, \textnormal{Chub O}, \textnormal{Virchenko
+O}}. \textbf{2016}. Modelling and prediction of crop losses from noaa
+polar-orbiting operational satellites. \emph{Geomatics, Natural Hazards
+and Risk} \textbf{7}: 886--900.
 
 \leavevmode\hypertarget{ref-lang2018using}{}%
-Lang PL, Willems FM, Scheepens J, Burbano HA, Bossdorf O. 2018. Using
-herbaria to study global environmental change. PeerJ Preprints.
+\textbf{\textnormal{Lang PL}, \textnormal{Willems FM},
+\textnormal{Scheepens J}, \textnormal{Burbano HA}, \textnormal{Bossdorf
+O}}. \textbf{2018}. \emph{Using herbaria to study global environmental
+change}. PeerJ Preprints.
 
 \leavevmode\hypertarget{ref-lelievre2009current}{}%
-Lelièvre F, Volaire F. 2009. Current and potential development of
-perennial grasses in rainfed mediterranean farming systems. \emph{Crop
-Science} \textbf{49}:2371--2378.
+\textbf{\textnormal{Lelièvre F}, \textnormal{Volaire F}}. \textbf{2009}.
+Current and potential development of perennial grasses in rainfed
+mediterranean farming systems. \emph{Crop Science} \textbf{49}:
+2371--2378.
 
 \leavevmode\hypertarget{ref-mandakova2012whole}{}%
-Mandáková T, Mummenhoff K, Al-Shehbaz IA, Mucina L, Mühlhausen A, Lysak
-MA. 2012. Whole-genome triplication and species radiation in the
-southern african tribe heliophileae (brassicaceae). \emph{Taxon}
-\textbf{61}:989--1000.
+\textbf{\textnormal{Mandáková T}, \textnormal{Mummenhoff K},
+\textnormal{Al-Shehbaz IA}, \textnormal{Mucina L},
+\textnormal{Mühlhausen A}, \textnormal{Lysak MA}}. \textbf{2012}.
+Whole-genome triplication and species radiation in the southern african
+tribe heliophileae (brassicaceae). \emph{Taxon} \textbf{61}: 989--1000.
 
 \leavevmode\hypertarget{ref-miller2010creating}{}%
-Miller M, Pfeiffer W, Schwartz T and. 2010. Creating the cipres science
+\textbf{\textnormal{Miller M}, \textnormal{Pfeiffer W},
+\textnormal{Schwartz T and}}. \textbf{2010}. Creating the cipres science
 gateway for inference of large phylogenetic trees. \emph{Proceedings of
-the Gateway Computing Environments Workshop} 1--8.
+the Gateway Computing Environments Workshop}: 1--8.
 
 \leavevmode\hypertarget{ref-Mojica2016}{}%
-Mojica JP, Mullen J, Lovell JT, Monroe JG, Paul JR, Oakley CG, McKay JK.
-2016. Genetics of water use physiology in locally adapted Arabidopsis
-thaliana. \emph{Plant Science}.
-doi:\href{https://doi.org/10.1016/j.plantsci.2016.03.015}{10.1016/j.plantsci.2016.03.015}
+\textbf{\textnormal{Mojica JP}, \textnormal{Mullen J},
+\textnormal{Lovell JT}, \textnormal{Monroe JG}, \textnormal{Paul JR},
+\textnormal{Oakley CG}, \textnormal{McKay JK}}. \textbf{2016}. Genetics
+of water use physiology in locally adapted Arabidopsis thaliana.
+\emph{Plant Science}.
 
 \leavevmode\hypertarget{ref-monroe2018ecoevolutionary}{}%
-Monroe J, Markman D, Beck W, Felton A, Vahsen M, Pressler Y. 2018.
-Ecoevolutionary dynamics of carbon cycling in the anthropocene.
-\emph{Trends in ecology \& evolution} \textbf{33}:213--225.
+\textbf{\textnormal{Monroe J}, \textnormal{Markman D}, \textnormal{Beck
+W}, \textnormal{Felton A}, \textnormal{Vahsen M}, \textnormal{Pressler
+Y}}. \textbf{2018a}. Ecoevolutionary dynamics of carbon cycling in the
+anthropocene. \emph{Trends in ecology \& evolution} \textbf{33}:
+213--225.
 
 \leavevmode\hypertarget{ref-monroe2018drought}{}%
-Monroe J, Powell T, Price N, Mullen J, Howard A, Evans K, Lovell J,
-McKay J. 2018. Drought adaptation in nature by extensive genetic
-loss-of-function. \emph{eLife} DOI: 10.7554/eLife.41038.
+\textbf{\textnormal{Monroe J}, \textnormal{Powell T}, \textnormal{Price
+N}, \textnormal{Mullen J}, \textnormal{Howard A}, \textnormal{Evans K},
+\textnormal{Lovell J}, \textnormal{McKay J}}. \textbf{2018b}. Drought
+adaptation in nature by extensive genetic loss-of-function.
+\emph{eLife}: DOI: 10.7554/eLife.41038.
 
 \leavevmode\hypertarget{ref-morishima1984differentiation}{}%
-Morishima H, Sano Y, Oka H. 1984. Differentiation of perennial and
-annual types due to habitat conditions in the wild riceOryza perennis.
-\emph{Plant Systematics and Evolution} \textbf{144}:119--135.
+\textbf{\textnormal{Morishima H}, \textnormal{Sano Y}, \textnormal{Oka
+H}}. \textbf{1984}. Differentiation of perennial and annual types due to
+habitat conditions in the wild riceOryza perennis. \emph{Plant
+Systematics and Evolution} \textbf{144}: 119--135.
 
 \leavevmode\hypertarget{ref-mummenhoff2005phylogeny}{}%
-Mummenhoff K, Al-Shehbaz IA, Bakker FT, Linder HP, Mühlhausen A. 2005.
-Phylogeny, morphological evolution, and speciation of endemic
-brassicaceae genera in the cape flora of southern africa. \emph{Annals
-of the Missouri Botanical Garden} 400--424.
+\textbf{\textnormal{Mummenhoff K}, \textnormal{Al-Shehbaz IA},
+\textnormal{Bakker FT}, \textnormal{Linder HP}, \textnormal{Mühlhausen
+A}}. \textbf{2005}. Phylogeny, morphological evolution, and speciation
+of endemic brassicaceae genera in the cape flora of southern africa.
+\emph{Annals of the Missouri Botanical Garden}: 400--424.
 
 \leavevmode\hypertarget{ref-R-bindrcpp}{}%
-Müller K. 2018. Bindrcpp: An 'rcpp' interface to active bindings.
+\textbf{\textnormal{Müller K}}. \textbf{2018}. \emph{Bindrcpp: An 'rcpp'
+interface to active bindings}.
 
 \leavevmode\hypertarget{ref-R-tibble}{}%
-Müller K, Wickham H. 2018. Tibble: Simple data frames.
+\textbf{\textnormal{Müller K}, \textnormal{Wickham H}}. \textbf{2018}.
+\emph{Tibble: Simple data frames}.
 
 \leavevmode\hypertarget{ref-R-caper}{}%
-Orme D, Freckleton R, Thomas G, Petzoldt T, Fritz S, Isaac N, Pearse W.
-2018. Caper: Comparative analyses of phylogenetics and evolution in r.
+\textbf{\textnormal{Orme D}, \textnormal{Freckleton R},
+\textnormal{Thomas G}, \textnormal{Petzoldt T}, \textnormal{Fritz S},
+\textnormal{Isaac N}, \textnormal{Pearse W}}. \textbf{2018}.
+\emph{Caper: Comparative analyses of phylogenetics and evolution in r}.
 
 \leavevmode\hypertarget{ref-R-ape}{}%
-Paradis E, Schliep K. 2018. Ape 5.0: An environment for modern
-phylogenetics and evolutionary analyses in R. \emph{Bioinformatics}
-\textbf{xx}:xxx--xxx.
+\textbf{\textnormal{Paradis E}, \textnormal{Schliep K}}. \textbf{2018}.
+Ape 5.0: An environment for modern phylogenetics and evolutionary
+analyses in R. \emph{Bioinformatics} \textbf{xx}: xxx--xxx.
 
 \leavevmode\hypertarget{ref-parry2005prospects}{}%
-Parry M, Flexas J, Medrano H. 2005. Prospects for crop production under
-drought: Research priorities and future directions. \emph{Annals of
-Applied Biology} \textbf{147}:211--226.
+\textbf{\textnormal{Parry M}, \textnormal{Flexas J}, \textnormal{Medrano
+H}}. \textbf{2005}. Prospects for crop production under drought:
+Research priorities and future directions. \emph{Annals of Applied
+Biology} \textbf{147}: 211--226.
 
 \leavevmode\hypertarget{ref-R-sp}{}%
-Pebesma EJ, Bivand RS. 2005. Classes and methods for spatial data in R.
-\emph{R News} \textbf{5}:9--13.
+\textbf{\textnormal{Pebesma EJ}, \textnormal{Bivand RS}}. \textbf{2005}.
+Classes and methods for spatial data in R. \emph{R News} \textbf{5}:
+9--13.
 
 \leavevmode\hypertarget{ref-R-coda}{}%
-Plummer M, Best N, Cowles K, Vines K. 2006. CODA: Convergence diagnosis
-and output analysis for mcmc. \emph{R News} \textbf{6}:7--11.
+\textbf{\textnormal{Plummer M}, \textnormal{Best N}, \textnormal{Cowles
+K}, \textnormal{Vines K}}. \textbf{2006}. CODA: Convergence diagnosis
+and output analysis for mcmc. \emph{R News} \textbf{6}: 7--11.
 
 \leavevmode\hypertarget{ref-R-gee}{}%
-R by Thomas Lumley VJCP to, author. BRN that maintainers are not
-available to give advice on using a package they did not. 2015. Gee:
-Generalized estimation equation solver.
+\textbf{\textnormal{R by Thomas Lumley VJCP to}, \textnormal{author. BRN
+that maintainers are not available to give advice on using a package
+they did not}}. \textbf{2015}. \emph{Gee: Generalized estimation
+equation solver}.
 
 \leavevmode\hypertarget{ref-R-base}{}%
-R Core Team. 2018. R: A language and environment for statistical
-computing. Vienna, Austria: R Foundation for Statistical Computing.
+\textbf{\textnormal{R Core Team}}. \textbf{2018}. \emph{R: A language
+and environment for statistical computing}. Vienna, Austria: R
+Foundation for Statistical Computing.
 
 \leavevmode\hypertarget{ref-R-phytools}{}%
-Revell LJ. 2012. Phytools: An r package for phylogenetic comparative
-biology (and other things). \emph{Methods in Ecology and Evolution}
-\textbf{3}:217--223.
+\textbf{\textnormal{Revell LJ}}. \textbf{2012}. Phytools: An r package
+for phylogenetic comparative biology (and other things). \emph{Methods
+in Ecology and Evolution} \textbf{3}: 217--223.
 
 \leavevmode\hypertarget{ref-R-maps}{}%
-Richard A. Becker OS code by, Ray Brownrigg. Enhancements by Thomas P
-Minka ARWR version by, Deckmyn. A. 2018. Maps: Draw geographical maps.
+\textbf{\textnormal{Richard A. Becker OS code by}, \textnormal{Ray
+Brownrigg. Enhancements by Thomas P Minka ARWR version by},
+\textnormal{Deckmyn. A}}. \textbf{2018}. \emph{Maps: Draw geographical
+maps}.
 
 \leavevmode\hypertarget{ref-rojas2011assessing}{}%
-Rojas O, Vrieling A, Rembold F. 2011. Assessing drought probability for
-agricultural areas in africa with coarse resolution remote sensing
-imagery. \emph{Remote sensing of Environment} \textbf{115}:343--352.
+\textbf{\textnormal{Rojas O}, \textnormal{Vrieling A},
+\textnormal{Rembold F}}. \textbf{2011}. Assessing drought probability
+for agricultural areas in africa with coarse resolution remote sensing
+imagery. \emph{Remote sensing of Environment} \textbf{115}: 343--352.
 
 \leavevmode\hypertarget{ref-roumet2006suites}{}%
-Roumet C, Urcelay C, Dı'az S. 2006. Suites of root traits differ between
-annual and perennial species growing in the field. \emph{New
-phytologist} \textbf{170}:357--368.
+\textbf{\textnormal{Roumet C}, \textnormal{Urcelay C}, \textnormal{Dı'az
+S}}. \textbf{2006}. Suites of root traits differ between annual and
+perennial species growing in the field. \emph{New phytologist}
+\textbf{170}: 357--368.
 
 \leavevmode\hypertarget{ref-sayre2013new}{}%
-Sayre RG, Comer P, Hak J, Josse C, Bow J, Warner H, Larwanou M, Kelbessa
-E, Bekele T, Kehl H, others. 2013. A new map of standardized terrestrial
-ecosystems of africa. \emph{African Geographical Review}.
+\textbf{\textnormal{Sayre RG}, \textnormal{Comer P}, \textnormal{Hak J},
+\textnormal{Josse C}, \textnormal{Bow J}, \textnormal{Warner H},
+\textnormal{Larwanou M}, \textnormal{Kelbessa E}, \textnormal{Bekele T},
+\textnormal{Kehl H} \emph{et al.}} \textbf{2013}. A new map of
+standardized terrestrial ecosystems of africa. \emph{African
+Geographical Review}.
 
 \leavevmode\hypertarget{ref-schaffer1975selection}{}%
-Schaffer W, Gadgil M. 1975. Selection for optimal life histories in
-plants. \emph{Ecology and evolution of communities} 142--157.
+\textbf{\textnormal{Schaffer W}, \textnormal{Gadgil M}}. \textbf{1975}.
+Selection for optimal life histories in plants. \emph{Ecology and
+evolution of communities.}: 142--157.
 
 \leavevmode\hypertarget{ref-R-geiger_c}{}%
-Slater G, Harmon L, Wegmann D, Joyce P, Revell L, Alfaro M. 2012.
-Fitting models of continuous trait evolution to incompletely sampled
-comparative data using approximate bayesian computation.
-\emph{Evolution} \textbf{66}:752--762.
+\textbf{\textnormal{Slater G}, \textnormal{Harmon L},
+\textnormal{Wegmann D}, \textnormal{Joyce P}, \textnormal{Revell L},
+\textnormal{Alfaro M}}. \textbf{2012}. Fitting models of continuous
+trait evolution to incompletely sampled comparative data using
+approximate bayesian computation. \emph{Evolution} \textbf{66}:
+752--762.
+
+\leavevmode\hypertarget{ref-R-diagram}{}%
+\textbf{\textnormal{Soetaert K}}. \textbf{2017}. \emph{Diagram:
+Functions for visualising simple graphs (networks), plotting flow
+diagrams}.
 
 \leavevmode\hypertarget{ref-R-shape}{}%
-Soetaert K. 2018. Shape: Functions for plotting graphical shapes,
-colors.
-
-\leavevmode\hypertarget{ref-R-diagram}{}%
-Soetaert K. 2017. Diagram: Functions for visualising simple graphs
-(networks), plotting flow diagrams.
+\textbf{\textnormal{Soetaert K}}. \textbf{2018}. \emph{Shape: Functions
+for plotting graphical shapes, colors}.
 
 \leavevmode\hypertarget{ref-soltis2017digitization}{}%
-Soltis PS. 2017. Digitization of herbaria enables novel research.
-\emph{American journal of botany} \textbf{104}:1281--1284.
+\textbf{\textnormal{Soltis PS}}. \textbf{2017}. Digitization of herbaria
+enables novel research. \emph{American journal of botany} \textbf{104}:
+1281--1284.
 
 \leavevmode\hypertarget{ref-stearns1992evolution}{}%
-Stearns SC. 1992. The evolution of life histories.
+\textbf{\textnormal{Stearns SC}}. \textbf{1992}. \emph{The evolution of
+life histories}.
 
 \leavevmode\hypertarget{ref-stropp2016mapping}{}%
-Stropp J, Ladle RJ, M. Malhado AC, Hortal J, Gaffuri J, H. Temperley W,
-Olav Skøien J, Mayaux P. 2016. Mapping ignorance: 300 years of
+\textbf{\textnormal{Stropp J}, \textnormal{Ladle RJ}, \textnormal{M.
+Malhado AC}, \textnormal{Hortal J}, \textnormal{Gaffuri J},
+\textnormal{H. Temperley W}, \textnormal{Olav Skøien J},
+\textnormal{Mayaux P}}. \textbf{2016}. Mapping ignorance: 300 years of
 collecting flowering plants in africa. \emph{Global Ecology and
-Biogeography} \textbf{25}:1085--1096.
+Biogeography} \textbf{25}: 1085--1096.
 
 \leavevmode\hypertarget{ref-tank2008annuals}{}%
-Tank DC, Olmstead RG. 2008. From annuals to perennials: Phylogeny of
-subtribe castillejinae (orobanchaceae). \emph{American Journal of
-Botany} \textbf{95}:608--625.
+\textbf{\textnormal{Tank DC}, \textnormal{Olmstead RG}}. \textbf{2008}.
+From annuals to perennials: Phylogeny of subtribe castillejinae
+(orobanchaceae). \emph{American Journal of Botany} \textbf{95}:
+608--625.
 
 \leavevmode\hypertarget{ref-vaclavik2017effects}{}%
-Václavı'k T, Beckmann M, Cord AF, Bindewald AM. 2017. Effects of uv-b
-radiation on leaf hair traits of invasive plants---combining historical
-herbarium records with novel remote sensing data. \emph{PloS one}
-\textbf{12}:e0175671.
+\textbf{\textnormal{Václavı'k T}, \textnormal{Beckmann M},
+\textnormal{Cord AF}, \textnormal{Bindewald AM}}. \textbf{2017}. Effects
+of uv-b radiation on leaf hair traits of invasive plants---combining
+historical herbarium records with novel remote sensing data. \emph{PloS
+one} \textbf{12}: e0175671.
 
 \leavevmode\hypertarget{ref-R-MASS}{}%
-Venables WN, Ripley BD. 2002. Modern applied statistics with s, Fourth.
-ed. New York: Springer.
+\textbf{\textnormal{Venables WN}, \textnormal{Ripley BD}}.
+\textbf{2002}. \emph{Modern applied statistics with s}. New York:
+Springer.
 
 \leavevmode\hypertarget{ref-verboom2004testing}{}%
-Verboom GA, Linder HP, Stock WD. 2004. Testing the adaptive nature of
+\textbf{\textnormal{Verboom GA}, \textnormal{Linder HP},
+\textnormal{Stock WD}}. \textbf{2004}. Testing the adaptive nature of
 radiation: Growth form and life history divergence in the african grass
 genus ehrharta (poaceae: Ehrhartoideae). \emph{American Journal of
-Botany} \textbf{91}:1364--1370.
+Botany} \textbf{91}: 1364--1370.
+
+\leavevmode\hypertarget{ref-R-ggplot2}{}%
+\textbf{\textnormal{Wickham H}}. \textbf{2016}. \emph{Ggplot2: Elegant
+graphics for data analysis}. Springer-Verlag New York.
+
+\leavevmode\hypertarget{ref-R-tidyverse}{}%
+\textbf{\textnormal{Wickham H}}. \textbf{2017}. \emph{Tidyverse: Easily
+install and load the 'tidyverse'}.
 
 \leavevmode\hypertarget{ref-R-forcats}{}%
-Wickham H. 2018a. Forcats: Tools for working with categorical variables
-(factors).
+\textbf{\textnormal{Wickham H}}. \textbf{2018a}. \emph{Forcats: Tools
+for working with categorical variables (factors)}.
 
 \leavevmode\hypertarget{ref-R-stringr}{}%
-Wickham H. 2018b. Stringr: Simple, consistent wrappers for common string
-operations.
-
-\leavevmode\hypertarget{ref-R-tidyverse}{}%
-Wickham H. 2017. Tidyverse: Easily install and load the 'tidyverse'.
-
-\leavevmode\hypertarget{ref-R-ggplot2}{}%
-Wickham H. 2016. Ggplot2: Elegant graphics for data analysis.
-Springer-Verlag New York.
+\textbf{\textnormal{Wickham H}}. \textbf{2018b}. \emph{Stringr: Simple,
+consistent wrappers for common string operations}.
 
 \leavevmode\hypertarget{ref-R-dplyr}{}%
-Wickham H, François R, Henry L, Müller K. 2018. Dplyr: A grammar of data
-manipulation.
+\textbf{\textnormal{Wickham H}, \textnormal{François R},
+\textnormal{Henry L}, \textnormal{Müller K}}. \textbf{2018}.
+\emph{Dplyr: A grammar of data manipulation}.
 
 \leavevmode\hypertarget{ref-R-tidyr}{}%
-Wickham H, Henry L. 2018. Tidyr: Easily tidy data with 'spread()' and
-'gather()' functions.
+\textbf{\textnormal{Wickham H}, \textnormal{Henry L}}. \textbf{2018}.
+\emph{Tidyr: Easily tidy data with 'spread()' and 'gather()' functions}.
 
 \leavevmode\hypertarget{ref-R-readr}{}%
-Wickham H, Hester J, Francois R. 2017. Readr: Read rectangular text
-data.
+\textbf{\textnormal{Wickham H}, \textnormal{Hester J},
+\textnormal{Francois R}}. \textbf{2017}. \emph{Readr: Read rectangular
+text data}.
 
 \leavevmode\hypertarget{ref-willis2017old}{}%
-Willis CG, Ellwood ER, Primack RB, Davis CC, Pearson KD, Gallinat AS,
-Yost JM, Nelson G, Mazer SJ, Rossington NL, others. 2017. Old plants,
-new tricks: Phenological research using herbarium specimens.
-\emph{Trends in ecology \& evolution} \textbf{32}:531--546.
+\textbf{\textnormal{Willis CG}, \textnormal{Ellwood ER},
+\textnormal{Primack RB}, \textnormal{Davis CC}, \textnormal{Pearson KD},
+\textnormal{Gallinat AS}, \textnormal{Yost JM}, \textnormal{Nelson G},
+\textnormal{Mazer SJ}, \textnormal{Rossington NL} \emph{et al.}}
+\textbf{2017}. Old plants, new tricks: Phenological research using
+herbarium specimens. \emph{Trends in ecology \& evolution} \textbf{32}:
+531--546.
 
 \leavevmode\hypertarget{ref-wolf2016altitudinal}{}%
-Wolf A, Zimmerman NB, Anderegg WR, Busby PE, Christensen J. 2016.
-Altitudinal shifts of the native and introduced flora of c alifornia in
-the context of 20th-century warming. \emph{Global ecology and
-biogeography} \textbf{25}:418--429.
+\textbf{\textnormal{Wolf A}, \textnormal{Zimmerman NB},
+\textnormal{Anderegg WR}, \textnormal{Busby PE}, \textnormal{Christensen
+J}}. \textbf{2016}. Altitudinal shifts of the native and introduced
+flora of c alifornia in the context of 20th-century warming.
+\emph{Global ecology and biogeography} \textbf{25}: 418--429.
 
@@
 We used R (Version 3.5.1; R Core Team, 2018) and the R-packages
-\emph{ape} (Version 5.2; Paradis and Schliep, 2018; Orme et al., 2018;
-Soetaert, 2018), \emph{bindrcpp} (Version 0.2.2; Müller, 2018),
-\emph{caper} (Version 1.0.1; Orme et al., 2018), \emph{coda} (Version
-0.19.2; Plummer et al., 2006), \emph{diagram} (Version 1.6.4; Soetaert,
-2017), \emph{dplyr} (Version 0.7.8; Wickham et al., 2018),
-\emph{forcats} (Version 0.3.0; Wickham, 2018a), \emph{gee} (Version
-4.13.19; R by Thomas Lumley and author., 2015), \emph{geiger} (Version
-2.0.6; Alfaro et al., 2009; Eastman et al., 2011; Harmon et al., 2008;
-Slater et al., 2012), \emph{ggplot2} (Version 3.1.0; Wickham, 2016),
-\emph{logistf} (Version 1.23; Heinze and Ploner, 2018), \emph{maps}
-(Version 3.3.0; Richard A. Becker et al., 2018), \emph{MASS} (Version
-7.3.51.1; Venables and Ripley, 2002), \emph{Matrix} (Version 1.2.15;
-Bates and Maechler, 2018), \emph{MCMCglmm} (Version 2.26; Hadfield,
-2010), \emph{mvtnorm} (Version 1.0.8; Genz and Bretz, 2009),
-\emph{papaja} (Version 0.1.0.9842; Aust and Barth, 2018), \emph{phylolm}
-(Version 2.6; Ho and Ane, 2014), \emph{phytools} (Version 0.6.60;
-Revell, 2012), \emph{purrr} (Version 0.2.5; Henry and Wickham, 2018),
-\emph{raster} (Version 2.8.4; Hijmans, 2018), \emph{readr} (Version
-1.2.1; Wickham et al., 2017), \emph{shape} (Version 1.4.4; Soetaert,
-2018), \emph{sp} (Version 1.3.1; Pebesma and Bivand, 2005),
-\emph{stringr} (Version 1.3.1; Wickham, 2018b), \emph{tibble} (Version
-1.4.2; Müller and Wickham, 2018), \emph{tidyr} (Version 0.8.2; Wickham
-and Henry, 2018), and \emph{tidyverse} (Version 1.2.1; Wickham, 2017)
-for all our analyses. \newpage
+\emph{ape} (Version 5.2; Paradis \& Schliep, 2018; Orme \emph{et al.},
+2018; Soetaert, 2018), \emph{bindrcpp} (Version 0.2.2; Müller, 2018),
+\emph{caper} (Version 1.0.1; Orme \emph{et al.}, 2018), \emph{coda}
+(Version 0.19.2; Plummer \emph{et al.}, 2006), \emph{diagram} (Version
+1.6.4; Soetaert, 2017), \emph{dplyr} (Version 0.7.8; Wickham \emph{et
+al.}, 2018), \emph{forcats} (Version 0.3.0; Wickham, 2018a), \emph{gee}
+(Version 4.13.19; R by Thomas Lumley \& author., 2015), \emph{geiger}
+(Version 2.0.6; Alfaro \emph{et al.}, 2009; Harmon \emph{et al.}, 2008;
+Eastman \emph{et al.}, 2011; Slater \emph{et al.}, 2012), \emph{ggplot2}
+(Version 3.1.0; Wickham, 2016), \emph{logistf} (Version 1.23; Heinze \&
+Ploner, 2018), \emph{maps} (Version 3.3.0; Richard A. Becker \emph{et
+al.}, 2018), \emph{MASS} (Version 7.3.51.1; Venables \& Ripley, 2002),
+\emph{Matrix} (Version 1.2.15; Bates \& Maechler, 2018), \emph{MCMCglmm}
+(Version 2.26; Hadfield, 2010), \emph{mvtnorm} (Version 1.0.8; Genz \&
+Bretz, 2009), \emph{papaja} (Version 0.1.0.9842; Aust \& Barth, 2018),
+\emph{phylolm} (Version 2.6; Ho \& Ane, 2014), \emph{phytools} (Version
+0.6.60; Revell, 2012), \emph{purrr} (Version 0.2.5; Henry \& Wickham,
+2018), \emph{raster} (Version 2.8.4; Hijmans, 2018), \emph{readr}
+(Version 1.2.1; Wickham \emph{et al.}, 2017), \emph{shape} (Version
+1.4.4; Soetaert, 2018), \emph{sp} (Version 1.3.1; Pebesma \& Bivand,
+2005), \emph{stringr} (Version 1.3.1; Wickham, 2018b), \emph{tibble}
+(Version 1.4.2; Müller \& Wickham, 2018), \emph{tidyr} (Version 0.8.2;
+Wickham \& Henry, 2018), and \emph{tidyverse} (Version 1.2.1; Wickham,
+2017) for all our analyses. \newpage
 

### 2018-12-12 manuscript/manuscript.Rmd
submission

@@
 Additionally, the collection dates of annual and perennial species indicate that annuals escape drought prone seasons during the seed phase of their life cycle.
-Together, these finding provide empirical support for classic hypotheses about the drivers of life history strategy in plants - that perennials out compete annuals in environments with less frequent drought and that annuals are adapted to environments with more frequent drought by escaping drought prone seasons as seeds.
+Together, these findings provide empirical support for classic hypotheses about the drivers of life history strategy in plants - that perennials out compete annuals in environments with less frequent drought and that annuals are adapted to environments with more frequent drought by escaping drought prone seasons as seeds.
 
@@
 In  plants, life histories are especially diverse, with herbaceous species that complete their life cycle in a number of weeks to trees that live for thousands of years [@brown1996oldlist].
-Along this continuum an important division exists, distinguishing annuals which complete their seed to seed life cycle within a single calendar year from perennials which can persist over multiple years.
+Along this continuum in angiosperms an important division exists, distinguishing annuals which complete their seed to seed life cycle within a single calendar year from perennials which can persist over multiple years.
 Annual plants flower once, set seed, senesce, and then die, spending at least some portion of the year as a seed, where they are relatively protected from environmental stress.
@@
 These represent fundamentally different life history strategies, but the ecological factors that explain their evolution and distributions remain empirically uresolved [@friedman2015all].
-Classical theory predict shorter life spans in environments where adult mortality is high [@charnov1973life; @stearns1992evolution;@franco1996life].
+Classical theory predicts shorter life spans in environments where adult mortality is high [@charnov1973life; @stearns1992evolution;@franco1996life].
 In plants, this has been extended to the hypothesis that annuality is adaptive when it allows plants to escape drought [@schaffer1975selection].
@@
 Thus, environments with greater seasonal drought frequency may select for annual life histories that complete reproduction prior to drought prone seasons.
-Conversely, environments with less frequent drought may select for perennial species, which may benefit from multiple bouts of reproduction and competitive advantage by preventing recruitment of annual species [@corbin2004competition].
+Conversely, environments with less frequent drought may select for perennial species, which benefit from multiple bouts of reproduction and competitive advantage by preventing recruitment of annual species [@corbin2004competition].
 These predictions have been supported by the observation of annuals in arid environments in _Oryza perennis_ [@morishima1984differentiation] and _Oenothera_ [@evans2005climate].
@@
 ```
-Botanists have collected and maintained over 350 million botanical specimens worldwide over the past 300 years.
+Botanists have collected and maintained over 350 million botanical specimens worldwide over the past 300 years [@thiers2016index].
 Herbarium specimens and their associated metadata have been used since the 1960s to study species’ geographical distributions (reviewed by  @willis2017old and @lang2018using).
@@
 * records without known geospatial coordinate issues (i.e., coordinates reported are those of herbarium)  
-* records from collection sites classified as land pixels   
+* records from collection sites classified as land pixels in the VHI dataset   
 * records from Africa (to exclude locations of cultivation)
@@
 To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimes, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency.
-First, we compared the frequency of drought during the winter, spring, summer, and fall between raw occurrence records of annual and perennial species by t-tests.
+First, we compared the frequency of drought during the winter, spring, summer, and fall between total occurrence records of annual and perennial species by t-tests.
 To account for variation in the number of occurrence records per species, we next calculated the mean drought frequency during the winter, spring, summer and fall for each species.
@@
 The number of samples varied between species, with a mean of `r mean(species_means$n)` samples per species.
-`r as.character(species_means$species[species_means$n==max(species_means$n)])` had the most records, `r max(species_means$n)`,  and  `r as.character(species_means$species[species_means$n==min(species_means$n)])` the fewest,  `r min(species_means$n)` (Table \@ref(tab:speciesmeanstable)).
+_H.
+rigidiuscula_ had the most records, `r max(species_means$n)`,  and  _H.
+cornellsbergia_ the fewest,  `r min(species_means$n)` (Table \@ref(tab:speciesmeanstable)).
 
@@
 We found that the frequency of drought was significantly higher at the locations of occurrence records for annual species.
-In terms of raw observation records
-(Figure \@ref(fig:mapsboxplots)g), the frequency of drought was significantly higher at the location of annuals during the winter (t = `r twin$statistic`, p = `r twin$p.value`), spring (t = `r tspr$statistic`, p = `r tspr$p.value`), summer (t = `r tsum$statistic`, p = `r tsum$p.value`), and fall (t = `r tfal$statistic`, p = `r tfal$p.value`).
-Because raw occurrence records do not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable)) or species relatedness (Figure \@ref(fig:phylogeny)a), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials using phylogenetic logistic regression.
+When comparing across all occurrence records
+(all records rather than species means, Figure \@ref(fig:mapsboxplots)g), the frequency of drought was significantly higher at the location of annuals during the winter (t = `r twin$statistic`, p = `r twin$p.value`), spring (t = `r tspr$statistic`, p = `r tspr$p.value`), summer (t = `r tsum$statistic`, p = `r tsum$p.value`), and fall (t = `r tfal$statistic`, p = `r tfal$p.value`).
+Because a comparison across all occurrence records does not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable)) or species relatedness (Figure \@ref(fig:phylogeny)a), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials using phylogenetic logistic regression.
 We found that the mean drought frequencies were significantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a).
-These findings indicate that common acestry alone does not explain differences the drought frequencies experienced in the environments of annual and perennial _Heliophila_.
+These findings indicate that common acestry alone does not explain differences the drought frequencies experienced between the environments of annual and perennial _Heliophila_.
 
@@
 apa_table(modelstable
-  , caption = "Phylogenetic logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species the winter, spring, summer, and fall."
+  , caption = "Phylogenetic logistic regressions between life history, and the mean drought frequency observed at specimen sites of Heliophila species the winter, spring, summer, and fall."
   , note = "Annual species were scored as 0 and perennial species as 1.",
@@
 ```
-The preceding results indicated that annual species are found in environments where droughts are significantly more frequent, especially in the summer and fall.
+The preceding results indicate that annual species are found in environments where droughts are significantly more frequent, especially in the summer and fall.
 Classic life history theory hypothesizes that annuality reflects adaptation to such environments because it allows species to escape stressful conditions.
@@
 While most previous work has focused on describing the evolutionary origins of annuality [@barrett1996comparative;@conti1999phylogenetic;@andreasen2001unequal;@verboom2004testing;@friedman2015all] there are at least a few other cases where perenniality appears to have arisen from an annual ancestor [@bena1998molecular;@tank2008annuals].
-And while early theory predicted selection for annuality when adult morality is high [@stearns1992evolution],  we also find evidence that the transition to perenniality could be explained by the historical frequency of drought, a likely cause of mortality in plants.
-The phylogeny reveals several transitions from annual to perennial life history (Figure \@ref(fig:phylogeny)a) and that the distributions of perennial _Heliophila_ extend into regions where drought frequency is low (Figure \@ref(fig:mapsboxplots)b, Figure \@ref(fig:speciesmaps)).
+And while early theory predicted selection for annuality when adult morality is high [@stearns1992evolution],  we also find evidence that perenniality could be explained by reduced frequency of drought.
+The phylogeny reveals several transitions from annual to perennial life history (Figure \@ref(fig:phylogeny)a) and the distributions of perennial _Heliophila_ extend into regions where drought frequency is low (Figure \@ref(fig:mapsboxplots)b, Figure \@ref(fig:speciesmaps)).
 Perennials may be able to out complete annual relatives in environments where the infrequency of drought favors strategies that allow plants to benefit from growth over many seasons.
-It may also indicate that annuals rely on drought as a source of disturbance for seedling recruitment when competing with perennials [@corbin2004competition].
+This also suggests that annuals rely on drought as a source of disturbance for seedling recruitment when competing with perennials [@corbin2004competition].
 Indeed, no annual species were observed in the low drought regions of eastern South Africa (Figure \@ref(fig:mapsboxplots), Figure \@ref(fig:speciesmaps)).
@@
 # Acknowledgments 
-We thank Jesse Lasky and members of the Sloan lab for generous feedback that improved the quality of this work.
-This research was supported by NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
+We thank members of the Sloan lab for generous feedback that improved the quality of this work. This research was supported by NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
 

### 2018-12-12 manuscript/manuscript.tex
submission

@@
 escape drought prone seasons during the seed phase of their life cycle.
-Together, these finding provide empirical support for classic hypotheses
-about the drivers of life history strategy in plants - that perennials
-out compete annuals in environments with less frequent drought and that
-annuals are adapted to environments with more frequent drought by
-escaping drought prone seasons as seeds. 
+Together, these findings provide empirical support for classic
+hypotheses about the drivers of life history strategy in plants - that
+perennials out compete annuals in environments with less frequent
+drought and that annuals are adapted to environments with more frequent
+drought by escaping drought prone seasons as seeds.
+
 }
@@
 that complete their life cycle in a number of weeks to trees that live
-for thousands of years (Brown, 1996). Along this continuum an important
-division exists, distinguishing annuals which complete their seed to
-seed life cycle within a single calendar year from perennials which can
-persist over multiple years. Annual plants flower once, set seed,
-senesce, and then die, spending at least some portion of the year as a
-seed, where they are relatively protected from environmental stress. In
-contrast, perennial plants can continue vegetative growth after
-reproduction and must survive conditions experienced during all seasons.
-These represent fundamentally different life history strategies, but the
-ecological factors that explain their evolution and distributions remain
-empirically uresolved (Friedman \& Rubin, 2015).\\
-Classical theory predict shorter life spans in environments where adult
+for thousands of years (Brown, 1996). Along this continuum in
+angiosperms an important division exists, distinguishing annuals which
+complete their seed to seed life cycle within a single calendar year
+from perennials which can persist over multiple years. Annual plants
+flower once, set seed, senesce, and then die, spending at least some
+portion of the year as a seed, where they are relatively protected from
+environmental stress. In contrast, perennial plants can continue
+vegetative growth after reproduction and must survive conditions
+experienced during all seasons. These represent fundamentally different
+life history strategies, but the ecological factors that explain their
+evolution and distributions remain empirically uresolved (Friedman \&
+Rubin, 2015).\\
+Classical theory predicts shorter life spans in environments where adult
 mortality is high (Charnov \& Schaffer, 1973; Stearns, 1992; Franco \&
@@
 seasons. Conversely, environments with less frequent drought may select
-for perennial species, which may benefit from multiple bouts of
-reproduction and competitive advantage by preventing recruitment of
-annual species (Corbin \& D'Antonio, 2004). These predictions have been
-supported by the observation of annuals in arid environments in
-\emph{Oryza perennis} (Morishima \emph{et al.}, 1984) and
-\emph{Oenothera} (Evans \emph{et al.}, 2005). Additionally, annual and
-perennial species of \emph{Nemesia} were qualitatively associated with
-winter rather and summer rainfall environments respectively (Datson
-\emph{et al.}, 2008) and annual species of \emph{Scorzoneroides} were
-associated with environments classified as unpredictable (Cruz-Mazo
-\emph{et al.}, 2009). However, whether the history frequency of drought
-events indeed predicts the distributions annual or perennial life
-history strategies has yet to be tested.\\
+for perennial species, which benefit from multiple bouts of reproduction
+and competitive advantage by preventing recruitment of annual species
+(Corbin \& D'Antonio, 2004). These predictions have been supported by
+the observation of annuals in arid environments in \emph{Oryza perennis}
+(Morishima \emph{et al.}, 1984) and \emph{Oenothera} (Evans \emph{et
+al.}, 2005). Additionally, annual and perennial species of
+\emph{Nemesia} were qualitatively associated with winter rather and
+summer rainfall environments respectively (Datson \emph{et al.}, 2008)
+and annual species of \emph{Scorzoneroides} were associated with
+environments classified as unpredictable (Cruz-Mazo \emph{et al.},
+2009). However, whether the history frequency of drought events indeed
+predicts the distributions annual or perennial life history strategies
+has yet to be tested.\\
 Here we combine a long-term global dataset of satellite detected drought
@@
 Botanists have collected and maintained over 350 million botanical
-specimens worldwide over the past 300 years. Herbarium specimens and
-their associated metadata have been used since the 1960s to study
-species' geographical distributions (reviewed by Willis \emph{et al.}
-(2017) and Lang \emph{et al.} (2018)). And as they become digitized
-(Soltis, 2017), these collections have been used to study relationships
-between trait distributions, geography, and climate (Davis \emph{et
-al.}, 2015; Stropp \emph{et al.}, 2016; Wolf \emph{et al.}, 2016;
-Václavı'k \emph{et al.}, 2017). To characterize the distributions of
-annual and perennial \emph{Heliophila} species, all records for the
+specimens worldwide over the past 300 years (Thiers, 2016). Herbarium
+specimens and their associated metadata have been used since the 1960s
+to study species' geographical distributions (reviewed by Willis
+\emph{et al.} (2017) and Lang \emph{et al.} (2018)). And as they become
+digitized (Soltis, 2017), these collections have been used to study
+relationships between trait distributions, geography, and climate (Davis
+\emph{et al.}, 2015; Stropp \emph{et al.}, 2016; Wolf \emph{et al.},
+2016; Václavı'k \emph{et al.}, 2017). To characterize the distributions
+of annual and perennial \emph{Heliophila} species, all records for the
 genus \emph{Heliophila} were downloaded from the Global Biodiversity
@@
 \item
-  records from collection sites classified as land pixels\\
+  records from collection sites classified as land pixels in the VHI
+  dataset\\
 \item
@@
 frequency of drought during the winter, spring, summer, and fall between
-raw occurrence records of annual and perennial species by t-tests. To
+total occurrence records of annual and perennial species by t-tests. To
 account for variation in the number of occurrence records per species,
@@
 samples varied between species, with a mean of 52.19 samples per
-species. Heliophila rigidiuscula had the most records, 201, and
-Heliophila cornellsbergia the fewest, 2 (Table
-\ref{tab:speciesmeanstable}).
+species. \emph{H. rigidiuscula} had the most records, 201, and \emph{H.
+cornellsbergia} the fewest, 2 (Table \ref{tab:speciesmeanstable}).
 
@@
 found that the frequency of drought was significantly higher at the
-locations of occurrence records for annual species. In terms of raw
-observation records (Figure \ref{fig:mapsboxplots}g), the frequency of
-drought was significantly higher at the location of annuals during the
-winter (t = 10.65, p = 0.00), spring (t = 10.73, p = 0.00), summer (t =
-12.67, p = 0.00), and fall (t = 15.26, p = 0.00). Because raw occurrence
-records do not account for variation in the number of records per
-species (Table \ref{tab:speciesmeanstable}) or species relatedness
-(Figure \ref{fig:phylogeny}a), we also tested whether mean drought
-frequency values of each species were significantly different between
-annuals and perennials using phylogenetic logistic regression. We found
-that the mean drought frequencies were significantly higher
+locations of occurrence records for annual species. When comparing
+across all occurrence records (all records rather than species means,
+Figure \ref{fig:mapsboxplots}g), the frequency of drought was
+significantly higher at the location of annuals during the winter (t =
+10.65, p = 0.00), spring (t = 10.73, p = 0.00), summer (t = 12.67, p =
+0.00), and fall (t = 15.26, p = 0.00). Because a comparison across all
+occurrence records does not account for variation in the number of
+records per species (Table \ref{tab:speciesmeanstable}) or species
+relatedness (Figure \ref{fig:phylogeny}a), we also tested whether mean
+drought frequency values of each species were significantly different
+between annuals and perennials using phylogenetic logistic regression.
+We found that the mean drought frequencies were significantly higher
 (\(\alpha = 0.05\)) in annual species during the spring, summer, and
@@
 findings indicate that common acestry alone does not explain differences
-the drought frequencies experienced in the environments of annual and
-perennial \emph{Heliophila}.
+the drought frequencies experienced between the environments of annual
+and perennial \emph{Heliophila}.
 
@@
 \begin{threeparttable}
-\caption{\label{tab:modelstable}Phylogenetic logistic regressions between life history, and the mean drought frequency observed at herbaria collection sites of Heliophila species the winter, spring, summer, and fall.}
+\caption{\label{tab:modelstable}Phylogenetic logistic regressions between life history, and the mean drought frequency observed at specimen sites of Heliophila species the winter, spring, summer, and fall.}
 \begin{tabular}{lll}
@@
 
-The preceding results indicated that annual species are found in
+The preceding results indicate that annual species are found in
 environments where droughts are significantly more frequent, especially
@@
 predicted selection for annuality when adult morality is high (Stearns,
-1992), we also find evidence that the transition to perenniality could
-be explained by the historical frequency of drought, a likely cause of
-mortality in plants. The phylogeny reveals several transitions from
-annual to perennial life history (Figure \ref{fig:phylogeny}a) and that
+1992), we also find evidence that perenniality could be explained by
+reduced frequency of drought. The phylogeny reveals several transitions
+from annual to perennial life history (Figure \ref{fig:phylogeny}a) and
 the distributions of perennial \emph{Heliophila} extend into regions
@@
 strategies that allow plants to benefit from growth over many seasons.
-It may also indicate that annuals rely on drought as a source of
+This also suggests that annuals rely on drought as a source of
 disturbance for seedling recruitment when competing with perennials
@@
 
-We thank Jesse Lasky and members of the Sloan lab for generous feedback
-that improved the quality of this work. This research was supported by
-NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
+We thank members of the Sloan lab for generous feedback that improved
+the quality of this work. This research was supported by NSF Award
+1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
 
@@
 H}}. \textbf{1984}. Differentiation of perennial and annual types due to
-habitat conditions in the wild riceOryza perennis. \emph{Plant
+habitat conditions in the wild rice oryza perennis. \emph{Plant
 Systematics and Evolution} \textbf{144}: 119--135.
@@
 608--625.
+
+\leavevmode\hypertarget{ref-thiers2016index}{}%
+\textbf{\textnormal{Thiers B}}. \textbf{2016}. Index herbariorum: A
+global directory of public herbaria and associated staff. New york
+botanical garden's virtual herbarium. \emph{http://sweetgum. nybg.
+org/ih}.
 

### 2019-03-08 manuscript/manuscript.Rmd
revisions

@@
 
-abstract: Explaining variation in life history strategies is a long-standing goal of evolutionary biology.
-For plants, annual and perennial life histories are thought to reflect adaptation to environments that differ in the frequency of stress events such as drought.
-Here we test this hypothesis in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa, by integrating 34 years of satellite-based drought measurements with 2192 herbaria occurrence records.
-Consistent with predictions from classic life history theory, we find that perennial \textit{Heliophila} species occur in environments where droughts are significantly less frequent compared to annuals.
-These associations are predictive while controlling for phylogeny, lending support to the hypothesis that drought related natural selection has influenced the distributions of these strategies.
-Additionally, the collection dates of annual and perennial species indicate that annuals escape drought prone seasons during the seed phase of their life cycle.
-Together, these findings provide empirical support for classic hypotheses about the drivers of life history strategy in plants - that perennials out compete annuals in environments with less frequent drought and that annuals are adapted to environments with more frequent drought by escaping drought prone seasons as seeds.
+abstract: \begin{center}\textbf{Summary}\end{center} \par Explaining variation in life history strategies is an enduring goal of evolutionary biology and ecology.
+For plants, annual and perennial life histories have long been thought to reflect adaptation to environments that experience alternative drought regimes.
+Neverthless, empirical support for this hypothesis from phylogeographic analyses remains lacking.
+\par Here, we tested classic life history theory in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa, with a phylogentically constrained analysis integrating 34 years of satellite-based drought detection with 2192 herbaria occurrence records.
+\par We found that annual and perennial life histories have evolved multiple times in \textit{Heliophila} and that compared to annuals, perennial species occur in environments where droughts are significantly less frequent.
+We also report evidence that annual plants adapt to predictable drought regimes by escaping drought prone seasons as seeds.
+\par These results provide compelling support for longstanding theoretical predictions and a more nuanced understanding of life history evolution.
+More broadly, this work outlines a scalable framework integrating herbaria records and remote sensing to address outstanding questions in evolutionary ecology.
 
-keywords          : "drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records"
+keywords          : "drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records, Heliophia"
 
@@
 library(png)
+library(ggimage)
 ```
@@
 Understanding the causes and consequences of life history variation is a longstanding goal of ecology and evolutionary biology  [@cole1954population].
-In  plants, life histories are especially diverse, with herbaceous species that complete their life cycle in a number of weeks to trees that live for thousands of years [@brown1996oldlist].
-Along this continuum in angiosperms an important division exists, distinguishing annuals which complete their seed to seed life cycle within a single calendar year from perennials which can persist over multiple years.
+In  plants, life histories are especially diverse, with some species completing their life cycle in a number of weeks to others that live for thousands of years [@brown1996oldlist].
+Along this continuum in angiosperms an important division exists between annuals which complete their seed to seed life cycle within a single calendar year from perennials which can persist over multiple years.
 Annual plants flower once, set seed, senesce, and then die, spending at least some portion of the year as a seed, where they are relatively protected from environmental stress.
-In contrast, perennial plants can continue vegetative growth after reproduction and must survive conditions experienced during all seasons.
-These represent fundamentally different life history strategies, but the ecological factors that explain their evolution and distributions remain empirically uresolved [@friedman2015all].
-Classical theory predicts shorter life spans in environments where adult mortality is high [@charnov1973life; @stearns1992evolution;@franco1996life].
-In plants, this has been extended to the hypothesis that annuality is adaptive when it allows plants to escape drought [@schaffer1975selection].
-Lack of water is perhaps the greatest threat to survival during vegetative or reproductive growth and annuals can remain dormant (and protected as a seed) during drought.
-Thus, environments with greater seasonal drought frequency may select for annual life histories that complete reproduction prior to drought prone seasons.
-Conversely, environments with less frequent drought may select for perennial species, which benefit from multiple bouts of reproduction and competitive advantage by preventing recruitment of annual species [@corbin2004competition].
-These predictions have been supported by the observation of annuals in arid environments in _Oryza perennis_ [@morishima1984differentiation] and _Oenothera_ [@evans2005climate].
-Additionally, annual and perennial species of _Nemesia_ were qualitatively associated with winter rather and summer rainfall environments respectively [@datson2008climate] and annual species of _Scorzoneroides_ were associated with environments classified as unpredictable [@cruz2009molecular].
-However, whether the history frequency of drought events indeed predicts the distributions annual or perennial life history strategies has yet to be tested.
+In contrast, perennial plants can continue vegetative growth after reproduction and must survive conditions experienced during all seasons but can also benefit from multiple bouts of reproduction and competative advantages.
+These represent fundamentally different life history strategies and predicting their occurance is important for community, ecosystem, and agricultural ecology.
+However, the environmental factors that explain their evolution and distributions remain empirically uresolved [@friedman2015all].
+
+Classical theory predicts shorter life spans in environments where adult mortality is high [@charnov1973life;@stearns1992evolution;@franco1996life].
+Because lack of water is perhaps the greatest threat to survival during vegetative or reproductive growth in plants, this theory has been extended to the hypothesis that annuality is adaptive when it allows plants to escape drought [@schaffer1975selection].
+Indeed, adaptation to drought,  defined as episodes of increased aridity causing plant sresss [@passioura1996drought], is often invoked as an explanation for the success of annual species.
+And while a few cases are cited where annuality appears to be more common in environments with greater aridity [@stebbins1952aridity; @morishima1984differentiation], this hypothesis has yet to be supported in a phylogenetic context.
+In one previous study where this question was addressed phylogenetically,  [@evans2005climate] annuals were not found to be associated with environments that experience more drought.
+This could be explained by the relatively small number of species studied and the reliance on a limited number of weather stations to characterize environments, highlighting the need to develop more scalable methods to study trait evolution phylogeographically.
+
+It is also critical to consider another dimension of drought adaptation: the expectation that annuality is most adaptive when droughts are not only frequent but also predictable.
+That is, when the frequency of drought is particularly high during certain seasons.
+Such predictability is important for selection to lead to escape from those seasons which are paricuarlary drought prone.
+While there has been at least one example of annuality associated with environments qualitatively classified as "predictable" in a general sense [@datson2008climate], the seasonal predictability of drought experienced by annuals has yet to be rigorously studied.
+As such, further empirical work is needed to support the model of annuality as a mechanism of drought adaptation via escape from drought prone seasons.
+
+In addition to drought escape in annuals as a mechanism of adaptation to frequent predictable droughts, droughts may be necessary for the success of annuals more generally by acting as episodes of disturbance that provide opportunites for annuals to establish and compete with sympatric perennial species.
+Indeed, there is considerable evidence that perennials dominate in envirnoments where distrurbance events are infrequent [@rees1992germination;@corbin2004competition;@clary2012determinants].
+The resulting prediction from this hypothesis is that in the absence of frequent drought, perenniality should evolve.
+However, little is known about this component of life history evolution because previous work has largely focused on the origins of annuality rather than perenniality [@friedman2015all].
+
 Here we combine a long-term global dataset of satellite detected drought events with metadata from natural history collections to test these classic hypotheses within the African endemic mustard genus, _Heliophila_ L.
@@
 If annuality is an adaptive strategy allowing plants to escape drought prone seasons, then drought frequency should predict the distribution of life history strategies across landscapes, and annual species should be more commonly associated with drought prone regions than perennial species.
-Furthermore, if annual species have adapted to escape drought prone seasons, observations of growing annual species (i.e.
-occurring in forms other than seed) should be rare during drought prone seasons.
-Phylogenetic relatedness can influence tests of associations between species’ traits and their environments [@barrett1996comparative;@felsenstein1985phylogenies], and therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetic context.
+Additionally, if perenniality offers competative advantage in the absence of drought, associations between life history and drought frequency should be signficant in even if phylogenies are enriched for transitions from annual to perenial life history strategy.
+Finally, if annual species have adapted to escape predictably drought prone seasons, observations of growing annual species (i.e.
+occurring in forms other than seed) should be rare during seasons when drought frequency is hightest.
+Phylogenetic relatedness can influence tests of associations between species’ traits and their environments [@barrett1996comparative;@felsenstein1985phylogenies].
+Therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetic context.
 
@@
 ## Data  
-### Availability
+### Data availability
 All analyses were performed using R. All data and the source code to produce this manuscript are available at https://github.com/greymonroe/heliophila. Software used is listed in the supplement.
 
-### Satellite-detected drought data
-```{r Droughtdata}
-Winter<-raster("../data/drought/Summer_drought_freq")
-Spring<-raster("../data/drought/Fall_drought_freq")
-Summer<-raster("../data/drought/Winter_drought_freq")
-Fall<-raster("../data/drought/Spring_drought_freq")
-```
-Remotely sensed data is a powerful tool for characterizing seasonal patterns in drought because it is less limited in spatial and temporal scope and resolution than weather stations or field observations [@aghakouchak2015remote].
-To quantify the frequency of drought during different seasons across landscapes, we used the remotely sensed Vegetative Health Index (VHI), which measures landscape scale reductions in plant cover and temperature conditions characteristic of drought [@kogan2001operational].
-Generated from data collected by NOAA AVHRR satellites since 1981, the VHI combines Normalized Difference Vegetation Index (NDVI) derived measures of vegetative stress (Vegetative Condition Index - VCI) with temperature stress indicated by anomalies in thermal spectra (Temperature Condition Index - TCI).
-The VHI of year $y$ during week $w$ of $[1,52]$ at pixel $i$ is derived from the following equations, where $n$ is the number of years observed.
-
-$$VCI_{y,w,i} = 100\frac{NDVI_{y,w,i} - NDVI_{min}}{NDVI_{max} - NDVI_{min}}$$
-
-$$TCI_{y,w,i} = 100\frac{T_{y,w,i} - T_{min}}{T_{max} - T_{min}}$$
-
-$$VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})$$
-where $NDVI_{min} = min(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})$ and $NDVI_{max} = max(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})$ and $T_{min} = min(T_{1981,w,i}...T_{1981+n,w,i})$ and $T_{max} = max(T_{1981,w,i}...T_{1981+n,w,i})$
-
-Thus, VHI measurements are standardized according to conditions historically observed at each locations.
-These measurements have been validated and generally used for evaluating drought risk and predicting crop yields in agriculture [e.g., @rojas2011assessing; @kogan2016modelling].
-But they also present a new tool to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes [@kerr2003space].
-As such, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting intraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought].
-Here, we accessed VHI data at $16km^2$ resolution from 1981 to 2015 (https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php) to characterize the seasonal drought frequencies experienced by annual and perennial _Heliophila_ species.
-
 ### Life history data for _Heliophila_
-_Heliophila_ is a genus of flowering plants endemic to the southern portion of Africa including the Cape Floristic and Succulent Karoo Regions.
+_Heliophila_ is a genus of flowering plants endemic to the southern Africa including the Cape Floristic and Succulent Karoo Regions.
 These are among the most botanically diverse environments on Earth and the _Heliophila_ species occurring there are considered to make up the most diverse genus of the family Brassicaceae [@mummenhoff2005phylogeny;@mandakova2012whole].
-This genus includes both perennial and annual species and this change in life history strategy has likely arisen multiple independent times [@appel1997generic;@mummenhoff2005phylogeny].
+This genus includes both perennial (n =21) and annual (n = 21) species and this change in life history strategy has likely arisen multiple independent times [@appel1997generic;@mummenhoff2005phylogeny].
 Furthermore, the fine scale climatic heterogeneity of Southern Africa is ideal for studying the distribution of traits in relation to environmental parameters [@sayre2013new].
@@
 ```
-Botanists have collected and maintained over 350 million botanical specimens worldwide over the past 300 years [@thiers2016index].
-Herbarium specimens and their associated metadata have been used since the 1960s to study species’ geographical distributions (reviewed by  @willis2017old and @lang2018using).
+To characterize the distributions of annual and perennial _Heliophila_ species, all (`r nrow(GBIF)`) records for the genus _Heliophila_ were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018 [@gbifdownload].
+herbaria records records such as these provide a rich data sources to characterize the geographical distributions of species [@thiers2016index; @willis2017old; @lang2018using].
 And as they become digitized [@soltis2017digitization], these collections have been used to study relationships between trait distributions, geography, and climate [@davis2015herbarium;@wolf2016altitudinal;@stropp2016mapping;@vaclavik2017effects].
-To characterize the distributions of annual and perennial _Heliophila_ species, all records for the genus _Heliophila_ were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018 [@gbifdownload].
 
 ### Sequence data for phylogeny
-An alignment of ITS I and II sequences for _Heliophila_ species was obtained from the authors of @mandakova2012whole.
+An alignment of ITS I and II sequences for 21 annual and 21 perennial _Heliophila_ species was obtained from the authors of @mandakova2012whole.
 Individual ITS I and II sequences for _Aethionema grandiflorum, Alliaria petiolata, Cardamine matthioli, Chamira circaeoides_, and _Rorippa amphibia_ were downloaded from Genbank.
+
+```{r Droughtdata}
+Winter<-raster("../data/drought/Summer_drought_freq")
+Spring<-raster("../data/drought/Fall_drought_freq")
+Summer<-raster("../data/drought/Winter_drought_freq")
+Fall<-raster("../data/drought/Spring_drought_freq")
+
+# Winter<-raster("~/VHI/droughtfreq/summer_mild")
+# Spring<-raster("~/VHI/droughtfreq/fall_mild")
+# Summer<-raster("~/VHI/droughtfreq/winter_mild")
+# Fall<-raster("~/VHI/droughtfreq/spring_mild")
+```
+
+### Satellite-detected drought data
+
+Remotely sensed data is a powerful tool for characterizing seasonal patterns in drought because it is less limited in spatial and temporal scope and resolution than weather stations or field observations [@aghakouchak2015remote].
+From an ecological perspective, droughts are best defined as episodes of plant stress caused by elevated aridity [@passioura1996drought].
+Thus remote sensing offers the additional benefit for studying drought as an agent of natural selection because plant stress caused by drought can be observed from space [@kogan1995droughts].
+The remotely sensed Vegetative Health Index (VHI) is one such metric, which detects landscape scale reductions in plant cover and temperature conditions characteristic of drought [@kogan2001operational].
+Generated from data collected by NOAA AVHRR satellites since 1981, the VHI is a composite index combining Normalized Difference Vegetation Index (NDVI) derived quantification of vegetative stress (Vegetative Condition Index - VCI) with temperature stress indicated by anomalies in thermal spectra (Temperature Condition Index - TCI).
+These indicies were developed to create an unbiased quantification of drought across ecosystem types.
+The VHI of year $y$ during week $w$ of $[1,52]$ at pixel $i$ is derived from the following equations, where $n$ is the number of years observed.
+
+$$VCI_{y,w,i} = 100\frac{NDVI_{y,w,i} - NDVI_{min,w,i}}{NDVI_{max,w,i} - NDVI_{min,w,i}}$$
+
+Low values of VCI indicate episodes when plant cover is particularly low for a given location during a given time of the year.
+Thus, it controls for the location and season in quantifying plant stress.
+
+$$TCI_{y,w,i} = 100\frac{T_{max,w,i}-T_{y,w,i}}{T_{max,w,i} - T_{min,w,i}}$$
+Similarly, low TCI values indicate episdoes of high thermal stress shown to be negatively correlated with precipition and soil moisture [@aghakouchak2015remote].
+
+$$VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})$$
+By combining VCI and TCI, the VHI is distinguishes drought from other forms of vegetative stress [@kogan1995application].
+The use of the VHI to detect drought has been validated globally and across ecosystem types [@aghakouchak2015remote], including in southern Africa, the focal region of this study (Figure \@ref:fig(mapsdroughtexamples)).
+To date, the VHI has most often been applied for evaluating drought risk for agricultural research [e.g., @rojas2011assessing; @kogan2016modelling].
+But it also presents a tool to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes [@kerr2003space].
+As such, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting intraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought].
+Here, we accessed VHI data at $16km^2$ resolution from 1981 to 2015 (https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php) to characterize the seasonal drought frequencies experienced by annual and perennial _Heliophila_ species across their native range of southern Africa.
 
@@
 To characterize drought regimens across the distributions of annual and perennial species of _Heliophila_, we calculated drought during different seasons at the location of observations for _Heliophila_ records using the VHI.
-Specifically, we created global maps of the frequencies of observing drought conditions (VHI<40, NOAA) during the winter (quarter surrounding winter solstice), spring (quarter surrounding spring equinox), summer (quarter surrounding summer solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015.
+Specifically, we created global maps of the frequencies of observing drought conditions between years (VHI<40, NOAA) during the winter (quarter surrounding winter solstice), spring (quarter surrounding spring equinox), summer (quarter surrounding summer solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015.
 From these maps, the drought frequency during the winter, spring, summer, and fall were extracted for the locations of all GBIF records.
@@
 
-### Phylogeny construction
-Out group ( _Aethionema grandiflorum, Alliaria petiolata, Cardamine matthioli, Chamira circaeoides_, and _Rorippa amphibia_) and ingroup _Heliophila_ ITS I and II sequences were aaligned using MAFFT [@katoh2002mafft] with strategy G-INS-I, offset value 0.1, and all other options set as default.
+### Phylogeny construction and ancestral state estimation
+Outgroup ( _Aethionema grandiflorum, Alliaria petiolata, Cardamine matthioli, Chamira circaeoides_, and _Rorippa amphibia_) and ingroup _Heliophila_ ITS I and II sequences were aligned using MAFFT [@katoh2002mafft] with strategy G-INS-I, offset value 0.1, and all other options set as default.
 The $GTR + \Gamma$ model of nucleotide substitution was determined to best fit the data based on AIC using jModelTest2 [@darriba2012jmodeltest;@guindon2003a].
@@
 Model selection and phylogenetic analyses were conducted through the CIPRES Science Gateway [@miller2010creating].
+Ancestral state estimation was performed in R using the package phytools [@revell2012phytools] to genearte 10,000 stochastic character maps simulated under an equal rates model of character evolution for the trait life habit (annual or perennial).
 
@@
 To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimes, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency.
-First, we compared the frequency of drought during the winter, spring, summer, and fall between total occurrence records of annual and perennial species by t-tests.
+First, we compared the frequency of drought across all occurrence records of between annual and perennial species by t-tests.
+Independent tests were performed to compare drought frequency between annuals and perennials during each season.
+We also tested for a relationship between drought frequency and life history, season, and their interaction by an analysis of variance while including species as a random effect.
 To account for variation in the number of occurrence records per species, we next calculated the mean drought frequency during the winter, spring, summer and fall for each species.
 Because shared evolutionary history of closely related species can lead to spurious associations between traits and environments  [@felsenstein1985phylogenies], we tested for a relationship between life history strategy and drought frequency while controlling for phylogeny using phylogenetic logistic regression [@ives2010phylogenetic].
+This statistical approach is designed to control for the confounding effects of geographic relationships explained by demographic history rather than processes of natural selection.
 
@@
 Orange circles at branch tips mark annual species and blue circles mark perennial species.
+At internal nodes, pie charts indicate the estimated posterior probability of being annual versus perennial.
 Example herbaria specimens accessed via GBIF of (a) _H.
@@
 
-pdf("../figures/phylogeny.pdf", height=6, width=5)
+# source("../src/ancestral state.R") #uncomment to rerun the ancestral state reconstruction. Takes a few minutes.
 
-tree<-read.nexus("../data/TrimmedMCCwoPolytomies.nex")
-tree<- rescale(tree, "depth", 1)
+pdf("../figures/phylogeny.pdf", height=5.5, width=4.5)
+
+tree <- read.nexus("../data/TrimmedMCCwoPolytomies.nex")
+ladderize(tree) -> tree
 tree$tip.label<-gsub("_"," ", tree$tip.label)
-tree<-keep.tip(tree, tip =as.character(species_means$species))
+rescale(tree, "depth", 1) -> tree
 
+data <- read.csv("../data/species_means_table.csv", header=T)
+data$A_P_num<-as.numeric(data$LH)-1
+
+pies<-read.csv("../data/ancestral_state.csv")
 treeplot<-ggtree(tree, layout = "rectangular")+
   xlim(c(0,2))+
-  geom_tippoint(size=3,col=ifelse(species_life_history$life_history[match(tree$tip.label,species_life_history$species)]=="a", "orange3", "dodgerblue3"))+
+  geom_tippoint(size=3,col=ifelse(data$A_P_num[match(tree$tip.label,data$Species)]=="0", "orange3", "dodgerblue3"))+
   geom_tiplab(offset = .05 , size=3)+
   theme(plot.background = element_rect(fill=alpha("white",0)))
+
+treeplot<-inset(treeplot, nodepie(pies, cols = c("X0", "X1" ), color  = c("orange3", "dodgerblue3" )), width=2.7, height=2.7)
 
@@
 
-plot_grid(treeplot, pics, ncol=2,labels=c("a",""), rel_widths = c(1,0.5))
+plot_grid(treeplot, pics, ncol=2,labels=c("a",""), rel_widths = c(1.9,0.5))
 
@@
 ```
+
 ```{r phylogeny, fig.cap='(ref:phylogeny)', out.width = "\\textwidth", fig.pos = "!h"}
@@
 ```
-(ref:mapsboxplots) Locations of occurrence records of (a) annual and (b) perennial _Heliophila._ Drought frequency during the (c) winter, (d) spring, (e) summer and (f) fall measured using the VHI.
-(g) Drought frequencies during each season at the observation locations of annual and perennial _Heliophila_ (t tests, \*\* = p < 0.01).
+(ref:mapsboxplots) Locations of occurrence records of (a) annual and (b) perennial _Heliophila._ Drought frequency during the (c) winter, (d) spring, (e) summer and (f) fall detected using the VHI.
+(g) Drought frequencies during each season at the observation locations of annual and perennial _Heliophila_ (t tests contrasting annuals and perennials, \*\* = p < 0.01).
 
@@
 ```
+The _Heliophila_ phylogeny was consistent with previous studies [@mummenhoff2005phylogeny;@mandakova2012whole].
+Based on 10,000 stochastic character maps simulated under an equal rates model of character evolution in life history, an average of approximately eight changes from annual to perennial and five changes from perennial to annual are observed per stochastic character map (Figure \@ref(fig:phylogeny)a).
+These results suggest that the ancestral state of _Heliophila_ was annual and that both charcter states have arisen independently mutliple times.
+
 Out of `r nrow(GBIF)` _Heliophila_ GBIF records,  `r sum(GBIF$species %in% species_life_history$species)` were for species with reported life history [@mummenhoff2005phylogeny], `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude))` had geospatial data, `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)))` did not have geospatial issues, `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)) & !(is.na(GBIF$Summer)) & !(is.na(GBIF$decimalLatitude)))` were located on pixels classified as land having drought measurements, `r sum(GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & (GBIF$species %in% species_life_history$species) & !(is.na(GBIF$Summer)))` were located in Africa, `r sum(!duplicated(GBIF$ymd_lat_lon_sp) & GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & (GBIF$species %in% species_life_history$species) & !(is.na(GBIF$Summer))) ` were not duplicated.
@@
 There were clear visual differences between the distributions of the `r sum(GBIF_cleaned$life_history=="a")` annual and the `r sum(GBIF_cleaned$life_history=="p")` perennial _Heliophila_ observation records (see Figure \@ref(fig:speciesmaps) for maps of individual species).
-While annual species were generally found in the western regions of South Africa and Namibia, primarily in the Cape Floristic Region and Succulent Karoo (Figure \@ref(fig:mapsboxplots)a), the occurrence of perennials extended to the east coast of South Africa (Figure \@ref(fig:mapsboxplots)b).
+While annual species were generally found in the western regions of South Africa and Namibia, primarily in the Cape Floristic Region and Succulent Karoo (Figure \@ref(fig:mapsboxplots)a), the occurrence of perennials extended to the southern and eastern coast of South Africa (Figure \@ref(fig:mapsboxplots)b).
 
@@
 tfal<-t.test(Fall~life_history, GBIF_cleaned)
+
+melted<-GBIF_cleaned %>% select(life_history, Winter, Spring, Summer,Fall, species) %>% gather(Winter, Spring, Summer,Fall, key="season",value="droughtfreq")
+library(nlme)
+fit<-lme(droughtfreq~life_history*season, random=~1|species, melted)
+anovatable<-data.frame(predictor=c("intercept","life history","season","life history x season"), anova(fit))
+row.names(anovatable)<-NULL
+
 ```
@@
 
-Theory predicts that annuality should be adaptive in places where stresses such as drought are more common.
-Conversely, perenniality should be adaptive in places where such stresses are less frequent.
-We found that the frequency of drought was significantly higher at the locations of occurrence records for annual species.
-When comparing across all occurrence records
-(all records rather than species means, Figure \@ref(fig:mapsboxplots)g), the frequency of drought was significantly higher at the location of annuals during the winter (t = `r twin$statistic`, p = `r twin$p.value`), spring (t = `r tspr$statistic`, p = `r tspr$p.value`), summer (t = `r tsum$statistic`, p = `r tsum$p.value`), and fall (t = `r tfal$statistic`, p = `r tfal$p.value`).
+We found that the frequency of drought was significantly higher at the locations of occurrence records for annual species. When comparing across all occurrence records
+(all records rather than species means, Figure \@ref(fig:mapsboxplots)g), the frequency of drought was significantly higher at the location of annuals during the winter (t = `r twin$statistic`, p < 0.01), spring (t = `r tspr$statistic`, p < 0.01), summer (t = `r tsum$statistic`, p = < 0.01), and fall (t = `r tfal$statistic`, p < 0.01).
+A mixed-model analysis of variace which included species as random effect also identified a signficant relationship between drought frequency and life history (p < 0.01), season (p < 0.01) and their interaction (p < 0.01) (Table \@ref(tab:anovatable)).
 Because a comparison across all occurrence records does not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable)) or species relatedness (Figure \@ref(fig:phylogeny)a), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials using phylogenetic logistic regression.
-We found that the mean drought frequencies were significantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a).
-These findings indicate that common acestry alone does not explain differences the drought frequencies experienced between the environments of annual and perennial _Heliophila_.
+We found that the mean drought frequencies were significantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a,b).
+These findings indicate that demographic history alone does not explain differences in the drought frequencies experienced between the environments of annual and perennial _Heliophila_.
 
@@
 
-(ref:lineplots) (a) Comparison (mean +- SE) of drought frequency across seasons measured at the GBIF records of annual and perennial species of _Heliophila_.
-(phylogenetic logistic regression, \* = p < 0.05, \*\* = p < 0.01) (b) Collection dates of GBIF records of annual and perennial species of _Heliophila_.
+(ref:lineplots) (a) Comparison (mean +- SE) of drought frequency across seasons calcualted at the occurance locations of GBIF records of annual and perennial species of _Heliophila_.
+(b) Results from phylogenetic logistic regression, where the line shows the model fit and \* = p < 0.05, \*\* = p < 0.01.
+Annuals were scored as 0 and perennials as 1.
+(c) Collection dates of GBIF records of annual and perennial species of _Heliophila_ in relation to the photoperiodic calender where day 1 is intermediate to the fall equinox and winter soltice.
 
@@
 
-pdf("../figures/line_and_dates.pdf", height=5, width=5)
+pdf("../figures/line_and_dates.pdf", height=6, width=5)
 
@@
 
-lines<-ggplot(life_history_drought_means, aes(x=variable, col=life_history,group=life_history, y=mean))+
-  theme_classic()+
+lines<-ggplot(life_history_drought_means, aes(x=life_history, col=life_history,group=life_history, y=mean))+
+  theme_bw()+
   geom_point(size=2)+
-  geom_line(lwd=0.5, lty=1)+
+  #geom_line(lwd=0.5, lty=1)+
+  facet_grid(~variable, labeller = as_labeller(c("Winter_mean"="Winter","Spring_mean"="Spring","Summer_mean"="Summer","Fall_mean"="Fall")))+
   geom_errorbar(width=0.25,lwd=1, aes(ymin=mean-se, ymax=mean+se))+
   scale_y_continuous(limits=c(0.3,0.45), labels = c("0.300", "0.350", "0.400", "0.450"))+
-  scale_x_discrete(labels=c("Winter","Spring","Summer","Fall"))+
+  scale_x_discrete(labels=NULL)+
   scale_color_manual(values=c( "orange3","dodgerblue3"), name="Life history", labels=c("Annual", "Prennial"))+
-  labs(x="Season", y="Drought Frequency")+
-  annotate("text", x=2, y=0.44, label="*", size=7) +
-  annotate("text", x=3:4, y=0.44, label="**", size=7)
+  labs(x=NULL,y="Drought Frequency")+
+  theme(axis.ticks.x =element_blank(), strip.background =element_rect(fill="white"))
 
@@
   scale_fill_manual(values=c("orange3","dodgerblue3"), name="Life history", labels=c("Annual", "Perennial"))+
-  theme_classic()+
-  scale_x_continuous(name="Season", breaks=c(45,135,225, 315), labels=c("Winter","Spring","Summer","Fall"))+
-  labs(title="Date of observation for occurrence records", y="Density")+
-  theme(plot.title = element_text(hjust = 0.5, size=10))
+  theme_bw()+
+  scale_x_continuous(name="Day of photoperiodic year", breaks=c(45,135,225, 315), labels=c("45\n(Winter)","135\n(Spring)","225\n(Summer)","315\n(Fall)"))+
+  labs(title="Timing of observation for occurrence records", y="Density")+
+  theme( plot.title = element_text(hjust = 0.5, size=10))
 
-plot_grid(lines, dates, ncol=1, labels="auto")
+tree<-read.nexus("../data/TrimmedMCCwoPolytomies.nex")
+tree<- rescale(tree, "depth", 1)
+tree<-keep.tip(tree, tip = gsub(" ","_", as.character(species_means$species)))
+
+species_means$life_history_num<-as.numeric(species_means$life_history)-1
+row.names(species_means)<-gsub(" ", "_", species_means$species)
+
+fit <-  phyloglm(species_means$life_history_num~species_means$Spring_mean, data=species_means, phy=tree)
+cc <- coef(fit)
+spcurve<-plogis(cc[1]+cc[2]*sort(species_means$Spring_mean))
+
+fit <-  phyloglm(species_means$life_history_num~species_means$Summer_mean, data=species_means, phy=tree)
+cc <- coef(fit)
+scurve<-plogis(cc[1]+cc[2]*sort(species_means$Summer_mean))
+
+fit <-  phyloglm(species_means$life_history_num~species_means$Fall_mean, data=species_means, phy=tree)
+cc <- coef(fit)
+fcurve<-plogis(cc[1]+cc[2]*sort(species_means$Fall_mean))
+
+fit <-  phyloglm(species_means$life_history_num~species_means$Winter_mean, data=species_means, phy=tree)
+cc <- coef(fit)
+wicurve<-plogis(cc[1]+cc[2]*sort(species_means$Winter_mean))
+
+curvedataframe<-data.frame(
+  droughtfreq=c(sort(species_means$Winter_mean),
+                sort(species_means$Spring_mean),
+                sort(species_means$Summer_mean),
+                sort(species_means$Fall_mean)),
+  season=rep(c("Winter","Spring","Summer","Fall"), each=42),
+  curves=c(wicurve, spcurve, scurve, fcurve),
+  lifehistory=c(species_means$life_history_num[order(species_means$Winter_mean)],
+                species_means$life_history_num[order(species_means$Spring_mean)],
+                species_means$life_history_num[order(species_means$Summer_mean)],
+                species_means$life_history_num[order(species_means$Fall_mean)])
+                           )
+curvedataframe$lifehistory<-curvedataframe$lifehistory
+curvedataframe$season<-factor(curvedataframe$season, levels=c("Winter","Spring","Summer","Fall"))
+
+fit_stats<-data.frame(season=factor(c("Winter","Spring","Summer","Fall"),levels=c("Winter","Spring","Summer","Fall")), label=c("","*","**","**"), lifehistory="1")
+
+fitplots<-ggplot(curvedataframe, aes(x=droughtfreq, col=as.factor(lifehistory), y=lifehistory))+
+  geom_jitter(height = .03, width=0, alpha=0.7)+
+  facet_grid(~season)+
+  theme_bw()+
+  scale_color_manual(values=c( "orange3","dodgerblue3"), name="Life history", labels=c("Annual", "Prennial"))+  scale_y_continuous(breaks = c(0,1), labels=c("       a","       p"))+
+  geom_line(aes(x=droughtfreq, y=curves), col="black")+
+  theme(axis.text.x = element_text(angle=90), strip.background = element_rect(fill="white"))+
+labs(y="Life history", x="Drought frequency")+
+  geom_text(data=fit_stats, aes(x=.4, y=.75, label=label), col="black", cex=5)
+
+plot_grid(lines, fitplots, dates, ncol=1, labels="auto", rel_heights = c(1,1.2,1.5))
 
@@
 ```
+
 The preceding results indicate that annual species are found in environments where droughts are significantly more frequent, especially in the summer and fall.
-Classic life history theory hypothesizes that annuality reflects adaptation to such environments because it allows species to escape stressful conditions.
+Classic life history theory hypothesizes that annuality reflects adaptation to such environments because it allows species to escape predictable stressful conditions.
 If this is the case, we would expect that annuals spend the drought prone seasons of summer and fall as seeds.
@@
 In comparison to perennials, which appear to be collected throughout the year, annuals are almost exclusively observed during the winter and spring (Figure \@ref(fig:lineplots)b).
-The differences between the distribution of collection dates were significant by all tests (ks.test D = `r  kout$statistic`, p = `r  kout$p.value`; bartlett.test K2 = `r bout$statistic`, p = `r bout$p.value`) This is consistent with a model of life history in which annual species flower in the spring, set seed, senesce, and die before the summer.
+The differences between the distribution of collection dates were significant by all tests (ks.test D = `r  kout$statistic`, p < 0.01; bartlett.test K2 = `r bout$statistic`, p < 0.01) This is consistent with a model of life history in which annual species flower in the spring, set seed, senesce, and die before the summer.
 Thus, these annual species are likely to remain dormant during the summer and fall, when drought is the strongest predictor of the distributions of annual and perennial life histories (Figure \@ref(fig:lineplots)a).
@@
 We cannot eliminate the possibility that confounding traits or environmental variables are the causative factors explaining variation in the distributions of annual and perennial species.
-Nevertheless, these results provide quantitative support for the classic prediction that annual species are found in environemnts that experience more frequent drought than perennial species.
-These findings complement previous reports of qualitative associations between annuality with enviroments characterized as having increased aridity [@evans2005climate], alternative precipitation defined habitats [@morishima1984differentiation;@datson2008climate], or greater unpredictability [@cruz2009molecular].
-However, to our knowledge this is the first study to demonstrate a significant association between life history and drought in a phylogenetic context informed by large scale species distribution data and long term drought measures.
+Nevertheless, these results provide quantitative support for the classic prediction that annual species are found in environments that experience more frequent drought than perennial species.
+These findings are a valuable contribution to previous investigations of associations between life history and climate [@evans2005climate;@morishima1984differentiation;@datson2008climate; @cruz2009molecular].
+To our knowledge this is the first study to demonstrate a significant association between life history and drought in a phylogenetic context informed by large scale species distribution data and long term drought detection.
 
@@
 
-These findings support classical theoretical predictions about the adaptive value of annual and perennial life history strategies.
+These findings empirical provide support classical theoretical predictions about the adaptive value of annual and perennial life history strategies.
 Taken together, they suggest that in _Heliophila_, annual species are adapted to environments with increased summer droughts by avoiding these seasons in a dormant seed phase of their life cycle.
@@
 While most previous work has focused on describing the evolutionary origins of annuality [@barrett1996comparative;@conti1999phylogenetic;@andreasen2001unequal;@verboom2004testing;@friedman2015all] there are at least a few other cases where perenniality appears to have arisen from an annual ancestor [@bena1998molecular;@tank2008annuals].
-And while early theory predicted selection for annuality when adult morality is high [@stearns1992evolution],  we also find evidence that perenniality could be explained by reduced frequency of drought.
+And while early theory predicted selection for annuality when adult mortality is high [@stearns1992evolution],  we also find evidence that perenniality could be explained by reduced frequency of drought.
+This is supported by the theoretical prediction that perenniality is advantageous is stable habitats.
 The phylogeny reveals several transitions from annual to perennial life history (Figure \@ref(fig:phylogeny)a) and the distributions of perennial _Heliophila_ extend into regions where drought frequency is low (Figure \@ref(fig:mapsboxplots)b, Figure \@ref(fig:speciesmaps)).
-Perennials may be able to out complete annual relatives in environments where the infrequency of drought favors strategies that allow plants to benefit from growth over many seasons.
+Perennials may be able to out compete annual relatives in environments where the infrequency of drought favors strategies that allow plants to benefit from growth over many seasons.
 This also suggests that annuals rely on drought as a source of disturbance for seedling recruitment when competing with perennials [@corbin2004competition].
@@
 These findings suggest that species with locally adaptive life history strategies could be threatened by rapidly changing drought regimens [@dai2011drought].
-This could have impacts on ecosystem functioning and processes such as carbon cycling if life history traits evolve or the composition of annual and perennial species changes in response [@garnier1997specific; @roumet2006suites;@monroe2018ecoevolutionary].
-Furthermore, the frequency of drought may be an important factor when considering the use of perennial cropping systems [@parry2005prospects;@lelievre2009current].
+In light of the findings here, changes in eastern South Africa which is forecast to experience signicantly lower rainfall [@south2017a] could be particularly impactful.
+In this study we found that this region is currently dominiated by derived perennial species of _Heliophila_.
+However, if drought become more frequent in this region, it may allow for the establishement of annuals.
+Such changes in selection patterns and shifts in plant functional diversity could have impacts on ecosystem functioning and processes such as carbon cycling [@garnier1997specific; @roumet2006suites;@monroe2018ecoevolutionary].
+Furthermore, the changes in frequency of drought may be an important factor when considering the use of perennial cropping systems [@parry2005prospects;@lelievre2009current].
 
@@
 # Acknowledgments 
-We thank members of the Sloan lab for generous feedback that improved the quality of this work. This research was supported by NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
+We thank Jesse Lasky, members of the Sloan lab at CSU and three ananymous reviewers for generous feedback that improved the quality of this work.
+This research was supported by NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
 
@@
 # References
-\newpage
+
 ```{r create_r-references}
@@
 
-### Software used
-We used `r cite_r("r-references.bib")` for all our analyses.
+### Supplementary tables and figures  
 \newpage
 
-### Supplementary tables and figures
+```{r anovatable, results='asis', echo = F}
+apa_table(anovatable
+          , caption = "Analysis of variance (ANOVA) to compare drought frequency as a function of life history, season, and their interaction while including species as a random effect.",
+          digits=4
+)
+```
+
 ```{r speciesmeanstable, results = 'asis', echo = F}
@@
 )
+
+write.csv(species_means_table, "../data/species_means_table.csv")
+```
+
+```{r results = 'hide', echo = F}
+write.csv(species_means_table, "../data/species_means_table.csv")
+```
+
+(ref:mapsdroughtexamples) Example years experiencing contrasting degrees of drought in southern Africa.
+Vegetative Health Index (VHI) values below 40 indicate remotely sensed drought.
+Drought detection during this years is validated by precipitation based estimates of drought ocurrance (Standardized Precipitation Index) which confirm that while (a) 2015 was one of the worst drought years on record, (b) 1999 was one of the wettest, and (c) 2012 was typical in terms of precipitation patterns [@monyela2017two].
+It is worth noting that drought can be detected using the VHI across ecosystems, including those inhabited by perennial rather than annual _Heliophila_ species.
+
+```{r make_drought_examples, results = 'hide', echo = F, warnings="hide"}
+pdf("../figures/maps_drought_examples.pdf", width=3, height=6)
+
+world <- map_data("world") # we already did this, but we can do it again
+col_scale<-c("white","black")
+
+drought<-raster::raster("~/VHI/raster/VHP.G16.C07.npp.P2015050.VH.nc.grd")
+drought <-crop(drought, rbind(c(12,37),c(-35, -18)))
+
+summer2015<-gplot(drought) + 
+  geom_tile(aes(fill=value<4000)) + 
+  scale_fill_manual(values = col_scale,  na.translate = F, name="VHI < 40") +
+  geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
+  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
+  theme_classic()+
+  theme(legend.position = c(0.92,.3),legend.text=element_text(size=rel(0.9)), legend.key.size=unit(.2,"line"), legend.background = element_rect(fill=alpha('gray90'), color = "black"), text = element_text(size=4.5))+
+  labs(title="VHI detected drought during summer 2015", x="Longitude (°)", y="Latitude (°)")
+
+nondrought<-raster::raster("~/VHI/raster/VHP.G16.C07.NJ.P1999050.VH.hdf.h5.grd")
+nondrought <-crop(nondrought, rbind(c(12,37),c(-35, -18)))
+
+summer1999<-gplot(nondrought) + 
+  geom_tile(aes(fill=value<4000)) + 
+  scale_fill_manual(values = col_scale,  na.translate = F, name="VHI < 40") +
+  geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
+  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
+  theme_classic()+
+  theme(legend.position = c(0.92,.3),legend.text=element_text(size=rel(0.9)), legend.key.size=unit(.2,"line"), legend.background = element_rect(fill=alpha('gray90'), color = "black"), text = element_text(size=4.5))+
+  labs(title="VHI detected drought during summer 1999", x="Longitude (°)", y="Latitude (°)")
+
+typical<-raster::raster("~/VHI/raster/VHP.G16.C07.NP.P2012050.VH.hdf.h5.grd")
+typical <-crop(typical, rbind(c(12,37),c(-35, -18)))
+
+summer2012<-gplot(typical) + 
+  geom_tile(aes(fill=value<4000)) + 
+  scale_fill_manual(values = col_scale,  na.translate = F, name="VHI < 40") +
+  geom_polygon(data = world, aes(x=long, y = lat, group=group),  fill = NA, color = "black") + 
+  coord_fixed(xlim=c(13,35), ylim=c(-35, -22)) + 
+  theme_classic()+
+  theme(legend.position = c(0.92,.3),legend.text=element_text(size=rel(0.9)), legend.key.size=unit(.2,"line"), legend.background = element_rect(fill=alpha('gray90'), color = "black"), text = element_text(size=4.5))+
+  labs(title="VHI detected drought during summer 2012", x="Longitude (°)", y="Latitude (°)")
+
+plot_grid(summer2015, 
+          summer1999,
+          summer2012, 
+          labels = "auto", ncol = 1)
+
+dev.off()
+
+```
+
+```{r mapsdroughtexamples, fig.cap='(ref:mapsdroughtexamples)'}
+knitr::include_graphics("../figures/maps_drought_examples.pdf", dpi = 108)
 ```

### 2019-03-08 manuscript/manuscript.tex
revisions

@@
             pdfkeywords={drought adaptation, life history evolution, remote sensing,
-phylogeography, herbaria records},
+phylogeography, herbaria records, Heliophia},
             pdfborder={0 0 0},
@@
 \textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{3} Institute for Environment and Society, Brown University, Providence, RI 02912, USA\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA 16802, USA}
-\keywords{drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records}
+\keywords{drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records, Heliophia}
 \usepackage{csquotes}
@@
 \abstract{
-Explaining variation in life history strategies is a long-standing goal
-of evolutionary biology. For plants, annual and perennial life histories
-are thought to reflect adaptation to environments that differ in the
-frequency of stress events such as drought. Here we test this hypothesis
-in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering
-plants native to Africa, by integrating 34 years of satellite-based
-drought measurements with 2192 herbaria occurrence records. Consistent
-with predictions from classic life history theory, we find that
-perennial \textit{Heliophila} species occur in environments where
-droughts are significantly less frequent compared to annuals. These
-associations are predictive while controlling for phylogeny, lending
-support to the hypothesis that drought related natural selection has
-influenced the distributions of these strategies. Additionally, the
-collection dates of annual and perennial species indicate that annuals
-escape drought prone seasons during the seed phase of their life cycle.
-Together, these findings provide empirical support for classic
-hypotheses about the drivers of life history strategy in plants - that
-perennials out compete annuals in environments with less frequent
-drought and that annuals are adapted to environments with more frequent
-drought by escaping drought prone seasons as seeds.
-
+\begin{center}\textbf{Summary}\end{center} \par
+
+Explaining variation in life history strategies is an enduring goal of
+evolutionary biology and ecology. For plants, annual and perennial life
+histories have long been thought to reflect adaptation to environments
+that experience alternative drought regimes. Neverthless, empirical
+support for this hypothesis from phylogeographic analyses remains
+lacking.
+
+\par
+
+Here, we tested classic life history theory in \textit{Heliophila}
+(Brassicaceae), a diverse genus of flowering plants native to Africa,
+with a phylogentically constrained analysis integrating 34 years of
+satellite-based drought detection with 2192 herbaria occurrence records.
+
+\par
+
+We found that annual and perennial life histories have evolved multiple
+times in \textit{Heliophila} and that compared to annuals, perennial
+species occur in environments where droughts are significantly less
+frequent. We also report evidence that annual plants adapt to
+predictable drought regimes by escaping drought prone seasons as seeds.
+
+\par
+
+These results provide compelling support for longstanding theoretical
+predictions and a more nuanced understanding of life history evolution.
+More broadly, this work outlines a scalable framework integrating
+herbaria records and remote sensing to address outstanding questions in
+evolutionary ecology. 
 }
@@
 longstanding goal of ecology and evolutionary biology (Cole, 1954). In
-plants, life histories are especially diverse, with herbaceous species
-that complete their life cycle in a number of weeks to trees that live
-for thousands of years (Brown, 1996). Along this continuum in
-angiosperms an important division exists, distinguishing annuals which
-complete their seed to seed life cycle within a single calendar year
-from perennials which can persist over multiple years. Annual plants
-flower once, set seed, senesce, and then die, spending at least some
-portion of the year as a seed, where they are relatively protected from
-environmental stress. In contrast, perennial plants can continue
-vegetative growth after reproduction and must survive conditions
-experienced during all seasons. These represent fundamentally different
-life history strategies, but the ecological factors that explain their
-evolution and distributions remain empirically uresolved (Friedman \&
-Rubin, 2015).\\
+plants, life histories are especially diverse, with some species
+completing their life cycle in a number of weeks to others that live for
+thousands of years (Brown, 1996). Along this continuum in angiosperms an
+important division exists between annuals which complete their seed to
+seed life cycle within a single calendar year from perennials which can
+persist over multiple years. Annual plants flower once, set seed,
+senesce, and then die, spending at least some portion of the year as a
+seed, where they are relatively protected from environmental stress. In
+contrast, perennial plants can continue vegetative growth after
+reproduction and must survive conditions experienced during all seasons
+but can also benefit from multiple bouts of reproduction and competative
+advantages. These represent fundamentally different life history
+strategies and predicting their occurance is important for community,
+ecosystem, and agricultural ecology. However, the environmental factors
+that explain their evolution and distributions remain empirically
+uresolved (Friedman \& Rubin, 2015).
+
 Classical theory predicts shorter life spans in environments where adult
 mortality is high (Charnov \& Schaffer, 1973; Stearns, 1992; Franco \&
-Silvertown, 1996). In plants, this has been extended to the hypothesis
-that annuality is adaptive when it allows plants to escape drought
-(Schaffer \& Gadgil, 1975). Lack of water is perhaps the greatest threat
-to survival during vegetative or reproductive growth and annuals can
-remain dormant (and protected as a seed) during drought. Thus,
-environments with greater seasonal drought frequency may select for
-annual life histories that complete reproduction prior to drought prone
-seasons. Conversely, environments with less frequent drought may select
-for perennial species, which benefit from multiple bouts of reproduction
-and competitive advantage by preventing recruitment of annual species
-(Corbin \& D'Antonio, 2004). These predictions have been supported by
-the observation of annuals in arid environments in \emph{Oryza perennis}
-(Morishima \emph{et al.}, 1984) and \emph{Oenothera} (Evans \emph{et
-al.}, 2005). Additionally, annual and perennial species of
-\emph{Nemesia} were qualitatively associated with winter rather and
-summer rainfall environments respectively (Datson \emph{et al.}, 2008)
-and annual species of \emph{Scorzoneroides} were associated with
-environments classified as unpredictable (Cruz-Mazo \emph{et al.},
-2009). However, whether the history frequency of drought events indeed
-predicts the distributions annual or perennial life history strategies
-has yet to be tested.\\
+Silvertown, 1996). Because lack of water is perhaps the greatest threat
+to survival during vegetative or reproductive growth in plants, this
+theory has been extended to the hypothesis that annuality is adaptive
+when it allows plants to escape drought (Schaffer \& Gadgil, 1975).
+Indeed, adaptation to drought, defined as episodes of increased aridity
+causing plant sresss (Passioura, 1996), is often invoked as an
+explanation for the success of annual species. And while a few cases are
+cited where annuality appears to be more common in environments with
+greater aridity (Stebbins Jr, 1952; Morishima \emph{et al.}, 1984), this
+hypothesis has yet to be supported in a phylogenetic context. In one
+previous study where this question was addressed phylogenetically,
+(Evans \emph{et al.}, 2005) annuals were not found to be associated with
+environments that experience more drought. This could be explained by
+the relatively small number of species studied and the reliance on a
+limited number of weather stations to characterize environments,
+highlighting the need to develop more scalable methods to study trait
+evolution phylogeographically.
+
+It is also critical to consider another dimension of drought adaptation:
+the expectation that annuality is most adaptive when droughts are not
+only frequent but also predictable. That is, when the frequency of
+drought is particularly high during certain seasons. Such predictability
+is important for selection to lead to escape from those seasons which
+are paricuarlary drought prone. While there has been at least one
+example of annuality associated with environments qualitatively
+classified as \enquote{predictable} in a general sense (Datson \emph{et
+al.}, 2008), the seasonal predictability of drought experienced by
+annuals has yet to be rigorously studied. As such, further empirical
+work is needed to support the model of annuality as a mechanism of
+drought adaptation via escape from drought prone seasons.
+
+In addition to drought escape in annuals as a mechanism of adaptation to
+frequent predictable droughts, droughts may be necessary for the success
+of annuals more generally by acting as episodes of disturbance that
+provide opportunites for annuals to establish and compete with sympatric
+perennial species. Indeed, there is considerable evidence that
+perennials dominate in envirnoments where distrurbance events are
+infrequent (Rees \& Long, 1992; Corbin \& D'Antonio, 2004; Clary, 2012).
+The resulting prediction from this hypothesis is that in the absence of
+frequent drought, perenniality should evolve. However, little is known
+about this component of life history evolution because previous work has
+largely focused on the origins of annuality rather than perenniality
+(Friedman \& Rubin, 2015).
+
 Here we combine a long-term global dataset of satellite detected drought
@@
 across landscapes, and annual species should be more commonly associated
-with drought prone regions than perennial species. Furthermore, if
-annual species have adapted to escape drought prone seasons,
-observations of growing annual species (i.e.~occurring in forms other
-than seed) should be rare during drought prone seasons. Phylogenetic
-relatedness can influence tests of associations between species' traits
-and their environments (Felsenstein, 1985; Barrett \emph{et al.}, 1996),
-and therefore we assessed the relationship between life history
-distribution and drought frequency in a phylogenetic context.
+with drought prone regions than perennial species. Additionally, if
+perenniality offers competative advantage in the absence of drought,
+associations between life history and drought frequency should be
+signficant in even if phylogenies are enriched for transitions from
+annual to perenial life history strategy. Finally, if annual species
+have adapted to escape predictably drought prone seasons, observations
+of growing annual species (i.e.~occurring in forms other than seed)
+should be rare during seasons when drought frequency is hightest.
+Phylogenetic relatedness can influence tests of associations between
+species' traits and their environments (Felsenstein, 1985; Barrett
+\emph{et al.}, 1996). Therefore we assessed the relationship between
+life history distribution and drought frequency in a phylogenetic
+context.
 
@@
 
-\hypertarget{availability}{%
-\subsubsection{Availability}\label{availability}}
+\hypertarget{data-availability}{%
+\subsubsection{Data availability}\label{data-availability}}
 
@@
 
-\hypertarget{satellite-detected-drought-data}{%
-\subsubsection{Satellite-detected drought
-data}\label{satellite-detected-drought-data}}
-
-Remotely sensed data is a powerful tool for characterizing seasonal
-patterns in drought because it is less limited in spatial and temporal
-scope and resolution than weather stations or field observations
-(AghaKouchak \emph{et al.}, 2015). To quantify the frequency of drought
-during different seasons across landscapes, we used the remotely sensed
-Vegetative Health Index (VHI), which measures landscape scale reductions
-in plant cover and temperature conditions characteristic of drought
-(Kogan, 2001). Generated from data collected by NOAA AVHRR satellites
-since 1981, the VHI combines Normalized Difference Vegetation Index
-(NDVI) derived measures of vegetative stress (Vegetative Condition Index
-- VCI) with temperature stress indicated by anomalies in thermal spectra
-(Temperature Condition Index - TCI). The VHI of year \(y\) during week
-\(w\) of \([1,52]\) at pixel \(i\) is derived from the following
-equations, where \(n\) is the number of years observed.
-
-\[VCI_{y,w,i} = 100\frac{NDVI_{y,w,i} - NDVI_{min}}{NDVI_{max} - NDVI_{min}}\]
-
-\[TCI_{y,w,i} = 100\frac{T_{y,w,i} - T_{min}}{T_{max} - T_{min}}\]
-
-\[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\] where
-\(NDVI_{min} = min(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})\) and
-\(NDVI_{max} = max(NDVI_{1981,w,i}...NDVI_{1981+n,w,i})\) and
-\(T_{min} = min(T_{1981,w,i}...T_{1981+n,w,i})\) and
-\(T_{max} = max(T_{1981,w,i}...T_{1981+n,w,i})\)
-
-Thus, VHI measurements are standardized according to conditions
-historically observed at each locations. These measurements have been
-validated and generally used for evaluating drought risk and predicting
-crop yields in agriculture (e.g., Rojas \emph{et al.}, 2011; Kogan
-\emph{et al.}, 2016). But they also present a new tool to study seasonal
-patterns in the frequency of drought across environments and to test
-hypotheses about the effect of drought on ecological and evolutionary
-processes (Kerr \& Ostrovsky, 2003). As such, the VHI has been applied
-recently to study drought related ecology of natural species and proven
-useful for predicting intraspecific variation in drought tolerance
-traits and genes (Mojica \emph{et al.}, 2016; Dittberner \emph{et al.},
-2018; Monroe \emph{et al.}, 2018b). Here, we accessed VHI data at
-\(16km^2\) resolution from 1981 to 2015
-(\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php}) to
-characterize the seasonal drought frequencies experienced by annual and
-perennial \emph{Heliophila} species.
-
 \hypertarget{life-history-data-for-heliophila}{%
@@
 \emph{Heliophila} is a genus of flowering plants endemic to the southern
-portion of Africa including the Cape Floristic and Succulent Karoo
-Regions. These are among the most botanically diverse environments on
-Earth and the \emph{Heliophila} species occurring there are considered
-to make up the most diverse genus of the family Brassicaceae (Mummenhoff
-\emph{et al.}, 2005; Mandáková \emph{et al.}, 2012). This genus includes
-both perennial and annual species and this change in life history
+Africa including the Cape Floristic and Succulent Karoo Regions. These
+are among the most botanically diverse environments on Earth and the
+\emph{Heliophila} species occurring there are considered to make up the
+most diverse genus of the family Brassicaceae (Mummenhoff \emph{et al.},
+2005; Mandáková \emph{et al.}, 2012). This genus includes both perennial
+(n =21) and annual (n = 21) species and this change in life history
 strategy has likely arisen multiple independent times (Appel \&
@@
 
-Botanists have collected and maintained over 350 million botanical
-specimens worldwide over the past 300 years (Thiers, 2016). Herbarium
-specimens and their associated metadata have been used since the 1960s
-to study species' geographical distributions (reviewed by Willis
-\emph{et al.} (2017) and Lang \emph{et al.} (2018)). And as they become
-digitized (Soltis, 2017), these collections have been used to study
-relationships between trait distributions, geography, and climate (Davis
-\emph{et al.}, 2015; Stropp \emph{et al.}, 2016; Wolf \emph{et al.},
-2016; Václavı'k \emph{et al.}, 2017). To characterize the distributions
-of annual and perennial \emph{Heliophila} species, all records for the
-genus \emph{Heliophila} were downloaded from the Global Biodiversity
-Information Facility (gbif.org) on July 21, 2018 (GBIF, 2018).
+To characterize the distributions of annual and perennial
+\emph{Heliophila} species, all (8670) records for the genus
+\emph{Heliophila} were downloaded from the Global Biodiversity
+Information Facility (gbif.org) on July 21, 2018 (GBIF, 2018). herbaria
+records records such as these provide a rich data sources to
+characterize the geographical distributions of species (Thiers, 2016;
+Willis \emph{et al.}, 2017; Lang \emph{et al.}, 2018). And as they
+become digitized (Soltis, 2017), these collections have been used to
+study relationships between trait distributions, geography, and climate
+(Davis \emph{et al.}, 2015; Stropp \emph{et al.}, 2016; Wolf \emph{et
+al.}, 2016; Václavı'k \emph{et al.}, 2017).
 
@@
 
-An alignment of ITS I and II sequences for \emph{Heliophila} species was
-obtained from the authors of Mandáková \emph{et al.} (2012). Individual
-ITS I and II sequences for \emph{Aethionema grandiflorum, Alliaria
-petiolata, Cardamine matthioli, Chamira circaeoides}, and \emph{Rorippa
-amphibia} were downloaded from Genbank.
+An alignment of ITS I and II sequences for 21 annual and 21 perennial
+\emph{Heliophila} species was obtained from the authors of Mandáková
+\emph{et al.} (2012). Individual ITS I and II sequences for
+\emph{Aethionema grandiflorum, Alliaria petiolata, Cardamine matthioli,
+Chamira circaeoides}, and \emph{Rorippa amphibia} were downloaded from
+Genbank.
+
+\hypertarget{satellite-detected-drought-data}{%
+\subsubsection{Satellite-detected drought
+data}\label{satellite-detected-drought-data}}
+
+Remotely sensed data is a powerful tool for characterizing seasonal
+patterns in drought because it is less limited in spatial and temporal
+scope and resolution than weather stations or field observations
+(AghaKouchak \emph{et al.}, 2015). From an ecological perspective,
+droughts are best defined as episodes of plant stress caused by elevated
+aridity (Passioura, 1996). Thus remote sensing offers the additional
+benefit for studying drought as an agent of natural selection because
+plant stress caused by drought can be observed from space (Kogan,
+1995a). The remotely sensed Vegetative Health Index (VHI) is one such
+metric, which detects landscape scale reductions in plant cover and
+temperature conditions characteristic of drought (Kogan, 2001).
+Generated from data collected by NOAA AVHRR satellites since 1981, the
+VHI is a composite index combining Normalized Difference Vegetation
+Index (NDVI) derived quantification of vegetative stress (Vegetative
+Condition Index - VCI) with temperature stress indicated by anomalies in
+thermal spectra (Temperature Condition Index - TCI). These indicies were
+developed to create an unbiased quantification of drought across
+ecosystem types. The VHI of year \(y\) during week \(w\) of \([1,52]\)
+at pixel \(i\) is derived from the following equations, where \(n\) is
+the number of years observed.
+
+\[VCI_{y,w,i} = 100\frac{NDVI_{y,w,i} - NDVI_{min,w,i}}{NDVI_{max,w,i} - NDVI_{min,w,i}}\]
+
+Low values of VCI indicate episodes when plant cover is particularly low
+for a given location during a given time of the year. Thus, it controls
+for the location and season in quantifying plant stress.
+
+\[TCI_{y,w,i} = 100\frac{T_{max,w,i}-T_{y,w,i}}{T_{max,w,i} - T_{min,w,i}}\]
+Similarly, low TCI values indicate episdoes of high thermal stress shown
+to be negatively correlated with precipition and soil moisture
+(AghaKouchak \emph{et al.}, 2015).
+
+\[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\] By combining VCI
+and TCI, the VHI is distinguishes drought from other forms of vegetative
+stress (Kogan, 1995b). The use of the VHI to detect drought has been
+validated globally and across ecosystem types (AghaKouchak \emph{et
+al.}, 2015), including in southern Africa, the focal region of this
+study (Figure @ref:fig(mapsdroughtexamples)). To date, the VHI has most
+often been applied for evaluating drought risk for agricultural research
+(e.g., Rojas \emph{et al.}, 2011; Kogan \emph{et al.}, 2016). But it
+also presents a tool to study seasonal patterns in the frequency of
+drought across environments and to test hypotheses about the effect of
+drought on ecological and evolutionary processes (Kerr \& Ostrovsky,
+2003). As such, the VHI has been applied recently to study drought
+related ecology of natural species and proven useful for predicting
+intraspecific variation in drought tolerance traits and genes (Mojica
+\emph{et al.}, 2016; Dittberner \emph{et al.}, 2018; Monroe \emph{et
+al.}, 2018b). Here, we accessed VHI data at \(16km^2\) resolution from
+1981 to 2015
+(\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php}) to
+characterize the seasonal drought frequencies experienced by annual and
+perennial \emph{Heliophila} species across their native range of
+southern Africa.
 
@@
 records using the VHI. Specifically, we created global maps of the
-frequencies of observing drought conditions (VHI\textless{}40, NOAA)
-during the winter (quarter surrounding winter solstice), spring (quarter
-surrounding spring equinox), summer (quarter surrounding summer
-solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015.
-From these maps, the drought frequency during the winter, spring,
-summer, and fall were extracted for the locations of all GBIF records.
+frequencies of observing drought conditions between years
+(VHI\textless{}40, NOAA) during the winter (quarter surrounding winter
+solstice), spring (quarter surrounding spring equinox), summer (quarter
+surrounding summer solstice) and fall (quarter surrounding fall equinox)
+from 1981 to 2015. From these maps, the drought frequency during the
+winter, spring, summer, and fall were extracted for the locations of all
+GBIF records.
 
@@
 
-\hypertarget{phylogeny-construction}{%
-\subsubsection{Phylogeny construction}\label{phylogeny-construction}}
-
-Out group ( \emph{Aethionema grandiflorum, Alliaria petiolata, Cardamine
+\hypertarget{phylogeny-construction-and-ancestral-state-estimation}{%
+\subsubsection{Phylogeny construction and ancestral state
+estimation}\label{phylogeny-construction-and-ancestral-state-estimation}}
+
+Outgroup ( \emph{Aethionema grandiflorum, Alliaria petiolata, Cardamine
 matthioli, Chamira circaeoides}, and \emph{Rorippa amphibia}) and
-ingroup \emph{Heliophila} ITS I and II sequences were aaligned using
+ingroup \emph{Heliophila} ITS I and II sequences were aligned using
 MAFFT (Katoh \emph{et al.}, 2002) with strategy G-INS-I, offset value
@@
 analyses were conducted through the CIPRES Science Gateway (Miller
-\emph{et al.}, 2010).
+\emph{et al.}, 2010). Ancestral state estimation was performed in R
+using the package phytools (Revell, 2012) to genearte 10,000 stochastic
+character maps simulated under an equal rates model of character
+evolution for the trait life habit (annual or perennial).
 
@@
 associated with historic drought frequency. First, we compared the
-frequency of drought during the winter, spring, summer, and fall between
-total occurrence records of annual and perennial species by t-tests. To
-account for variation in the number of occurrence records per species,
-we next calculated the mean drought frequency during the winter, spring,
-summer and fall for each species. Because shared evolutionary history of
-closely related species can lead to spurious associations between traits
-and environments (Felsenstein, 1985), we tested for a relationship
-between life history strategy and drought frequency while controlling
-for phylogeny using phylogenetic logistic regression (Ives \& Garland,
-2010).
+frequency of drought across all occurrence records of between annual and
+perennial species by t-tests. Independent tests were performed to
+compare drought frequency between annuals and perennials during each
+season. We also tested for a relationship between drought frequency and
+life history, season, and their interaction by an analysis of variance
+while including species as a random effect. To account for variation in
+the number of occurrence records per species, we next calculated the
+mean drought frequency during the winter, spring, summer and fall for
+each species. Because shared evolutionary history of closely related
+species can lead to spurious associations between traits and
+environments (Felsenstein, 1985), we tested for a relationship between
+life history strategy and drought frequency while controlling for
+phylogeny using phylogenetic logistic regression (Ives \& Garland,
+2010). This statistical approach is designed to control for the
+confounding effects of geographic relationships explained by demographic
+history rather than processes of natural selection.
 
@@
 studied. Orange circles at branch tips mark annual species and blue
-circles mark perennial species. Example herbaria specimens accessed via
-GBIF of (a) \emph{H. minima}, (b) \emph{H. deserticola}, (c) \emph{H.
-coronopifolia} and (d) \emph{H. ephemera}. Images (a,c,d) courtesy of
-The Bavarian Natural History Collections (CC BY-SA 4.0) and (b) The
-London Natural History Museum (CC BY 4.0). Links to images are found in
-the supplement.}\label{fig:phylogeny}
+circles mark perennial species. At internal nodes, pie charts indicate
+the estimated posterior probability of being annual versus perennial.
+Example herbaria specimens accessed via GBIF of (a) \emph{H. minima},
+(b) \emph{H. deserticola}, (c) \emph{H. coronopifolia} and (d) \emph{H.
+ephemera}. Images (a,c,d) courtesy of The Bavarian Natural History
+Collections (CC BY-SA 4.0) and (b) The London Natural History Museum (CC
+BY 4.0). Links to images are found in the supplement.}\label{fig:phylogeny}
 \end{figure}
@@
 perennial \emph{Heliophila.} Drought frequency during the (c) winter,
-(d) spring, (e) summer and (f) fall measured using the VHI. (g) Drought
+(d) spring, (e) summer and (f) fall detected using the VHI. (g) Drought
 frequencies during each season at the observation locations of annual
-and perennial \emph{Heliophila} (t tests, ** = p \textless{} 0.01).}\label{fig:mapsboxplots}
+and perennial \emph{Heliophila} (t tests contrasting annuals and
+perennials, ** = p \textless{} 0.01).}\label{fig:mapsboxplots}
 \end{figure}
+
+The \emph{Heliophila} phylogeny was consistent with previous studies
+(Mummenhoff \emph{et al.}, 2005; Mandáková \emph{et al.}, 2012). Based
+on 10,000 stochastic character maps simulated under an equal rates model
+of character evolution in life history, an average of approximately
+eight changes from annual to perennial and five changes from perennial
+to annual are observed per stochastic character map (Figure
+\ref{fig:phylogeny}a). These results suggest that the ancestral state of
+\emph{Heliophila} was annual and that both charcter states have arisen
+independently mutliple times.
 
@@
 Karoo (Figure \ref{fig:mapsboxplots}a), the occurrence of perennials
-extended to the east coast of South Africa (Figure
+extended to the southern and eastern coast of South Africa (Figure
 \ref{fig:mapsboxplots}b).
@@
 
-Theory predicts that annuality should be adaptive in places where
-stresses such as drought are more common. Conversely, perenniality
-should be adaptive in places where such stresses are less frequent. We
-found that the frequency of drought was significantly higher at the
+We found that the frequency of drought was significantly higher at the
 locations of occurrence records for annual species. When comparing
@@
 significantly higher at the location of annuals during the winter (t =
-10.65, p = 0.00), spring (t = 10.73, p = 0.00), summer (t = 12.67, p =
-0.00), and fall (t = 15.26, p = 0.00). Because a comparison across all
-occurrence records does not account for variation in the number of
-records per species (Table \ref{tab:speciesmeanstable}) or species
-relatedness (Figure \ref{fig:phylogeny}a), we also tested whether mean
-drought frequency values of each species were significantly different
-between annuals and perennials using phylogenetic logistic regression.
-We found that the mean drought frequencies were significantly higher
+10.65, p \textless{} 0.01), spring (t = 10.73, p \textless{} 0.01),
+summer (t = 12.67, p = \textless{} 0.01), and fall (t = 15.26, p
+\textless{} 0.01). A mixed-model analysis of variace which included
+species as random effect also identified a signficant relationship
+between drought frequency and life history (p \textless{} 0.01), season
+(p \textless{} 0.01) and their interaction (p \textless{} 0.01) (Table
+\ref{tab:anovatable}). Because a comparison across all occurrence
+records does not account for variation in the number of records per
+species (Table \ref{tab:speciesmeanstable}) or species relatedness
+(Figure \ref{fig:phylogeny}a), we also tested whether mean drought
+frequency values of each species were significantly different between
+annuals and perennials using phylogenetic logistic regression. We found
+that the mean drought frequencies were significantly higher
 (\(\alpha = 0.05\)) in annual species during the spring, summer, and
-fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a). These
-findings indicate that common acestry alone does not explain differences
-the drought frequencies experienced between the environments of annual
-and perennial \emph{Heliophila}.
+fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a,b). These
+findings indicate that demographic history alone does not explain
+differences in the drought frequencies experienced between the
+environments of annual and perennial \emph{Heliophila}.
 
@@
 \includegraphics[width=\textwidth]{../figures/line_and_dates} \caption{(a) Comparison (mean +- SE) of drought frequency across
-seasons measured at the GBIF records of annual and perennial species of
-\emph{Heliophila}. (phylogenetic logistic regression, * = p \textless{}
-0.05, ** = p \textless{} 0.01) (b) Collection dates of GBIF records of
-annual and perennial species of \emph{Heliophila}.}\label{fig:lineplots}
+seasons calcualted at the occurance locations of GBIF records of annual
+and perennial species of \emph{Heliophila}. (b) Results from
+phylogenetic logistic regression, where the line shows the model fit and
+* = p \textless{} 0.05, ** = p \textless{} 0.01. Annuals were scored as
+0 and perennials as 1. (c) Collection dates of GBIF records of annual
+and perennial species of \emph{Heliophila} in relation to the
+photoperiodic calender where day 1 is intermediate to the fall equinox
+and winter soltice.}\label{fig:lineplots}
 \end{figure}
@@
 annuality reflects adaptation to such environments because it allows
-species to escape stressful conditions. If this is the case, we would
-expect that annuals spend the drought prone seasons of summer and fall
-as seeds. To test this hypothesis, we compared the dates of occurrence
-records between annual and perennial \emph{Heliophila} species. The
-distributions reveal a considerable difference in the timing of
-observation of these two life histories. In comparison to perennials,
-which appear to be collected throughout the year, annuals are almost
-exclusively observed during the winter and spring (Figure
+species to escape predictable stressful conditions. If this is the case,
+we would expect that annuals spend the drought prone seasons of summer
+and fall as seeds. To test this hypothesis, we compared the dates of
+occurrence records between annual and perennial \emph{Heliophila}
+species. The distributions reveal a considerable difference in the
+timing of observation of these two life histories. In comparison to
+perennials, which appear to be collected throughout the year, annuals
+are almost exclusively observed during the winter and spring (Figure
 \ref{fig:lineplots}b). The differences between the distribution of
-collection dates were significant by all tests (ks.test D = 0.25, p = 0;
-bartlett.test K2 = 503.18, p = 0.00) This is consistent with a model of
-life history in which annual species flower in the spring, set seed,
-senesce, and die before the summer. Thus, these annual species are
-likely to remain dormant during the summer and fall, when drought is the
-strongest predictor of the distributions of annual and perennial life
-histories (Figure \ref{fig:lineplots}a).
+collection dates were significant by all tests (ks.test D = 0.25, p
+\textless{} 0.01; bartlett.test K2 = 503.18, p \textless{} 0.01) This is
+consistent with a model of life history in which annual species flower
+in the spring, set seed, senesce, and die before the summer. Thus, these
+annual species are likely to remain dormant during the summer and fall,
+when drought is the strongest predictor of the distributions of annual
+and perennial life histories (Figure \ref{fig:lineplots}a).
 
@@
 these results provide quantitative support for the classic prediction
-that annual species are found in environemnts that experience more
-frequent drought than perennial species. These findings complement
-previous reports of qualitative associations between annuality with
-enviroments characterized as having increased aridity (Evans \emph{et
-al.}, 2005), alternative precipitation defined habitats (Morishima
-\emph{et al.}, 1984; Datson \emph{et al.}, 2008), or greater
-unpredictability (Cruz-Mazo \emph{et al.}, 2009). However, to our
+that annual species are found in environments that experience more
+frequent drought than perennial species. These findings are a valuable
+contribution to previous investigations of associations between life
+history and climate (Morishima \emph{et al.}, 1984; Evans \emph{et al.},
+2005; Datson \emph{et al.}, 2008; Cruz-Mazo \emph{et al.}, 2009). To our
 knowledge this is the first study to demonstrate a significant
@@
 informed by large scale species distribution data and long term drought
-measures.
+detection.
 
@@
 
-These findings support classical theoretical predictions about the
-adaptive value of annual and perennial life history strategies. Taken
-together, they suggest that in \emph{Heliophila}, annual species are
-adapted to environments with increased summer droughts by avoiding these
-seasons in a dormant seed phase of their life cycle. They also suggest
-that perenniality is adaptive in environemnts where droughts are less
-frequent. While most previous work has focused on describing the
-evolutionary origins of annuality (Barrett \emph{et al.}, 1996; Conti
-\emph{et al.}, 1999; Andreasen \& Baldwin, 2001; Verboom \emph{et al.},
-2004; Friedman \& Rubin, 2015) there are at least a few other cases
-where perenniality appears to have arisen from an annual ancestor (Bena
-\emph{et al.}, 1998; Tank \& Olmstead, 2008). And while early theory
-predicted selection for annuality when adult morality is high (Stearns,
-1992), we also find evidence that perenniality could be explained by
-reduced frequency of drought. The phylogeny reveals several transitions
-from annual to perennial life history (Figure \ref{fig:phylogeny}a) and
-the distributions of perennial \emph{Heliophila} extend into regions
-where drought frequency is low (Figure \ref{fig:mapsboxplots}b, Figure
-\ref{fig:speciesmaps}). Perennials may be able to out complete annual
-relatives in environments where the infrequency of drought favors
-strategies that allow plants to benefit from growth over many seasons.
-This also suggests that annuals rely on drought as a source of
-disturbance for seedling recruitment when competing with perennials
-(Corbin \& D'Antonio, 2004). Indeed, no annual species were observed in
-the low drought regions of eastern South Africa (Figure
-\ref{fig:mapsboxplots}, Figure \ref{fig:speciesmaps}).
+These findings empirical provide support classical theoretical
+predictions about the adaptive value of annual and perennial life
+history strategies. Taken together, they suggest that in
+\emph{Heliophila}, annual species are adapted to environments with
+increased summer droughts by avoiding these seasons in a dormant seed
+phase of their life cycle. They also suggest that perenniality is
+adaptive in environemnts where droughts are less frequent. While most
+previous work has focused on describing the evolutionary origins of
+annuality (Barrett \emph{et al.}, 1996; Conti \emph{et al.}, 1999;
+Andreasen \& Baldwin, 2001; Verboom \emph{et al.}, 2004; Friedman \&
+Rubin, 2015) there are at least a few other cases where perenniality
+appears to have arisen from an annual ancestor (Bena \emph{et al.},
+1998; Tank \& Olmstead, 2008). And while early theory predicted
+selection for annuality when adult mortality is high (Stearns, 1992), we
+also find evidence that perenniality could be explained by reduced
+frequency of drought. This is supported by the theoretical prediction
+that perenniality is advantageous is stable habitats. The phylogeny
+reveals several transitions from annual to perennial life history
+(Figure \ref{fig:phylogeny}a) and the distributions of perennial
+\emph{Heliophila} extend into regions where drought frequency is low
+(Figure \ref{fig:mapsboxplots}b, Figure \ref{fig:speciesmaps}).
+Perennials may be able to out compete annual relatives in environments
+where the infrequency of drought favors strategies that allow plants to
+benefit from growth over many seasons. This also suggests that annuals
+rely on drought as a source of disturbance for seedling recruitment when
+competing with perennials (Corbin \& D'Antonio, 2004). Indeed, no annual
+species were observed in the low drought regions of eastern South Africa
+(Figure \ref{fig:mapsboxplots}, Figure \ref{fig:speciesmaps}).
 
@@
 strategies could be threatened by rapidly changing drought regimens
-(Dai, 2011). This could have impacts on ecosystem functioning and
-processes such as carbon cycling if life history traits evolve or the
-composition of annual and perennial species changes in response (Garnier
-\emph{et al.}, 1997; Roumet \emph{et al.}, 2006; Monroe \emph{et al.},
-2018a). Furthermore, the frequency of drought may be an important factor
-when considering the use of perennial cropping systems (Parry \emph{et
-al.}, 2005; Lelièvre \& Volaire, 2009).
+(Dai, 2011). In light of the findings here, changes in eastern South
+Africa which is forecast to experience signicantly lower rainfall
+(Service \& Comission, 2017) could be particularly impactful. In this
+study we found that this region is currently dominiated by derived
+perennial species of \emph{Heliophila}. However, if drought become more
+frequent in this region, it may allow for the establishement of annuals.
+Such changes in selection patterns and shifts in plant functional
+diversity could have impacts on ecosystem functioning and processes such
+as carbon cycling (Garnier \emph{et al.}, 1997; Roumet \emph{et al.},
+2006; Monroe \emph{et al.}, 2018a). Furthermore, the changes in
+frequency of drought may be an important factor when considering the use
+of perennial cropping systems (Parry \emph{et al.}, 2005; Lelièvre \&
+Volaire, 2009).
 
@@
 
-We thank members of the Sloan lab for generous feedback that improved
-the quality of this work. This research was supported by NSF Award
-1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
+We thank Jesse Lasky, members of the Sloan lab at CSU and three
+ananymous reviewers for generous feedback that improved the quality of
+this work. This research was supported by NSF Award 1701918 and
+USDA-NIFA Award 2014-38420-21801 to JGM.
 
@@
 \section{References}\label{references}}
-
-\newpage
 
@@
 
-\leavevmode\hypertarget{ref-R-geiger_a}{}%
-\textbf{\textnormal{Alfaro M}, \textnormal{Santini F}, \textnormal{Brock
-C}, \textnormal{Alamillo H}, \textnormal{Dornburg A},
-\textnormal{Rabosky D}, \textnormal{Carnevale G}, \textnormal{Harmon
-L}}. \textbf{2009}. Nine exceptional radiations plus high turnover
-explain species diversity in jawed vertebrates. \emph{Proceedings of the
-National Academy of Sciences of the United States of America}
-\textbf{106}: 13410--13414.
-
 \leavevmode\hypertarget{ref-andreasen2001unequal}{}%
@@
 
-\leavevmode\hypertarget{ref-R-papaja}{}%
-\textbf{\textnormal{Aust F}, \textnormal{Barth M}}. \textbf{2018}.
-\emph{papaja: Create APA manuscripts with R Markdown}.
-
 \leavevmode\hypertarget{ref-barrett1996comparative}{}%
@@
 Lond. B} \textbf{351}: 1271--1280.
-
-\leavevmode\hypertarget{ref-R-Matrix}{}%
-\textbf{\textnormal{Bates D}, \textnormal{Maechler M}}. \textbf{2018}.
-\emph{Matrix: Sparse and dense matrix classes and methods}.
 
@@
 
+\leavevmode\hypertarget{ref-clary2012determinants}{}%
+\textbf{\textnormal{Clary J}}. \textbf{2012}. Determinants of perennial
+and annual grass distribution in mediterranean-climate california.
+\emph{Plant Ecology} \textbf{213}: 1203--1208.
+
 \leavevmode\hypertarget{ref-cole1954population}{}%
@@
 
-\leavevmode\hypertarget{ref-R-geiger_b}{}%
-\textbf{\textnormal{Eastman J}, \textnormal{Alfaro M}, \textnormal{Joyce
-P}, \textnormal{Hipp A}, \textnormal{Harmon L}}. \textbf{2011}. A novel
-comparative method for identifying shifts in the rate of character
-evolution on trees. \emph{Evolution} \textbf{65}: 3578--3589.
-
 \leavevmode\hypertarget{ref-evans2005climate}{}%
@@
 
-\leavevmode\hypertarget{ref-R-mvtnorm}{}%
-\textbf{\textnormal{Genz A}, \textnormal{Bretz F}}. \textbf{2009}.
-\emph{Computation of multivariate normal and t probabilities}.
-Heidelberg: Springer-Verlag.
-
 \leavevmode\hypertarget{ref-guindon2003a}{}%
@@
 maximum-likelihood. \emph{Systematic Biology} \textbf{52}: 696--704.
-
-\leavevmode\hypertarget{ref-R-MCMCglmm}{}%
-\textbf{\textnormal{Hadfield JD}}. \textbf{2010}. MCMC methods for
-multi-response generalized linear mixed models: The MCMCglmm R package.
-\emph{Journal of Statistical Software} \textbf{33}: 1--22.
-
-\leavevmode\hypertarget{ref-R-geiger_d}{}%
-\textbf{\textnormal{Harmon L}, \textnormal{Weir J}, \textnormal{Brock
-C}, \textnormal{Glor R}, \textnormal{Challenger W}}. \textbf{2008}.
-GEIGER: Investigating evolutionary radiations. \emph{Bioinformatics}
-\textbf{24}: 129--131.
-
-\leavevmode\hypertarget{ref-R-logistf}{}%
-\textbf{\textnormal{Heinze G}, \textnormal{Ploner M}}. \textbf{2018}.
-\emph{Logistf: Firth's bias-reduced logistic regression}.
-
-\leavevmode\hypertarget{ref-R-purrr}{}%
-\textbf{\textnormal{Henry L}, \textnormal{Wickham H}}. \textbf{2018}.
-\emph{Purrr: Functional programming tools}.
-
-\leavevmode\hypertarget{ref-R-raster}{}%
-\textbf{\textnormal{Hijmans RJ}}. \textbf{2018}. \emph{Raster:
-Geographic data analysis and modeling}.
-
-\leavevmode\hypertarget{ref-R-phylolm}{}%
-\textbf{\textnormal{Ho LST}, \textnormal{Ane C}}. \textbf{2014}. A
-linear-time algorithm for gaussian and non-gaussian trait evolution
-models. \emph{Systematic Biology} \textbf{63}: 397--408.
 
@@
 \emph{Trends in ecology \& evolution} \textbf{18}: 299--305.
+
+\leavevmode\hypertarget{ref-kogan1995droughts}{}%
+\textbf{\textnormal{Kogan FN}}. \textbf{1995a}. Droughts of the late
+1980s in the united states as derived from noaa polar-orbiting satellite
+data. \emph{Bulletin of the American Meteorological Society}
+\textbf{76}: 655--668.
+
+\leavevmode\hypertarget{ref-kogan1995application}{}%
+\textbf{\textnormal{Kogan FN}}. \textbf{1995b}. Application of
+vegetation index and brightness temperature for drought detection.
+\emph{Advances in space research} \textbf{15}: 91--100.
 
@@
 
+\leavevmode\hypertarget{ref-monyela2017two}{}%
+\textbf{\textnormal{Monyela BM}}. \textbf{2017}. A two-year long drought
+in summer 2014/2015 and 2015/2016 over south africa.
+
 \leavevmode\hypertarget{ref-morishima1984differentiation}{}%
@@
 
-\leavevmode\hypertarget{ref-R-bindrcpp}{}%
-\textbf{\textnormal{Müller K}}. \textbf{2018}. \emph{Bindrcpp: An 'rcpp'
-interface to active bindings}.
-
-\leavevmode\hypertarget{ref-R-tibble}{}%
-\textbf{\textnormal{Müller K}, \textnormal{Wickham H}}. \textbf{2018}.
-\emph{Tibble: Simple data frames}.
-
-\leavevmode\hypertarget{ref-R-caper}{}%
-\textbf{\textnormal{Orme D}, \textnormal{Freckleton R},
-\textnormal{Thomas G}, \textnormal{Petzoldt T}, \textnormal{Fritz S},
-\textnormal{Isaac N}, \textnormal{Pearse W}}. \textbf{2018}.
-\emph{Caper: Comparative analyses of phylogenetics and evolution in r}.
-
-\leavevmode\hypertarget{ref-R-ape}{}%
-\textbf{\textnormal{Paradis E}, \textnormal{Schliep K}}. \textbf{2018}.
-Ape 5.0: An environment for modern phylogenetics and evolutionary
-analyses in R. \emph{Bioinformatics} \textbf{xx}: xxx--xxx.
-
 \leavevmode\hypertarget{ref-parry2005prospects}{}%
@@
 
-\leavevmode\hypertarget{ref-R-sp}{}%
-\textbf{\textnormal{Pebesma EJ}, \textnormal{Bivand RS}}. \textbf{2005}.
-Classes and methods for spatial data in R. \emph{R News} \textbf{5}:
-9--13.
-
-\leavevmode\hypertarget{ref-R-coda}{}%
-\textbf{\textnormal{Plummer M}, \textnormal{Best N}, \textnormal{Cowles
-K}, \textnormal{Vines K}}. \textbf{2006}. CODA: Convergence diagnosis
-and output analysis for mcmc. \emph{R News} \textbf{6}: 7--11.
-
-\leavevmode\hypertarget{ref-R-gee}{}%
-\textbf{\textnormal{R by Thomas Lumley VJCP to}, \textnormal{author. BRN
-that maintainers are not available to give advice on using a package
-they did not}}. \textbf{2015}. \emph{Gee: Generalized estimation
-equation solver}.
-
-\leavevmode\hypertarget{ref-R-base}{}%
-\textbf{\textnormal{R Core Team}}. \textbf{2018}. \emph{R: A language
-and environment for statistical computing}. Vienna, Austria: R
-Foundation for Statistical Computing.
-
-\leavevmode\hypertarget{ref-R-phytools}{}%
+\leavevmode\hypertarget{ref-passioura1996drought}{}%
+\textbf{\textnormal{Passioura J}}. \textbf{1996}. Drought and drought
+tolerance. \emph{Plant growth regulation} \textbf{20}: 79--83.
+
+\leavevmode\hypertarget{ref-rees1992germination}{}%
+\textbf{\textnormal{Rees M}, \textnormal{Long MJ}}. \textbf{1992}.
+Germination biology and the ecology of annual plants. \emph{The American
+Naturalist} \textbf{139}: 484--508.
+
+\leavevmode\hypertarget{ref-revell2012phytools}{}%
 \textbf{\textnormal{Revell LJ}}. \textbf{2012}. Phytools: An r package
@@
 in Ecology and Evolution} \textbf{3}: 217--223.
-
-\leavevmode\hypertarget{ref-R-maps}{}%
-\textbf{\textnormal{Richard A. Becker OS code by}, \textnormal{Ray
-Brownrigg. Enhancements by Thomas P Minka ARWR version by},
-\textnormal{Deckmyn. A}}. \textbf{2018}. \emph{Maps: Draw geographical
-maps}.
 
@@
 
-\leavevmode\hypertarget{ref-R-geiger_c}{}%
-\textbf{\textnormal{Slater G}, \textnormal{Harmon L},
-\textnormal{Wegmann D}, \textnormal{Joyce P}, \textnormal{Revell L},
-\textnormal{Alfaro M}}. \textbf{2012}. Fitting models of continuous
-trait evolution to incompletely sampled comparative data using
-approximate bayesian computation. \emph{Evolution} \textbf{66}:
-752--762.
-
-\leavevmode\hypertarget{ref-R-diagram}{}%
-\textbf{\textnormal{Soetaert K}}. \textbf{2017}. \emph{Diagram:
-Functions for visualising simple graphs (networks), plotting flow
-diagrams}.
-
-\leavevmode\hypertarget{ref-R-shape}{}%
-\textbf{\textnormal{Soetaert K}}. \textbf{2018}. \emph{Shape: Functions
-for plotting graphical shapes, colors}.
+\leavevmode\hypertarget{ref-south2017a}{}%
+\textbf{\textnormal{Service SAW}, \textnormal{Comission WR}}.
+\textbf{2017}. \emph{A climate change reference atlas}.
 
@@
 life histories}.
+
+\leavevmode\hypertarget{ref-stebbins1952aridity}{}%
+\textbf{\textnormal{Stebbins Jr GL}}. \textbf{1952}. Aridity as a
+stimulus to plant evolution. \emph{The American Naturalist} \textbf{86}:
+33--44.
 
@@
 
-\leavevmode\hypertarget{ref-R-MASS}{}%
-\textbf{\textnormal{Venables WN}, \textnormal{Ripley BD}}.
-\textbf{2002}. \emph{Modern applied statistics with s}. New York:
-Springer.
-
 \leavevmode\hypertarget{ref-verboom2004testing}{}%
@@
 Botany} \textbf{91}: 1364--1370.
-
-\leavevmode\hypertarget{ref-R-ggplot2}{}%
-\textbf{\textnormal{Wickham H}}. \textbf{2016}. \emph{Ggplot2: Elegant
-graphics for data analysis}. Springer-Verlag New York.
-
-\leavevmode\hypertarget{ref-R-tidyverse}{}%
-\textbf{\textnormal{Wickham H}}. \textbf{2017}. \emph{Tidyverse: Easily
-install and load the 'tidyverse'}.
-
-\leavevmode\hypertarget{ref-R-forcats}{}%
-\textbf{\textnormal{Wickham H}}. \textbf{2018a}. \emph{Forcats: Tools
-for working with categorical variables (factors)}.
-
-\leavevmode\hypertarget{ref-R-stringr}{}%
-\textbf{\textnormal{Wickham H}}. \textbf{2018b}. \emph{Stringr: Simple,
-consistent wrappers for common string operations}.
-
-\leavevmode\hypertarget{ref-R-dplyr}{}%
-\textbf{\textnormal{Wickham H}, \textnormal{François R},
-\textnormal{Henry L}, \textnormal{Müller K}}. \textbf{2018}.
-\emph{Dplyr: A grammar of data manipulation}.
-
-\leavevmode\hypertarget{ref-R-tidyr}{}%
-\textbf{\textnormal{Wickham H}, \textnormal{Henry L}}. \textbf{2018}.
-\emph{Tidyr: Easily tidy data with 'spread()' and 'gather()' functions}.
-
-\leavevmode\hypertarget{ref-R-readr}{}%
-\textbf{\textnormal{Wickham H}, \textnormal{Hester J},
-\textnormal{Francois R}}. \textbf{2017}. \emph{Readr: Read rectangular
-text data}.
 
@@
 
-\url{https://www.gbif.org/occurrence/1099023487}
-\url{https://www.gbif.org/occurrence/1057389408}
-\url{https://www.gbif.org/occurrence/1099023562}
+\url{https://www.gbif.org/occurrence/1099023487}\\
+\url{https://www.gbif.org/occurrence/1057389408}~\\
+\url{https://www.gbif.org/occurrence/1099023562}~\\
 \url{https://www.gbif.org/occurrence/1099023490}
-
-\hypertarget{software-used}{%
-\subsubsection{Software used}\label{software-used}}
-
-We used R (Version 3.5.1; R Core Team, 2018) and the R-packages
-\emph{ape} (Version 5.2; Paradis \& Schliep, 2018; Orme \emph{et al.},
-2018; Soetaert, 2018), \emph{bindrcpp} (Version 0.2.2; Müller, 2018),
-\emph{caper} (Version 1.0.1; Orme \emph{et al.}, 2018), \emph{coda}
-(Version 0.19.2; Plummer \emph{et al.}, 2006), \emph{diagram} (Version
-1.6.4; Soetaert, 2017), \emph{dplyr} (Version 0.7.8; Wickham \emph{et
-al.}, 2018), \emph{forcats} (Version 0.3.0; Wickham, 2018a), \emph{gee}
-(Version 4.13.19; R by Thomas Lumley \& author., 2015), \emph{geiger}
-(Version 2.0.6; Alfaro \emph{et al.}, 2009; Harmon \emph{et al.}, 2008;
-Eastman \emph{et al.}, 2011; Slater \emph{et al.}, 2012), \emph{ggplot2}
-(Version 3.1.0; Wickham, 2016), \emph{logistf} (Version 1.23; Heinze \&
-Ploner, 2018), \emph{maps} (Version 3.3.0; Richard A. Becker \emph{et
-al.}, 2018), \emph{MASS} (Version 7.3.51.1; Venables \& Ripley, 2002),
-\emph{Matrix} (Version 1.2.15; Bates \& Maechler, 2018), \emph{MCMCglmm}
-(Version 2.26; Hadfield, 2010), \emph{mvtnorm} (Version 1.0.8; Genz \&
-Bretz, 2009), \emph{papaja} (Version 0.1.0.9842; Aust \& Barth, 2018),
-\emph{phylolm} (Version 2.6; Ho \& Ane, 2014), \emph{phytools} (Version
-0.6.60; Revell, 2012), \emph{purrr} (Version 0.2.5; Henry \& Wickham,
-2018), \emph{raster} (Version 2.8.4; Hijmans, 2018), \emph{readr}
-(Version 1.2.1; Wickham \emph{et al.}, 2017), \emph{shape} (Version
-1.4.4; Soetaert, 2018), \emph{sp} (Version 1.3.1; Pebesma \& Bivand,
-2005), \emph{stringr} (Version 1.3.1; Wickham, 2018b), \emph{tibble}
-(Version 1.4.2; Müller \& Wickham, 2018), \emph{tidyr} (Version 0.8.2;
-Wickham \& Henry, 2018), and \emph{tidyverse} (Version 1.2.1; Wickham,
-2017) for all our analyses. \newpage
 
@@
 figures}\label{supplementary-tables-and-figures}}
+
+\newpage
+
+\begin{table}[tbp]
+\begin{center}
+\begin{threeparttable}
+\caption{\label{tab:anovatable}Analysis of variance (ANOVA) to compare drought frequency as a function of life history, season, and their interaction while including species as a random effect.}
+\begin{tabular}{lllll}
+\toprule
+predictor & \multicolumn{1}{c}{numDF} & \multicolumn{1}{c}{denDF} & \multicolumn{1}{c}{F.value} & \multicolumn{1}{c}{p.value}\\
+\midrule
+intercept & 1 & 8,720.0000 & 2,833.9567 & 0.0000\\
+life history & 1 & 40.0000 & 12.2117 & 0.0012\\
+season & 3 & 8,720.0000 & 106.4947 & 0.0000\\
+life history x season & 3 & 8,720.0000 & 4.8301 & 0.0023\\
+\bottomrule
+\end{tabular}
+\end{threeparttable}
+\end{center}
+\end{table}
 
@@
 \centering
+\includegraphics{../figures/maps_drought_examples.pdf}
+\caption{\label{fig:mapsdroughtexamples}Example years experiencing contrasting degrees
+of drought in southern Africa. Vegetative Health Index (VHI) values
+below 40 indicate remotely sensed drought. Drought detection during this
+years is validated by precipitation based estimates of drought ocurrance
+(Standardized Precipitation Index) which confirm that while (a) 2015 was
+one of the worst drought years on record, (b) 1999 was one of the
+wettest, and (c) 2012 was typical in terms of precipitation patterns
+(Monyela, 2017). It is worth noting that drought can be detected using
+the VHI across ecosystems, including those inhabited by perennial rather
+than annual \emph{Heliophila} species.}
+\end{figure}
+
+\begin{figure}
+\centering
 \includegraphics{../figures/speciesmaps.pdf}

### 2019-03-08 manuscript/manuscript.R

@@ -1,0 +1,488 @@
+## ----setup, include = FALSE----------------------------------------------
+library("papaja")
+library(raster)
+library(tidyverse)
+library(ggplot2)
+library(logistf)
+library(geiger) 
+library(ape) 
+library(phylolm)
+library(gridExtra)
+library(cowplot)
+library(rasterVis)
+library(ggtree)
+library(grid)
+library(png)
+library(ggimage)
+
+## ----raw_GBIF------------------------------------------------------------
+GBIF<-read.csv("../data/0006178-160311141623029/occurrence.csv")
+
+## ----Droughtdata---------------------------------------------------------
+Winter<-raster("../data/drought/Summer_drought_freq")
+Spring<-raster("../data/drought/Fall_drought_freq")
+Summer<-raster("../data/drought/Winter_drought_freq")
+Fall<-raster("../data/drought/Spring_drought_freq")
+
+# Winter<-raster("~/VHI/droughtfreq/summer_mild")
+# Spring<-raster("~/VHI/droughtfreq/fall_mild")
+# Summer<-raster("~/VHI/droughtfreq/winter_mild")
+# Fall<-raster("~/VHI/droughtfreq/spring_mild")
+
+## ----cleanup-------------------------------------------------------------
+species_life_history<-read.csv("../data/Heliophila_life_history.csv")
+species_life_history$species<-gsub("_", " ", species_life_history$species)
+species_life_history$species <- as.factor(species_life_history$species)
+
+GBIF <- GBIF %>%
+  mutate(Winter=raster::extract(Winter, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(Spring=raster::extract(Spring, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(Summer=raster::extract(Summer, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(Fall=raster::extract(Fall, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(ymd_lat_lon_sp = paste0(year,"_",month,"_",day,"_", decimalLatitude,"_", decimalLongitude,"_", species))
+  
+GBIF_cleaned <- GBIF %>%
+  filter(species %in% species_life_history$species) %>% 
+  filter(!is.na(decimalLatitude)) %>%
[…]

### 2019-03-08 src/GBIF_records_drought_frequency.R

@@ -1,395 +1,0 @@
-#Heliophila data cleaning
-#9/6/2018
-
-#accompanies/partially replicates JGM script "GBIF_records_drought_frequency.R" rows 1-19
-#things to clean/check:
-#basis of specimen
-#names through TNRS
-#duplicates?
-
-library(raster)
-library(tidyverse)
-library(ggplot2)
-#library(plotly)
-#library(rgbif) #0.9.8
-#library(taxize)
-# library(dplyr)
-
-# Read occurance data 
-GBIF<-read.csv("data/0006178-160311141623029/occurrence.csv");cat("\nrows in raw GBIF data file:", nrow(GBIF))
-# GBIF<-read.csv("Heliophila_occurrence.csv");cat("\nrows in raw GBIF data file:", nrow(GBIF), file = "analysis log")
-# summary(GBIF$species)
-#261 rows have blanks for species, NA for speciesKey
-# colnames(GBIF)
-
-# select species with life history data 
-species_life_history<-read.csv("data/Heliophila_life_history.csv")
-species_life_history$species<-gsub("_", " ", species_life_history$species)
-# unique(species_life_history$species)#42
-species_life_history$species <- as.factor(species_life_history$species)
-# levels(species_life_history$species) #42
-
-# class(GBIF$species)
-# [1] "factor"
-# class(species_life_history$species)
-# [1] "factor"
-GBIF_JGMcleaned<-subset(GBIF, species %in% species_life_history$species);cat("\nrows in raw GBIF data after choosing species with life history JGM:", nrow(GBIF_JGMcleaned), file = "analysis log", append=T)
-#6634
-GBIF_JGMcleaned<-merge(GBIF_JGMcleaned, species_life_history)
-
-####check names through tnrs, see how JGM cleaned and tnrs cleaned lists vary####
-# # RESULT: original JGM cleaned list fine for this stage
-# tnrsnames <- sapply(species_life_history$species, function(x) name_suggest(x)$key[1], USE.NAMES=FALSE)
-# unique(tnrsnames)#42
-# class(tnrsnames)#integer
-# class(GBIF$speciesKey)#integer
-# 
[…]

### 2019-03-08 src/ancestral state.R

@@ -1,0 +1,34 @@
+#ancestral state
+
+require(phytools)
+require(ape)
+require(geiger)
+require(extRemes)
+
+tree <- read.nexus("../data/TrimmedMCCwoPolytomies.nex")
+ladderize(tree) -> tree
+rescale(tree, "depth", 1) -> tree
+
+data <- read.csv("../data/species_means_table.csv", header=T)
+data$A_P_num<-as.numeric(data$LH)-1
+
+data.frame(data$A_P_num) -> df
+rownames(df) <- gsub(" ","_", data$Species)
+colnames(df) <- "Habit"
+attach(df)
+
+names(Habit) <- rownames(df)
+name.check(tree, df)
+
+mtreesER <- make.simmap(tree, Habit, model = "ER")
+mtreesARD <- make.simmap(tree, Habit, model = "ARD")
+
+lr.test(mtreesER$logL, mtreesARD$logL, df=100)
+
+Maps <- make.simmap(tree, Habit, model = "ER", nsim=10000)
+XX <- describe.simmap(Maps, plot = F)
+pies<-cbind(data.frame(node=42+(1:nrow(XX$ace))), XX$ace)
+write.csv(pies, "../data/ancestral_state.csv")
+
+XX
+         

### 2019-03-25 manuscript/manuscript.Rmd
MS updated

@@
 Finally, if annual species have adapted to escape predictably drought prone seasons, observations of growing annual species (i.e.
-occurring in forms other than seed) should be rare during seasons when drought frequency is hightest.
+occurring in forms other than seed) should be rare during seasons when drought frequency is highest.
 Phylogenetic relatedness can influence tests of associations between species’ traits and their environments [@barrett1996comparative;@felsenstein1985phylogenies].
@@
 By combining VCI and TCI, the VHI is distinguishes drought from other forms of vegetative stress [@kogan1995application].
-The use of the VHI to detect drought has been validated globally and across ecosystem types [@aghakouchak2015remote], including in southern Africa, the focal region of this study (Figure \@ref:fig(mapsdroughtexamples)).
+The use of the VHI to detect drought has been validated globally and across ecosystem types [@aghakouchak2015remote], including in southern Africa, the focal region of this study (Figure \@ref(fig:mapsdroughtexamples)).
 To date, the VHI has most often been applied for evaluating drought risk for agricultural research [e.g., @rojas2011assessing; @kogan2016modelling].

### 2019-03-25 manuscript/manuscript.tex
MS updated

@@
 of growing annual species (i.e.~occurring in forms other than seed)
-should be rare during seasons when drought frequency is hightest.
+should be rare during seasons when drought frequency is highest.
 Phylogenetic relatedness can influence tests of associations between
@@
 al.}, 2015), including in southern Africa, the focal region of this
-study (Figure @ref:fig(mapsdroughtexamples)). To date, the VHI has most
+study (Figure \ref{fig:mapsdroughtexamples}). To date, the VHI has most
 often been applied for evaluating drought risk for agricultural research

### 2019-03-31 manuscript/manuscript.Rmd
revised

@@
 ---
-title             : "Drought frequency predicts life history strategies in _Heliophila_"
+title             : "Drought regimens predict life history strategies in _Heliophila_"
 shorttitle        : "Drought and life history"
@@
 abstract: \begin{center}\textbf{Summary}\end{center} \par Explaining variation in life history strategies is an enduring goal of evolutionary biology and ecology.
-For plants, annual and perennial life histories have long been thought to reflect adaptation to environments that experience alternative drought regimes.
-Neverthless, empirical support for this hypothesis from phylogeographic analyses remains lacking.
-\par Here, we tested classic life history theory in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa, with a phylogentically constrained analysis integrating 34 years of satellite-based drought detection with 2192 herbaria occurrence records.
-\par We found that annual and perennial life histories have evolved multiple times in \textit{Heliophila} and that compared to annuals, perennial species occur in environments where droughts are significantly less frequent.
-We also report evidence that annual plants adapt to predictable drought regimes by escaping drought prone seasons as seeds.
-\par These results provide compelling support for longstanding theoretical predictions and a more nuanced understanding of life history evolution.
-More broadly, this work outlines a scalable framework integrating herbaria records and remote sensing to address outstanding questions in evolutionary ecology.
+Early theory predicted that for plants, annual and perennial life histories reflect adaptation to environments that experience alternative drought regimens.
+Nevertheless, empirical support for this hypothesis from comparative analyses remains lacking.
+\par Here, we test classic life history theory in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa, controlling for phylogeny and integrating 34 years of satellite-based drought detection with 2,192 herbaria occurrence records.
+\par We find that the common ancestor of \textit{Heliophila} species was likely an annual, and that perenniality and annuality have repeatedly evolved, eight and five times, respectively.
+By comparing historical drought regimens, we show that annuals occur in environments where droughts are significantly more frequent than perennial species.
+We also provide evidence that annual plants adapt to predictable drought regimens by escaping drought prone seasons as seeds.
+\par These results yield compelling support for longstanding theoretical predictions by revealing the importance of drought frequency and predictability to explain plant life history.
+More broadly, this work highlights scalable approaches integrating herbaria records and remote sensing to address outstanding questions in evolutionary ecology.
 
-keywords          : "drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records, Heliophia"
+keywords          : "drought adaptation, herbaria records, Heliophila, life history evolution, phylogeography, remote sensing "
 
@@
 library(ggimage)
+library(lmerTest)
+library(lme4)
+library(emmeans)
 ```
@@
 In  plants, life histories are especially diverse, with some species completing their life cycle in a number of weeks to others that live for thousands of years [@brown1996oldlist].
-Along this continuum in angiosperms an important division exists between annuals which complete their seed to seed life cycle within a single calendar year from perennials which can persist over multiple years.
-Annual plants flower once, set seed, senesce, and then die, spending at least some portion of the year as a seed, where they are relatively protected from environmental stress.
-In contrast, perennial plants can continue vegetative growth after reproduction and must survive conditions experienced during all seasons but can also benefit from multiple bouts of reproduction and competative advantages.
-These represent fundamentally different life history strategies and predicting their occurance is important for community, ecosystem, and agricultural ecology.
-However, the environmental factors that explain their evolution and distributions remain empirically uresolved [@friedman2015all].
+Along this continuum in angiosperms an important division exists distinguishing annuals which complete their seed to seed life cycle within a single calendar year from perennials which can persist over multiple years.
+Annual plants do not need to survive through the full range of seasonal environmental variation and spend at least some portion of the year as a seed where they are relatively protected from environmental stress.
+In contrast, perennial plants can continue vegetative growth over multiple years and must survive conditions experienced during all seasons but can also benefit from competitive advantages and, if iteroparous, multiple bouts of reproduction.
+These represent fundamentally different life history strategies and predicting their occurrence is important for community, ecosystem, and agricultural ecology.
+However, the environmental factors that explain their evolution and distributions remain empirically unresolved [@friedman2015all].
 
@@
 Because lack of water is perhaps the greatest threat to survival during vegetative or reproductive growth in plants, this theory has been extended to the hypothesis that annuality is adaptive when it allows plants to escape drought [@schaffer1975selection].
-Indeed, adaptation to drought,  defined as episodes of increased aridity causing plant sresss [@passioura1996drought], is often invoked as an explanation for the success of annual species.
-And while a few cases are cited where annuality appears to be more common in environments with greater aridity [@stebbins1952aridity; @morishima1984differentiation], this hypothesis has yet to be supported in a phylogenetic context.
+Indeed, adaptation to drought,  defined as episodes of increased aridity causing plant sress [@passioura1996drought], is often invoked as an explanation for the success of annual species.
+And while a few cases are cited where annuality appears to be more common in environments with greater aridity [@stebbins1952aridity; @morishima1984differentiation], this hypothesis has yet to be supported while controlling for the effect of common ancestry (phylogeny) on life habit.
 In one previous study where this question was addressed phylogenetically,  [@evans2005climate] annuals were not found to be associated with environments that experience more drought.
-This could be explained by the relatively small number of species studied and the reliance on a limited number of weather stations to characterize environments, highlighting the need to develop more scalable methods to study trait evolution phylogeographically.
+This could be explained by the relatively small number of species studied and the reliance on a limited number of weather stations to characterize environments, highlighting the need to develop more scalable methods to study the geographic distributions of traits such as life history.
+Thus, in this study we leverage thousands of herbaria specimens among dozens of species and high-resolution remote sensing to study the distributions and environmental factors potentially driving the evolution and distribution of life history.
 
@@
 That is, when the frequency of drought is particularly high during certain seasons.
-Such predictability is important for selection to lead to escape from those seasons which are paricuarlary drought prone.
+Such predictability is important for selection to favor and escape strategy during those seasons which are particularly drought prone.
 While there has been at least one example of annuality associated with environments qualitatively classified as "predictable" in a general sense [@datson2008climate], the seasonal predictability of drought experienced by annuals has yet to be rigorously studied.
 As such, further empirical work is needed to support the model of annuality as a mechanism of drought adaptation via escape from drought prone seasons.
+Here we study herbarium collection dates to ask whether annuals indeed exhibit evidence of an escape strategy from seasons with elevated drought frequency.
 
-In addition to drought escape in annuals as a mechanism of adaptation to frequent predictable droughts, droughts may be necessary for the success of annuals more generally by acting as episodes of disturbance that provide opportunites for annuals to establish and compete with sympatric perennial species.
-Indeed, there is considerable evidence that perennials dominate in envirnoments where distrurbance events are infrequent [@rees1992germination;@corbin2004competition;@clary2012determinants].
+In addition to drought escape in annuals as a mechanism of adaptation to frequent and predictable droughts, droughts may be necessary for the success of annuals more generally by acting as episodes of disturbance that provide opportunities for annuals to establish and compete with sympatric perennial species.
+Indeed, there is evidence that perennials dominate in environments where disturbance events are infrequent [@rees1992germination;@corbin2004competition;@clary2012determinants].
 The resulting prediction from this hypothesis is that in the absence of frequent drought, perenniality should evolve.
-However, little is known about this component of life history evolution because previous work has largely focused on the origins of annuality rather than perenniality [@friedman2015all].
+However, little is known about this component of life history evolution because previous work has almost entirely focused on the origins of annuality rather than perenniality [@friedman2015all].
+This highlights the need to study taxa which have seen transitions from annual to perennial life histories as well.
 
@@
 If annuality is an adaptive strategy allowing plants to escape drought prone seasons, then drought frequency should predict the distribution of life history strategies across landscapes, and annual species should be more commonly associated with drought prone regions than perennial species.
-Additionally, if perenniality offers competative advantage in the absence of drought, associations between life history and drought frequency should be signficant in even if phylogenies are enriched for transitions from annual to perenial life history strategy.
+Additionally, if perenniality offers competitive advantage in the absence of drought, associations between life history and drought frequency should be significant when phylogenies include transitions from annual to perennial life history strategy.
 Finally, if annual species have adapted to escape predictably drought prone seasons, observations of growing annual species (i.e.
 occurring in forms other than seed) should be rare during seasons when drought frequency is highest.
-Phylogenetic relatedness can influence tests of associations between species’ traits and their environments [@barrett1996comparative;@felsenstein1985phylogenies].
-Therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetic context.
+Phylogenetic relatedness can influence tests of associations between species’ traits and their environments by confounding common environments caused by selection from common environments caused my ancestry.
+[@barrett1996comparative;@felsenstein1985phylogenies].
+Therefore, we assessed the relationship between life history distribution and drought frequency while controlling for phylogeny.
 
@@
 ### Data availability
-All analyses were performed using R. All data and the source code to produce this manuscript are available at https://github.com/greymonroe/heliophila. Software used is listed in the supplement.
+All analyses were performed using R. All data and the source code to produce this manuscript are available at https://github.com/greymonroe/heliophila.
 
 ### Life history data for _Heliophila_
-_Heliophila_ is a genus of flowering plants endemic to the southern Africa including the Cape Floristic and Succulent Karoo Regions.
-These are among the most botanically diverse environments on Earth and the _Heliophila_ species occurring there are considered to make up the most diverse genus of the family Brassicaceae [@mummenhoff2005phylogeny;@mandakova2012whole].
-This genus includes both perennial (n =21) and annual (n = 21) species and this change in life history strategy has likely arisen multiple independent times [@appel1997generic;@mummenhoff2005phylogeny].
-Furthermore, the fine scale climatic heterogeneity of Southern Africa is ideal for studying the distribution of traits in relation to environmental parameters [@sayre2013new].
-We used life histories reported by @mummenhoff2005phylogeny, grouping species with annual or perennial life histories.
-Perenniality was defined based any form of perennial life history (e.g., herbs, shrubs, mixed, etc).
+_Heliophila_ is a genus of flowering plants endemic to  southern Africa including the Cape Floristic and Succulent Karoo Regions.
+These are among the most botanically diverse environments on Earth and the _Heliophila_ species occurring there are considered to be among the most phenotypically diverse genera of the family Brassicaceae [@mummenhoff2005phylogeny;@mandakova2012whole].
+This genus includes both perennial and annual species, and transitions between life history strategy may have occurred multiple independent times [@appel1997generic;@mummenhoff2005phylogeny].
+Furthermore, the fine scale climatic heterogeneity of southern Africa is ideal for studying the distribution of traits in relation to environmental parameters [@sayre2013new].
+We used life histories reported by @mummenhoff2005phylogeny, grouping species into annual or perennial life history categories.
+Perenniality was defined as any form of perennial life history (e.g., herbs, shrubs, mixed, etc).
 Because the nature of species reported with mixed traits were unknown (i.e.
 plasticity vs.
-genetic variation), we classified these species here as perennial since they can maintain vegetative growth after reproduction at least to some capacity.
+genetic variation), we classified these species here as perennial since they can maintain vegetative across multiple years at least to some capacity.
 
@@
 ```
-To characterize the distributions of annual and perennial _Heliophila_ species, all (`r nrow(GBIF)`) records for the genus _Heliophila_ were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018 [@gbifdownload].
-herbaria records records such as these provide a rich data sources to characterize the geographical distributions of species [@thiers2016index; @willis2017old; @lang2018using].
-And as they become digitized [@soltis2017digitization], these collections have been used to study relationships between trait distributions, geography, and climate [@davis2015herbarium;@wolf2016altitudinal;@stropp2016mapping;@vaclavik2017effects].
+To characterize the distributions of annual (studied here, n = 21) and perennial (studied here, n = 21) _Heliophila_ species, all (`r nrow(GBIF)`) records for the genus _Heliophila_ were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018 [@gbifdownload].
+Herbaria records such as these provide a rich data sources to characterize the geographical distributions of species [@thiers2016index; @willis2017old; @lang2018using].
+And as they become digitized [@soltis2017digitization], herbaria collections have been used to study relationships between trait distributions, geography, and climate [@davis2015herbarium;@wolf2016altitudinal;@stropp2016mapping;@vaclavik2017effects].
 
@@
 Generated from data collected by NOAA AVHRR satellites since 1981, the VHI is a composite index combining Normalized Difference Vegetation Index (NDVI) derived quantification of vegetative stress (Vegetative Condition Index - VCI) with temperature stress indicated by anomalies in thermal spectra (Temperature Condition Index - TCI).
-These indicies were developed to create an unbiased quantification of drought across ecosystem types.
+These indices were developed to create an unbiased quantification of drought across ecosystem types.
 The VHI of year $y$ during week $w$ of $[1,52]$ at pixel $i$ is derived from the following equations, where $n$ is the number of years observed.
@@
 $$TCI_{y,w,i} = 100\frac{T_{max,w,i}-T_{y,w,i}}{T_{max,w,i} - T_{min,w,i}}$$
-Similarly, low TCI values indicate episdoes of high thermal stress shown to be negatively correlated with precipition and soil moisture [@aghakouchak2015remote].
+Similarly, low TCI values indicate episdoes of high thermal stress shown to be negatively correlated with precipitation and soil moisture [@aghakouchak2015remote].
 
 $$VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})$$
-By combining VCI and TCI, the VHI is distinguishes drought from other forms of vegetative stress [@kogan1995application].
-The use of the VHI to detect drought has been validated globally and across ecosystem types [@aghakouchak2015remote], including in southern Africa, the focal region of this study (Figure \@ref(fig:mapsdroughtexamples)).
+By combining VCI and TCI, the VHI distinguishes drought from other forms of vegetative stress [@kogan1995application].
+The use of the VHI to detect drought has been validated globally and across ecosystem types [@aghakouchak2015remote], including in southern Africa, the focal region of this study (e.g.
+Figure \@ref(fig:mapsdroughtexamples)).
 To date, the VHI has most often been applied for evaluating drought risk for agricultural research [e.g., @rojas2011assessing; @kogan2016modelling].
@@
 To characterize drought regimens across the distributions of annual and perennial species of _Heliophila_, we calculated drought during different seasons at the location of observations for _Heliophila_ records using the VHI.
-Specifically, we created global maps of the frequencies of observing drought conditions between years (VHI<40, NOAA) during the winter (quarter surrounding winter solstice), spring (quarter surrounding spring equinox), summer (quarter surrounding summer solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015.
-From these maps, the drought frequency during the winter, spring, summer, and fall were extracted for the locations of all GBIF records.
+Specifically, we created maps of the frequencies of observing drought conditions between years (VHI<40, NOAA) during the winter (quarter surrounding winter solstice), spring (quarter surrounding spring equinox), summer (quarter surrounding summer solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015 across the range of _Heliophila_.
+From these maps, the drought frequency (the number of times drought is observed divided by the total number of years, 34) during the winter, spring, summer, and fall were extracted for the locations of all GBIF records.
 
@@
 
-To avoid instances with spurious location data, we filtered raw GBIF by restricting our analyses to include only:
+To avoid instances with spurious location data, we filtered raw GBIF records by restricting our analyses to include only:
 
@@
 Model selection and phylogenetic analyses were conducted through the CIPRES Science Gateway [@miller2010creating].
-Ancestral state estimation was performed in R using the package phytools [@revell2012phytools] to genearte 10,000 stochastic character maps simulated under an equal rates model of character evolution for the trait life habit (annual or perennial).
+Ancestral state estimation was performed in R using the package phytools [@revell2012phytools] to generate 10,000 stochastic character maps simulated under an equal rates model of character evolution for the trait life habit (annual or perennial).
 
@@
 ```
-To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimes, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency.
+To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimens, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency.
 First, we compared the frequency of drought across all occurrence records of between annual and perennial species by t-tests.
 Independent tests were performed to compare drought frequency between annuals and perennials during each season.
-We also tested for a relationship between drought frequency and life history, season, and their interaction by an analysis of variance while including species as a random effect.
+We also tested for a relationship between drought frequency and life history, season, and their interaction by analysis of variance while including species as a random effect.
 To account for variation in the number of occurrence records per species, we next calculated the mean drought frequency during the winter, spring, summer and fall for each species.
 Because shared evolutionary history of closely related species can lead to spurious associations between traits and environments  [@felsenstein1985phylogenies], we tested for a relationship between life history strategy and drought frequency while controlling for phylogeny using phylogenetic logistic regression [@ives2010phylogenetic].
-This statistical approach is designed to control for the confounding effects of geographic relationships explained by demographic history rather than processes of natural selection.
+This statistical approach is designed to control for the confounding effects of common ancestry's influence on demographic features such as geospatial relationships when addressing hypotheses about the role of natural selection on trait distributions.
 
@@
 ephemera_.
-Images (a,c,d) courtesy of The Bavarian Natural History Collections (CC BY-SA 4.0) and (b) The London  Natural History Museum (CC BY 4.0).
+Images (b,d,e) courtesy of The Bavarian Natural History Collections (CC BY-SA 4.0) and (c) The London  Natural History Museum (CC BY 4.0).
 Links to images are found in the supplement.
@@
 ```
+
 (ref:mapsboxplots) Locations of occurrence records of (a) annual and (b) perennial _Heliophila._ Drought frequency during the (c) winter, (d) spring, (e) summer and (f) fall detected using the VHI.
-(g) Drought frequencies during each season at the observation locations of annual and perennial _Heliophila_ (t tests contrasting annuals and perennials, \*\* = p < 0.01).
+(g) Drought frequencies during each season at the observation locations of annual and perennial _Heliophila_ (Post-hoc contrasts annuals and perennials from ANOVA, \*\* = p < 0.01).
 
@@
 
+ann_text <- data.frame(x = 1.5,drought_freq = 0.8,life_history="a", lab = "**",
+                       season = factor(c("Summer","Fall"), levels=c("Winter","Spring","Summer","Fall")))
+
 boxplots<-ggplot(GBIF_cleaned_long, aes(x=life_history, col=life_history, y=drought_freq))+
@@
   labs(y="Drought frequency")+
-  annotate("text", x=1.5, y=0.8, label= "**")
+  geom_text(data=ann_text, label="**", col="black", nudge_x = 0.5, cex=4)
 
@@
 ```
-The _Heliophila_ phylogeny was consistent with previous studies [@mummenhoff2005phylogeny;@mandakova2012whole].
+
+The topology of the estimated _Heliophila_ phylogeny was consistent with previous studies [@mummenhoff2005phylogeny;@mandakova2012whole].
 Based on 10,000 stochastic character maps simulated under an equal rates model of character evolution in life history, an average of approximately eight changes from annual to perennial and five changes from perennial to annual are observed per stochastic character map (Figure \@ref(fig:phylogeny)a).
-These results suggest that the ancestral state of _Heliophila_ was annual and that both charcter states have arisen independently mutliple times.
+These results suggest that the ancestral state of _Heliophila_ was annual and that both character states have arisen independently mutliple times.
 
@@
 
-```{r ttests, eval=T, message=F, warning="hide", results="hide"}
-twin<-t.test(Winter~life_history, GBIF_cleaned)
-tspr<-t.test(Spring~life_history, GBIF_cleaned)
-tsum<-t.test(Summer~life_history, GBIF_cleaned)
-tfal<-t.test(Fall~life_history, GBIF_cleaned)
+```{r anova, eval=T, message=F, warning="hide", results="hide"}
 
 melted<-GBIF_cleaned %>% select(life_history, Winter, Spring, Summer,Fall, species) %>% gather(Winter, Spring, Summer,Fall, key="season",value="droughtfreq")
-library(nlme)
-fit<-lme(droughtfreq~life_history*season, random=~1|species, melted)
-anovatable<-data.frame(predictor=c("intercept","life history","season","life history x season"), anova(fit))
+
+fit<-lmer(droughtfreq~life_history*season+(1|species), melted)
+anovatable<-data.frame(predictor=c("life history","season","life history x season"), anova(fit))
+colnames(anovatable)[7]<-"p-value"
 row.names(anovatable)<-NULL
+
+post_hoc<-emmeans(fit, list(pairwise ~ life_history*season), adjust = "tukey")
+post_hoc<-as.data.frame(post_hoc$`pairwise differences of life_history, season`)
+post_hoc<-post_hoc[c(1,14,23,28), ]
+post_hoc$contrast<-c("Fall","Spring","Summer","Winter")
 
@@
 
-We found that the frequency of drought was significantly higher at the locations of occurrence records for annual species. When comparing across all occurrence records
-(all records rather than species means, Figure \@ref(fig:mapsboxplots)g), the frequency of drought was significantly higher at the location of annuals during the winter (t = `r twin$statistic`, p < 0.01), spring (t = `r tspr$statistic`, p < 0.01), summer (t = `r tsum$statistic`, p = < 0.01), and fall (t = `r tfal$statistic`, p < 0.01).
-A mixed-model analysis of variace which included species as random effect also identified a signficant relationship between drought frequency and life history (p < 0.01), season (p < 0.01) and their interaction (p < 0.01) (Table \@ref(tab:anovatable)).
+We found that the frequency of drought was significantly higher at the locations of occurrence records for annual species.
+When comparing across all occurrence records (all records rather than species means, Figure \@ref(fig:mapsboxplots)g), a mixed-model analysis of variance which included species as random effect revealed a significant relationship between drought frequency and life history (p < 0.01), season (p < 0.01) and their interaction (p < 0.01) (Table \@ref(tab:anovatable)).
+Post-hoc contrasts showed that the frequency of drought was significantly higher at the location of annuals during the summer (z ratio = `r post_hoc$z.ratio[post_hoc$contrast=="Summer"]`, p = < 0.01), and fall (z ratio = `r post_hoc$z.ratio[post_hoc$contrast=="Fall"]`, p < 0.01).
 Because a comparison across all occurrence records does not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable)) or species relatedness (Figure \@ref(fig:phylogeny)a), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials using phylogenetic logistic regression.
 We found that the mean drought frequencies were significantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a,b).
-These findings indicate that demographic history alone does not explain differences in the drought frequencies experienced between the environments of annual and perennial _Heliophila_.
+These findings indicate that common ancestry alone does not explain differences in the drought frequencies experienced between the environments of annual and perennial _Heliophila_.
 
@@
 
-(ref:lineplots) (a) Comparison (mean +- SE) of drought frequency across seasons calcualted at the occurance locations of GBIF records of annual and perennial species of _Heliophila_.
+(ref:lineplots) (a) Comparison (mean +- SE) of drought frequency across seasons calculated at the occurrence locations of GBIF records of annual and perennial species of _Heliophila_.
 (b) Results from phylogenetic logistic regression, where the line shows the model fit and \* = p < 0.05, \*\* = p < 0.01.
 Annuals were scored as 0 and perennials as 1.
-(c) Collection dates of GBIF records of annual and perennial species of _Heliophila_ in relation to the photoperiodic calender where day 1 is intermediate to the fall equinox and winter soltice.
+(c) Collection dates of GBIF records of annual and perennial species of _Heliophila_ in relation to the photoperiodic calendar where day 1 is intermediate to the fall equinox and winter solstice.
 
@@
 We cannot eliminate the possibility that confounding traits or environmental variables are the causative factors explaining variation in the distributions of annual and perennial species.
-Nevertheless, these results provide quantitative support for the classic prediction that annual species are found in environments that experience more frequent drought than perennial species.
-These findings are a valuable contribution to previous investigations of associations between life history and climate [@evans2005climate;@morishima1984differentiation;@datson2008climate; @cruz2009molecular].
+Nevertheless, these results provide quantitative support for the classic prediction that annual species are found in environments that experience more frequent drought than perennial species, building on previous investigations of associations between life history and climate [@evans2005climate;@morishima1984differentiation;@datson2008climate; @cruz2009molecular].
 To our knowledge this is the first study to demonstrate a significant association between life history and drought in a phylogenetic context informed by large scale species distribution data and long term drought detection.
@@
 
-These findings empirical provide support classical theoretical predictions about the adaptive value of annual and perennial life history strategies.
-Taken together, they suggest that in _Heliophila_, annual species are adapted to environments with increased summer droughts by avoiding these seasons in a dormant seed phase of their life cycle.
-They also suggest that perenniality is adaptive in environemnts where droughts are less frequent.
+These findings empirically  support classical theoretical predictions about the adaptive value of annual and perennial life history strategies.
+Taken together, they suggest that in _Heliophila_, annual species are adapted to environments with predictable droughts by escaping drought prone seasons during the dormant seed phase of their life cycle.
+They also suggest that perenniality is adaptive in environments where droughts are less frequent.
 While most previous work has focused on describing the evolutionary origins of annuality [@barrett1996comparative;@conti1999phylogenetic;@andreasen2001unequal;@verboom2004testing;@friedman2015all] there are at least a few other cases where perenniality appears to have arisen from an annual ancestor [@bena1998molecular;@tank2008annuals].
 And while early theory predicted selection for annuality when adult mortality is high [@stearns1992evolution],  we also find evidence that perenniality could be explained by reduced frequency of drought.
-This is supported by the theoretical prediction that perenniality is advantageous is stable habitats.
+This is supported by the theoretical prediction that perenniality is advantageous in stable habitats.
 The phylogeny reveals several transitions from annual to perennial life history (Figure \@ref(fig:phylogeny)a) and the distributions of perennial _Heliophila_ extend into regions where drought frequency is low (Figure \@ref(fig:mapsboxplots)b, Figure \@ref(fig:speciesmaps)).
@@
 These findings suggest that species with locally adaptive life history strategies could be threatened by rapidly changing drought regimens [@dai2011drought].
-In light of the findings here, changes in eastern South Africa which is forecast to experience signicantly lower rainfall [@south2017a] could be particularly impactful.
-In this study we found that this region is currently dominiated by derived perennial species of _Heliophila_.
-However, if drought become more frequent in this region, it may allow for the establishement of annuals.
+In light of the findings here, forecasted reductions in rainfall across eastern South Africa [@south2017a] could be particularly impactful to plant community compositions.
+Here we found that this region is currently dominated by derived perennial species of _Heliophila_.
+However, a scenario in which droughts become more frequent in this region may allow for the establishment of annuals.
 Such changes in selection patterns and shifts in plant functional diversity could have impacts on ecosystem functioning and processes such as carbon cycling [@garnier1997specific; @roumet2006suites;@monroe2018ecoevolutionary].
@@
 
-In conclusion, we find strong support for classic life history theory which predicts that annuality is adaptive in environments where droughts occur more frequently.
-Additionally, we report evidence consistent with a life history model in annuals in which they escape drought prone seasons during the seed phase of their life cycle.
+In conclusion, we find strong support for classic life history theory that predicts that annuality is adaptive in environments with frequency and predictable droughts.
+We report evidence consistent with a life history model in annuals in which they escape drought prone seasons during the seed phase of their life cycle.
 Finally, we find evidence that the distributions of perennial lineages may indicate a competitive advantage in areas where droughts are infrequent.
@@
 # Acknowledgments 
-We thank Jesse Lasky, members of the Sloan lab at CSU and three ananymous reviewers for generous feedback that improved the quality of this work.
+We thank Jesse Lasky, members of the Sloan lab, and three anonymous reviewers for generous feedback that improved the quality of this work.
 This research was supported by NSF Award 1701918 and USDA-NIFA Award 2014-38420-21801 to JGM.
@@
 
-```{r anovatable, results='asis', echo = F}
-apa_table(anovatable
-          , caption = "Analysis of variance (ANOVA) to compare drought frequency as a function of life history, season, and their interaction while including species as a random effect.",
-          digits=4
-)
-```
-
 ```{r speciesmeanstable, results = 'asis', echo = F}
@@
 
+```{r anovatable, results='asis', echo = F}
+apa_table(anovatable
+          , caption = "Analysis of variance (ANOVA) to compare drought frequency as a function of life history, season, and their interaction while including species as a random effect.",
+          note = "Type III Analysis of Variance Table with Satterthwaite's method",
+          digits=4
+)
+```
+
 ```{r results = 'hide', echo = F}
@@
 Vegetative Health Index (VHI) values below 40 indicate remotely sensed drought.
-Drought detection during this years is validated by precipitation based estimates of drought ocurrance (Standardized Precipitation Index) which confirm that while (a) 2015 was one of the worst drought years on record, (b) 1999 was one of the wettest, and (c) 2012 was typical in terms of precipitation patterns [@monyela2017two].
+Drought detection during these years is validated by previously reported precipitation based estimates of drought ocurrence [@monyela2017two] which confirm that while (a) 2015 was one of the worst drought years on record, (b) 1999 was one of the wettest, and (c) 2012 was typical in terms of precipitation patterns.
 It is worth noting that drought can be detected using the VHI across ecosystems, including those inhabited by perennial rather than annual _Heliophila_ species.
@@
   theme_classic()+
-  theme(legend.position = c(0.92,.3),legend.text=element_text(size=rel(0.9)), legend.key.size=unit(.2,"line"), legend.background = element_rect(fill=alpha('gray90'), color = "black"), text = element_text(size=4.5))+
-  labs(title="VHI detected drought during summer 2015", x="Longitude (°)", y="Latitude (°)")
+   theme(legend.position = c(1,.3),legend.text=element_text(size=rel(0.6)), legend.title = element_text(size=rel(0.6)), legend.key.size=unit(.2,"line"), legend.background = element_rect(fill=alpha('gray90'), color = "black"), text = element_text(size=8))+
+  labs(title="VHI detected drought \n during summer 2015", x="Longitude (°)", y="Latitude (°)")
 
@@
   theme_classic()+
-  theme(legend.position = c(0.92,.3),legend.text=element_text(size=rel(0.9)), legend.key.size=unit(.2,"line"), legend.background = element_rect(fill=alpha('gray90'), color = "black"), text = element_text(size=4.5))+
-  labs(title="VHI detected drought during summer 1999", x="Longitude (°)", y="Latitude (°)")
+  theme(legend.position = c(1,.3),legend.text=element_text(size=rel(0.6)), legend.title = element_text(size=rel(0.6)), legend.key.size=unit(.2,"line"), legend.background = element_rect(fill=alpha('gray90'), color = "black"), text = element_text(size=8))+
+  labs(title="VHI detected drought \n during summer 1999", x="Longitude (°)", y="Latitude (°)")
 
@@
   theme_classic()+
-  theme(legend.position = c(0.92,.3),legend.text=element_text(size=rel(0.9)), legend.key.size=unit(.2,"line"), legend.background = element_rect(fill=alpha('gray90'), color = "black"), text = element_text(size=4.5))+
-  labs(title="VHI detected drought during summer 2012", x="Longitude (°)", y="Latitude (°)")
+  theme(legend.position = c(1,.3),legend.text=element_text(size=rel(0.6)), legend.title = element_text(size=rel(0.6)), legend.key.size=unit(.2,"line"), legend.background = element_rect(fill=alpha('gray90'), color = "black"), text = element_text(size=8))+
+  labs(title="VHI detected drought \n during summer 2012", x="Longitude (°)", y="Latitude (°)")
 

### 2019-03-31 manuscript/manuscript.tex
revised

@@
 \hypersetup{unicode=true,
-            pdftitle={Drought frequency predicts life history strategies in Heliophila},
+            pdftitle={Drought regimens predict life history strategies in Heliophila},
             pdfauthor={J. Grey Monroe, Brian Gill, Kathryn G. Turner, \& John K. McKay},
-            pdfkeywords={drought adaptation, life history evolution, remote sensing,
-phylogeography, herbaria records, Heliophia},
+            pdfkeywords={drought adaptation, herbaria records, Heliophila, life history
+evolution, phylogeography, remote sensing},
             pdfborder={0 0 0},
@@
 
-  \title{Drought frequency predicts life history strategies in \emph{Heliophila}}
+  \title{Drought regimens predict life history strategies in \emph{Heliophila}}
     \author{J. Grey Monroe\textsuperscript{1,2}, Brian Gill\textsuperscript{3},
@@
 \textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{3} Institute for Environment and Society, Brown University, Providence, RI 02912, USA\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA 16802, USA}
-\keywords{drought adaptation, life history evolution, remote sensing, phylogeography, herbaria records, Heliophia}
+\keywords{drought adaptation, herbaria records, Heliophila, life history evolution, phylogeography, remote sensing }
 \usepackage{csquotes}
@@
 Explaining variation in life history strategies is an enduring goal of
-evolutionary biology and ecology. For plants, annual and perennial life
-histories have long been thought to reflect adaptation to environments
-that experience alternative drought regimes. Neverthless, empirical
-support for this hypothesis from phylogeographic analyses remains
+evolutionary biology and ecology. Early theory predicted that for
+plants, annual and perennial life histories reflect adaptation to
+environments that experience alternative drought regimens. Nevertheless,
+empirical support for this hypothesis from comparative analyses remains
 lacking.
@@
 
-Here, we tested classic life history theory in \textit{Heliophila}
+Here, we test classic life history theory in \textit{Heliophila}
 (Brassicaceae), a diverse genus of flowering plants native to Africa,
-with a phylogentically constrained analysis integrating 34 years of
-satellite-based drought detection with 2192 herbaria occurrence records.
+controlling for phylogeny and integrating 34 years of satellite-based
+drought detection with 2,192 herbaria occurrence records.
 
@@
 
-We found that annual and perennial life histories have evolved multiple
-times in \textit{Heliophila} and that compared to annuals, perennial
-species occur in environments where droughts are significantly less
-frequent. We also report evidence that annual plants adapt to
-predictable drought regimes by escaping drought prone seasons as seeds.
+We find that the common ancestor of \textit{Heliophila} species was
+likely an annual, and that perenniality and annuality have repeatedly
+evolved, eight and five times, respectively. By comparing historical
+drought regimens, we show that annuals occur in environments where
+droughts are significantly more frequent than perennial species. We also
+provide evidence that annual plants adapt to predictable drought
+regimens by escaping drought prone seasons as seeds.
 
@@
 
-These results provide compelling support for longstanding theoretical
-predictions and a more nuanced understanding of life history evolution.
-More broadly, this work outlines a scalable framework integrating
-herbaria records and remote sensing to address outstanding questions in
-evolutionary ecology. 
+These results yield compelling support for longstanding theoretical
+predictions by revealing the importance of drought frequency and
+predictability to explain plant life history. More broadly, this work
+highlights scalable approaches integrating herbaria records and remote
+sensing to address outstanding questions in evolutionary ecology.
+
 }
@@
 thousands of years (Brown, 1996). Along this continuum in angiosperms an
-important division exists between annuals which complete their seed to
-seed life cycle within a single calendar year from perennials which can
-persist over multiple years. Annual plants flower once, set seed,
-senesce, and then die, spending at least some portion of the year as a
-seed, where they are relatively protected from environmental stress. In
-contrast, perennial plants can continue vegetative growth after
-reproduction and must survive conditions experienced during all seasons
-but can also benefit from multiple bouts of reproduction and competative
-advantages. These represent fundamentally different life history
-strategies and predicting their occurance is important for community,
+important division exists distinguishing annuals which complete their
+seed to seed life cycle within a single calendar year from perennials
+which can persist over multiple years. Annual plants do not need to
+survive through the full range of seasonal environmental variation and
+spend at least some portion of the year as a seed where they are
+relatively protected from environmental stress. In contrast, perennial
+plants can continue vegetative growth over multiple years and must
+survive conditions experienced during all seasons but can also benefit
+from competitive advantages and, if iteroparous, multiple bouts of
+reproduction. These represent fundamentally different life history
+strategies and predicting their occurrence is important for community,
 ecosystem, and agricultural ecology. However, the environmental factors
 that explain their evolution and distributions remain empirically
-uresolved (Friedman \& Rubin, 2015).
+unresolved (Friedman \& Rubin, 2015).
 
@@
 Indeed, adaptation to drought, defined as episodes of increased aridity
-causing plant sresss (Passioura, 1996), is often invoked as an
+causing plant sress (Passioura, 1996), is often invoked as an
 explanation for the success of annual species. And while a few cases are
@@
 greater aridity (Stebbins Jr, 1952; Morishima \emph{et al.}, 1984), this
-hypothesis has yet to be supported in a phylogenetic context. In one
-previous study where this question was addressed phylogenetically,
-(Evans \emph{et al.}, 2005) annuals were not found to be associated with
-environments that experience more drought. This could be explained by
-the relatively small number of species studied and the reliance on a
-limited number of weather stations to characterize environments,
-highlighting the need to develop more scalable methods to study trait
-evolution phylogeographically.
+hypothesis has yet to be supported while controlling for the effect of
+common ancestry (phylogeny) on life habit. In one previous study where
+this question was addressed phylogenetically, (Evans \emph{et al.},
+2005) annuals were not found to be associated with environments that
+experience more drought. This could be explained by the relatively small
+number of species studied and the reliance on a limited number of
+weather stations to characterize environments, highlighting the need to
+develop more scalable methods to study the geographic distributions of
+traits such as life history. Thus, in this study we leverage thousands
+of herbaria specimens among dozens of species and high-resolution remote
+sensing to study the distributions and environmental factors potentially
+driving the evolution and distribution of life history.
 
@@
 drought is particularly high during certain seasons. Such predictability
-is important for selection to lead to escape from those seasons which
-are paricuarlary drought prone. While there has been at least one
-example of annuality associated with environments qualitatively
-classified as \enquote{predictable} in a general sense (Datson \emph{et
-al.}, 2008), the seasonal predictability of drought experienced by
-annuals has yet to be rigorously studied. As such, further empirical
-work is needed to support the model of annuality as a mechanism of
-drought adaptation via escape from drought prone seasons.
+is important for selection to favor and escape strategy during those
+seasons which are particularly drought prone. While there has been at
+least one example of annuality associated with environments
+qualitatively classified as \enquote{predictable} in a general sense
+(Datson \emph{et al.}, 2008), the seasonal predictability of drought
+experienced by annuals has yet to be rigorously studied. As such,
+further empirical work is needed to support the model of annuality as a
+mechanism of drought adaptation via escape from drought prone seasons.
+Here we study herbarium collection dates to ask whether annuals indeed
+exhibit evidence of an escape strategy from seasons with elevated
+drought frequency.
 
 In addition to drought escape in annuals as a mechanism of adaptation to
-frequent predictable droughts, droughts may be necessary for the success
-of annuals more generally by acting as episodes of disturbance that
-provide opportunites for annuals to establish and compete with sympatric
-perennial species. Indeed, there is considerable evidence that
-perennials dominate in envirnoments where distrurbance events are
-infrequent (Rees \& Long, 1992; Corbin \& D'Antonio, 2004; Clary, 2012).
-The resulting prediction from this hypothesis is that in the absence of
-frequent drought, perenniality should evolve. However, little is known
-about this component of life history evolution because previous work has
-largely focused on the origins of annuality rather than perenniality
-(Friedman \& Rubin, 2015).
+frequent and predictable droughts, droughts may be necessary for the
+success of annuals more generally by acting as episodes of disturbance
+that provide opportunities for annuals to establish and compete with
+sympatric perennial species. Indeed, there is evidence that perennials
+dominate in environments where disturbance events are infrequent (Rees
+\& Long, 1992; Corbin \& D'Antonio, 2004; Clary, 2012). The resulting
+prediction from this hypothesis is that in the absence of frequent
+drought, perenniality should evolve. However, little is known about this
+component of life history evolution because previous work has almost
+entirely focused on the origins of annuality rather than perenniality
+(Friedman \& Rubin, 2015). This highlights the need to study taxa which
+have seen transitions from annual to perennial life histories as well.
 
@@
 with drought prone regions than perennial species. Additionally, if
-perenniality offers competative advantage in the absence of drought,
+perenniality offers competitive advantage in the absence of drought,
 associations between life history and drought frequency should be
-signficant in even if phylogenies are enriched for transitions from
-annual to perenial life history strategy. Finally, if annual species
-have adapted to escape predictably drought prone seasons, observations
-of growing annual species (i.e.~occurring in forms other than seed)
-should be rare during seasons when drought frequency is highest.
-Phylogenetic relatedness can influence tests of associations between
-species' traits and their environments (Felsenstein, 1985; Barrett
-\emph{et al.}, 1996). Therefore we assessed the relationship between
-life history distribution and drought frequency in a phylogenetic
-context.
+significant when phylogenies include transitions from annual to
+perennial life history strategy. Finally, if annual species have adapted
+to escape predictably drought prone seasons, observations of growing
+annual species (i.e.~occurring in forms other than seed) should be rare
+during seasons when drought frequency is highest. Phylogenetic
+relatedness can influence tests of associations between species' traits
+and their environments by confounding common environments caused by
+selection from common environments caused my ancestry. (Felsenstein,
+1985; Barrett \emph{et al.}, 1996). Therefore, we assessed the
+relationship between life history distribution and drought frequency
+while controlling for phylogeny.
 
@@
 produce this manuscript are available at
-\url{https://github.com/greymonroe/heliophila}. Software used is listed
-in the supplement.
+\url{https://github.com/greymonroe/heliophila}.
 
@@
 
-\emph{Heliophila} is a genus of flowering plants endemic to the southern
+\emph{Heliophila} is a genus of flowering plants endemic to southern
 Africa including the Cape Floristic and Succulent Karoo Regions. These
 are among the most botanically diverse environments on Earth and the
-\emph{Heliophila} species occurring there are considered to make up the
-most diverse genus of the family Brassicaceae (Mummenhoff \emph{et al.},
-2005; Mandáková \emph{et al.}, 2012). This genus includes both perennial
-(n =21) and annual (n = 21) species and this change in life history
-strategy has likely arisen multiple independent times (Appel \&
-Al-Shehbaz, 1997; Mummenhoff \emph{et al.}, 2005). Furthermore, the fine
-scale climatic heterogeneity of Southern Africa is ideal for studying
-the distribution of traits in relation to environmental parameters
-(Sayre \emph{et al.}, 2013). We used life histories reported by
-Mummenhoff \emph{et al.} (2005), grouping species with annual or
-perennial life histories. Perenniality was defined based any form of
-perennial life history (e.g., herbs, shrubs, mixed, etc). Because the
-nature of species reported with mixed traits were unknown
-(i.e.~plasticity vs.~genetic variation), we classified these species
-here as perennial since they can maintain vegetative growth after
-reproduction at least to some capacity.
+\emph{Heliophila} species occurring there are considered to be among the
+most phenotypically diverse genera of the family Brassicaceae
+(Mummenhoff \emph{et al.}, 2005; Mandáková \emph{et al.}, 2012). This
+genus includes both perennial and annual species, and transitions
+between life history strategy may have occurred multiple independent
+times (Appel \& Al-Shehbaz, 1997; Mummenhoff \emph{et al.}, 2005).
+Furthermore, the fine scale climatic heterogeneity of southern Africa is
+ideal for studying the distribution of traits in relation to
+environmental parameters (Sayre \emph{et al.}, 2013). We used life
+histories reported by Mummenhoff \emph{et al.} (2005), grouping species
+into annual or perennial life history categories. Perenniality was
+defined as any form of perennial life history (e.g., herbs, shrubs,
+mixed, etc). Because the nature of species reported with mixed traits
+were unknown (i.e.~plasticity vs.~genetic variation), we classified
+these species here as perennial since they can maintain vegetative
+across multiple years at least to some capacity.
 
@@
 
-To characterize the distributions of annual and perennial
-\emph{Heliophila} species, all (8670) records for the genus
-\emph{Heliophila} were downloaded from the Global Biodiversity
-Information Facility (gbif.org) on July 21, 2018 (GBIF, 2018). herbaria
-records records such as these provide a rich data sources to
+To characterize the distributions of annual (studied here, n = 21) and
+perennial (studied here, n = 21) \emph{Heliophila} species, all (8670)
+records for the genus \emph{Heliophila} were downloaded from the Global
+Biodiversity Information Facility (gbif.org) on July 21, 2018 (GBIF,
+2018). Herbaria records such as these provide a rich data sources to
 characterize the geographical distributions of species (Thiers, 2016;
 Willis \emph{et al.}, 2017; Lang \emph{et al.}, 2018). And as they
-become digitized (Soltis, 2017), these collections have been used to
+become digitized (Soltis, 2017), herbaria collections have been used to
 study relationships between trait distributions, geography, and climate
@@
 Condition Index - VCI) with temperature stress indicated by anomalies in
-thermal spectra (Temperature Condition Index - TCI). These indicies were
+thermal spectra (Temperature Condition Index - TCI). These indices were
 developed to create an unbiased quantification of drought across
@@
 Similarly, low TCI values indicate episdoes of high thermal stress shown
-to be negatively correlated with precipition and soil moisture
+to be negatively correlated with precipitation and soil moisture
 (AghaKouchak \emph{et al.}, 2015).
@@
 \[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\] By combining VCI
-and TCI, the VHI is distinguishes drought from other forms of vegetative
+and TCI, the VHI distinguishes drought from other forms of vegetative
 stress (Kogan, 1995b). The use of the VHI to detect drought has been
@@
 al.}, 2015), including in southern Africa, the focal region of this
-study (Figure \ref{fig:mapsdroughtexamples}). To date, the VHI has most
-often been applied for evaluating drought risk for agricultural research
-(e.g., Rojas \emph{et al.}, 2011; Kogan \emph{et al.}, 2016). But it
-also presents a tool to study seasonal patterns in the frequency of
-drought across environments and to test hypotheses about the effect of
-drought on ecological and evolutionary processes (Kerr \& Ostrovsky,
+study (e.g.~Figure \ref{fig:mapsdroughtexamples}). To date, the VHI has
+most often been applied for evaluating drought risk for agricultural
+research (e.g., Rojas \emph{et al.}, 2011; Kogan \emph{et al.}, 2016).
+But it also presents a tool to study seasonal patterns in the frequency
+of drought across environments and to test hypotheses about the effect
+of drought on ecological and evolutionary processes (Kerr \& Ostrovsky,
 2003). As such, the VHI has been applied recently to study drought
@@
 different seasons at the location of observations for \emph{Heliophila}
-records using the VHI. Specifically, we created global maps of the
-frequencies of observing drought conditions between years
-(VHI\textless{}40, NOAA) during the winter (quarter surrounding winter
-solstice), spring (quarter surrounding spring equinox), summer (quarter
-surrounding summer solstice) and fall (quarter surrounding fall equinox)
-from 1981 to 2015. From these maps, the drought frequency during the
-winter, spring, summer, and fall were extracted for the locations of all
-GBIF records.
+records using the VHI. Specifically, we created maps of the frequencies
+of observing drought conditions between years (VHI\textless{}40, NOAA)
+during the winter (quarter surrounding winter solstice), spring (quarter
+surrounding spring equinox), summer (quarter surrounding summer
+solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015
+across the range of \emph{Heliophila}. From these maps, the drought
+frequency (the number of times drought is observed divided by the total
+number of years, 34) during the winter, spring, summer, and fall were
+extracted for the locations of all GBIF records.
 
@@
 
-To avoid instances with spurious location data, we filtered raw GBIF by
-restricting our analyses to include only:
+To avoid instances with spurious location data, we filtered raw GBIF
+records by restricting our analyses to include only:
 
@@
 \emph{et al.}, 2010). Ancestral state estimation was performed in R
-using the package phytools (Revell, 2012) to genearte 10,000 stochastic
+using the package phytools (Revell, 2012) to generate 10,000 stochastic
 character maps simulated under an equal rates model of character
@@
 To evaluate the hypothesis that annual and perennial life history
-strategies reflect adaptations to alternative drought regimes, we tested
-the corresponding prediction that the observed distributions of annual
-and perennial \emph{Heliophila} species would be significantly
+strategies reflect adaptations to alternative drought regimens, we
+tested the corresponding prediction that the observed distributions of
+annual and perennial \emph{Heliophila} species would be significantly
 associated with historic drought frequency. First, we compared the
@@
 season. We also tested for a relationship between drought frequency and
-life history, season, and their interaction by an analysis of variance
+life history, season, and their interaction by analysis of variance
 while including species as a random effect. To account for variation in
@@
 2010). This statistical approach is designed to control for the
-confounding effects of geographic relationships explained by demographic
-history rather than processes of natural selection.
+confounding effects of common ancestry's influence on demographic
+features such as geospatial relationships when addressing hypotheses
+about the role of natural selection on trait distributions.
 
@@
 (b) \emph{H. deserticola}, (c) \emph{H. coronopifolia} and (d) \emph{H.
-ephemera}. Images (a,c,d) courtesy of The Bavarian Natural History
-Collections (CC BY-SA 4.0) and (b) The London Natural History Museum (CC
+ephemera}. Images (b,d,e) courtesy of The Bavarian Natural History
+Collections (CC BY-SA 4.0) and (c) The London Natural History Museum (CC
 BY 4.0). Links to images are found in the supplement.}\label{fig:phylogeny}
@@
 frequencies during each season at the observation locations of annual
-and perennial \emph{Heliophila} (t tests contrasting annuals and
-perennials, ** = p \textless{} 0.01).}\label{fig:mapsboxplots}
+and perennial \emph{Heliophila} (Post-hoc contrasts annuals and
+perennials from ANOVA, ** = p \textless{} 0.01).}\label{fig:mapsboxplots}
 \end{figure}
 
-The \emph{Heliophila} phylogeny was consistent with previous studies
-(Mummenhoff \emph{et al.}, 2005; Mandáková \emph{et al.}, 2012). Based
-on 10,000 stochastic character maps simulated under an equal rates model
-of character evolution in life history, an average of approximately
-eight changes from annual to perennial and five changes from perennial
-to annual are observed per stochastic character map (Figure
-\ref{fig:phylogeny}a). These results suggest that the ancestral state of
-\emph{Heliophila} was annual and that both charcter states have arisen
-independently mutliple times.
+The topology of the estimated \emph{Heliophila} phylogeny was consistent
+with previous studies (Mummenhoff \emph{et al.}, 2005; Mandáková
+\emph{et al.}, 2012). Based on 10,000 stochastic character maps
+simulated under an equal rates model of character evolution in life
+history, an average of approximately eight changes from annual to
+perennial and five changes from perennial to annual are observed per
+stochastic character map (Figure \ref{fig:phylogeny}a). These results
+suggest that the ancestral state of \emph{Heliophila} was annual and
+that both character states have arisen independently mutliple times.
 
@@
 across all occurrence records (all records rather than species means,
-Figure \ref{fig:mapsboxplots}g), the frequency of drought was
-significantly higher at the location of annuals during the winter (t =
-10.65, p \textless{} 0.01), spring (t = 10.73, p \textless{} 0.01),
-summer (t = 12.67, p = \textless{} 0.01), and fall (t = 15.26, p
-\textless{} 0.01). A mixed-model analysis of variace which included
-species as random effect also identified a signficant relationship
-between drought frequency and life history (p \textless{} 0.01), season
-(p \textless{} 0.01) and their interaction (p \textless{} 0.01) (Table
-\ref{tab:anovatable}). Because a comparison across all occurrence
-records does not account for variation in the number of records per
-species (Table \ref{tab:speciesmeanstable}) or species relatedness
-(Figure \ref{fig:phylogeny}a), we also tested whether mean drought
-frequency values of each species were significantly different between
-annuals and perennials using phylogenetic logistic regression. We found
-that the mean drought frequencies were significantly higher
+Figure \ref{fig:mapsboxplots}g), a mixed-model analysis of variance
+which included species as random effect revealed a significant
+relationship between drought frequency and life history (p \textless{}
+0.01), season (p \textless{} 0.01) and their interaction (p \textless{}
+0.01) (Table \ref{tab:anovatable}). Post-hoc contrasts showed that the
+frequency of drought was significantly higher at the location of annuals
+during the summer (z ratio = 3.93, p = \textless{} 0.01), and fall (z
+ratio = 4.06, p \textless{} 0.01). Because a comparison across all
+occurrence records does not account for variation in the number of
+records per species (Table \ref{tab:speciesmeanstable}) or species
+relatedness (Figure \ref{fig:phylogeny}a), we also tested whether mean
+drought frequency values of each species were significantly different
+between annuals and perennials using phylogenetic logistic regression.
+We found that the mean drought frequencies were significantly higher
 (\(\alpha = 0.05\)) in annual species during the spring, summer, and
 fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a,b). These
-findings indicate that demographic history alone does not explain
+findings indicate that common ancestry alone does not explain
 differences in the drought frequencies experienced between the
@@
 \includegraphics[width=\textwidth]{../figures/line_and_dates} \caption{(a) Comparison (mean +- SE) of drought frequency across
-seasons calcualted at the occurance locations of GBIF records of annual
+seasons calculated at the occurrence locations of GBIF records of annual
 and perennial species of \emph{Heliophila}. (b) Results from
@@
 and perennial species of \emph{Heliophila} in relation to the
-photoperiodic calender where day 1 is intermediate to the fall equinox
-and winter soltice.}\label{fig:lineplots}
+photoperiodic calendar where day 1 is intermediate to the fall equinox
+and winter solstice.}\label{fig:lineplots}
 \end{figure}
@@
 that annual species are found in environments that experience more
-frequent drought than perennial species. These findings are a valuable
-contribution to previous investigations of associations between life
-history and climate (Morishima \emph{et al.}, 1984; Evans \emph{et al.},
-2005; Datson \emph{et al.}, 2008; Cruz-Mazo \emph{et al.}, 2009). To our
-knowledge this is the first study to demonstrate a significant
-association between life history and drought in a phylogenetic context
-informed by large scale species distribution data and long term drought
-detection.
+frequent drought than perennial species, building on previous
+investigations of associations between life history and climate
+(Morishima \emph{et al.}, 1984; Evans \emph{et al.}, 2005; Datson
+\emph{et al.}, 2008; Cruz-Mazo \emph{et al.}, 2009). To our knowledge
+this is the first study to demonstrate a significant association between
+life history and drought in a phylogenetic context informed by large
+scale species distribution data and long term drought detection.
 
@@
 
-These findings empirical provide support classical theoretical
-predictions about the adaptive value of annual and perennial life
-history strategies. Taken together, they suggest that in
-\emph{Heliophila}, annual species are adapted to environments with
-increased summer droughts by avoiding these seasons in a dormant seed
-phase of their life cycle. They also suggest that perenniality is
-adaptive in environemnts where droughts are less frequent. While most
-previous work has focused on describing the evolutionary origins of
-annuality (Barrett \emph{et al.}, 1996; Conti \emph{et al.}, 1999;
-Andreasen \& Baldwin, 2001; Verboom \emph{et al.}, 2004; Friedman \&
-Rubin, 2015) there are at least a few other cases where perenniality
-appears to have arisen from an annual ancestor (Bena \emph{et al.},
-1998; Tank \& Olmstead, 2008). And while early theory predicted
-selection for annuality when adult mortality is high (Stearns, 1992), we
-also find evidence that perenniality could be explained by reduced
-frequency of drought. This is supported by the theoretical prediction
-that perenniality is advantageous is stable habitats. The phylogeny
-reveals several transitions from annual to perennial life history
-(Figure \ref{fig:phylogeny}a) and the distributions of perennial
+These findings empirically support classical theoretical predictions
+about the adaptive value of annual and perennial life history
+strategies. Taken together, they suggest that in \emph{Heliophila},
+annual species are adapted to environments with predictable droughts by
+escaping drought prone seasons during the dormant seed phase of their
+life cycle. They also suggest that perenniality is adaptive in
+environments where droughts are less frequent. While most previous work
+has focused on describing the evolutionary origins of annuality (Barrett
+\emph{et al.}, 1996; Conti \emph{et al.}, 1999; Andreasen \& Baldwin,
+2001; Verboom \emph{et al.}, 2004; Friedman \& Rubin, 2015) there are at
+least a few other cases where perenniality appears to have arisen from
+an annual ancestor (Bena \emph{et al.}, 1998; Tank \& Olmstead, 2008).
+And while early theory predicted selection for annuality when adult
+mortality is high (Stearns, 1992), we also find evidence that
+perenniality could be explained by reduced frequency of drought. This is
+supported by the theoretical prediction that perenniality is
+advantageous in stable habitats. The phylogeny reveals several
+transitions from annual to perennial life history (Figure
+\ref{fig:phylogeny}a) and the distributions of perennial
 \emph{Heliophila} extend into regions where drought frequency is low
@@
 strategies could be threatened by rapidly changing drought regimens
-(Dai, 2011). In light of the findings here, changes in eastern South
-Africa which is forecast to experience signicantly lower rainfall
-(Service \& Comission, 2017) could be particularly impactful. In this
-study we found that this region is currently dominiated by derived
-perennial species of \emph{Heliophila}. However, if drought become more
-frequent in this region, it may allow for the establishement of annuals.
-Such changes in selection patterns and shifts in plant functional
-diversity could have impacts on ecosystem functioning and processes such
-as carbon cycling (Garnier \emph{et al.}, 1997; Roumet \emph{et al.},
-2006; Monroe \emph{et al.}, 2018a). Furthermore, the changes in
-frequency of drought may be an important factor when considering the use
-of perennial cropping systems (Parry \emph{et al.}, 2005; Lelièvre \&
-Volaire, 2009).
+(Dai, 2011). In light of the findings here, forecasted reductions in
+rainfall across eastern South Africa (Service \& Comission, 2017) could
+be particularly impactful to plant community compositions. Here we found
+that this region is currently dominated by derived perennial species of
+\emph{Heliophila}. However, a scenario in which droughts become more
+frequent in this region may allow for the establishment of annuals. Such
+changes in selection patterns and shifts in plant functional diversity
+could have impacts on ecosystem functioning and processes such as carbon
+cycling (Garnier \emph{et al.}, 1997; Roumet \emph{et al.}, 2006; Monroe
+\emph{et al.}, 2018a). Furthermore, the changes in frequency of drought
+may be an important factor when considering the use of perennial
+cropping systems (Parry \emph{et al.}, 2005; Lelièvre \& Volaire, 2009).
 
 In conclusion, we find strong support for classic life history theory
-which predicts that annuality is adaptive in environments where droughts
-occur more frequently. Additionally, we report evidence consistent with
-a life history model in annuals in which they escape drought prone
-seasons during the seed phase of their life cycle. Finally, we find
-evidence that the distributions of perennial lineages may indicate a
-competitive advantage in areas where droughts are infrequent. More
-broadly, this work highlights the irreplaceable value of natural history
-collections and demonstrates the power of combining such information
-with large scale remote sensing data to address outstanding classic
-hypotheses in ecology and evolution.
+that predicts that annuality is adaptive in environments with frequency
+and predictable droughts. We report evidence consistent with a life
+history model in annuals in which they escape drought prone seasons
+during the seed phase of their life cycle. Finally, we find evidence
+that the distributions of perennial lineages may indicate a competitive
+advantage in areas where droughts are infrequent. More broadly, this
+work highlights the irreplaceable value of natural history collections
+and demonstrates the power of combining such information with large
+scale remote sensing data to address outstanding classic hypotheses in
+ecology and evolution.
 
@@
 
-We thank Jesse Lasky, members of the Sloan lab at CSU and three
-ananymous reviewers for generous feedback that improved the quality of
-this work. This research was supported by NSF Award 1701918 and
-USDA-NIFA Award 2014-38420-21801 to JGM.
+We thank Jesse Lasky, members of the Sloan lab, and three anonymous
+reviewers for generous feedback that improved the quality of this work.
+This research was supported by NSF Award 1701918 and USDA-NIFA Award
+2014-38420-21801 to JGM.
 
@@
 \newpage
-
-\begin{table}[tbp]
-\begin{center}
-\begin{threeparttable}
-\caption{\label{tab:anovatable}Analysis of variance (ANOVA) to compare drought frequency as a function of life history, season, and their interaction while including species as a random effect.}
-\begin{tabular}{lllll}
-\toprule
-predictor & \multicolumn{1}{c}{numDF} & \multicolumn{1}{c}{denDF} & \multicolumn{1}{c}{F.value} & \multicolumn{1}{c}{p.value}\\
-\midrule
-intercept & 1 & 8,720.0000 & 2,833.9567 & 0.0000\\
-life history & 1 & 40.0000 & 12.2117 & 0.0012\\
-season & 3 & 8,720.0000 & 106.4947 & 0.0000\\
-life history x season & 3 & 8,720.0000 & 4.8301 & 0.0023\\
-\bottomrule
-\end{tabular}
-\end{threeparttable}
-\end{center}
-\end{table}
 
@@
 
+\begin{table}[tbp]
+\begin{center}
+\begin{threeparttable}
+\caption{\label{tab:anovatable}Analysis of variance (ANOVA) to compare drought frequency as a function of life history, season, and their interaction while including species as a random effect.}
+\begin{tabular}{lllllll}
+\toprule
+predictor & \multicolumn{1}{c}{Sum.Sq} & \multicolumn{1}{c}{Mean.Sq} & \multicolumn{1}{c}{NumDF} & \multicolumn{1}{c}{DenDF} & \multicolumn{1}{c}{F.value} & \multicolumn{1}{c}{p-value}\\
+\midrule
+life history & 0.1715 & 0.1715 & 1 & 36.9156 & 12.2117 & 0.0013\\
+season & 4.3906 & 1.4635 & 3 & 8,718.5058 & 104.2028 & 0.0000\\
+life history x season & 0.2035 & 0.0678 & 3 & 8,718.5058 & 4.8301 & 0.0023\\
+\bottomrule
+\addlinespace
+\end{tabular}
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Type III Analysis of Variance Table with Satterthwaite's method}
+\end{tablenotes}
+\end{threeparttable}
+\end{center}
+\end{table}
+
 \begin{figure}
@@
 of drought in southern Africa. Vegetative Health Index (VHI) values
-below 40 indicate remotely sensed drought. Drought detection during this
-years is validated by precipitation based estimates of drought ocurrance
-(Standardized Precipitation Index) which confirm that while (a) 2015 was
-one of the worst drought years on record, (b) 1999 was one of the
-wettest, and (c) 2012 was typical in terms of precipitation patterns
-(Monyela, 2017). It is worth noting that drought can be detected using
-the VHI across ecosystems, including those inhabited by perennial rather
-than annual \emph{Heliophila} species.}
+below 40 indicate remotely sensed drought. Drought detection during
+these years is validated by previously reported precipitation based
+estimates of drought ocurrence (Monyela, 2017) which confirm that while
+(a) 2015 was one of the worst drought years on record, (b) 1999 was one
+of the wettest, and (c) 2012 was typical in terms of precipitation
+patterns. It is worth noting that drought can be detected using the VHI
+across ecosystems, including those inhabited by perennial rather than
+annual \emph{Heliophila} species.}
 \end{figure}

### 2019-03-31 manuscript/manuscript.R

@@ -14,6 +14,7 @@ library(ggtree)
 library(grid)
 library(png)
 library(ggimage)
+setwd("~/github/heliophila/manuscript/")
 
 ## ----raw_GBIF------------------------------------------------------------
 GBIF<-read.csv("../data/0006178-160311141623029/occurrence.csv")
@@ -249,10 +250,17 @@ tsum<-t.test(Summer~life_history, GBIF_cleaned)
 tfal<-t.test(Fall~life_history, GBIF_cleaned)
 
 melted<-GBIF_cleaned %>% select(life_history, Winter, Spring, Summer,Fall, species) %>% gather(Winter, Spring, Summer,Fall, key="season",value="droughtfreq")
-library(nlme)
-fit<-lme(droughtfreq~life_history*season, random=~1|species, melted)
-anovatable<-data.frame(predictor=c("intercept","life history","season","life history x season"), anova(fit))
-row.names(anovatable)<-NULL
+
+library(lmerTest)
+library(lme4)
+fit<-lmer(droughtfreq~life_history*season+(1|species), melted)
+anovatable<-data.frame(predictor=c("life history","season","life history x season"), anova(fit))
+
+library(emmeans)
+post_hoc<-emmeans(fit, list(pairwise ~ life_history*season), adjust = "tukey")
+post_hoc<-as.data.frame(post_hoc$`pairwise differences of life_history, season`)
+post_hoc[c(1,14,23,28), ]
+
 
 
 ## ----modelstable, results = 'asis', echo = F-----------------------------

### 2019-04-05 manuscript/manuscript.Rmd
update

@@
 \par Here, we test classic life history theory in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa, controlling for phylogeny and integrating 34 years of satellite-based drought detection with 2,192 herbaria occurrence records.
-\par We find that the common ancestor of \textit{Heliophila} species was likely an annual, and that perenniality and annuality have repeatedly evolved, eight and five times, respectively.
+\par We find that the common ancestor of \textit{Heliophila} species was likely an annual, and that perenniality and annuality have repeatedly evolved, an estimated seven and five times, respectively.
 By comparing historical drought regimens, we show that annuals occur in environments where droughts are significantly more frequent than perennial species.
@@
 To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimens, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency.
-First, we compared the frequency of drought across all occurrence records of between annual and perennial species by t-tests.
-Independent tests were performed to compare drought frequency between annuals and perennials during each season.
-We also tested for a relationship between drought frequency and life history, season, and their interaction by analysis of variance while including species as a random effect.
-To account for variation in the number of occurrence records per species, we next calculated the mean drought frequency during the winter, spring, summer and fall for each species.
+We  tested for a relationship between drought frequency and life history, season, and their interaction by analysis of variance while including species as a random effect using the lme4 package in R [@bates2014fitting] and compared annuals and perennials using by Tukey adjusted post-hoc contrasts.
+We next calculated the mean drought frequency during the winter, spring, summer and fall for each species.
 Because shared evolutionary history of closely related species can lead to spurious associations between traits and environments  [@felsenstein1985phylogenies], we tested for a relationship between life history strategy and drought frequency while controlling for phylogeny using phylogenetic logistic regression [@ives2010phylogenetic].
@@
 The topology of the estimated _Heliophila_ phylogeny was consistent with previous studies [@mummenhoff2005phylogeny;@mandakova2012whole].
-Based on 10,000 stochastic character maps simulated under an equal rates model of character evolution in life history, an average of approximately eight changes from annual to perennial and five changes from perennial to annual are observed per stochastic character map (Figure \@ref(fig:phylogeny)a).
+Based on 10,000 stochastic character maps simulated under an equal rates model of character evolution in life history, an average of approximately seven changes from annual to perennial and five changes from perennial to annual are observed per stochastic character map (Figure \@ref(fig:phylogeny)a).
 These results suggest that the ancestral state of _Heliophila_ was annual and that both character states have arisen independently mutliple times.

### 2019-04-05 manuscript/manuscript.tex
update

@@
 likely an annual, and that perenniality and annuality have repeatedly
-evolved, eight and five times, respectively. By comparing historical
-drought regimens, we show that annuals occur in environments where
-droughts are significantly more frequent than perennial species. We also
-provide evidence that annual plants adapt to predictable drought
+evolved, an estimated seven and five times, respectively. By comparing
+historical drought regimens, we show that annuals occur in environments
+where droughts are significantly more frequent than perennial species.
+We also provide evidence that annual plants adapt to predictable drought
 regimens by escaping drought prone seasons as seeds.
@@
 annual and perennial \emph{Heliophila} species would be significantly
-associated with historic drought frequency. First, we compared the
-frequency of drought across all occurrence records of between annual and
-perennial species by t-tests. Independent tests were performed to
-compare drought frequency between annuals and perennials during each
-season. We also tested for a relationship between drought frequency and
-life history, season, and their interaction by analysis of variance
-while including species as a random effect. To account for variation in
-the number of occurrence records per species, we next calculated the
-mean drought frequency during the winter, spring, summer and fall for
-each species. Because shared evolutionary history of closely related
-species can lead to spurious associations between traits and
-environments (Felsenstein, 1985), we tested for a relationship between
-life history strategy and drought frequency while controlling for
-phylogeny using phylogenetic logistic regression (Ives \& Garland,
+associated with historic drought frequency. We tested for a relationship
+between drought frequency and life history, season, and their
+interaction by analysis of variance while including species as a random
+effect using the lme4 package in R ({\textbf{???}}) and compared annuals
+and perennials using by Tukey adjusted post-hoc contrasts. We next
+calculated the mean drought frequency during the winter, spring, summer
+and fall for each species. Because shared evolutionary history of
+closely related species can lead to spurious associations between traits
+and environments (Felsenstein, 1985), we tested for a relationship
+between life history strategy and drought frequency while controlling
+for phylogeny using phylogenetic logistic regression (Ives \& Garland,
 2010). This statistical approach is designed to control for the
@@
 simulated under an equal rates model of character evolution in life
-history, an average of approximately eight changes from annual to
+history, an average of approximately seven changes from annual to
 perennial and five changes from perennial to annual are observed per

### 2019-06-08 manuscript/manuscript.Rmd
proofing

@@
     email         : "monroejg@colostate.edu"
+    phone         : "919-810-8800"
+    orcid         : 0000-0002-4025-5572
+    twitter       : "@Grey_Monroe"
   - name          : "Brian Gill"
     affiliation   : "3"
+    twitter       : "@BrianGillPhD"
+    orcid         : 0000-0001-9567-8989
   - name          : "Kathryn G. Turner"
     affiliation   : "4"
+    twitter       : "@KTInvasion"
+    orcid         : 0000-0001-8982-0301
   - name          : "John K. McKay"
     affiliation   : "2"
+    orcid         : 0000-0003-4311-5513
 
@@
 
-keywords          : "drought adaptation, herbaria records, Heliophila, life history evolution, phylogeography, remote sensing "
+keywords          : "drought adaptation, herbaria records, Heliophila, life history evolution, phylogeography, remote sensing"
 
@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output            : papaja::apa6_word
 
@@
 
-Understanding the causes and consequences of life history variation is a longstanding goal of ecology and evolutionary biology  [@cole1954population].
+Understanding the causes of life history variation and climate adaptation are longstanding goals of ecology and evolutionary biology  [@raunkiaer1905types;@turesson1925plant;@clausen1948experimental;@cole1954population].
 In  plants, life histories are especially diverse, with some species completing their life cycle in a number of weeks to others that live for thousands of years [@brown1996oldlist].
-Along this continuum in angiosperms an important division exists distinguishing annuals which complete their seed to seed life cycle within a single calendar year from perennials which can persist over multiple years.
+Along this continuum in angiosperms an important division exists distinguishing annuals which complete their seed to seed life cycle within a single year from perennials which can persist over multiple years.
 Annual plants do not need to survive through the full range of seasonal environmental variation and spend at least some portion of the year as a seed where they are relatively protected from environmental stress.
@@
 Classical theory predicts shorter life spans in environments where adult mortality is high [@charnov1973life;@stearns1992evolution;@franco1996life].
-Because lack of water is perhaps the greatest threat to survival during vegetative or reproductive growth in plants, this theory has been extended to the hypothesis that annuality is adaptive when it allows plants to escape drought [@schaffer1975selection].
-Indeed, adaptation to drought,  defined as episodes of increased aridity causing plant sress [@passioura1996drought], is often invoked as an explanation for the success of annual species.
+Because lack of water is one of the greatest threats to survival during vegetative or reproductive growth in plants, this theory has been extended to the hypothesis that annuality is adaptive when it allows plants to escape drought [@cody1975ecology;@schaffer1975selection].
+Indeed, adaptation to drought,  defined as episodes of increased aridity causing plant stress [@passioura1996drought], is often invoked as an explanation for the success of annual species, though drought adapted perennials are also well known [@raunkiaer1905types].
 And while a few cases are cited where annuality appears to be more common in environments with greater aridity [@stebbins1952aridity; @morishima1984differentiation], this hypothesis has yet to be supported while controlling for the effect of common ancestry (phylogeny) on life habit.
@@
 That is, when the frequency of drought is particularly high during certain seasons.
-Such predictability is important for selection to favor and escape strategy during those seasons which are particularly drought prone.
+Such predictability is important for selection to favor an escape strategy during those seasons which are particularly drought prone.
 While there has been at least one example of annuality associated with environments qualitatively classified as "predictable" in a general sense [@datson2008climate], the seasonal predictability of drought experienced by annuals has yet to be rigorously studied.
 As such, further empirical work is needed to support the model of annuality as a mechanism of drought adaptation via escape from drought prone seasons.
-Here we study herbarium collection dates to ask whether annuals indeed exhibit evidence of an escape strategy from seasons with elevated drought frequency.
+Here we examine herbarium collection dates to ask whether annuals indeed exhibit evidence of an escape strategy from seasons with elevated drought frequency.
 
@@
 occurring in forms other than seed) should be rare during seasons when drought frequency is highest.
-Phylogenetic relatedness can influence tests of associations between species’ traits and their environments by confounding common environments caused by selection from common environments caused my ancestry.
+Phylogenetic relatedness can influence tests of associations between species’ traits and their environments by confounding common environments caused by selection from common environments caused by ancestry.
 [@barrett1996comparative;@felsenstein1985phylogenies].
@@
 To characterize the distributions of annual (studied here, n = 21) and perennial (studied here, n = 21) _Heliophila_ species, all (`r nrow(GBIF)`) records for the genus _Heliophila_ were downloaded from the Global Biodiversity Information Facility (gbif.org) on July 21, 2018 [@gbifdownload].
-Herbaria records such as these provide a rich data sources to characterize the geographical distributions of species [@thiers2016index; @willis2017old; @lang2018using].
+Herbaria records such as these provide a rich data sources to characterize the geographical distributions of species [@thiers2016index; @willis2017old; @lang2019using].
 And as they become digitized [@soltis2017digitization], herbaria collections have been used to study relationships between trait distributions, geography, and climate [@davis2015herbarium;@wolf2016altitudinal;@stropp2016mapping;@vaclavik2017effects].
@@
 Summer<-raster("../data/drought/Winter_drought_freq")
-Fall<-raster("../data/drought/Spring_drought_freq")
+Autumn<-raster("../data/drought/Spring_drought_freq")
 
-# Winter<-raster("~/VHI/droughtfreq/summer_mild")
-# Spring<-raster("~/VHI/droughtfreq/fall_mild")
-# Summer<-raster("~/VHI/droughtfreq/winter_mild")
-# Fall<-raster("~/VHI/droughtfreq/spring_mild")
 ```
@@
 But it also presents a tool to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes [@kerr2003space].
-As such, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting intraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought].
+As such, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting intraspecific variation in drought tolerance traits and genes [@mojica2016genetics;@dittberner2018natural;@monroe2018drought].
 Here, we accessed VHI data at $16km^2$ resolution from 1981 to 2015 (https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php) to characterize the seasonal drought frequencies experienced by annual and perennial _Heliophila_ species across their native range of southern Africa.
@@
 To characterize drought regimens across the distributions of annual and perennial species of _Heliophila_, we calculated drought during different seasons at the location of observations for _Heliophila_ records using the VHI.
-Specifically, we created maps of the frequencies of observing drought conditions between years (VHI<40, NOAA) during the winter (quarter surrounding winter solstice), spring (quarter surrounding spring equinox), summer (quarter surrounding summer solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015 across the range of _Heliophila_.
-From these maps, the drought frequency (the number of times drought is observed divided by the total number of years, 34) during the winter, spring, summer, and fall were extracted for the locations of all GBIF records.
+Specifically, we created maps of the frequencies of observing drought conditions between years (VHI<40, NOAA) during the winter (quarter surrounding winter solstice), spring (quarter surrounding spring equinox), summer (quarter surrounding summer solstice) and autumn (quarter surrounding autumn equinox) from 1981 to 2015 across the range of _Heliophila_.
+From these maps, the drought frequency (the number of times drought is observed divided by the total number of years, 34) during the winter, spring, summer, and autumn were extracted for the locations of all GBIF records.
 
@@
   mutate(Summer=raster::extract(Summer, cbind(decimalLongitude, decimalLatitude))) %>%
-  mutate(Fall=raster::extract(Fall, cbind(decimalLongitude, decimalLatitude))) %>%
+  mutate(Autumn=raster::extract(Autumn, cbind(decimalLongitude, decimalLatitude))) %>%
   mutate(ymd_lat_lon_sp = paste0(year,"_",month,"_",day,"_", decimalLatitude,"_", decimalLongitude,"_", species))
@@
 
+Out of `r nrow(GBIF)` _Heliophila_ GBIF records,  `r sum(GBIF$species %in% species_life_history$species)` were for species with reported life history [@mummenhoff2005phylogeny], `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude))` had geospatial data, `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)))` did not have geospatial issues, `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)) & !(is.na(GBIF$Summer)) & !(is.na(GBIF$decimalLatitude)))` were located on pixels classified as land having drought measurements, `r sum(GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & (GBIF$species %in% species_life_history$species) & !(is.na(GBIF$Summer)))` were located in Africa, `r sum(!duplicated(GBIF$ymd_lat_lon_sp) & GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & (GBIF$species %in% species_life_history$species) & !(is.na(GBIF$Summer))) ` were not duplicated.
+
 ### Phylogeny construction and ancestral state estimation
@@
 species_means <- GBIF_cleaned %>%
-  dplyr::select(species, life_history, Winter, Summer, Spring, Fall) %>%
+  dplyr::select(species, life_history, Winter, Summer, Spring, Autumn) %>%
   group_by(species, life_history) %>%
-  summarise(n=n(), Winter_mean=mean(Winter), Spring_mean=mean(Spring), Summer_mean=mean(Summer), Fall_mean=mean(Fall)) %>% 
+  summarise(n=n(), Winter_mean=mean(Winter), Spring_mean=mean(Spring), Summer_mean=mean(Summer), Autumn_mean=mean(Autumn)) %>% 
   as.data.frame()
@@
 row.names(species_means)<-gsub(" ", "_", species_means$species)
-seasons<-c("Winter_mean", "Spring_mean", "Summer_mean","Fall_mean")
+seasons<-c("Winter_mean", "Spring_mean", "Summer_mean","Autumn_mean")
 
@@
 To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimens, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency.
-We  tested for a relationship between drought frequency and life history, season, and their interaction by analysis of variance while including species as a random effect using the lme4 package in R [@bates2014fitting] and compared annuals and perennials using by Tukey adjusted post-hoc contrasts.
-We next calculated the mean drought frequency during the winter, spring, summer and fall for each species.
+We  tested for a relationship between drought frequency and life history, season, and their interaction by analysis of variance while including species as a random effect using the lme4 package in R [@bates2014fitting] and compared annuals and perennials using Tukey adjusted post-hoc contrasts.
+We next calculated the mean drought frequency during the winter, spring, summer and autumn for each species.
 Because shared evolutionary history of closely related species can lead to spurious associations between traits and environments  [@felsenstein1985phylogenies], we tested for a relationship between life history strategy and drought frequency while controlling for phylogeny using phylogenetic logistic regression [@ives2010phylogenetic].
@@
 At internal nodes, pie charts indicate the estimated posterior probability of being annual versus perennial.
-Example herbaria specimens accessed via GBIF of (a) _H.
-minima_, (b) _H.
-deserticola_, (c) _H.
-coronopifolia_ and (d) _H.
+Example herbaria specimens accessed via GBIF of (b) _H.
+minima_, (c) _H.
+deserticola_, (d) _H.
+coronopifolia_ and (e) _H.
 ephemera_.
@@
 
-(ref:mapsboxplots) Locations of occurrence records of (a) annual and (b) perennial _Heliophila._ Drought frequency during the (c) winter, (d) spring, (e) summer and (f) fall detected using the VHI.
+(ref:mapsboxplots) Locations of occurrence records of (a) annual and (b) perennial _Heliophila._ Drought frequency during the (c) winter, (d) spring, (e) summer and (f) autumn detected using the VHI.
 (g) Drought frequencies during each season at the observation locations of annual and perennial _Heliophila_ (Post-hoc contrasts annuals and perennials from ANOVA, \*\* = p < 0.01).
+Boxplots indicate median (horizontal line), first and third quartiles (upper and lower limits of box), and 1.5 times the interquartile range (whiskers).
+Individual observations are shown by crosshairs.
+Annual and perennial species are indicated by data colored blue and oranage, respectively.
 
@@
 
-Fall_SA <-crop(Fall, rbind(c(12,37),c(-35, -18)))
+Fall_SA <-crop(Autumn, rbind(c(12,37),c(-35, -18)))
 
@@
   theme(legend.position = c(0.95,.3),legend.text=element_text(size=rel(0.7)), legend.key.size=unit(.4,"line"), legend.background = element_rect(fill=alpha('white', 0)), text = element_text(size=8))+
-  labs(title="Fall drought freq.", x="Longitude (°)", y="Latitude (°)")
+  labs(title="Autumn drought freq.", x="Longitude (°)", y="Latitude (°)")
 
 GBIF_cleaned_long<- GBIF_cleaned %>%
-  dplyr::select(species, life_history, Winter, Spring, Summer,Fall) %>%
-  gather(season, drought_freq, Winter, Spring, Summer,Fall)
+  dplyr::select(species, life_history, Winter, Spring, Summer,Autumn) %>%
+  gather(season, drought_freq, Winter, Spring, Summer,Autumn)
 
-GBIF_cleaned_long$season<-factor(GBIF_cleaned_long$season, levels=c("Winter","Spring","Summer","Fall"))
+GBIF_cleaned_long$season<-factor(GBIF_cleaned_long$season, levels=c("Winter","Spring","Summer","Autumn"))
 
 ann_text <- data.frame(x = 1.5,drought_freq = 0.8,life_history="a", lab = "**",
-                       season = factor(c("Summer","Fall"), levels=c("Winter","Spring","Summer","Fall")))
+                       season = factor(c("Summer","Autumn"), levels=c("Winter","Spring","Summer","Autumn")))
 
@@
 
-Out of `r nrow(GBIF)` _Heliophila_ GBIF records,  `r sum(GBIF$species %in% species_life_history$species)` were for species with reported life history [@mummenhoff2005phylogeny], `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude))` had geospatial data, `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)))` did not have geospatial issues, `r sum((GBIF$species %in% species_life_history$species) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & !(is.na(GBIF$decimalLatitude)) & !(is.na(GBIF$Summer)) & !(is.na(GBIF$decimalLatitude)))` were located on pixels classified as land having drought measurements, `r sum(GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & (GBIF$species %in% species_life_history$species) & !(is.na(GBIF$Summer)))` were located in Africa, `r sum(!duplicated(GBIF$ymd_lat_lon_sp) & GBIF$decimalLatitude < (-20) & GBIF$decimalLongitude < (40) & !is.na(GBIF$decimalLatitude) & !(GBIF$hasGeospatialIssues==T) & (GBIF$species %in% species_life_history$species) & !(is.na(GBIF$Summer))) ` were not duplicated.
-Thus, after all filtering steps, `r nrow(GBIF_cleaned)` records for `r length(unique(GBIF_cleaned$species))` species (Figure \@ref(fig:phylogeny), Table \@ref(tab:speciesmeanstable)) passed for further analyses.
+After all filtering steps, `r nrow(GBIF_cleaned)` records for `r length(unique(GBIF_cleaned$species))` species (Figure \@ref(fig:phylogeny), Table \@ref(tab:speciesmeanstable)) passed for further analyses.
 The number of samples varied between species, with a mean of `r mean(species_means$n)` samples per species.
@@
 
-There were clear visual differences between the distributions of the `r sum(GBIF_cleaned$life_history=="a")` annual and the `r sum(GBIF_cleaned$life_history=="p")` perennial _Heliophila_ observation records (see Figure \@ref(fig:speciesmaps) for maps of individual species).
+There were clear visual differences between the distributions of the `r sum(GBIF_cleaned$life_history=="a")` annual and the `r sum(GBIF_cleaned$life_history=="p")` perennial _Heliophila_ observation records (Figure \@ref(fig:mapsboxplots), Figure \@ref(fig:speciesmaps)).
 While annual species were generally found in the western regions of South Africa and Namibia, primarily in the Cape Floristic Region and Succulent Karoo (Figure \@ref(fig:mapsboxplots)a), the occurrence of perennials extended to the southern and eastern coast of South Africa (Figure \@ref(fig:mapsboxplots)b).
@@
 
-melted<-GBIF_cleaned %>% select(life_history, Winter, Spring, Summer,Fall, species) %>% gather(Winter, Spring, Summer,Fall, key="season",value="droughtfreq")
+melted<-GBIF_cleaned %>% select(life_history, Winter, Spring, Summer,Autumn, species) %>% gather(Winter, Spring, Summer,Autumn, key="season",value="droughtfreq")
 
@@
 post_hoc<-post_hoc[c(1,14,23,28), ]
-post_hoc$contrast<-c("Fall","Spring","Summer","Winter")
+post_hoc$contrast<-c("Autumn","Spring","Summer","Winter")
 
@@
 When comparing across all occurrence records (all records rather than species means, Figure \@ref(fig:mapsboxplots)g), a mixed-model analysis of variance which included species as random effect revealed a significant relationship between drought frequency and life history (p < 0.01), season (p < 0.01) and their interaction (p < 0.01) (Table \@ref(tab:anovatable)).
-Post-hoc contrasts showed that the frequency of drought was significantly higher at the location of annuals during the summer (z ratio = `r post_hoc$z.ratio[post_hoc$contrast=="Summer"]`, p = < 0.01), and fall (z ratio = `r post_hoc$z.ratio[post_hoc$contrast=="Fall"]`, p < 0.01).
+Post-hoc contrasts showed that the frequency of drought was significantly higher at the location of annuals during the summer (z ratio = `r post_hoc$z.ratio[post_hoc$contrast=="Summer"]`, p = < 0.01), and autumn (z ratio = `r post_hoc$z.ratio[post_hoc$contrast=="Autumn"]`, p < 0.01).
 Because a comparison across all occurrence records does not account for variation in the number of records per species (Table \@ref(tab:speciesmeanstable)) or species relatedness (Figure \@ref(fig:phylogeny)a), we also tested whether mean drought frequency values of each species were significantly different between annuals and perennials using phylogenetic logistic regression.
-We found that the mean drought frequencies were significantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and fall (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a,b).
+We found that the mean drought frequencies were significantly higher ($\alpha = 0.05$) in annual species during the spring, summer, and autumn (Table \@ref(tab:modelstable), Figure \@ref(fig:lineplots)a,b).
 These findings indicate that common ancestry alone does not explain differences in the drought frequencies experienced between the environments of annual and perennial _Heliophila_.
@@
 apa_table(modelstable
-  , caption = "Phylogenetic logistic regressions between life history, and the mean drought frequency observed at specimen sites of Heliophila species the winter, spring, summer, and fall."
+  , caption = "Phylogenetic logistic regressions between life history, and the mean drought frequency observed at specimen sites of Heliophila species the winter, spring, summer, and autumn"
   , note = "Annual species were scored as 0 and perennial species as 1.",
@@
 Annuals were scored as 0 and perennials as 1.
-(c) Collection dates of GBIF records of annual and perennial species of _Heliophila_ in relation to the photoperiodic calendar where day 1 is intermediate to the fall equinox and winter solstice.
+(c) Collection dates of GBIF records of annual and perennial species of _Heliophila_ in relation to the photoperiodic calendar where day 1 is intermediate to the autumn equinox and winter solstice.
 
@@
 life_history_drought_means<- species_means %>% 
-  dplyr::select(life_history, Winter_mean, Spring_mean, Summer_mean,Fall_mean)  %>%
+  dplyr::select(life_history, Winter_mean, Spring_mean, Summer_mean,Autumn_mean)  %>%
   gather(variable, value, -life_history) %>%
@@
   #geom_line(lwd=0.5, lty=1)+
-  facet_grid(~variable, labeller = as_labeller(c("Winter_mean"="Winter","Spring_mean"="Spring","Summer_mean"="Summer","Fall_mean"="Fall")))+
+  facet_grid(~variable, labeller = as_labeller(c("Winter_mean"="Winter","Spring_mean"="Spring","Summer_mean"="Summer","Autumn_mean"="Autumn")))+
   geom_errorbar(width=0.25,lwd=1, aes(ymin=mean-se, ymax=mean+se))+
@@
   theme_bw()+
-  scale_x_continuous(name="Day of photoperiodic year", breaks=c(45,135,225, 315), labels=c("45\n(Winter)","135\n(Spring)","225\n(Summer)","315\n(Fall)"))+
+  scale_x_continuous(name="Day of photoperiodic year", breaks=c(45,135,225, 315), labels=c("45\n(Winter)","135\n(Spring)","225\n(Summer)","315\n(Autumn)"))+
   labs(title="Timing of observation for occurrence records", y="Density")+
@@
 
-fit <-  phyloglm(species_means$life_history_num~species_means$Fall_mean, data=species_means, phy=tree)
+fit <-  phyloglm(species_means$life_history_num~species_means$Autumn_mean, data=species_means, phy=tree)
 cc <- coef(fit)
-fcurve<-plogis(cc[1]+cc[2]*sort(species_means$Fall_mean))
+fcurve<-plogis(cc[1]+cc[2]*sort(species_means$Autumn_mean))
 
@@
                 sort(species_means$Summer_mean),
-                sort(species_means$Fall_mean)),
-  season=rep(c("Winter","Spring","Summer","Fall"), each=42),
+                sort(species_means$Autumn_mean)),
+  season=rep(c("Winter","Spring","Summer","Autumn"), each=42),
   curves=c(wicurve, spcurve, scurve, fcurve),
@@
                 species_means$life_history_num[order(species_means$Summer_mean)],
-                species_means$life_history_num[order(species_means$Fall_mean)])
+                species_means$life_history_num[order(species_means$Autumn_mean)])
                            )
 curvedataframe$lifehistory<-curvedataframe$lifehistory
-curvedataframe$season<-factor(curvedataframe$season, levels=c("Winter","Spring","Summer","Fall"))
+curvedataframe$season<-factor(curvedataframe$season, levels=c("Winter","Spring","Summer","Autumn"))
 
-fit_stats<-data.frame(season=factor(c("Winter","Spring","Summer","Fall"),levels=c("Winter","Spring","Summer","Fall")), label=c("","*","**","**"), lifehistory="1")
+fit_stats<-data.frame(season=factor(c("Winter","Spring","Summer","Autumn"),levels=c("Winter","Spring","Summer","Autumn")), label=c("","*","**","**"), lifehistory="1")
 
@@
 
-The preceding results indicate that annual species are found in environments where droughts are significantly more frequent, especially in the summer and fall.
+The preceding results indicate that annual species are found in environments where droughts are significantly more frequent, especially in the summer and autumn.
 Classic life history theory hypothesizes that annuality reflects adaptation to such environments because it allows species to escape predictable stressful conditions.
-If this is the case, we would expect that annuals spend the drought prone seasons of summer and fall as seeds.
+If this is the case, we would expect that annuals spend the drought prone seasons of summer and autumn as seeds.
 To test this hypothesis, we compared the dates of occurrence records between annual and perennial _Heliophila_ species.
@@
 The differences between the distribution of collection dates were significant by all tests (ks.test D = `r  kout$statistic`, p < 0.01; bartlett.test K2 = `r bout$statistic`, p < 0.01) This is consistent with a model of life history in which annual species flower in the spring, set seed, senesce, and die before the summer.
-Thus, these annual species are likely to remain dormant during the summer and fall, when drought is the strongest predictor of the distributions of annual and perennial life histories (Figure \@ref(fig:lineplots)a).
+Thus, these annual species are likely to remain dormant during the summer and autumn, when drought is the strongest predictor of the distributions of annual and perennial life histories (Figure \@ref(fig:lineplots)a).
 
@@
 
-To test the hypothesis that annual and perennial plants reflect adaptation to alternative drought environments we examined the landscape distribution of life history strategies in the large and diverse mustard genus, _Heliophila_.
+To test the hypothesis that annual and perennial plants reflect adaptation to alternative drought environments we examined the distributions of life history strategies in the large and diverse mustard genus, _Heliophila_.
 Using metadata of `r nrow(GBIF_cleaned)` occurrence records and a 34 year dataset of satellite-detected droughts, we tested the prediction that annual species are more often observed in drought-prone locations than perennial species, when controlling for phylogenetic relatedness.
@@
 
-In conclusion, we find strong support for classic life history theory that predicts that annuality is adaptive in environments with frequency and predictable droughts.
+In conclusion, we find strong support for classic life history theory that predicts that annuality is adaptive in environments with frequent and predictable droughts.
 We report evidence consistent with a life history model in annuals in which they escape drought prone seasons during the seed phase of their life cycle.
-Finally, we find evidence that the distributions of perennial lineages may indicate a competitive advantage in areas where droughts are infrequent.
+We also find evidence that the distributions of perennial lineages may indicate a competitive advantage in areas where droughts are infrequent.
 More broadly, this work highlights the irreplaceable value of natural history collections and demonstrates the power of combining such information with large scale remote sensing data to address outstanding classic hypotheses in ecology and evolution.
@@
 
-### Images used
+### Notes S1
+Images of herbarium specimens in Figure 1. Used under creative commons licenses, permissions not required. 
 https://www.gbif.org/occurrence/1099023487  
@@
 species_means_table<-species_means_table %>% dplyr::select(-life_history_num)
-colnames(species_means_table)<-c("Species","LH","n","Winter","Spring","Summer","Fall")
+colnames(species_means_table)<-c("Species","LH","n","Winter","Spring","Summer","Autumn")
 apa_table(
@@
 )
-```
-
-```{r results = 'hide', echo = F}
-write.csv(species_means_table, "../data/species_means_table.csv")
 ```

### 2019-06-08 manuscript/manuscript.tex
proofing

@@
 \textsuperscript{1} Graduate Degree Program in Ecology, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{2} College of Agriculture, Colorado State University, Fort Collins, CO 80521, USA\\\textsuperscript{3} Institute for Environment and Society, Brown University, Providence, RI 02912, USA\\\textsuperscript{4} Biology Department, Pennsylvania State University, State College, PA 16802, USA}
-\keywords{drought adaptation, herbaria records, Heliophila, life history evolution, phylogeography, remote sensing }
+\keywords{drought adaptation, herbaria records, Heliophila, life history evolution, phylogeography, remote sensing}
 \usepackage{csquotes}
@@
 
-Understanding the causes and consequences of life history variation is a
-longstanding goal of ecology and evolutionary biology (Cole, 1954). In
-plants, life histories are especially diverse, with some species
-completing their life cycle in a number of weeks to others that live for
-thousands of years (Brown, 1996). Along this continuum in angiosperms an
-important division exists distinguishing annuals which complete their
-seed to seed life cycle within a single calendar year from perennials
-which can persist over multiple years. Annual plants do not need to
-survive through the full range of seasonal environmental variation and
-spend at least some portion of the year as a seed where they are
-relatively protected from environmental stress. In contrast, perennial
-plants can continue vegetative growth over multiple years and must
-survive conditions experienced during all seasons but can also benefit
-from competitive advantages and, if iteroparous, multiple bouts of
-reproduction. These represent fundamentally different life history
+Understanding the causes of life history variation and climate
+adaptation are longstanding goals of ecology and evolutionary biology
+(Raunkiaer, 1905; Turesson, 1925; Clausen \emph{et al.}, 1948; Cole,
+1954). In plants, life histories are especially diverse, with some
+species completing their life cycle in a number of weeks to others that
+live for thousands of years (Brown, 1996). Along this continuum in
+angiosperms an important division exists distinguishing annuals which
+complete their seed to seed life cycle within a single year from
+perennials which can persist over multiple years. Annual plants do not
+need to survive through the full range of seasonal environmental
+variation and spend at least some portion of the year as a seed where
+they are relatively protected from environmental stress. In contrast,
+perennial plants can continue vegetative growth over multiple years and
+must survive conditions experienced during all seasons but can also
+benefit from competitive advantages and, if iteroparous, multiple bouts
+of reproduction. These represent fundamentally different life history
 strategies and predicting their occurrence is important for community,
@@
 mortality is high (Charnov \& Schaffer, 1973; Stearns, 1992; Franco \&
-Silvertown, 1996). Because lack of water is perhaps the greatest threat
+Silvertown, 1996). Because lack of water is one of the greatest threats
 to survival during vegetative or reproductive growth in plants, this
 theory has been extended to the hypothesis that annuality is adaptive
-when it allows plants to escape drought (Schaffer \& Gadgil, 1975).
-Indeed, adaptation to drought, defined as episodes of increased aridity
-causing plant sress (Passioura, 1996), is often invoked as an
-explanation for the success of annual species. And while a few cases are
-cited where annuality appears to be more common in environments with
-greater aridity (Stebbins Jr, 1952; Morishima \emph{et al.}, 1984), this
-hypothesis has yet to be supported while controlling for the effect of
-common ancestry (phylogeny) on life habit. In one previous study where
-this question was addressed phylogenetically, (Evans \emph{et al.},
-2005) annuals were not found to be associated with environments that
-experience more drought. This could be explained by the relatively small
-number of species studied and the reliance on a limited number of
-weather stations to characterize environments, highlighting the need to
-develop more scalable methods to study the geographic distributions of
-traits such as life history. Thus, in this study we leverage thousands
-of herbaria specimens among dozens of species and high-resolution remote
-sensing to study the distributions and environmental factors potentially
-driving the evolution and distribution of life history.
+when it allows plants to escape drought (Cody \emph{et al.}, 1975;
+Schaffer \& Gadgil, 1975). Indeed, adaptation to drought, defined as
+episodes of increased aridity causing plant stress (Passioura, 1996), is
+often invoked as an explanation for the success of annual species,
+though drought adapted perennials are also well known (Raunkiaer, 1905).
+And while a few cases are cited where annuality appears to be more
+common in environments with greater aridity (Stebbins Jr, 1952;
+Morishima \emph{et al.}, 1984), this hypothesis has yet to be supported
+while controlling for the effect of common ancestry (phylogeny) on life
+habit. In one previous study where this question was addressed
+phylogenetically, (Evans \emph{et al.}, 2005) annuals were not found to
+be associated with environments that experience more drought. This could
+be explained by the relatively small number of species studied and the
+reliance on a limited number of weather stations to characterize
+environments, highlighting the need to develop more scalable methods to
+study the geographic distributions of traits such as life history. Thus,
+in this study we leverage thousands of herbaria specimens among dozens
+of species and high-resolution remote sensing to study the distributions
+and environmental factors potentially driving the evolution and
+distribution of life history.
 
@@
 drought is particularly high during certain seasons. Such predictability
-is important for selection to favor and escape strategy during those
+is important for selection to favor an escape strategy during those
 seasons which are particularly drought prone. While there has been at
@@
 mechanism of drought adaptation via escape from drought prone seasons.
-Here we study herbarium collection dates to ask whether annuals indeed
+Here we examine herbarium collection dates to ask whether annuals indeed
 exhibit evidence of an escape strategy from seasons with elevated
@@
 and their environments by confounding common environments caused by
-selection from common environments caused my ancestry. (Felsenstein,
+selection from common environments caused by ancestry. (Felsenstein,
 1985; Barrett \emph{et al.}, 1996). Therefore, we assessed the
@@
 surrounding spring equinox), summer (quarter surrounding summer
-solstice) and fall (quarter surrounding fall equinox) from 1981 to 2015
-across the range of \emph{Heliophila}. From these maps, the drought
+solstice) and autumn (quarter surrounding autumn equinox) from 1981 to
+2015 across the range of \emph{Heliophila}. From these maps, the drought
 frequency (the number of times drought is observed divided by the total
-number of years, 34) during the winter, spring, summer, and fall were
+number of years, 34) during the winter, spring, summer, and autumn were
 extracted for the locations of all GBIF records.
@@
 
+Out of 8670 \emph{Heliophila} GBIF records, 6634 were for species with
+reported life history (Mummenhoff \emph{et al.}, 2005), 2856 had
+geospatial data, 2833 did not have geospatial issues, 2684 were located
+on pixels classified as land having drought measurements, 2543 were
+located in Africa, 2192 were not duplicated.
+
 \hypertarget{phylogeny-construction-and-ancestral-state-estimation}{%
@@
 interaction by analysis of variance while including species as a random
-effect using the lme4 package in R ({\textbf{???}}) and compared annuals
-and perennials using by Tukey adjusted post-hoc contrasts. We next
-calculated the mean drought frequency during the winter, spring, summer
-and fall for each species. Because shared evolutionary history of
-closely related species can lead to spurious associations between traits
-and environments (Felsenstein, 1985), we tested for a relationship
-between life history strategy and drought frequency while controlling
-for phylogeny using phylogenetic logistic regression (Ives \& Garland,
-2010). This statistical approach is designed to control for the
-confounding effects of common ancestry's influence on demographic
+effect using the lme4 package in R (Bates \emph{et al.}, 2014) and
+compared annuals and perennials using Tukey adjusted post-hoc contrasts.
+We next calculated the mean drought frequency during the winter, spring,
+summer and autumn for each species. Because shared evolutionary history
+of closely related species can lead to spurious associations between
+traits and environments (Felsenstein, 1985), we tested for a
+relationship between life history strategy and drought frequency while
+controlling for phylogeny using phylogenetic logistic regression (Ives
+\& Garland, 2010). This statistical approach is designed to control for
+the confounding effects of common ancestry's influence on demographic
 features such as geospatial relationships when addressing hypotheses
@@
 the estimated posterior probability of being annual versus perennial.
-Example herbaria specimens accessed via GBIF of (a) \emph{H. minima},
-(b) \emph{H. deserticola}, (c) \emph{H. coronopifolia} and (d) \emph{H.
+Example herbaria specimens accessed via GBIF of (b) \emph{H. minima},
+(c) \emph{H. deserticola}, (d) \emph{H. coronopifolia} and (e) \emph{H.
 ephemera}. Images (b,d,e) courtesy of The Bavarian Natural History
@@
 perennial \emph{Heliophila.} Drought frequency during the (c) winter,
-(d) spring, (e) summer and (f) fall detected using the VHI. (g) Drought
-frequencies during each season at the observation locations of annual
-and perennial \emph{Heliophila} (Post-hoc contrasts annuals and
+(d) spring, (e) summer and (f) autumn detected using the VHI. (g)
+Drought frequencies during each season at the observation locations of
+annual and perennial \emph{Heliophila} (Post-hoc contrasts annuals and
 perennials from ANOVA, ** = p \textless{} 0.01).}\label{fig:mapsboxplots}
@@
 
-Out of 8670 \emph{Heliophila} GBIF records, 6634 were for species with
-reported life history (Mummenhoff \emph{et al.}, 2005), 2856 had
-geospatial data, 2833 did not have geospatial issues, 2684 were located
-on pixels classified as land having drought measurements, 2543 were
-located in Africa, 2192 were not duplicated. Thus, after all filtering
-steps, 2192 records for 42 species (Figure \ref{fig:phylogeny}, Table
-\ref{tab:speciesmeanstable}) passed for further analyses. The number of
-samples varied between species, with a mean of 52.19 samples per
-species. \emph{H. rigidiuscula} had the most records, 201, and \emph{H.
-cornellsbergia} the fewest, 2 (Table \ref{tab:speciesmeanstable}).
+After all filtering steps, 2192 records for 42 species (Figure
+\ref{fig:phylogeny}, Table \ref{tab:speciesmeanstable}) passed for
+further analyses. The number of samples varied between species, with a
+mean of 52.19 samples per species. \emph{H. rigidiuscula} had the most
+records, 201, and \emph{H. cornellsbergia} the fewest, 2 (Table
+\ref{tab:speciesmeanstable}).
 
 There were clear visual differences between the distributions of the 960
-annual and the 1232 perennial \emph{Heliophila} observation records (see
-Figure \ref{fig:speciesmaps} for maps of individual species). While
+annual and the 1232 perennial \emph{Heliophila} observation records
+(Figure \ref{fig:mapsboxplots}, Figure \ref{fig:speciesmaps}). While
 annual species were generally found in the western regions of South
@@
 frequency of drought was significantly higher at the location of annuals
-during the summer (z ratio = 3.93, p = \textless{} 0.01), and fall (z
+during the summer (z ratio = 3.93, p = \textless{} 0.01), and autumn (z
 ratio = 4.06, p \textless{} 0.01). Because a comparison across all
@@
 (\(\alpha = 0.05\)) in annual species during the spring, summer, and
-fall (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a,b). These
-findings indicate that common ancestry alone does not explain
+autumn (Table \ref{tab:modelstable}, Figure \ref{fig:lineplots}a,b).
+These findings indicate that common ancestry alone does not explain
 differences in the drought frequencies experienced between the
@@
 \begin{threeparttable}
-\caption{\label{tab:modelstable}Phylogenetic logistic regressions between life history, and the mean drought frequency observed at specimen sites of Heliophila species the winter, spring, summer, and fall.}
+\caption{\label{tab:modelstable}Phylogenetic logistic regressions between life history, and the mean drought frequency observed at specimen sites of Heliophila species the winter, spring, summer, and autumn}
 \begin{tabular}{lll}
@@
 and perennial species of \emph{Heliophila} in relation to the
-photoperiodic calendar where day 1 is intermediate to the fall equinox
+photoperiodic calendar where day 1 is intermediate to the autumn equinox
 and winter solstice.}\label{fig:lineplots}
@@
 environments where droughts are significantly more frequent, especially
-in the summer and fall. Classic life history theory hypothesizes that
+in the summer and autumn. Classic life history theory hypothesizes that
 annuality reflects adaptation to such environments because it allows
@@
 we would expect that annuals spend the drought prone seasons of summer
-and fall as seeds. To test this hypothesis, we compared the dates of
+and autumn as seeds. To test this hypothesis, we compared the dates of
 occurrence records between annual and perennial \emph{Heliophila}
@@
 in the spring, set seed, senesce, and die before the summer. Thus, these
-annual species are likely to remain dormant during the summer and fall,
-when drought is the strongest predictor of the distributions of annual
-and perennial life histories (Figure \ref{fig:lineplots}a).
+annual species are likely to remain dormant during the summer and
+autumn, when drought is the strongest predictor of the distributions of
+annual and perennial life histories (Figure \ref{fig:lineplots}a).
 
@@
 To test the hypothesis that annual and perennial plants reflect
-adaptation to alternative drought environments we examined the landscape
-distribution of life history strategies in the large and diverse mustard
-genus, \emph{Heliophila}. Using metadata of 2192 occurrence records and
-a 34 year dataset of satellite-detected droughts, we tested the
-prediction that annual species are more often observed in drought-prone
-locations than perennial species, when controlling for phylogenetic
-relatedness. We found that drought frequency is significantly different
-between the distributions of annual and perennial species, with annuals
-being found in environments with more frequent drought, and that this
-signal is strongest during the seasons when annuals are likely escaping
-via seed dormancy. These results remain significant while controlling
-for the phylogenetic relationships of \emph{Heliophila} species,
-yielding support for the role that natural selection has played in
-driving contemporary distributions of these alternatives strategies in
-relation to drought regimens.
+adaptation to alternative drought environments we examined the
+distributions of life history strategies in the large and diverse
+mustard genus, \emph{Heliophila}. Using metadata of 2192 occurrence
+records and a 34 year dataset of satellite-detected droughts, we tested
+the prediction that annual species are more often observed in
+drought-prone locations than perennial species, when controlling for
+phylogenetic relatedness. We found that drought frequency is
+significantly different between the distributions of annual and
+perennial species, with annuals being found in environments with more
+frequent drought, and that this signal is strongest during the seasons
+when annuals are likely escaping via seed dormancy. These results remain
+significant while controlling for the phylogenetic relationships of
+\emph{Heliophila} species, yielding support for the role that natural
+selection has played in driving contemporary distributions of these
+alternatives strategies in relation to drought regimens.
 
@@
 In conclusion, we find strong support for classic life history theory
-that predicts that annuality is adaptive in environments with frequency
+that predicts that annuality is adaptive in environments with frequent
 and predictable droughts. We report evidence consistent with a life
 history model in annuals in which they escape drought prone seasons
-during the seed phase of their life cycle. Finally, we find evidence
-that the distributions of perennial lineages may indicate a competitive
+during the seed phase of their life cycle. We also find evidence that
+the distributions of perennial lineages may indicate a competitive
 advantage in areas where droughts are infrequent. More broadly, this
@@
 
+\leavevmode\hypertarget{ref-bates2014fitting}{}%
+\textbf{\textnormal{Bates D}, \textnormal{Mächler M}, \textnormal{Bolker
+B}, \textnormal{Walker S}}. \textbf{2014}. Fitting linear mixed-effects
+models using lme4. \emph{arXiv preprint arXiv:1406.5823}.
+
 \leavevmode\hypertarget{ref-bena1998molecular}{}%
@@
 
+\leavevmode\hypertarget{ref-clausen1948experimental}{}%
+\textbf{\textnormal{Clausen J}, \textnormal{Keck DD}, \textnormal{Hiesey
+WM}, \textnormal{others}}. \textbf{1948}. Experimental studies on the
+nature of species. III. Environresponses of climatic races of achillea.
+\emph{Experimental studies on the nature of species. III.
+Environresponses of climatic races of Achillea.}
+
+\leavevmode\hypertarget{ref-cody1975ecology}{}%
+\textbf{\textnormal{Cody ML}, \textnormal{MacArthur RH},
+\textnormal{Diamond JM}, \textnormal{others}}. \textbf{1975}.
+\emph{Ecology and evolution of communities}. Harvard University Press.
+
 \leavevmode\hypertarget{ref-cole1954population}{}%
@@
 \textbf{\textnormal{Dittberner H}, \textnormal{Korte A},
-\textnormal{Mettler-Altmann T}, \textnormal{Weber A}, \textnormal{Monroe
-G}, \textnormal{Meaux J de}}. \textbf{2018}. Natural variation in
-stomata size contributes to the local adaptation of water-use efficiency
-in arabidopsis thaliana. \emph{bioRxiv}: 253021.
+\textnormal{Mettler-Altmann T}, \textnormal{Weber AP},
+\textnormal{Monroe G}, \textnormal{Meaux J de}}. \textbf{2018}. Natural
+variation in stomata size contributes to the local adaptation of
+water-use efficiency in arabidopsis thaliana. \emph{Molecular ecology}
+\textbf{27}: 4052--4065.
 
@@
 tolerance. \emph{Plant growth regulation} \textbf{20}: 79--83.
+
+\leavevmode\hypertarget{ref-raunkiaer1905types}{}%
+\textbf{\textnormal{Raunkiaer CC}}. \textbf{1905}. \emph{Types
+biologiques pour la géographie botanique}.
 
@@
 org/ih}.
+
+\leavevmode\hypertarget{ref-turesson1925plant}{}%
+\textbf{\textnormal{Turesson G}}. \textbf{1925}. The plant species in
+relation to habitat and climate: Contributions to the knowledge of
+genecological units. \emph{Hereditas} \textbf{6}: 147--236.
 
@@
 \toprule
-Species & \multicolumn{1}{c}{LH} & \multicolumn{1}{c}{n} & \multicolumn{1}{c}{Winter} & \multicolumn{1}{c}{Spring} & \multicolumn{1}{c}{Summer} & \multicolumn{1}{c}{Fall}\\
+Species & \multicolumn{1}{c}{LH} & \multicolumn{1}{c}{n} & \multicolumn{1}{c}{Winter} & \multicolumn{1}{c}{Spring} & \multicolumn{1}{c}{Summer} & \multicolumn{1}{c}{Autumn}\\
 \midrule

### 2019-06-08 manuscript/manuscript.R

@@ -1,496 +1,0 @@
-## ----setup, include = FALSE----------------------------------------------
-library("papaja")
-library(raster)
-library(tidyverse)
-library(ggplot2)
-library(logistf)
-library(geiger) 
-library(ape) 
-library(phylolm)
-library(gridExtra)
-library(cowplot)
-library(rasterVis)
-library(ggtree)
-library(grid)
-library(png)
-library(ggimage)
-setwd("~/github/heliophila/manuscript/")
-
-## ----raw_GBIF------------------------------------------------------------
-GBIF<-read.csv("../data/0006178-160311141623029/occurrence.csv")
-
-## ----Droughtdata---------------------------------------------------------
-Winter<-raster("../data/drought/Summer_drought_freq")
-Spring<-raster("../data/drought/Fall_drought_freq")
-Summer<-raster("../data/drought/Winter_drought_freq")
-Fall<-raster("../data/drought/Spring_drought_freq")
-
-# Winter<-raster("~/VHI/droughtfreq/summer_mild")
-# Spring<-raster("~/VHI/droughtfreq/fall_mild")
-# Summer<-raster("~/VHI/droughtfreq/winter_mild")
-# Fall<-raster("~/VHI/droughtfreq/spring_mild")
-
-## ----cleanup-------------------------------------------------------------
-species_life_history<-read.csv("../data/Heliophila_life_history.csv")
-species_life_history$species<-gsub("_", " ", species_life_history$species)
-species_life_history$species <- as.factor(species_life_history$species)
-
-GBIF <- GBIF %>%
-  mutate(Winter=raster::extract(Winter, cbind(decimalLongitude, decimalLatitude))) %>%
-  mutate(Spring=raster::extract(Spring, cbind(decimalLongitude, decimalLatitude))) %>%
-  mutate(Summer=raster::extract(Summer, cbind(decimalLongitude, decimalLatitude))) %>%
-  mutate(Fall=raster::extract(Fall, cbind(decimalLongitude, decimalLatitude))) %>%
-  mutate(ymd_lat_lon_sp = paste0(year,"_",month,"_",day,"_", decimalLatitude,"_", decimalLongitude,"_", species))
-  
-GBIF_cleaned <- GBIF %>%
[…]

### 2019-06-22 manuscript/manuscript.tex
minor spelling

@@
 characterize the geographical distributions of species (Thiers, 2016;
-Willis \emph{et al.}, 2017; Lang \emph{et al.}, 2018). And as they
+Willis \emph{et al.}, 2017; Lang \emph{et al.}, 2019). And as they
 become digitized (Soltis, 2017), herbaria collections have been used to
@@
 annual and perennial \emph{Heliophila} (Post-hoc contrasts annuals and
-perennials from ANOVA, ** = p \textless{} 0.01).}\label{fig:mapsboxplots}
+perennials from ANOVA, ** = p \textless{} 0.01). Boxplots indicate
+median (horizontal line), first and third quartiles (upper and lower
+limits of box), and 1.5 times the interquartile range (whiskers).
+Individual observations are shown by crosshairs. Annual and perennial
+species are indicated by data colored blue and oranage, respectively.}\label{fig:mapsboxplots}
 \end{figure}
@@
 Intercept & 7.0162 & 0.0082\\
-Fall drought freq. & -20.8174 & 0.0067\\ \midrule
+Autumn drought freq. & -20.8174 & 0.0067\\ \midrule
 \bottomrule
@@
 (Dai, 2011). In light of the findings here, forecasted reductions in
-rainfall across eastern South Africa (Service \& Comission, 2017) could
-be particularly impactful to plant community compositions. Here we found
+rainfall across eastern South Africa (SAWS \& WRC, 2017) could be
+particularly impactful to plant community compositions. Here we found
 that this region is currently dominated by derived perennial species of
@@
 \textbf{\textnormal{Clausen J}, \textnormal{Keck DD}, \textnormal{Hiesey
-WM}, \textnormal{others}}. \textbf{1948}. Experimental studies on the
-nature of species. III. Environresponses of climatic races of achillea.
-\emph{Experimental studies on the nature of species. III.
-Environresponses of climatic races of Achillea.}
+WM}}. \textbf{1948}. Experimental studies on the nature of species. III.
+Environresponses of climatic races of achillea. \emph{Experimental
+studies on the nature of species. III. Environresponses of climatic
+races of Achillea.}
 
@@
 \textbf{\textnormal{Cody ML}, \textnormal{MacArthur RH},
-\textnormal{Diamond JM}, \textnormal{others}}. \textbf{1975}.
-\emph{Ecology and evolution of communities}. Harvard University Press.
+\textnormal{Diamond JM}}. \textbf{1975}. \emph{Ecology and evolution of
+communities}. Harvard University Press.
 
@@
 
-\leavevmode\hypertarget{ref-lang2018using}{}%
+\leavevmode\hypertarget{ref-lang2019using}{}%
 \textbf{\textnormal{Lang PL}, \textnormal{Willems FM},
 \textnormal{Scheepens J}, \textnormal{Burbano HA}, \textnormal{Bossdorf
-O}}. \textbf{2018}. \emph{Using herbaria to study global environmental
-change}. PeerJ Preprints.
+O}}. \textbf{2019}. Using herbaria to study global environmental change.
+\emph{New Phytologist} \textbf{221}: 110--122.
 
@@
 
-\leavevmode\hypertarget{ref-Mojica2016}{}%
+\leavevmode\hypertarget{ref-mojica2016genetics}{}%
 \textbf{\textnormal{Mojica JP}, \textnormal{Mullen J},
@@
 \textnormal{Oakley CG}, \textnormal{McKay JK}}. \textbf{2016}. Genetics
-of water use physiology in locally adapted Arabidopsis thaliana.
-\emph{Plant Science}.
+of water use physiology in locally adapted arabidopsis thaliana.
+\emph{Plant Science} \textbf{251}: 12--22.
 
@@
 \textnormal{Lovell J}, \textnormal{McKay J}}. \textbf{2018b}. Drought
-adaptation in nature by extensive genetic loss-of-function.
-\emph{eLife}: DOI: 10.7554/eLife.41038.
+adaptation in arabidopsis thaliana by extensive genetic
+loss-of-function. \emph{eLife}: DOI: 10.7554/eLife.41038.
 
@@
 of endemic brassicaceae genera in the cape flora of southern africa.
-\emph{Annals of the Missouri Botanical Garden}: 400--424.
+\emph{Annals of the Missouri Botanical Garden} \textbf{92}: 400--424.
 
@@
 \leavevmode\hypertarget{ref-raunkiaer1905types}{}%
-\textbf{\textnormal{Raunkiaer CC}}. \textbf{1905}. \emph{Types
-biologiques pour la géographie botanique}.
+\textbf{\textnormal{Raunkiaer CC}}. \textbf{1905}. Types biologiques
+pour la géographie botanique. \emph{Oversigt over Det Kongelige Danske
+Videnskabernes Selskabs Forhandlinger}: 347--438.
 
@@
 
+\leavevmode\hypertarget{ref-south2017a}{}%
+\textbf{\textnormal{SAWS}, \textnormal{WRC}}. \textbf{2017}. \emph{A
+climate change reference atlas}.
+
 \leavevmode\hypertarget{ref-sayre2013new}{}%
@@
 \textnormal{Larwanou M}, \textnormal{Kelbessa E}, \textnormal{Bekele T},
-\textnormal{Kehl H} \emph{et al.}} \textbf{2013}. A new map of
-standardized terrestrial ecosystems of africa. \emph{African
-Geographical Review}.
+\textnormal{Kehl H et a}}. \textbf{2013}. \emph{A new map of
+standardized terrestrial ecosystems of africa}.
 
@@
 \textbf{\textnormal{Schaffer W}, \textnormal{Gadgil M}}. \textbf{1975}.
-Selection for optimal life histories in plants. \emph{Ecology and
-evolution of communities.}: 142--157.
-
-\leavevmode\hypertarget{ref-south2017a}{}%
-\textbf{\textnormal{Service SAW}, \textnormal{Comission WR}}.
-\textbf{2017}. \emph{A climate change reference atlas}.
+Selection for optimal life histories in plants. In: Cody M, Diamond J,
+eds. The ecology and evolution of communities. Cambridge, MA: Belknap,
+142--157.
 
@@
 \textbf{\textnormal{Stearns SC}}. \textbf{1992}. \emph{The evolution of
-life histories}.
+life histories}. Oxford: Oxford University Press.
 
@@
 \textnormal{Gallinat AS}, \textnormal{Yost JM}, \textnormal{Nelson G},
-\textnormal{Mazer SJ}, \textnormal{Rossington NL} \emph{et al.}}
-\textbf{2017}. Old plants, new tricks: Phenological research using
-herbarium specimens. \emph{Trends in ecology \& evolution} \textbf{32}:
-531--546.
+\textnormal{Mazer SJ}, \textnormal{Rossington NL et a}}. \textbf{2017}.
+Old plants, new tricks: Phenological research using herbarium specimens.
+\emph{Trends in ecology \& evolution} \textbf{32}: 531--546.
 
@@
 
-\hypertarget{images-used}{%
-\subsubsection{Images used}\label{images-used}}
-
+\hypertarget{notes-s1}{%
+\subsubsection{Notes S1}\label{notes-s1}}
+
+Images of herbarium specimens in Figure 1. Used under creative commons
+licenses, permissions not required.
 \url{https://www.gbif.org/occurrence/1099023487}\\

### 2019-06-22 src/ancestral state.R

@@ -4,6 +4,7 @@ require(phytools)
 require(ape)
 require(geiger)
 require(extRemes)
+#setwd("manuscript/") #unhash if running outisde of manuscript build
 
 tree <- read.nexus("../data/TrimmedMCCwoPolytomies.nex")
 ladderize(tree) -> tree

### 2019-08-19 manuscript/manuscript.Rmd
minor spelling
moved example VHI datasets to data folder

@@
 classoption       : "man"
-output            : papaja::apa6_word
+output            : papaja::apa6_pdf
 
@@
 
-drought<-raster::raster("~/VHI/raster/VHP.G16.C07.npp.P2015050.VH.nc.grd")
+drought<-raster::raster("../data/VHP.G16.C07.npp.P2015050.VH.nc.grd")
 drought <-crop(drought, rbind(c(12,37),c(-35, -18)))
@@
 
-nondrought<-raster::raster("~/VHI/raster/VHP.G16.C07.NJ.P1999050.VH.hdf.h5.grd")
+nondrought<-raster::raster("../data/VHP.G16.C07.NJ.P1999050.VH.hdf.h5.grd")
 nondrought <-crop(nondrought, rbind(c(12,37),c(-35, -18)))
@@
 
-typical<-raster::raster("~/VHI/raster/VHP.G16.C07.NP.P2012050.VH.hdf.h5.grd")
+typical<-raster::raster("../data/VHP.G16.C07.NP.P2012050.VH.hdf.h5.grd")
 typical <-crop(typical, rbind(c(12,37),c(-35, -18)))
